### 2018-03-22 morus.tex
tex stub with figure

\documentclass[a4paper, 11pt]{article}

\usepackage{geometry}
\usepackage{amsmath,amssymb}
\usepackage{ifthen}

\usepackage{tikz}
\usetikzlibrary{calc,math}

\makeatletter
\pgfdeclareshape{oplus}{%{{{
  \inheritsavedanchors[from=circle]
  \inheritanchorborder[from=circle]
  \foreach \s in {center,mid,base,text, north,south,west,east,
                  mid west,mid east,base west,base east,
                  north west,south west,north east,south east} {
    \inheritanchor[from=circle]{\s}
  }
  \backgroundpath{%
    \pgfutil@tempdima=\radius%
    \pgfmathsetlength{\pgf@xb}{\pgfkeysvalueof{/pgf/outer xsep}}%  
    \pgfmathsetlength{\pgf@yb}{\pgfkeysvalueof{/pgf/outer ysep}}%  
    \ifdim\pgf@xb<\pgf@yb%
      \advance\pgfutil@tempdima by-\pgf@yb%
    \else%
      \advance\pgfutil@tempdima by-\pgf@xb%
    \fi%
    \pgfpathcircle{\centerpoint}{\pgfutil@tempdima}%
    % north-south
    \centerpoint\advance\pgf@y by\radius  \pgf@xa=\pgf@x \pgf@ya=\pgf@y
    \centerpoint\advance\pgf@y by-\radius \pgf@xb=\pgf@x \pgf@yb=\pgf@y
    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
    % east-west
    \centerpoint\advance\pgf@x by\radius  \pgf@xa=\pgf@x \pgf@ya=\pgf@y
    \centerpoint\advance\pgf@x by-\radius \pgf@xb=\pgf@x \pgf@yb=\pgf@y
    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
  }
}%}}}
\makeatother
\tikzset{op/.style={draw, fill=none, minimum size=1.25ex, inner sep=0pt}}
\tikzset{opxor/.style={op, shape=oplus}}
\tikzset{opand/.style={op, rectangle, rounded corners=2pt, minimum height=3ex}}
\tikzset{oprot/.style={op, rectangle, rounded corners=2pt, inner sep=2pt, minimum width=3ex, font=\scriptsize}}
\tikzset{optee/.style={shape=circle, fill, draw, inner sep=0pt, minimum size=2pt}}
\tikzset{next/.style={->, >=latex}}

\newcommand{\cipher}[1]{\textsf{#1}}

\newif\ifsubstates\substatesfalse

\begin{document}

\begin{figure}
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
  \ifsubstates \pgfmathsetmacro{\roundsep}{1.25}
  \else        \pgfmathsetmacro{\roundsep}{0.80} \fi
  \pgfmathsetmacro{\opoffset}{.1}

  \ifsubstates
    \foreach \r in {-2,...,4} {
      \foreach \w in {0,...,4} {
        \draw[thick] (\w-.5,-\r*\roundsep) -- ++(0,-.25);
        \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W\r\w) at (\w,-\r*\roundsep-.125) {};
      }
      \draw[thick] (-.5,-\r*\roundsep) node[below left, inner sep=0pt, xshift=-3pt] {$S_{\r,*}^{\r}$} rectangle ++(5,-.25);
    }
    \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W-1-1) at (-1,--1*\roundsep-.125) {};
    \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W-2-1) at (-2,--1*\roundsep-.125) {};
  \else
    \foreach \r in {-1,...,4} {
      \foreach \w in {0,...,4} {
        \coordinate (W\r\w) at (\w,-\r*\roundsep-.125);
      }
    }
    \foreach \w in {0,...,4} {
      \coordinate (W-2\w) at (\w,--2.375*\roundsep-.125);
    }
    \coordinate (W-1-1) at (-1,--1*\roundsep-.125);
    \coordinate (W-2-1) at (-2,--1*\roundsep-.125);
  \fi

  \foreach \r/\rot in {-1/0,0/5,1/31,2/7,3/22,4/13} { % round
    \ifthenelse{\equal{\r}{-1}}{
      \pgfmathsetmacro{\txorx}{int(mod(\r+2,5))}
      \pgfmathsetmacro{\tanAx}{int(mod(\r+4,5))}
      \pgfmathsetmacro{\tanBx}{int(mod(\r+3,5))}
      \pgfmathsetmacro{\tlllx}{int(mod(\r+1,5))}
    }{
      \pgfmathsetmacro{\txorx}{int(mod(\r+3,5))}
      \pgfmathsetmacro{\tanAx}{int(mod(\r+2,5))}
      \pgfmathsetmacro{\tanBx}{int(mod(\r+1,5))}
    }
    \pgfmathsetmacro{\rprev}{int(\r-1)}
    \ifthenelse{\equal{\r}{-1}}%
    { \node[opxor] (lll\r) at ($(W\r\r) +(0,\opoffset+.25)$)  {}; }%
    { \node[oprot] (lll\r) at ($(W\r\r) +(0,\opoffset+.25)$)  {$\lll\!\rot$}; }
      \node[opxor] (xor\r) at ($(W\r\r) +(0,\opoffset+.50)$)  {};
      \node[opxor] (xnd\r) at ($(W\r\r) +(0,\opoffset+.75)$)  {};
    \ifthenelse{\equal{\r}{4}}
    { \node[opand] (and\r) at ($(W\r3) +(.5,\opoffset+.75)$)  {$\cdot$}; }
    { \node[opand] (and\r) at ($(W\r\r) +(.5,\opoffset+.75)$) {$\cdot$}; }
    \ifthenelse{\equal{\r}{-1}}%
    { \node[inner sep=1pt] (M) at ($(xnd\r) +(-.6,0)$) {$M$};
      \node[inner sep=1pt] (C) at ($(lll\r) +(-.6,0)$) {$C$};
      \draw[next] (M) -- (xnd\r);
      \draw[next] (lll\r) -- (C);
      \coordinate[optee] (tlll\r) at ($(W\r\tlllx) +(0,\opoffset+.25)$); }{}
      \coordinate[optee] (txor\r) at ($(W\r\txorx) +(0,\opoffset+.50)$);
      \coordinate[optee] (tanA\r) at ($(W\r\tanAx) +(0,\opoffset+.675)$);
      \coordinate[optee] (tanB\r) at ($(W\r\tanBx) +(0,\opoffset+.825)$);

    \ifthenelse{\equal{\r}{-1}}%
    { \draw[next] (tlll\r) -- (lll\r); }{}
      \draw[next] (txor\r) -- (xor\r);
    \ifthenelse{\tanAx < \r}%
    { \draw[next] (tanA\r) -- (tanA\r-|and\r.west); }%
    { \draw[next] (tanA\r) -- (tanA\r-|and\r.east); }
    \ifthenelse{\tanBx < \r}%
    { \draw[next] (tanB\r) -- (tanB\r-|and\r.west); }%
    { \draw[next] (tanB\r) -- (tanB\r-|and\r.east); }
    \draw[next] (and\r) -- (xnd\r);
    \draw[    ] (xnd\r) -- (xor\r);
    \draw[    ] (xor\r) -- (lll\r);

    \foreach \w in {0,...,4} {
      \ifthenelse{\equal{\w}{\r}}{
        \draw (W\rprev\w) -- (xnd\r);
        \draw (lll\r) -- (W\r\w);
      }{
        \draw (W\rprev\w) -- (W\r\w);
      }
    }
  }
  \end{tikzpicture}%}}}
  \caption{\cipher{MiniMorus} state update function (1 step = 5 rounds).}
  \label{fig:minimorus}
\end{figure}

\end{document}

### 2018-03-22 morus.tex
added approximations
linear approximation fragments...

@@
 
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
+\noindent
+2 approximations for $S^{2,2}_{0}$:
+\begin{align*}
+  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
+  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus
+  C^4_{19} \tag{corr. $2^{-9}$} \\
+\intertext{Combined ciphertext-only approximation:}
+  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
+  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $2^{-16}$}
+\end{align*}
+
+\newcommand{\printstate}[1]{
   \ifsubstates \pgfmathsetmacro{\roundsep}{1.25}
@@
   }
+}
+
+\begin{figure}
+  \substatesfalse
+  % \substatesfalse to label state words and/or masks
+  \centering
+  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
+    \printstate{}
   \end{tikzpicture}%}}}
@@
 
+\begin{figure}
+  \substatesfalse
+  % \substatesfalse to label state words and/or masks
+  \centering
+  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
+    \printstate{}
+
+    \draw[very thick, green!50!black]
+      (C) -- (lll-1)
+      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
+      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) --
+      (tanB0-|and0.east)
+      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
+      (and-1.east|-tanA-1) -- (tanA-1)
+      (tanA-1) -- (txor0) (txor0) -- (xor0)
+      (xor0) -- (lll0) (lll0) -- (W40)
+      ;
+
+  \end{tikzpicture}%}}}
+  \caption{\cipher{MiniMorus} linear approximation fragments.}
+  \label{fig:minimorus}
+\end{figure}
 \end{document}

### 2018-03-22 morus.tex
...

@@
 \usepackage{ifthen}
+\usepackage{subcaption}
 
@@
 \tikzset{next/.style={->, >=latex}}
+\tikzset{trail/.style={line width=2pt, green!50!black}}
 
@@
 
-\newcommand{\printstate}[1]{
+\newcommand{\printstate}{
   \ifsubstates \pgfmathsetmacro{\roundsep}{1.25}
@@
   \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate{}
+    \printstate
   \end{tikzpicture}%}}}
@@
   \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate{}
-
-    \draw[very thick, green!50!black]
-      (C) -- (lll-1)
+  \begin{subfigure}{.3\textwidth}
+  \centering
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail]
+      (C) -- node[below] {$i$} (lll-1)
       (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) --
-      (tanB0-|and0.east)
+      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
       (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1)
-      (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40)
+      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
+      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+5$}
+      (W-20) node[above] {\phantom{$i$}}
       ;
-
   \end{tikzpicture}%}}}
+  \caption{$\alpha_i$: $C^j \to S^{j-1,0}$}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{.3\textwidth}
+  \centering
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail]
+      (C) -- node[below] {$i$} (lll-1)
+      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
+      (tlll-1) -- (W-20) node[above] {$i$}
+      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
+      (xnd-1) -- (and-1)
+      (W40) node[below] {\phantom{$i$}}
+      ;
+  \end{tikzpicture}%}}}
+  \caption{$\beta_i$: $C^j \to S^{j,0}, S^{j,1}$}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{.3\textwidth}
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail]
+      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
+      (xnd1) -- (and1)
+      (xor1) -- (txor1)
+      (txor1) -- (W-24) node[above] {$i$}
+      (lll1) -- (W41) node[below] {$i-1$}
+      ;
+  \end{tikzpicture}%}}}
+  \caption{$\gamma_i$: $S^{j,1}, S^{j,4} \to S^{j+1,1}$}
+  \end{subfigure}
   \caption{\cipher{MiniMorus} linear approximation fragments.}

### 2018-03-22 morus.tex
linear approximation fragments

@@
   \end{subfigure}
+  \bigskip
+
+  \begin{subfigure}{.3\textwidth}
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail]
+      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
+      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
+      (lll4) -- (W44) node[below] {$i+13$}
+      ;
+  \end{tikzpicture}%}}}
+  \caption{$\delta_i$: $S^{j,4} \to S^{j+1,2}, S^{j+1,4}$}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{.3\textwidth}
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail]
+      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
+      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
+      (lll2) -- (W42) node[below] {$i+7$}
+      ;
+  \end{tikzpicture}%}}}
+  \caption{$\varepsilon_i$: $S^{j,2} \to S^{j+1,0}, S^{j+1,2}$}
+  \end{subfigure}
   \caption{\cipher{MiniMorus} linear approximation fragments.}

### 2018-03-22 aes-rng.c
AES-based RNG

@@ -1,0 +1,113 @@
+// Use AES-based RNG
+
+#ifdef __AES__
+
+#include <smmintrin.h>
+#include <wmmintrin.h>
+#include <stdint.h>
+#include <assert.h>
+
+__m128i key, count;
+int next;
+
+void init_aesrand(unsigned long long seed) {
+  srand(seed);
+  key   = _mm_setr_epi32(rand(), rand(), rand(), rand());
+  count = _mm_setr_epi32(rand(), rand(), rand(), rand());
+  next = 0;
+}
+
+unsigned int aesrand_int32() {
+  static __m128i t;
+  switch(next++) {
+  case 0:
+    // Increment counter
+    count = _mm_add_epi64(count, _mm_setr_epi32(0,0,1,0));
+    
+    // 6 AES rounds
+    t = count;
+    t = _mm_aesenc_si128(t, key);
+    t = _mm_aesenc_si128(t, key);
+    t = _mm_aesenc_si128(t, key);
+    t = _mm_aesenc_si128(t, key);
+    t = _mm_aesenc_si128(t, key);
+    t = _mm_aesenc_si128(t, key);
+    
+    return _mm_extract_epi32(t,0);
+  case 1:
+    return _mm_extract_epi32(t,1);
+  case 2:
+    return _mm_extract_epi32(t,2);
+  case 3:
+    next = 0;
+    return _mm_extract_epi32(t,3);
+  default:
+    assert(0);
+  }
+}
+
+struct RNG_state {
+  __m128i count;
+  __m128i t;
+  int next;
+};
+
+void* init_aesrand_r(uint32_t seed1, uint32_t seed2) {
+  // Internal state: count + t
+  struct RNG_state *st = malloc(sizeof(struct RNG_state));
+  assert(st);
+  st->count = _mm_setr_epi32(seed1, seed2, 0, 0);
+  st->next = 0;
+  return st;
+}
+
+unsigned int aesrand_int32_r(struct RNG_state* st) {
+  switch(st->next++) {
+  case 0:
+    // Increment counter
+    st->count = _mm_add_epi64(st->count, _mm_setr_epi32(0,0,1,0));
+    
+    // 6 AES rounds
+    st->t = st->count;
+    st->t = _mm_aesenc_si128(st->t, key);
+    st->t = _mm_aesenc_si128(st->t, key);
+    st->t = _mm_aesenc_si128(st->t, key);
[…]

### 2018-03-22 morus.tex
text version of trail
fixed trail
...

--- 2018-03-22 morus.tex
@@
 \tikzset{next/.style={->, >=latex}}
+\tikzset{trail/.style={line width=2pt, line cap=round}}
+\definecolor{delta}{HTML}{B58900}
+\definecolor{epsilon}{HTML}{CB4B16}
+\definecolor{gamma}{HTML}{6C71C4}
+\definecolor{beta}{HTML}{268BD2}
+\definecolor{alpha}{HTML}{859900}
+
+%\definecolor{mybase0}{HTML}{839496}
+%\definecolor{mybase1}{HTML}{93A1A1}
+%\definecolor{mybase2}{HTML}{EEE8D5}
+%\definecolor{mybase3}{HTML}{FDF6E3}
+%\definecolor{myred}{HTML}{DC322F}
+%\definecolor{mymagenta}{HTML}{D33682}
+%\definecolor{mycyan}{HTML}{2AA198}
 
@@
 
-  \foreach \r/\rot in {-1/0,0/5,1/31,2/7,3/22,4/13} { % round
+  %\foreach \r/\rot in {-1/0,0/5,1/31,2/7,3/22,4/13} { % round
+  \foreach \r/\rot in {-1/0,0/b_0,1/b_1,2/b_2,3/b_3,4/b_4} { % round
     \ifthenelse{\equal{\r}{-1}}{
@@
     \ifthenelse{\equal{\r}{-1}}%
-    { \node[inner sep=1pt] (M) at ($(xnd\r) +(-.6,0)$) {$M$};
-      \node[inner sep=1pt] (C) at ($(lll\r) +(-.6,0)$) {$C$};
+    { \node[inner sep=1pt] (M) at ($(xnd\r) +(0,.45)$) {$M$};
+      \node[inner sep=1pt] (C) at ($(lll\r) +(0,-.5)$) {$C$};
       \draw[next] (M) -- (xnd\r);
@@
   \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate{}
+    \printstate
   \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMorus} state update function (1 step = 5 rounds).}
+  \caption{\cipher{MiniMORUS} state update function (1 step = 5 rounds).}
   \label{fig:minimorus}
@@
   \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate{}
+  \begin{subfigure}{.32\textwidth}
+  \centering
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail, alpha]
+      (C) -- node[right] {$i$} (lll-1)
+      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
+      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
+      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
+      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
+      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
+      ;
+  \end{tikzpicture}%}}}
+  \caption*{$\alpha_i$: weight 1} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{.32\textwidth}
+  \centering
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail, beta]
+      (C) -- node[right] {$i$} (lll-1)
+      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
+      (tlll-1) -- (W-20) node[above] {$i$}
+      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
+      (xnd-1) -- (and-1)
+      (W40) node[below] {\phantom{$i$}}
+      ;
+  \end{tikzpicture}%}}}
+  \caption*{$\beta_i$: weight 0} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{.32\textwidth}
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail, gamma]
+      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
+      (xnd1) -- (and1)
+      (xor1) -- (txor1)
+      (txor1) -- (W-24) node[above] {$i$}
+      (lll1) -- (W41) node[below] {$i+b_1$}
+      ;
+  \end{tikzpicture}%}}}
+  \caption*{$\gamma_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
+  \end{subfigure}
+  \bigskip
 
-    \draw[very thick, green!50!black]
-      (C) -- (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) --
-      (tanB0-|and0.east)
-      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1)
-      (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40)
+  \begin{subfigure}{.32\textwidth}
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail, delta]
+      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
+      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
+      (lll4) -- (W44) node[below] {$i+b_4$}
       ;
-
   \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMorus} linear approximation fragments.}
+  \caption*{$\delta_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{.32\textwidth}
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail, epsilon]
+      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
+      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
+      (lll2) -- (W42) node[below] {$i+b_2$}
+      ;
+  \end{tikzpicture}%}}}
+  \caption*{$\varepsilon_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{.32\textwidth}
+    \small
+    \begin{align*}
+      \intertext{\cipher{MiniMORUS-640}:}
+      \alpha_i:      && C_i &\to S^{0}_{i+5} \\
+      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
+      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+31} \\
+      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+13} \\
+      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+7}
+      \intertext{\cipher{MiniMORUS-1280}:}
+      \alpha_i:      && C_i &\to S^{0}_{i+13} \\
+      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
+      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+46} \\
+      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+4} \\
+      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+38}
+    \end{align*}
+    \caption*{\cipher{MiniMORUS} instances}
+  \end{subfigure}
+  \caption{\cipher{MiniMORUS} linear approximation fragments.}
   \label{fig:minimorus}
 \end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{\textwidth}
+
+    Approximation 1 for $S^{2,2}_0$ (weight 7):
+
+    Step 0: $\alpha_{27}$ (weight 1)
+
+    Step 1: $\alpha_{8,26}, \beta_0, \gamma_0, \delta_0$ (weight 4)
+
+    Step 2: $\alpha_7, \beta_{13,31}, \gamma_{13}$ (weight 2)
+
+    Step 3: $\beta_{12}$ (weight 0)
+
+    Step 4: ---
+    \bigskip
+
+    Approximation 2 for $S^{2,2}_0$ (weight 9):
+
+    Step 0: ---
+
+    Step 1: $\alpha_2$ (weight 1)
+
+    St.2: $\alpha_{1,15,27}, \beta_7, \gamma_7, \delta_7, \varepsilon_0$ (weight 6)
+
+    Step 3: $\alpha_{14}, \beta_{6, 20}, \gamma_{20}$ (weight 2)
+
+    Step 4: $\beta_{19}$ (weight 0)
+    \caption{\cipher{MiniMORUS-640}}
+  \end{subfigure}
+  \bigskip
+
+  \begin{subfigure}{\textwidth}
+
+    \caption{\cipher{MiniMORUS-1280}}
+  \end{subfigure}
+  \caption{\cipher{MiniMORUS} approximations for $S^{2,2}_0$}
+\end{figure}
+
 \end{document}

### 2018-03-22 minimorus.py
update minimorus.py

@@ -2,7 +2,6 @@
 
 import random
 import math
-import sys
 
 def randword():
     return random.randrange(0,2**32)
@@ -13,9 +12,9 @@ def rotl(word, rotations=1, width=32):
     return ((word << rotations) & ones) | (word >> (width - rotations))
 
 def printstates(statelist, desc=""):
-    print(desc)
+    print desc
     for state in statelist:
-        print("[" + " ".join([hex(word)[2:].zfill(8) for word in state]) + "]")
+        print "[" + " ".join([hex(word)[2:].zfill(32/4) for word in state]) + "]"
 
 def xor(l):
     return reduce((lambda x, y: x ^ y), l, 0)
@@ -36,16 +35,9 @@ def minimorus_stateupdate(S, C=[], M=0):
 def minimorus_example():
     S = []
     C = []
-    # S.append([[randword() for w in range(5)]]); printstates(S[0], "init")
-    S.append([[1 for w in range(5)]]);
-    printstates(S[0], "init")
-    minimorus_stateupdate(S, C);
-    printstates(S[-1], "S^1")
-    minimorus_stateupdate(S, C);
-    printstates(S[-1], "S^2")
-
-minimorus_example()
-sys.exit("Stop")
+    S.append([[randword() for w in range(5)]]); printstates(S[0], "init")
+    minimorus_stateupdate(S, C);                printstates(S[-1], "S^1")
+    minimorus_stateupdate(S, C);                printstates(S[-1], "S^2")
 
 def minimorus_linearsample():
     S = []
@@ -58,10 +50,10 @@ def minimorus_linearsample():
     minimorus_stateupdate(S, C)
     bit = 0
 
-    #return xor([mask(C[0], [27]),
-    #            mask(C[1], [0,26]),
-    #            mask(C[2], [31]),
-    #            mask(S[1][-1][4], [0])]) # 2^-3
+    return xor([mask(C[0], [27]),
+                mask(C[1], [0,26]),
+                mask(C[2], [31]),
+                mask(S[1][-1][4], [0])]) # 2^-3
 
     # Approx 1 for S2
     #return xor([mask(C[0], [27]),
@@ -72,99 +64,25 @@ def minimorus_linearsample():
 
     # Approx 2 for S2
     # TESTED WITH 2^-22
-    return xor([mask(C[1], [2]),
-                mask(C[2], [1,7,15,27]),
[…]

### 2018-03-22 minimorus.py
debugging correlations

@@ -1,4 +1,4 @@
-#!/usr/bin/env python
+#!/usr/bin/env python3
 
 import random
 import math
@@ -24,27 +24,29 @@ def xor(l):
 def mask(word, bits):
     return xor([(word & (1 << bit)) >> bit for bit in bits])
 
-def minimorus_stateupdate(S, C=[], M=0):
-    Si = list(S[-1][-1])
+def minimorus_stateupdate(S, C=[], M=0, wordsize=32):
+    Si = list(S[-1])
     C.append(M ^ Si[0] ^ Si[1] ^ (Si[2] & Si[3]))
-    S.append([]);                                            #S[-1].append(list(Si))  # S^i_0,*
-    Si[0] = rotl((Si[0] ^ (Si[1] & Si[2]) ^ Si[3]),      5); #S[-1].append(list(Si))  # S^i_1,*
-    Si[1] = rotl((Si[1] ^ (Si[2] & Si[3]) ^ Si[4] ^ M), 31); #S[-1].append(list(Si))  # S^i_2,*
-    Si[2] = rotl((Si[2] ^ (Si[3] & Si[4]) ^ Si[0] ^ M),  7); #S[-1].append(list(Si))  # S^i_3,*
-    Si[3] = rotl((Si[3] ^ (Si[4] & Si[0]) ^ Si[1] ^ M), 22); #S[-1].append(list(Si))  # S^i_4,*
-    Si[4] = rotl((Si[4] ^ (Si[0] & Si[1]) ^ Si[2] ^ M), 13); S[-1].append(list(Si))  # S^i_5,* = S^i+1_0,*
+    if wordsize == 32:
+        Si[0] = rotl((Si[0] ^ (Si[1] & Si[2]) ^ Si[3]),      5);
+        Si[1] = rotl((Si[1] ^ (Si[2] & Si[3]) ^ Si[4] ^ M), 31);
+        Si[2] = rotl((Si[2] ^ (Si[3] & Si[4]) ^ Si[0] ^ M),  7);
+        Si[3] = rotl((Si[3] ^ (Si[4] & Si[0]) ^ Si[1] ^ M), 22);
+        Si[4] = rotl((Si[4] ^ (Si[0] & Si[1]) ^ Si[2] ^ M), 13); S.append(list(Si))
+    else:
+        pass
 
 def minimorus_example():
     S = []
     C = []
-    S.append([[randword() for w in range(5)]]); printstates(S[0], "init")
[…]

### 2018-03-22 little-morus.c
correlation ready?

@@ -1,61 +1,0 @@
-#include <stdint.h>
-
-typedef uint32_t S[5];
-
-uint32_t rotate_left(uint32_t x, int bits)
-{
-  if (bits == 0) return x;
-  return (x << bits) | (x >> (32 - bits));
-}
-
-void round(S state, uint32_t message) {
-	state[0] ^= state[1] & state[2];
-	state[0] ^= state[3];
-	state[0] = rotate_left(state[0], 5);
-
-	state[1] ^= message;
-	state[1] ^= state[2] & state[3];
-	state[1] ^= state[4];
-	state[1] = rotate_left(state[1], 31);
-
-	state[2] ^= message;
-	state[2] ^= state[3] & state[4];
-	state[2] ^= state[0];
-	state[2] = rotate_left(state[2], 7);
-
-	state[3] ^= message;
-	state[3] ^= state[4] & state[0];
-	state[3] ^= state[1];
-	state[3] = rotate_left(state[3], 22);
-
-	state[4] ^= message;
-	state[4] ^= state[0] ^ state[1];
-	state[4] ^= state[2];
-	state[4] = rotate_left(state[4], 13);
-}
-
-void init(S state, uint32_t iv, uint32_t key, uint32_t const1, uint32_t const2){
-	state[0] = iv;
-	state[1] = key;
-	state[2] = 0xffffffff;
-	state[3] = const1;
-	state[4] = const2;
-	for (int i = 0; i < 16; ++i)
-	{
-		round(state,0);
-	}
-}
-
-void encrypt(uint32_t message, S state)
-{
-	uint32_t mem = message;
-	message ^= state[0];
-	message ^= state[1];
-	message ^= state[2] & state[3];
-	round(state, mem);
-	return message;
-}
-
-
-
-

### 2018-03-22 mini_morus.c

@@ -1,0 +1,236 @@
+#include "mini_morus.h"
+
+// int log2_fast(double d) {
+//     int result;
+//     frexp(d, &result);
+//     return result-1;
+// }
+
+inline uint32_t mask (uint32_t* mask, unsigned int i)
+{
+  *mask |= (1 << i);
+  // return mask | (1 << i);
+  return 0;
+}
+
+// Compute the Hamming Weight
+inline uint32_t HW(uint32_t x)
+{
+    x = x - ((x >> 1) & 0x55555555);
+    x = (x & 0x33333333) + ((x >> 2) & 0x33333333);
+    x = (x + (x >> 4)) & 0x0F0F0F0F;
+    x = x + (x >> 8);
+    x = x + (x >> 16);
+    return x & 0x0000003F;
+}
+
+// Remark the XOR of the bits are equal to the parity of the Hamming Weight !
+
+inline void save_state(S* state_saved, S state, int i)
+{
+  state_saved[i][0] = state[0];
+  state_saved[i][1] = state[1];
+  state_saved[i][2] = state[2];
+  state_saved[i][3] = state[3];
+  state_saved[i][4] = state[4];
+}
+
+inline void print_state(S state){
+  printf("%08x %08x %08x %08x %08x\n", state[0], state[1], state[2], state[3], state[4]);
+}
+
+inline uint32_t rotate_left(uint32_t x, int bits)
+{
+  if (bits == 0) return x;
+  return (x << bits) | (x >> (32 - bits));
+}
+
+void mini_morus_round(S state, uint32_t message) {
+	state[0] ^= state[1] & state[2];
+	state[0] ^= state[3];
+	state[0] = rotate_left(state[0], 5);
+
+	state[1] ^= message;
+	state[1] ^= state[2] & state[3];
+	state[1] ^= state[4];
+	state[1] = rotate_left(state[1], 31);
+
+	state[2] ^= message;
+	state[2] ^= state[3] & state[4];
+	state[2] ^= state[0];
+	state[2] = rotate_left(state[2], 7);
+
+	state[3] ^= message;
+	state[3] ^= state[4] & state[0];
+	state[3] ^= state[1];
+	state[3] = rotate_left(state[3], 22);
+
+	state[4] ^= message;
+	state[4] ^= state[0] ^ state[1];
+	state[4] ^= state[2];
+	state[4] = rotate_left(state[4], 13);
+}
+
+void init(S state, uint32_t iv, uint32_t key, uint32_t const1, uint32_t const2){
[…]

### 2018-03-22 mini_morus.h

@@ -1,0 +1,31 @@
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <math.h>
+#include <omp.h>
+
+typedef uint32_t S[5];
+
+// Helper function to generate the mask for Linear correlation
+
+inline uint32_t mask (uint32_t* mask, unsigned int i);
+
+// Compute the Hamming Weight
+inline uint32_t HW(uint32_t x);
+
+// Remark the XOR of the bits are equal to the parity of the Hamming Weight !
+
+void print_state(S state);
+
+inline uint32_t rotate_left(uint32_t x, int bits);
+
+void mini_morus_round(S state, uint32_t message);
+
+void init(S state, uint32_t iv, uint32_t key, uint32_t const1, uint32_t const2);
+
+uint32_t encrypt(uint32_t message, S state);
+
+long long minimorus_linearsample();
+
+void minimorus_linearstats(unsigned long long num);

### 2018-03-22 minimorus.py

@@ -1,0 +1,162 @@
+#!/usr/bin/env python
+
+import random
+import math
+
+def randword():
+    return random.randrange(0,2**32)
+
+def rotl(word, rotations=1, width=32):
+    ones = 2**width-1
+    word &= ones
+    return ((word << rotations) & ones) | (word >> (width - rotations))
+
+def printstates(statelist, desc=""):
+    print desc
+    for state in statelist:
+        print "[" + " ".join([hex(word)[2:].zfill(32/4) for word in state]) + "]"
+
+def xor(l):
+    return reduce((lambda x, y: x ^ y), l, 0)
+
+def mask(word, bits):
+    return xor([(word & (1 << bit)) >> bit for bit in bits])
+
+def minimorus_stateupdate(S, C=[], M=0):
+    Si = list(S[-1][-1])
+    C.append(M ^ Si[0] ^ Si[1] ^ (Si[2] & Si[3]))
+    S.append([]);                                            #S[-1].append(list(Si))  # S^i_0,*
+    Si[0] = rotl((Si[0] ^ (Si[1] & Si[2]) ^ Si[3]),      5); #S[-1].append(list(Si))  # S^i_1,*
+    Si[1] = rotl((Si[1] ^ (Si[2] & Si[3]) ^ Si[4] ^ M), 31); #S[-1].append(list(Si))  # S^i_2,*
+    Si[2] = rotl((Si[2] ^ (Si[3] & Si[4]) ^ Si[0] ^ M),  7); #S[-1].append(list(Si))  # S^i_3,*
+    Si[3] = rotl((Si[3] ^ (Si[4] & Si[0]) ^ Si[1] ^ M), 22); #S[-1].append(list(Si))  # S^i_4,*
+    Si[4] = rotl((Si[4] ^ (Si[0] & Si[1]) ^ Si[2] ^ M), 13); S[-1].append(list(Si))  # S^i_5,* = S^i+1_0,*
+
+def minimorus_example():
+    S = []
+    C = []
+    S.append([[randword() for w in range(5)]]); printstates(S[0], "init")
+    minimorus_stateupdate(S, C);                printstates(S[-1], "S^1")
+    minimorus_stateupdate(S, C);                printstates(S[-1], "S^2")
+
+def minimorus_linearsample():
+    S = []
+    C = []
+    S.append([[randword() for w in range(5)]])
+    minimorus_stateupdate(S, C)
+    minimorus_stateupdate(S, C)
+    minimorus_stateupdate(S, C)
[…]

### 2018-03-22 morus.c

@@ -28,7 +28,7 @@ void and_state(S o, S s1 , S s2)
 {
 	for (int i = 0; i < 4; ++i)
 	{
-		o[i] = s1[i] ^ s2[i]
+		o[i] = s1[i] ^ s2[i];
 	}
 }
 
@@ -36,7 +36,7 @@ void xor_state(S o, S s)
 {
 	for (int i = 0; i < 4; ++i)
 	{
-		o[i] ^= s[i]
+		o[i] ^= s[i];
 	}
 }
 
@@ -56,7 +56,7 @@ void permute_state(S s, int bits)
 		t = s[0];
 		s[0] = s[2];
 		s[2] = t;
-		t = s[1]
+		t = s[1];
 		s[1] = s[3];
 		s[3] = t;
 	}
@@ -69,8 +69,3 @@ void permute_state(S s, int bits)
 		s[0] = t;
 	}
 }
-
-
-
-
-

### 2018-03-22 mini_morus.c
some more tests

@@ -66,7 +66,7 @@ void mini_morus_round(S state, uint32_t message) {
 	state[3] = rotate_left(state[3], 22);
 
 	state[4] ^= message;
-	state[4] ^= state[0] ^ state[1];
+	state[4] ^= state[0] & state[1];
 	state[4] ^= state[2];
 	state[4] = rotate_left(state[4], 13);
 }
@@ -84,12 +84,6 @@ void init(S state, uint32_t iv, uint32_t key, uint32_t const1, uint32_t const2){
 }
 
 void rand_init(S state){
-  // FILE * f = fopen("/dev/urandom", "rb");
-  // fread(&state[0], 5*sizeof(uint32_t), 1, f);
-  // fread(&state[1], sizeof(uint32_t), 1, f);
-  // fread(&state[2], sizeof(uint32_t), 1, f);
-  // fread(&state[3], sizeof(uint32_t), 1, f);
-  // fread(&state[4], sizeof(uint32_t), 1, f);
 	state[0] = rand();
 	state[1] = rand();
 	state[2] = rand();
@@ -117,11 +111,12 @@ inline void minimorus_sample(S* saved_state, uint32_t* saved_cipher, int num)
     save_state(saved_state,state,i);
     saved_cipher[i] = encrypt(0, state);
   }
+  save_state(saved_state,state,i);
 }
 
 long long minimorus_linearsample()
 {
-  S saved_state[5];
+  S saved_state[6];
   uint32_t saved_cipher[5];
 
   uint32_t mask1 = 0;
@@ -139,28 +134,106 @@ long long minimorus_linearsample()
   //             mask(C[2], [1,7,15,27]),
   //             mask(C[3], [6,20,14]),
   //             mask(C[4], [19]),
-  //             mask(S[2][-1][2], [0])]) # 2^-9
+  //             mask(S[2][2], [0])]) # 2^-9
 
+  // --------------------------------------------
+  //
+  // mask(&mask1,27);
+  // res ^= saved_cipher[0] & mask1;
+  //
+  // mask(&mask2,0);
+  // mask(&mask2,26);
+  // mask(&mask2,8);
+  // res ^= saved_cipher[1] & mask2;
+  //
+  // mask(&mask3,31);
+  // mask(&mask3,13);
+  // mask(&mask3,7);
+  // res ^= saved_cipher[2] & mask3;
+  //
+  // mask(&mask4,12);
+  // res ^= saved_cipher[3] & mask4;
+  //
+  // mask(&mask5,0);
[…]

### 2018-03-22 minimorus.py

@@ -2,6 +2,7 @@
 
 import random
 import math
+import sys
 
 def randword():
     return random.randrange(0,2**32)
@@ -12,9 +13,9 @@ def rotl(word, rotations=1, width=32):
     return ((word << rotations) & ones) | (word >> (width - rotations))
 
 def printstates(statelist, desc=""):
-    print desc
+    print(desc)
     for state in statelist:
-        print "[" + " ".join([hex(word)[2:].zfill(32/4) for word in state]) + "]"
+        print("[" + " ".join([hex(word)[2:].zfill(8) for word in state]) + "]")
 
 def xor(l):
     return reduce((lambda x, y: x ^ y), l, 0)
@@ -35,9 +36,16 @@ def minimorus_stateupdate(S, C=[], M=0):
 def minimorus_example():
     S = []
     C = []
-    S.append([[randword() for w in range(5)]]); printstates(S[0], "init")
-    minimorus_stateupdate(S, C);                printstates(S[-1], "S^1")
-    minimorus_stateupdate(S, C);                printstates(S[-1], "S^2")
+    # S.append([[randword() for w in range(5)]]); printstates(S[0], "init")
+    S.append([[1 for w in range(5)]]);
+    printstates(S[0], "init")
+    minimorus_stateupdate(S, C);
+    printstates(S[-1], "S^1")
+    minimorus_stateupdate(S, C);
+    printstates(S[-1], "S^2")
+
+minimorus_example()
+sys.exit("Stop")
 
 def minimorus_linearsample():
     S = []
@@ -147,12 +155,12 @@ def minimorus_linearstats(num):
     bias = (bias - num/2)/float(num)
     potential = correlation**2
     #print "prob", probability
-    print "corr", correlation
+    print("corr", correlation)
     #print "bias", bias
     #print "pote", potential, math.log(potential, 2)
-    print "logD", math.log(bias**(-2), 2)
+    print("logD", math.log(bias**(-2), 2))
     biascost = math.fabs(math.log(2*math.fabs(bias), 2))  # math.copysign(1,bias)
-    print "bias: +- 2^-", biascost
+    print("bias: +- 2^-", biascost)
 
 
     # p = 1/4

### 2018-03-22 morus.tex
MiniMORUS trail

@@
 \definecolor{delta}{HTML}{B58900}
-\definecolor{epsilon}{HTML}{CB4B16}
+\definecolor{epsil}{HTML}{CB4B16}
 \definecolor{gamma}{HTML}{6C71C4}
@@
   0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $2^{-16}$}
+  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-16.5}$}
 \end{align*}
@@
       ;
+    \draw[trail, alpha, dashed]
+      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
+      ;
   \end{tikzpicture}%}}}
-  \caption*{$\alpha_i$: weight 1} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
+  \caption*{$\alpha_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
   \end{subfigure}
@@
   \end{tikzpicture}%}}}
-  \caption*{$\beta_i$: weight 0} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
+  \caption*{$\beta_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
   \end{subfigure}
@@
     \printstate
-    \draw[trail, epsilon]
+    \draw[trail, epsil]
       (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
@@
 \begin{figure}
+  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
+  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
+  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
   \centering
   \begin{subfigure}{\textwidth}
+    \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
+      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
+      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
 
-    Approximation 1 for $S^{2,2}_0$ (weight 7):
+      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
+                   (\M{0}{0}) node[right] {$\alpha_{27}$};
 
-    Step 0: $\alpha_{27}$ (weight 1)
+      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
+                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
+                   (\M{1}{0}) node[right] {$\beta_0$};
+      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
+                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
+      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
+      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
+                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
+                   (\M{1}{2}) node[right] {$\gamma_0$};
+      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
+                   (\S{1}{4}) ++(0,-1.0) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
+                   (\M{1}{3}) node[right] {$\delta_0$};
 
-    Step 1: $\alpha_{8,26}, \beta_0, \gamma_0, \delta_0$ (weight 4)
+      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
+                   (\C{2}-|[xshift=-1pt]\S{2}{0}) -| ([xshift=0pt]\S{2}{1}) node[below left] {31\!\!}
+                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
+      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
+                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
+      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
+                   (\M{2}{1}) node[right] {$\alpha_7$};
+      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
+                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
+                   (\M{2}{2}) node[right] {$\gamma_{13}$};
 
-    Step 2: $\alpha_7, \beta_{13,31}, \gamma_{13}$ (weight 2)
+      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
+                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
+                   (\M{3}{0}) node[right] {$\beta_{12}$};
 
-    Step 3: $\beta_{12}$ (weight 0)
+      \draw (\S{4}{2}) node[above] {\normalsize weight 7};
 
-    Step 4: ---
-    \bigskip
+      \begin{scope}[xshift=2.5cm]
+        \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+        \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
+        \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
 
-    Approximation 2 for $S^{2,2}_0$ (weight 9):
+        \draw[alpha] (\M{1}{1}) node[right] {$\alpha_2$};
 
-    Step 0: ---
+        \draw[beta]  (\M{2}{0}) node[right] {$\beta_7$};
+        \draw[alpha] (\M{2}{1}) node[right] {$\alpha_{1,15,27}$};
+        \draw[gamma] (\M{2}{2}) node[right] {$\gamma_7$};
+        \draw[delta] (\M{2}{3}) node[right] {$\delta_7$};
+        \draw[epsil] (\M{2}{4}) node[right] {$\varepsilon_0$};
 
-    Step 1: $\alpha_2$ (weight 1)
+        \draw[beta]  (\M{3}{0}) node[right] {$\beta_{6,20}$};
+        \draw[alpha] (\M{3}{1}) node[right] {$\alpha_{14}$};
+        \draw[gamma] (\M{3}{2}) node[right] {$\gamma_{20}$};
 
-    St.2: $\alpha_{1,15,27}, \beta_7, \gamma_7, \delta_7, \varepsilon_0$ (weight 6)
+        \draw[beta]  (\M{4}{0}) node[right] {$\beta_{19}$};
 
-    Step 3: $\alpha_{14}, \beta_{6, 20}, \gamma_{20}$ (weight 2)
-
-    Step 4: $\beta_{19}$ (weight 0)
+        \draw (\S{5}{2}) node[above] {\normalsize weight 9};
+      \end{scope}
+    \end{tikzpicture}
     \caption{\cipher{MiniMORUS-640}}
@@
   \end{subfigure}
-  \caption{\cipher{MiniMORUS} approximations for $S^{2,2}_0$}
+  \caption{\cipher{MiniMORUS}: Two approximations for $S^{2,2}_0$}
 \end{figure}

### 2018-03-23 mini_morus.c
moving sources into src directory

@@ -1,317 +1,0 @@
-#include "mini_morus.h"
-
-// int log2_fast(double d) {
-//     int result;
-//     frexp(d, &result);
-//     return result-1;
-// }
-
-inline uint32_t mask (uint32_t* mask, unsigned int i)
-{
-  *mask |= (1 << i);
-  // return mask | (1 << i);
-  return 0;
-}
-
-// Compute the Hamming Weight
-inline uint32_t HW(uint32_t x)
-{
-    x = x - ((x >> 1) & 0x55555555);
-    x = (x & 0x33333333) + ((x >> 2) & 0x33333333);
-    x = (x + (x >> 4)) & 0x0F0F0F0F;
-    x = x + (x >> 8);
-    x = x + (x >> 16);
-    return x & 0x0000003F;
-}
-
-// Remark the XOR of the bits are equal to the parity of the Hamming Weight !
-
-inline void save_state(S* state_saved, S state, int i)
-{
-  state_saved[i][0] = state[0];
-  state_saved[i][1] = state[1];
-  state_saved[i][2] = state[2];
-  state_saved[i][3] = state[3];
-  state_saved[i][4] = state[4];
-}
-
-inline void print_state(S state){
-  printf("%08x %08x %08x %08x %08x\n", state[0], state[1], state[2], state[3], state[4]);
-}
-
-inline uint32_t rotate_left(uint32_t x, int bits)
-{
-  if (bits == 0) return x;
-  return (x << bits) | (x >> (32 - bits));
-}
-
-void mini_morus_round(S state, uint32_t message) {
-	state[0] ^= state[1] & state[2];
-	state[0] ^= state[3];
-	state[0] = rotate_left(state[0], 5);
-
-	state[1] ^= message;
-	state[1] ^= state[2] & state[3];
-	state[1] ^= state[4];
-	state[1] = rotate_left(state[1], 31);
-
-	state[2] ^= message;
-	state[2] ^= state[3] & state[4];
-	state[2] ^= state[0];
-	state[2] = rotate_left(state[2], 7);
-
-	state[3] ^= message;
-	state[3] ^= state[4] & state[0];
-	state[3] ^= state[1];
-	state[3] = rotate_left(state[3], 22);
-
-	state[4] ^= message;
-	state[4] ^= state[0] & state[1];
-	state[4] ^= state[2];
-	state[4] = rotate_left(state[4], 13);
-}
-
-void init(S state, uint32_t iv, uint32_t key, uint32_t const1, uint32_t const2){
[…]

### 2018-03-23 mini_morus.h

@@ -1,31 +1,0 @@
-#include <stdint.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <time.h>
-#include <math.h>
-#include <omp.h>
-
-typedef uint32_t S[5];
-
-// Helper function to generate the mask for Linear correlation
-
-inline uint32_t mask (uint32_t* mask, unsigned int i);
-
-// Compute the Hamming Weight
-inline uint32_t HW(uint32_t x);
-
-// Remark the XOR of the bits are equal to the parity of the Hamming Weight !
-
-void print_state(S state);
-
-inline uint32_t rotate_left(uint32_t x, int bits);
-
-void mini_morus_round(S state, uint32_t message);
-
-void init(S state, uint32_t iv, uint32_t key, uint32_t const1, uint32_t const2);
-
-uint32_t encrypt(uint32_t message, S state);
-
-long long minimorus_linearsample();
-
-void minimorus_linearstats(unsigned long long num);

### 2018-03-23 morus.c

@@ -1,71 +1,0 @@
-#include <stdint.h>
-
-typedef uint32_t S[4];
-
-uint32_t rotate_left(uint32_t x, int bits)
-{
-  if (bits == 0) return x;
-  return (x << bits) | (x >> (32 - bits));
-}
-
-void rotate_left_state(S state, int bits)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		state[i] = rotate_left(state[i],bits);
-	}
-}
-
-void rotate_right_state(S state, int bits)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		state[i] = rotate_left(state[i],32-bits);
-	}
-}
-
-void and_state(S o, S s1 , S s2)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		o[i] = s1[i] ^ s2[i];
-	}
-}
-
-void xor_state(S o, S s)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		o[i] ^= s[i];
-	}
-}
-
-void permute_state(S s, int bits)
-{
-	uint32_t t;
-	if (bits == 96)
-	{
-		t = s[0];
-		s[0] = s[1];
-		s[1] = s[2];
-		s[2] = s[3];
-		s[3] = t;
-	}
-	if (bits == 64)
-	{
-		t = s[0];
-		s[0] = s[2];
-		s[2] = t;
-		t = s[1];
-		s[1] = s[3];
-		s[3] = t;
-	}
-	if (bits == 32)
-	{
-		t = s[3];
-		s[3] = s[2];
-		s[2] = s[1];
-		s[1] = s[0];
-		s[0] = t;
-	}
-}

### 2018-03-23 src/main.c

@@ -1,0 +1,82 @@
+#include "main.h"
+
+inline void copy_word(state_words* to, const state_words* from)
+{
+  *to = *from;
+}
+
+inline void save_state(state* state_saved, state state, int i)
+{
+  copy_word(&state_saved[i][0], &state[0]);
+  copy_word(&state_saved[i][1], &state[1]);
+  copy_word(&state_saved[i][2], &state[2]);
+  copy_word(&state_saved[i][3], &state[3]);
+  copy_word(&state_saved[i][4], &state[4]);
+}
+
+inline void sample(state* saved_state, state_words* saved_cipher, int num, unsigned int* seed)
+{
+  state state;
+  int i;
+  rand_init(state, seed);
+  for (i = 0; i < num; i++) {
+    save_state(saved_state, state, i);
+    saved_cipher[i] = encrypt(0, state);
+  }
+  save_state(saved_state, state, i);
+}
+
+inline int linear_sample(unsigned int* seed)
+{
+  state saved_state[6];
+  state_words saved_cipher[5];
+
+  sample(saved_state, saved_cipher, 5, seed);
+  return linear(saved_state, saved_cipher);
+}
+
+void linear_stats(unsigned long long num) {
+    long long res = 0;
+    long long inbalance = 0;
+    unsigned long long bias = 0;
+    unsigned long long i = 0;
+
+    int tid;
+    unsigned int seed;
+    printf("--------------------------\n");
+    printf("num:  %llu\n", (num* omp_get_max_threads()));
+    printf("--------------------------\n");
+
+    # pragma omp parallel private(i, tid, seed) reduction(+:bias,inbalance)
+    {
+      tid = omp_get_thread_num();
+      seed = 123456789 ^ tid;
+      printf ( "  %6d  %12d\n", tid, seed);
+
+      // try using openmp to speed things up
+      #pragma omp parallel for reduction(+:bias,inbalance)
+      for(i = 0 ; i < num; ++i)
+      {
+        res = linear_sample(&seed);
+        inbalance += 1 - 2*res;
+        bias += res;
+        if((i & 0x3ffffff) == 0)
+        {
+          printf("sampling, %llu -- thread %d\n", i, tid);
+        }
+      }
+    }
+
+    printf("--------------------------\n");
+    printf("num:  %llu\n", (num* omp_get_max_threads()));
+    printf("inba: %lld\n", inbalance);
+    printf("bias: %llu\n", bias);
[…]

### 2018-03-23 src/main.h

@@ -1,0 +1,5 @@
+#include <stdint.h>
+#include <stdio.h>
+#include <time.h>
+#include <omp.h>
+#include "mini_morus.h"

### 2018-03-23 src/mini_morus.c

@@ -1,0 +1,115 @@
+#include "mini_morus.h"
+
+inline void print_state(state state) {
+  printf("%08x %08x %08x %08x %08x\n", state[0], state[1], state[2], state[3], state[4]);
+}
+
+inline uint32_t rotate_left(uint32_t x, int bits)
+{
+  if (bits == 0) return x;
+  return (x << bits) | (x >> (32 - bits));
+}
+
+void rand_init(state st, unsigned int* seed){
+	st[0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  st[1] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  st[2] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  st[3] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  st[4] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+}
+
+void iterate(state st, state_words message) {
+	st[0] ^= st[1] & st[2];
+	st[0] ^= st[3];
+	st[0] = rotate_left(st[0], 5);
+
+	st[1] ^= message;
+	st[1] ^= st[2] & st[3];
+	st[1] ^= st[4];
+	st[1] = rotate_left(st[1], 31);
+
+	st[2] ^= message;
+	st[2] ^= st[3] & st[4];
+	st[2] ^= st[0];
+	st[2] = rotate_left(st[2], 7);
+
+	st[3] ^= message;
+	st[3] ^= st[4] & st[0];
+	st[3] ^= st[1];
+	st[3] = rotate_left(st[3], 22);
+
+	st[4] ^= message;
+	st[4] ^= st[0] & st[1];
+	st[4] ^= st[2];
+	st[4] = rotate_left(st[4], 13);
+}
+
+state_words encrypt(state_words message, state st)
+{
+	state_words mem = message;
+	message ^= st[0];
+	message ^= st[1];
+	message ^= st[2] & st[3];
+	iterate(st, mem);
+	return message;
+}
+
+
+
+
+
+
+inline uint32_t gen_mask(uint32_t* mask, unsigned int i)
+{
+  *mask |= (1 << i);
+  return 0;
+}
+
+// Compute the Hamming Weight
[…]

### 2018-03-23 src/mini_morus.h

@@ -1,0 +1,16 @@
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+
+typedef uint32_t state_words;
+typedef state_words state[5];
+
+void print_state(state st);
+
+void rand_init(state st, unsigned int* seed);
+
+void iterate(state st, state_words message);
+
+state_words encrypt(state_words message, state state);
+
+int linear(state* saved_state, state_words* saved_cipher);

### 2018-03-23 src/morus.c

@@ -1,0 +1,138 @@
+#include "morus.h"
+
+inline void print_state(state state) {
+  printf("%08x %08x %08x %08x %08x\n", state[0][0], state[1][0], state[2][0], state[3][0], state[4][0]);
+  printf("%08x %08x %08x %08x %08x\n", state[0][1], state[1][1], state[2][1], state[3][1], state[4][1]);
+  printf("%08x %08x %08x %08x %08x\n", state[0][2], state[1][2], state[2][2], state[3][2], state[4][2]);
+  printf("%08x %08x %08x %08x %08x\n", state[0][3], state[1][3], state[2][3], state[3][3], state[4][3]);
+}
+
+uint32_t rotate_left(uint32_t x, int bits)
+{
+  if (bits == 0) return x;
+  return (x << bits) | (x >> (32 - bits));
+}
+
+void rotate_left_state(state_words word, int bits)
+{
+	for (int i = 0; i < 4; ++i)
+	{
+		word[i] = rotate_left(word[i],bits);
+	}
+}
+
+void rotate_right_state(state_words words, int bits)
+{
+	for (int i = 0; i < 4; ++i)
+	{
+		words[i] = rotate_left(words[i],32-bits);
+	}
+}
+
+void and_xor_to_state(state_words o, const state_words s1 , const state_words s2)
+{
+	for (int i = 0; i < 4; ++i)
+	{
+		o[i] ^= s1[i] & s2[i];
+	}
+}
+
+void xor_to_state(state_words o, const state_words s)
+{
+	for (int i = 0; i < 4; ++i)
+	{
+		o[i] ^= s[i];
+	}
+}
+
+void permute_state(state_words s, int bits)
+{
+	uint32_t t;
+	if (bits == 96)
+	{
+		t = s[0];
+		s[0] = s[1];
+		s[1] = s[2];
+		s[2] = s[3];
+		s[3] = t;
+	}
+	if (bits == 64)
+	{
+		t = s[0];
+		s[0] = s[2];
+		s[2] = t;
+		t = s[1];
+		s[1] = s[3];
+		s[3] = t;
+	}
+	if (bits == 32)
+	{
+		t = s[3];
+		s[3] = s[2];
+		s[2] = s[1];
+		s[1] = s[0];
+		s[0] = t;
+	}
+}
+
[…]

### 2018-03-23 src/morus.h

@@ -1,0 +1,16 @@
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+
+typedef uint32_t state_words[4];
+typedef state_words state[5];
+
+void print_state(state st);
+
+void rand_init(state st, unsigned int* seed);
+
+void iterate(state st, state_words message);
+
+state_words encrypt(state_words message, state state);
+
+int linear(state* saved_state, state_words* saved_cipher);

### 2018-03-23 src/main.c
fix limit to number of threads

@@ -47,6 +47,7 @@ void linear_stats(unsigned long long num) {
     printf("num:  %llu\n", (num* omp_get_max_threads()));
     printf("--------------------------\n");
 
+    omp_set_num_threads(16);
     # pragma omp parallel private(i, tid, seed) reduction(+:bias,inbalance)
     {
       tid = omp_get_thread_num();
@@ -76,7 +77,7 @@ void linear_stats(unsigned long long num) {
 
 int main(int argc, char const *argv[]) {
   long long int num = 1;
-  num <<= 20;
+  num <<= 33;
   linear_stats(num);
   return 0;
 }

### 2018-03-23 src/morus.c

@@ -1,138 +1,138 @@
-#include "morus.h"
-
-inline void print_state(state state) {
-  printf("%08x %08x %08x %08x %08x\n", state[0][0], state[1][0], state[2][0], state[3][0], state[4][0]);
-  printf("%08x %08x %08x %08x %08x\n", state[0][1], state[1][1], state[2][1], state[3][1], state[4][1]);
-  printf("%08x %08x %08x %08x %08x\n", state[0][2], state[1][2], state[2][2], state[3][2], state[4][2]);
-  printf("%08x %08x %08x %08x %08x\n", state[0][3], state[1][3], state[2][3], state[3][3], state[4][3]);
-}
-
-uint32_t rotate_left(uint32_t x, int bits)
-{
-  if (bits == 0) return x;
-  return (x << bits) | (x >> (32 - bits));
-}
-
-void rotate_left_state(state_words word, int bits)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		word[i] = rotate_left(word[i],bits);
-	}
-}
-
-void rotate_right_state(state_words words, int bits)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		words[i] = rotate_left(words[i],32-bits);
-	}
-}
-
-void and_xor_to_state(state_words o, const state_words s1 , const state_words s2)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		o[i] ^= s1[i] & s2[i];
-	}
-}
-
-void xor_to_state(state_words o, const state_words s)
-{
-	for (int i = 0; i < 4; ++i)
-	{
-		o[i] ^= s[i];
-	}
-}
-
-void permute_state(state_words s, int bits)
-{
-	uint32_t t;
-	if (bits == 96)
-	{
-		t = s[0];
-		s[0] = s[1];
-		s[1] = s[2];
-		s[2] = s[3];
-		s[3] = t;
-	}
-	if (bits == 64)
-	{
-		t = s[0];
-		s[0] = s[2];
-		s[2] = t;
-		t = s[1];
-		s[1] = s[3];
-		s[3] = t;
-	}
-	if (bits == 32)
-	{
-		t = s[3];
-		s[3] = s[2];
-		s[2] = s[1];
-		s[1] = s[0];
-		s[0] = t;
-	}
-}
-
[…]

### 2018-03-23 minimorus.py
segfault in Morus and python3 on the minimorus.py

@@ -2,6 +2,8 @@
 
 import random
 import math
+# import sys
+from functools import reduce
 
 def randword():
     return random.randrange(0,2**32)
@@ -12,9 +14,9 @@ def rotl(word, rotations=1, width=32):
     return ((word << rotations) & ones) | (word >> (width - rotations))
 
 def printstates(statelist, desc=""):
-    print desc
+    print(desc)
     for state in statelist:
-        print "[" + " ".join([hex(word)[2:].zfill(32/4) for word in state]) + "]"
+        print("[" + " ".join([hex(word)[2:].zfill(8) for word in state]) + "]")
 
 def xor(l):
     return reduce((lambda x, y: x ^ y), l, 0)
@@ -73,16 +75,16 @@ def minimorus_linearsample():
 def minimorus_linearstats(num):
     count = 0
     imbalance2 = 0
-    for i in xrange(num):
+    for i in range(num):
         res = minimorus_linearsample()
         count += res
         imbalance2 += (-1)**res  # imbalance: 1/2 (#0 - #1)
     bias = (count - num/2)/float(num)
     correlation = imbalance2 / float(num)
     weight = math.fabs(math.log(math.fabs(correlation), 2))
-    print "correl", correlation
-    print "weight", weight
-    print "logDat", 2*weight + 2
+    print ("correl", correlation)
+    print ("weight", weight)
+    print ("logDat", 2*weight + 2)
 
 if __name__ == "__main__":
     minimorus_linearstats(2**16)

### 2018-03-23 src/main.c

@@ -17,11 +17,14 @@ inline void save_state(state* state_saved, state state, int i)
 inline void sample(state* saved_state, state_words* saved_cipher, int num, unsigned int* seed)
 {
   state state;
+  state_words null_words;
+  // state_words output;
   int i;
   rand_init(state, seed);
+  null_word(&null_words);
   for (i = 0; i < num; i++) {
     save_state(saved_state, state, i);
-    saved_cipher[i] = encrypt(0, state);
+    encrypt(&saved_cipher[i], null_words, state);
   }
   save_state(saved_state, state, i);
 }
@@ -47,7 +50,7 @@ void linear_stats(unsigned long long num) {
     printf("num:  %llu\n", (num* omp_get_max_threads()));
     printf("--------------------------\n");
 
-    omp_set_num_threads(16);
+    // omp_set_num_threads(16);
     # pragma omp parallel private(i, tid, seed) reduction(+:bias,inbalance)
     {
       tid = omp_get_thread_num();

### 2018-03-23 src/mini_morus.c

@@ -1,5 +1,9 @@
 #include "mini_morus.h"
 
+void null_word(state_words* w) {
+   *w = 0;
+}
+
 inline void print_state(state state) {
   printf("%08x %08x %08x %08x %08x\n", state[0], state[1], state[2], state[3], state[4]);
 }
@@ -44,20 +48,15 @@ void iterate(state st, state_words message) {
 	st[4] = rotate_left(st[4], 13);
 }
 
-state_words encrypt(state_words message, state st)
+void encrypt(state_words* out, state_words message, state st)
 {
 	state_words mem = message;
 	message ^= st[0];
 	message ^= st[1];
 	message ^= st[2] & st[3];
 	iterate(st, mem);
-	return message;
+  *out = message;
 }
-
-
-
-
-
 
 inline uint32_t gen_mask(uint32_t* mask, unsigned int i)
 {

### 2018-03-23 src/mini_morus.h

@@ -11,6 +11,8 @@ void rand_init(state st, unsigned int* seed);
 
 void iterate(state st, state_words message);
 
-state_words encrypt(state_words message, state state);
+void encrypt(state_words* out, state_words message, state state);
 
 int linear(state* saved_state, state_words* saved_cipher);
+
+void null_word(state_words* w);

### 2018-03-23 src/morus.c

@@ -1,138 +1,217 @@
-// #include "morus.h"
-//
-// inline void print_state(state state) {
-//   printf("%08x %08x %08x %08x %08x\n", state[0][0], state[1][0], state[2][0], state[3][0], state[4][0]);
-//   printf("%08x %08x %08x %08x %08x\n", state[0][1], state[1][1], state[2][1], state[3][1], state[4][1]);
-//   printf("%08x %08x %08x %08x %08x\n", state[0][2], state[1][2], state[2][2], state[3][2], state[4][2]);
-//   printf("%08x %08x %08x %08x %08x\n", state[0][3], state[1][3], state[2][3], state[3][3], state[4][3]);
-// }
-//
-// uint32_t rotate_left(uint32_t x, int bits)
-// {
-//   if (bits == 0) return x;
-//   return (x << bits) | (x >> (32 - bits));
-// }
-//
-// void rotate_left_state(state_words word, int bits)
-// {
-// 	for (int i = 0; i < 4; ++i)
-// 	{
-// 		word[i] = rotate_left(word[i],bits);
-// 	}
-// }
-//
-// void rotate_right_state(state_words words, int bits)
-// {
-// 	for (int i = 0; i < 4; ++i)
-// 	{
-// 		words[i] = rotate_left(words[i],32-bits);
-// 	}
-// }
-//
-// void and_xor_to_state(state_words o, const state_words s1 , const state_words s2)
-// {
-// 	for (int i = 0; i < 4; ++i)
-// 	{
-// 		o[i] ^= s1[i] & s2[i];
-// 	}
-// }
-//
-// void xor_to_state(state_words o, const state_words s)
-// {
-// 	for (int i = 0; i < 4; ++i)
-// 	{
-// 		o[i] ^= s[i];
-// 	}
-// }
-//
-// void permute_state(state_words s, int bits)
-// {
-// 	uint32_t t;
-// 	if (bits == 96)
-// 	{
-// 		t = s[0];
-// 		s[0] = s[1];
-// 		s[1] = s[2];
-// 		s[2] = s[3];
-// 		s[3] = t;
-// 	}
-// 	if (bits == 64)
-// 	{
-// 		t = s[0];
-// 		s[0] = s[2];
-// 		s[2] = t;
-// 		t = s[1];
[…]

### 2018-03-23 src/morus.h

@@ -11,6 +11,8 @@ void rand_init(state st, unsigned int* seed);
 
 void iterate(state st, state_words message);
 
-state_words encrypt(state_words message, state state);
+void encrypt(state_words* out, state_words message, state state);
 
 int linear(state* saved_state, state_words* saved_cipher);
+
+inline void null_word(state_words* w);

### 2018-03-23 src/main_mini_morus.c
now have two version, second maybe not working

@@ -1,4 +1,4 @@
-#include "main.h"
+#include "main_mini_morus.h"
 
 inline void copy_word(state_words* to, const state_words* from)
 {

### 2018-03-23 src/main_morus.c

@@ -1,0 +1,89 @@
+#include "main_morus.h"
+
+inline void copy_word(state_words* to, const state_words* from)
+{
+  (*to)[0] = (*from)[0];
+  (*to)[1] = (*from)[1];
+  (*to)[2] = (*from)[2];
+  (*to)[3] = (*from)[3];
+}
+
+inline void save_state(state* state_saved, state state, int i)
+{
+  copy_word(&state_saved[i][0], &state[0]);
+  copy_word(&state_saved[i][1], &state[1]);
+  copy_word(&state_saved[i][2], &state[2]);
+  copy_word(&state_saved[i][3], &state[3]);
+  copy_word(&state_saved[i][4], &state[4]);
+}
+
+inline void sample(state* saved_state, state_words* saved_cipher, int num, unsigned int* seed)
+{
+  state state;
+  state_words null_words;
+  // state_words output;
+  int i;
+  rand_init(state, seed);
+  null_word(&null_words);
+  for (i = 0; i < num; i++) {
+    save_state(saved_state, state, i);
+    encrypt(&saved_cipher[i], null_words, state);
+  }
+  save_state(saved_state, state, i);
+}
+
+inline int linear_sample(unsigned int* seed)
+{
+  state saved_state[6];
+  state_words saved_cipher[5];
+
+  sample(saved_state, saved_cipher, 5, seed);
+  return linear(saved_state, saved_cipher);
+}
+
+void linear_stats(unsigned long long num) {
+    long long res = 0;
+    long long inbalance = 0;
+    unsigned long long bias = 0;
+    unsigned long long i = 0;
+
+    int tid;
+    unsigned int seed;
+    printf("--------------------------\n");
+    printf("num:  %llu\n", (num* omp_get_max_threads()));
+    printf("--------------------------\n");
+
+    // omp_set_num_threads(16);
+    # pragma omp parallel private(i, tid, seed) reduction(+:bias,inbalance)
+    {
+      tid = omp_get_thread_num();
+      seed = 123456789 ^ tid;
+      printf ( "  %6d  %12d\n", tid, seed);
+
+      // try using openmp to speed things up
+      #pragma omp parallel for reduction(+:bias,inbalance)
+      for(i = 0 ; i < num; ++i)
+      {
+        res = linear_sample(&seed);
+        inbalance += 1 - 2*res;
+        bias += res;
+        if((i & 0x3ffffff) == 0)
+        {
[…]

### 2018-03-23 src/main_morus.h

@@ -1,0 +1,5 @@
+#include <stdint.h>
+#include <stdio.h>
+#include <time.h>
+#include <omp.h>
+#include "morus.h"

### 2018-03-23 src/morus.h

@@ -15,4 +15,4 @@ void encrypt(state_words* out, state_words message, state state);
 
 int linear(state* saved_state, state_words* saved_cipher);
 
-inline void null_word(state_words* w);
+void null_word(state_words* w);

### 2018-03-23 morus.tex
completed tex trail

@@
     \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
+      \begin{scope}
         \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
@@
         \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.0) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
+                     (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
                      (\M{1}{3}) node[right] {$\delta_0$};
@@
         \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   (\C{2}-|[xshift=-1pt]\S{2}{0}) -| ([xshift=0pt]\S{2}{1}) node[below left] {31\!\!}
+                     ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
                      (\M{2}{0}) node[right] {$\beta_{13,31}$};
@@
 
-      \draw (\S{4}{2}) node[above] {\normalsize weight 7};
+        \draw (\S{4}{2}) node[above] {\normalsize weight 7 (not 9)};
+      \end{scope}
 
@@
 
-        \draw[alpha] (\M{1}{1}) node[right] {$\alpha_2$};
+        \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
+                     (\M{1}{0}) node[right] {$\alpha_{2}$};
 
-        \draw[beta]  (\M{2}{0}) node[right] {$\beta_7$};
-        \draw[alpha] (\M{2}{1}) node[right] {$\alpha_{1,15,27}$};
-        \draw[gamma] (\M{2}{2}) node[right] {$\gamma_7$};
-        \draw[delta] (\M{2}{3}) node[right] {$\delta_7$};
-        \draw[epsil] (\M{2}{4}) node[right] {$\varepsilon_0$};
+        \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
+                     (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
+                     (\M{2}{0}) node[right] {$\beta_7$};
+        \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
+                     (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
+        \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
+        \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
+        \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
+                     (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
+                     (\M{2}{2}) node[right] {$\gamma_7$};
+        \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
+                     (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
+                     (\M{2}{4}) node[right] {$\delta_7$};
+        \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
+                     (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
+                     (\M{2}{3}) node[right] {$\varepsilon_0$};
 
-        \draw[beta]  (\M{3}{0}) node[right] {$\beta_{6,20}$};
-        \draw[alpha] (\M{3}{1}) node[right] {$\alpha_{14}$};
-        \draw[gamma] (\M{3}{2}) node[right] {$\gamma_{20}$};
+        \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
+                     ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
+                     (\M{3}{0}) node[right] {$\beta_{20,6}$};
+        \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
+                     ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
+        \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
+                     (\M{3}{1}) node[right] {$\alpha_{14}$};
+        \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
+                     ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
+                     (\M{3}{2}) node[right] {$\gamma_{20}$};
 
-        \draw[beta]  (\M{4}{0}) node[right] {$\beta_{19}$};
+        \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
+                     (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
+                     (\M{4}{0}) node[right] {$\beta_{19}$};
 
-        \draw (\S{5}{2}) node[above] {\normalsize weight 9};
+        \draw (\S{5}{2}) node[above] {\normalsize weight 9 (not 11)};
       \end{scope}
@@
 
+\begin{figure}
+
+  \caption{Suspicious things I'd like to verify with full \cipher{MORUS}}
+\end{figure}
+
 \end{document}

### 2018-03-23 morus.tex
morus1280 version of trail

@@
   \centering
-  \begin{subfigure}{\textwidth}
   \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
@@
       \draw (\S{5}{2}) node[above] {\normalsize weight 9 (not 11)};
+      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
+    \end{scope}
+
+    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
+
+    \begin{scope}[yshift=-7.0cm]
+      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
+      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
+
+      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
+                   (\M{0}{0}) node[right] {$\alpha_{51}$};
+
+      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
+                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
+                   (\M{1}{0}) node[right] {$\beta_0$};
+      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
+                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
+      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
+      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
+                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
+                   (\M{1}{2}) node[right] {$\gamma_0$};
+      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
+                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
+                   (\M{1}{3}) node[right] {$\delta_0$};
+
+      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
+                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
+                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
+      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
+                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
+      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
+                   (\M{2}{1}) node[right] {$\alpha_{37}$};
+      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
+                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
+                   (\M{2}{2}) node[right] {$\gamma_{4}$};
+
+      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
+                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
+                   (\M{3}{0}) node[right] {$\beta_{50}$};
+
+      \draw (\S{4}{2}) node[above] {\normalsize weight 7 (not 9)};
+      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
+    \end{scope}
+
+    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
+      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
+      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
+
+      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
+                   (\M{1}{0}) node[right] {$\alpha_{25}$};
+
+      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
+                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
+                   (\M{2}{0}) node[right] {$\beta_{38}$};
+      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
+                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
+      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
+      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
+      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
+                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
+                   (\M{2}{2}) node[right] {$\gamma_{38}$};
+      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
+                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
+                   (\M{2}{4}) node[right] {$\delta_{38}$};
+      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
+                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
+                   (\M{2}{3}) node[right] {$\varepsilon_0$};
+
+      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
+                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
+                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
+      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
+                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
+      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
+                   (\M{3}{1}) node[right] {$\alpha_{11}$};
+      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
+                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
+                   (\M{3}{2}) node[right] {$\gamma_{42}$};
+
+      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
+                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
+                   (\M{4}{0}) node[right] {$\beta_{24}$};
+
+      \draw (\S{5}{2}) node[above] {\normalsize weight 9 (not 11)};
     \end{scope}
   \end{tikzpicture}
-    \caption{\cipher{MiniMORUS-640}}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{\textwidth}
-
-    \caption{\cipher{MiniMORUS-1280}}
-  \end{subfigure}
   \caption{\cipher{MiniMORUS}: Two approximations for $S^{2,2}_0$}

### 2018-03-23 minimorus.py
extend .py to 64-bit version

@@ -2,21 +2,74 @@
 
 import random
 import math
-# import sys
+import sys
 from functools import reduce
 
-def randword():
-    return random.randrange(0,2**32)
+width=64
+approximations = dict( \
+        alpha=dict(w=1, masks=lambda S, C, width: []) if width==32 else \
+              dict(w=1, masks=lambda S, C, width: []),
+        appr1=dict(w=7, masks=lambda S, C, width: [mask(C[0], [27]),
+                                                   mask(C[1], [0,26,8]),
+                                                   mask(C[2], [31,13,7]),
+                                                   mask(C[3], [12]),
+                                                   mask(S[2][2], [0])]) if width==32 else \
+              dict(w=7, masks=lambda S, C, width: [mask(C[0], [51]),
+                                                   mask(C[1], [0,33,55]),
+                                                   mask(C[2], [4,37,46]),
+                                                   mask(C[3], [50]),
+                                                   mask(S[2][2], [0])]),
+        appr2=dict(w=9, masks=lambda S, C, width: [mask(C[1], [2]),
+                                                   mask(C[2], [1,7,15,27]),
+                                                   mask(C[3], [6,20,14]),
+                                                   mask(C[4], [19]),
+                                                   mask(S[2][2], [0])]) if width==32 else \
+              dict(w=9, masks=lambda S, C, width: [mask(C[1], [25]),
+                                                   mask(C[2], [7,29,38,51]),
+                                                   mask(C[3], [11,20,42]),
+                                                   mask(C[4], [24]),
+                                                   mask(S[2][2], [0])]),
+        )
+#           xor([mask(S[1][0], [0]),
+#                mask(S[1][4], [0]),
+#                mask(C[1], [0,8,26]),
+#                mask(C[2], [13,31]),
+#                # 2^-3 version
[…]

### 2018-03-23 minimorus.py
Confirmed correlation for two halves of 1280 trail

@@ -7,8 +7,10 @@ from functools import reduce
 
 width=64
 approximations = dict( \
-        alpha=dict(w=1, masks=lambda S, C, width: []) if width==32 else \
-              dict(w=1, masks=lambda S, C, width: []),
+        alpha=dict(w=1, masks=lambda S, C, width: [mask(C[0], [27]),
+                                                   mask(S[1][0], [0])]) if width==32 else \
+              dict(w=1, masks=lambda S, C, width: [mask(C[0], [51]),
+                                                   mask(S[1][0], [0])]),
         appr1=dict(w=7, masks=lambda S, C, width: [mask(C[0], [27]),
                                                    mask(C[1], [0,26,8]),
                                                    mask(C[2], [31,13,7]),
@@ -56,7 +58,8 @@ approximations = dict( \
 #                mask(C[2], [31]),
 #                mask(S[1][4], [0])]) # 2^-3
 
-approximation = approximations['appr1']
+#approximation = approximations['appr1']
+approximation = approximations['appr2']
 
 def randword(width=32):
     return random.randrange(0,2**width)
@@ -97,11 +100,11 @@ def minimorus_linearsample(masks, width=32):
     S = []
     C = []
     S.append([randword(width) for w in range(5)])
-    minimorus_stateupdate(S, C, width)
-    minimorus_stateupdate(S, C, width)
-    minimorus_stateupdate(S, C, width)
-    minimorus_stateupdate(S, C, width)
-    minimorus_stateupdate(S, C, width)
+    minimorus_stateupdate(S, C, 0, width)
+    minimorus_stateupdate(S, C, 0, width)
+    minimorus_stateupdate(S, C, 0, width)
+    minimorus_stateupdate(S, C, 0, width)
+    minimorus_stateupdate(S, C, 0, width)
     bit = 0
 
     return xor(masks(S, C, width))
@@ -109,7 +112,7 @@ def minimorus_linearsample(masks, width=32):
 def minimorus_linearstats(approximation, width):
     expweight = approximation['w']
     masks = approximation['masks']
-    num = 2**(2*expweight+7)
+    num = 2**(2*expweight+5)
 
     count = 0
     imbalance = 0

### 2018-03-23 src/main_morus.c
trail in morus (WIP)

@@ -83,7 +83,7 @@ void linear_stats(unsigned long long num) {
 
 int main(int argc, char const *argv[]) {
   long long int num = 1;
-  num <<= 33;
+  num <<= 30;
   linear_stats(num);
   return 0;
 }

### 2018-03-23 src/morus.c

@@ -180,38 +180,70 @@ inline uint32_t HW(uint32_t x)
     return x & 0x0000003F;
 }
 
+// return xor([mask(S[1][0], [0]),
+//             mask(S[1][4], [0]),
+//             mask(C[1], [0,8,26]),
+//             mask(C[2], [13,31]),
+//             # 2^-3 version
+//             mask(S[2][4], [13]),
+//             mask(S[3][1], [12]),
+//             # 2^-4 version
+//             #mask(S[2][1], [13]),
+//             ]) # 2^-3
+
 int linear(state* saved_state, state_words* saved_cipher) {
-  // uint32_t mask = 0;
+  uint32_t mask = 0;
   uint32_t res = 0;
 
-  // gen_mask(&mask,27);
-  // res ^= saved_cipher[0] & mask;
-  //
-  // mask = 0;
-  // gen_mask(&mask,0);
-  // gen_mask(&mask,26);
-  // gen_mask(&mask,8);
-  // gen_mask(&mask,2);
-  // res ^= saved_cipher[1] & mask;
-  //
-  // mask = 0;
-  // gen_mask(&mask,31);
-  // gen_mask(&mask,13);
-  // gen_mask(&mask,1);
-  // gen_mask(&mask,15);
-  // gen_mask(&mask,27);
-  // res ^= saved_cipher[2] & mask;
-  //
-  // mask = 0;
-  // gen_mask(&mask,6);
-  // gen_mask(&mask,12);
-  // gen_mask(&mask,20);
-  // gen_mask(&mask,14);
-  // res ^= saved_cipher[3] & mask;
-  //
-  // mask = 0;
-  // gen_mask(&mask,19);
-  // res ^= saved_cipher[4] & mask;
+  mask = 0;
+  gen_mask(&mask,0);
+  res ^= saved_state[1][0][0] & mask;
+  res ^= saved_state[1][0][1] & mask;
+  res ^= saved_state[1][0][2] & mask;
+  res ^= saved_state[1][0][3] & mask;
+
+  res ^= saved_state[1][4][0] & mask;
+  res ^= saved_state[1][4][1] & mask;
+  res ^= saved_state[1][4][2] & mask;
+  res ^= saved_state[1][4][3] & mask;
+
+  mask = 0;
+  gen_mask(&mask,0);
+  gen_mask(&mask,8);
+  gen_mask(&mask,26);
[…]

### 2018-03-23 src/main_morus.c
WIP

@@ -10,11 +10,11 @@ inline void copy_word(state_words* to, const state_words* from)
 
 inline void save_state(state* state_saved, state state, int i)
 {
-  copy_word(&state_saved[i][0], &state[0]);
-  copy_word(&state_saved[i][1], &state[1]);
-  copy_word(&state_saved[i][2], &state[2]);
-  copy_word(&state_saved[i][3], &state[3]);
-  copy_word(&state_saved[i][4], &state[4]);
+  copy_word(&(state_saved[i][0]), &(state[0]));
+  copy_word(&(state_saved[i][1]), &(state[1]));
+  copy_word(&(state_saved[i][2]), &(state[2]));
+  copy_word(&(state_saved[i][3]), &(state[3]));
+  copy_word(&(state_saved[i][4]), &(state[4]));
 }
 
 inline void sample(state* saved_state, state_words* saved_cipher, int num, unsigned int* seed)

### 2018-03-23 src/morus.c

@@ -85,29 +85,52 @@ void null_word(state_words* w){
 
 void rand_init(state st, unsigned int* seed){
   // need to be faster !!!
-	st[0][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[1][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[2][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[3][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[4][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+	// st[0][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[1][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[2][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[3][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[4][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  //
+  // st[0][1] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[1][1] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[2][1] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[3][1] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[4][1] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  //
+  // st[0][2] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[1][2] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+  // st[2][2] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
[…]

### 2018-03-23 src/main_morus.c
minimorus in bigmorus

@@ -53,7 +53,7 @@ void linear_stats(unsigned long long num) {
     printf("num:  %llu\n", (num* omp_get_max_threads()));
     printf("--------------------------\n");
 
-    // omp_set_num_threads(16);
+    // omp_set_num_threads(1);
     # pragma omp parallel private(i, tid, seed) reduction(+:bias,inbalance)
     {
       tid = omp_get_thread_num();

### 2018-03-23 src/morus.c

@@ -1,6 +1,7 @@
 #include "morus.h"
 
 inline void print_state(state state) {
+  printf("--------------------------------------------\n");
   printf("   S0   |   S1   |   S3   |   S3   |   S4   \n");
   printf("%08x %08x %08x %08x %08x\n", state[0][3], state[1][3], state[2][3], state[3][3], state[4][3]);
   printf("%08x %08x %08x %08x %08x\n", state[0][2], state[1][2], state[2][2], state[3][2], state[4][2]);
@@ -8,6 +9,22 @@ inline void print_state(state state) {
   printf("%08x %08x %08x %08x %08x\n", state[0][0], state[1][0], state[2][0], state[3][0], state[4][0]);
 }
 
+inline void print_word(state_words word) {
+  printf("%08x %08x %08x %08x\n", word[0], word[1], word[2], word[3]);
+}
+
+inline void print(state* saved_state, state_words* saved_cipher)
+{
+  int i;
+  printf("-------------------------------------------- Ciphers\n");
+  for ( i = 0; i < 5; i++) {
+    print_word(saved_cipher[i]);
+  }
+  printf("-------------------------------------------- States\n");
+  for ( i = 0; i < 6; i++) {
+    print_state(saved_state[i]);
+  }
+}
 inline uint32_t rotate_left(uint32_t x, int bits)
 {
   if (bits == 0) return x;
@@ -24,10 +41,10 @@ inline void copy_state_words(state_words out, const state_words word)
 
 inline void rotate_left_word(state_words word, int bits)
 {
-  word[0] = rotate_left(word[0],bits);
-  word[1] = rotate_left(word[1],bits);
-  word[2] = rotate_left(word[2],bits);
-  word[3] = rotate_left(word[3],bits);
+  word[0] = rotate_left(word[0], bits);
+  word[1] = rotate_left(word[1], bits);
+  word[2] = rotate_left(word[2], bits);
+  word[3] = rotate_left(word[3], bits);
 }
 
 inline void and_xor_to_word(state_words o, const state_words s1 , const state_words s2)
@@ -59,12 +76,8 @@ inline void permute_words(state_words s, int bits)
 	}
 	if (bits == 64)
 	{
-		t = s[0];
[…]

### 2018-03-23 aes-rng.c
use AES random number generator

@@ -1,113 +1,0 @@
-// Use AES-based RNG
-
-#ifdef __AES__
-
-#include <smmintrin.h>
-#include <wmmintrin.h>
-#include <stdint.h>
-#include <assert.h>
-
-__m128i key, count;
-int next;
-
-void init_aesrand(unsigned long long seed) {
-  srand(seed);
-  key   = _mm_setr_epi32(rand(), rand(), rand(), rand());
-  count = _mm_setr_epi32(rand(), rand(), rand(), rand());
-  next = 0;
-}
-
-unsigned int aesrand_int32() {
-  static __m128i t;
-  switch(next++) {
-  case 0:
-    // Increment counter
-    count = _mm_add_epi64(count, _mm_setr_epi32(0,0,1,0));
-    
-    // 6 AES rounds
-    t = count;
-    t = _mm_aesenc_si128(t, key);
-    t = _mm_aesenc_si128(t, key);
-    t = _mm_aesenc_si128(t, key);
-    t = _mm_aesenc_si128(t, key);
-    t = _mm_aesenc_si128(t, key);
-    t = _mm_aesenc_si128(t, key);
-    
-    return _mm_extract_epi32(t,0);
-  case 1:
-    return _mm_extract_epi32(t,1);
-  case 2:
-    return _mm_extract_epi32(t,2);
-  case 3:
-    next = 0;
-    return _mm_extract_epi32(t,3);
-  default:
-    assert(0);
-  }
-}
-
-struct RNG_state {
-  __m128i count;
-  __m128i t;
-  int next;
-};
-
-void* init_aesrand_r(uint32_t seed1, uint32_t seed2) {
-  // Internal state: count + t
-  struct RNG_state *st = malloc(sizeof(struct RNG_state));
-  assert(st);
-  st->count = _mm_setr_epi32(seed1, seed2, 0, 0);
-  st->next = 0;
-  return st;
-}
-
-unsigned int aesrand_int32_r(struct RNG_state* st) {
-  switch(st->next++) {
-  case 0:
-    // Increment counter
-    st->count = _mm_add_epi64(st->count, _mm_setr_epi32(0,0,1,0));
-    
-    // 6 AES rounds
-    st->t = st->count;
-    st->t = _mm_aesenc_si128(st->t, key);
-    st->t = _mm_aesenc_si128(st->t, key);
-    st->t = _mm_aesenc_si128(st->t, key);
[…]

### 2018-03-23 src/aes-rng.c

@@ -1,0 +1,104 @@
+// Use AES-based RNG
+#include "aes-rng.h"
+#ifdef __AES__
+
+// __m128i key, count;
+// int next;
+//
+// void init_aesrand(unsigned long long seed) {
+//   srand(seed);
+//   key   = _mm_setr_epi32(rand(), rand(), rand(), rand());
+//   count = _mm_setr_epi32(rand(), rand(), rand(), rand());
+//   next = 0;
+// }
+//
+// unsigned int aesrand_int32() {
+//   static __m128i t;
+//   switch(next++) {
+//   case 0:
+//     // Increment counter
+//     count = _mm_add_epi64(count, _mm_setr_epi32(0,0,1,0));
+//
+//     // 6 AES rounds
+//     t = count;
+//     t = _mm_aesenc_si128(t, key);
+//     t = _mm_aesenc_si128(t, key);
+//     t = _mm_aesenc_si128(t, key);
+//     t = _mm_aesenc_si128(t, key);
+//     t = _mm_aesenc_si128(t, key);
+//     t = _mm_aesenc_si128(t, key);
+//
+//     return _mm_extract_epi32(t,0);
+//   case 1:
+//     return _mm_extract_epi32(t,1);
+//   case 2:
+//     return _mm_extract_epi32(t,2);
+//   case 3:
+//     next = 0;
+//     return _mm_extract_epi32(t,3);
+//   default:
+//     assert(0);
+//   }
+// }
+
+
+struct RNG_state* init_aesrand_r(uint32_t seed1, uint32_t seed2) {
+  // Internal state: count + t
+  struct RNG_state *st = malloc(sizeof(struct RNG_state));
+  assert(st);
+  st->count = _mm_setr_epi32(seed1, seed2, 0, 0);
+  st->k = _mm_setr_epi32(0, 0, seed1, seed2);
+  st->next = 0;
+  return st;
+}
+
+uint32_t aesrand_int32_r(struct RNG_state* rng) {
+  switch(rng->next++) {
+  case 0:
+    // Increment counter
+    rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0,0,1,0));
+
+    // 6 AES rounds
+    rng->t = rng->count;
+    rng->t = _mm_aesenc_si128(rng->t, rng->k);
+    rng->t = _mm_aesenc_si128(rng->t, rng->k);
+    rng->t = _mm_aesenc_si128(rng->t, rng->k);
[…]

### 2018-03-23 src/main_morus.c

@@ -17,7 +17,7 @@ inline void save_state(state* state_saved, state state, int i)
   copy_word(&(state_saved[i][4]), &(state[4]));
 }
 
-inline void sample(state* saved_state, state_words* saved_cipher, int num, unsigned int* seed)
+inline void sample(state* saved_state, state_words* saved_cipher, int num, struct RNG_state* seed)
 {
   state state;
   state_words null_words;
@@ -32,7 +32,7 @@ inline void sample(state* saved_state, state_words* saved_cipher, int num, unsig
   save_state(saved_state, state, i);
 }
 
-inline int linear_sample(unsigned int* seed)
+inline int linear_sample(struct RNG_state* seed)
 {
   state saved_state[6];
   state_words saved_cipher[5];
@@ -46,25 +46,25 @@ void linear_stats(unsigned long long num) {
     long long inbalance = 0;
     unsigned long long bias = 0;
     unsigned long long i = 0;
+    struct RNG_state* seed;
+    int small_seed = rand();
+    int tid = 0;
 
-    int tid;
-    unsigned int seed;
     printf("--------------------------\n");
     printf("num:  %llu\n", (num* omp_get_max_threads()));
     printf("--------------------------\n");
 
     // omp_set_num_threads(1);
     # pragma omp parallel private(i, tid, seed) reduction(+:bias,inbalance)
-    {
       tid = omp_get_thread_num();
-      seed = 123456789 ^ tid;
-      printf ( "  %6d  %12d\n", tid, seed);
+      seed = init_aesrand_r(small_seed,tid);
+      // printf ( "  %6d  %12d\n", tid, seed);
 
       // try using openmp to speed things up
       #pragma omp parallel for reduction(+:bias,inbalance)
       for(i = 0 ; i < num; ++i)
       {
-        res = linear_sample(&seed);
+        res = linear_sample(seed);
         inbalance += 1 - 2*res;
         bias += res;
         if((i & 0x3ffffff) == 0)
@@ -72,7 +72,6 @@ void linear_stats(unsigned long long num) {
           printf("sampling, %llu -- thread %d\n", i, tid);
         }
       }
-    }
 
     printf("--------------------------\n");
     printf("num:  %llu\n", (num* omp_get_max_threads()));
@@ -83,7 +82,8 @@ void linear_stats(unsigned long long num) {
 
 int main(int argc, char const *argv[]) {
   long long int num = 1;
-  num <<= 30;
[…]

### 2018-03-23 src/main_morus.h

@@ -3,3 +3,4 @@
 #include <time.h>
 #include <omp.h>
 #include "morus.h"
+#include "aes-rng.h"

### 2018-03-23 src/mini_morus.c

@@ -14,12 +14,12 @@ inline uint32_t rotate_left(uint32_t x, int bits)
   return (x << bits) | (x >> (32 - bits));
 }
 
-void rand_init(state st, unsigned int* seed){
-	st[0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[1] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[2] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[3] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[4] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
+void rand_init(state st, struct RNG_state* seed){
+  st[0] = aesrand_int32_r(seed);
+  st[1] = aesrand_int32_r(seed);
+  st[2] = aesrand_int32_r(seed);
+  st[3] = aesrand_int32_r(seed);
+  st[4] = aesrand_int32_r(seed);
 }
 
 void iterate(state st, state_words message) {

### 2018-03-23 src/mini_morus.h

@@ -1,13 +1,14 @@
 #include <stdint.h>
 #include <stdio.h>
 #include <stdlib.h>
+#include "aes-rng.h"
 
 typedef uint32_t state_words;
 typedef state_words state[5];
 
 void print_state(state st);
 
-void rand_init(state st, unsigned int* seed);
+void rand_init(state st, struct RNG_state* rng);
 
 void iterate(state st, state_words message);
 

### 2018-03-23 src/morus.c

@@ -96,7 +96,7 @@ void null_word(state_words* w){
   (*w)[3] = 0;
 }
 
-void rand_init(state st, unsigned int* seed){
+void rand_init(state st, struct RNG_state* seed){
   // need to be faster !!!
 	// st[0][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
   // st[1][0] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
@@ -121,29 +121,29 @@ void rand_init(state st, unsigned int* seed){
   // st[2][3] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
   // st[3][3] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
   // st[4][3] = (rand_r(seed) & 0xffff) | (rand_r(seed) & 0xffff) << 16;
-  st[0][0] = rand_r(seed);
-  st[1][0] = rand_r(seed);
-  st[2][0] = rand_r(seed);
-  st[3][0] = rand_r(seed);
-  st[4][0] = rand_r(seed);
+  st[0][0] = aesrand_int32_r(seed);
+  st[1][0] = aesrand_int32_r(seed);
+  st[2][0] = aesrand_int32_r(seed);
+  st[3][0] = aesrand_int32_r(seed);
+  st[4][0] = aesrand_int32_r(seed);
 
-  st[0][1] = rand_r(seed);
-  st[1][1] = rand_r(seed);
-  st[2][1] = rand_r(seed);
-  st[3][1] = rand_r(seed);
-  st[4][1] = rand_r(seed);
+  st[0][1] = aesrand_int32_r(seed);
+  st[1][1] = aesrand_int32_r(seed);
+  st[2][1] = aesrand_int32_r(seed);
+  st[3][1] = aesrand_int32_r(seed);
+  st[4][1] = aesrand_int32_r(seed);
 
-  st[0][2] = rand_r(seed);
-  st[1][2] = rand_r(seed);
-  st[2][2] = rand_r(seed);
-  st[3][2] = rand_r(seed);
-  st[4][2] = rand_r(seed);
+  st[0][2] = aesrand_int32_r(seed);
+  st[1][2] = aesrand_int32_r(seed);
+  st[2][2] = aesrand_int32_r(seed);
+  st[3][2] = aesrand_int32_r(seed);
[…]

### 2018-03-23 src/main_medium_morus.c
add medium morus

@@ -1,0 +1,87 @@
+#include "main_medium_morus.h"
+
+inline void copy_word(state_words* to, const state_words* from)
+{
+  *to = *from;
+}
+
+inline void save_state(state* state_saved, state state, int i)
+{
+  copy_word(&state_saved[i][0], &state[0]);
+  copy_word(&state_saved[i][1], &state[1]);
+  copy_word(&state_saved[i][2], &state[2]);
+  copy_word(&state_saved[i][3], &state[3]);
+  copy_word(&state_saved[i][4], &state[4]);
+}
+
+inline void sample(state* saved_state, state_words* saved_cipher, int num, struct RNG_state* seed)
+{
+  state state;
+  state_words null_words;
+  // state_words output;
+  int i;
+  rand_init(state, seed);
+  null_word(&null_words);
+  for (i = 0; i < num; i++) {
+    save_state(saved_state, state, i);
+    encrypt(&saved_cipher[i], null_words, state);
+  }
+  save_state(saved_state, state, i);
+}
+
+inline int linear_sample(struct RNG_state* seed)
+{
+  state saved_state[6];
+  state_words saved_cipher[5];
+
+  sample(saved_state, saved_cipher, 5, seed);
+  return linear(saved_state, saved_cipher);
+}
+
+void linear_stats(unsigned long long num) {
+    long long res = 0;
+    long long inbalance = 0;
+    unsigned long long bias = 0;
+    unsigned long long i = 0;
+    struct RNG_state* seed;
+    unsigned int small_seed = rand();
+    int tid;
+    printf("--------------------------\n");
+    printf("num:  %llu\n", (num* omp_get_max_threads()));
+    printf("--------------------------\n");
+
+    // omp_set_num_threads(16);
+    # pragma omp parallel private(i, tid, seed) reduction(+:bias,inbalance)
+    {
+      tid = omp_get_thread_num();
+      seed = init_aesrand_r(small_seed,tid);
+      // printf ( "  %6d  %12d\n", tid, seed);
+
+      // try using openmp to speed things up
+      #pragma omp parallel for reduction(+:bias,inbalance)
+      for(i = 0 ; i < num; ++i)
+      {
+        res = linear_sample(seed);
+        inbalance += 1 - 2*res;
+        bias += res;
+        if((i & 0x3ffffff) == 0)
+        {
+          printf("sampling, %llu -- thread %d\n", i, tid);
+        }
+      }
+    }
+
[…]

### 2018-03-23 src/main_medium_morus.h

@@ -1,0 +1,6 @@
+#include <stdint.h>
+#include <stdio.h>
+#include <time.h>
+#include <omp.h>
+#include "medium_morus.h"
+#include "aes-rng.h"

### 2018-03-23 src/medium_morus.c

@@ -1,0 +1,126 @@
+#include "medium_morus.h"
+
+void null_word(state_words* w) {
+   *w = 0;
+}
+
+inline void print_state(state state) {
+  printf("%016lx %016lx %016lx %016lx %016lx\n", state[0], state[1], state[2], state[3], state[4]);
+}
+
+inline uint64_t rotate_left(uint64_t x, int bits)
+{
+  if (bits == 0) return x;
+  return (x << bits) | (x >> (64 - bits));
+}
+
+void rand_init(state st, struct RNG_state* seed){
+  st[0] = aesrand_int32_r(seed);
+  st[0] <<= 32;
+  st[0] |= aesrand_int32_r(seed);
+  st[1] = aesrand_int32_r(seed);
+  st[1] <<= 32;
+  st[1] |= aesrand_int32_r(seed);
+  st[2] = aesrand_int32_r(seed);
+  st[2] <<= 32;
+  st[2] |= aesrand_int32_r(seed);
+  st[3] = aesrand_int32_r(seed);
+  st[3] <<= 32;
+  st[3] |= aesrand_int32_r(seed);
+  st[4] = aesrand_int32_r(seed);
+  st[4] <<= 32;
+  st[4] |= aesrand_int32_r(seed);
+}
+
+void iterate(state st, state_words message) {
+	st[0] ^= st[1] & st[2];
+	st[0] ^= st[3];
+	st[0] = rotate_left(st[0], 13);
+
+	st[1] ^= message;
+	st[1] ^= st[2] & st[3];
+	st[1] ^= st[4];
+	st[1] = rotate_left(st[1], 46);
+
+	st[2] ^= message;
+	st[2] ^= st[3] & st[4];
+	st[2] ^= st[0];
+	st[2] = rotate_left(st[2], 38);
+
+	st[3] ^= message;
+	st[3] ^= st[4] & st[0];
+	st[3] ^= st[1];
+	st[3] = rotate_left(st[3], 7);
+
+	st[4] ^= message;
+	st[4] ^= st[0] & st[1];
+	st[4] ^= st[2];
+	st[4] = rotate_left(st[4], 4);
+}
+
+void encrypt(state_words* out, state_words message, state st)
+{
+	state_words mem = message;
+	message ^= st[0];
+	message ^= st[1];
+	message ^= st[2] & st[3];
+	iterate(st, mem);
+  *out = message;
+}
+
[…]

### 2018-03-23 src/medium_morus.h

@@ -1,0 +1,19 @@
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include "aes-rng.h"
+
+typedef uint64_t state_words;
+typedef state_words state[5];
+
+void print_state(state st);
+
+void rand_init(state st, struct RNG_state* rng);
+
+void iterate(state st, state_words message);
+
+void encrypt(state_words* out, state_words message, state state);
+
+int linear(state* saved_state, state_words* saved_cipher);
+
+void null_word(state_words* w);

### 2018-03-24 morus.tex
...
Confirmed correlation for two halves of 1280 trail
added part of Yu's observations to tex

@@
 \usepackage{subcaption}
+\usepackage{booktabs}
+\usepackage{siunitx}
+\usepackage[section]{placeins}
+\usepackage[hidelinks]{hyperref}
 
@@
 \newif\ifsubstates\substatesfalse
-
-\begin{document}
-
-\noindent
-2 approximations for $S^{2,2}_{0}$:
-\begin{align*}
-  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus
-  C^4_{19} \tag{corr. $2^{-9}$} \\
-\intertext{Combined ciphertext-only approximation:}
-  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-16.5}$}
-\end{align*}
 
@@
 
-\begin{figure}
+\begin{document}
+
+\section{Description of \cipher{MORUS} and \cipher{MiniMORUS}}
+
+\begin{figure}[h]
   \substatesfalse
@@
 \end{figure}
+
+\clearpage
+
+\section{Linear Approximations for \cipher{MiniMORUS}}
+
+Basic idea using the fragments in \autoref{fig:minimorusfragments} and a fixed message $M=0$:
+\begin{itemize}
+    \item $\alpha_i$ approximates (one bit of) state word $S^0$ with the ciphertext
+    \item $\beta_i$ approximates $S^1$ using $S^0$ and the ciphertext
+    \item $\gamma_i$ approximates $S^4$ using two approximations of $S^1$ in consecutive steps
+    \item $\delta_i$ approximates $S^2$ using two approximations of $S^4$ in consecutive steps
+    \item $\varepsilon_i$ approximates $S^0$ using two approximations of $S^2$ in consecutive steps
+\end{itemize}
+Thus, we obtain a ciphertext-only approximation by appropriately connecting
+$1 \times \alpha$ with $1 \times \varepsilon$, $2 \times \delta$, $4 \times \gamma$, $8 \times \beta$, $8 \times \alpha$.
+The attack is essentially independent of the rotation constants $b_i$.
+The resulting biased ciphertext expression is listed for both
+\cipher{MiniMORUS-640} and \cipher{MiniMORUS-1280} in \autoref{tab:miniapproximations}.
+Since each fragment has weight 1, we would expect a total weight of 24;
+however, the approximations are not independent (\autoref{fig:todo}).
+The actual weight is 16, and experiments confirm an even higher correlation of $2^{-15.5}$.
+
+%\noindent
+%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
+%\begin{align*}
+%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
+%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
+%\intertext{Combined ciphertext-only approximation:}
+%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
+%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
+%\end{align*}
+%\bigskip
+%
+%\noindent
+%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
+%\begin{align*}
+%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
+%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
+%\intertext{Combined ciphertext-only approximation:}
+%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
+%\end{align*}
+
+\begin{table}[hb!]
+  \caption{Verification of approximations}
+  \label{tab:miniapproximations}
+  \begin{tabular}{@{}llSSS@{}}
+    \toprule
+    & & \multicolumn{3}{@{}c@{}}{Weight} \\
+    \cmidrule{3-5}
+    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS-640}}
+             & {Exp.} & {Bool.} & {Meas.} \\
+    \midrule
+    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
+             & 7    & 7     & 7 \\
+    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
+             & 9    & 9     & 9 \\
+    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
+             & 16   & 16    & 15.5 \\
+    \midrule
+    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS-1280}} \\
+    \midrule
+    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
+             & 7    & 7     & 7 \\
+    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
+             & 9    & 9     & 9 \\
+    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
+             & 16   & 16    & {??} \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
   \caption{\cipher{MiniMORUS} linear approximation fragments.}
-  \label{fig:minimorus}
+  \label{fig:minimorusfragments}
 \end{figure}
+
+\begin{figure}
+  \substatesfalse
+  % \substatesfalse to label state words and/or masks
+  \centering
+  \begin{subfigure}{.32\textwidth}
+  \centering
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail, beta]
+      (C) -- node[right] {$i$} (lll-1)
+      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
+      (tlll-1) -- (W-20) node[above] {$i$}
+      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
+      %(xnd-1) -- (and-1)
+      (W40) node[below] {\phantom{$i$}}
+      ;
+    \draw[trail, gamma]
+      %(W-21) node[above] {$i$}
+      (txor-1) -- (xnd1)
+      (xnd1) -- (xor1) (xor1) -- (lll1)
+      %(xnd1) -- (and1)
+      (xor1) -- (txor1)
+      (txor1) -- (W-24) node[above] {$i$}
+      (lll1) -- (W41) node[below] {$i+b_1$}
+      ;
+    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
+    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
+    \draw[trail, beta]  (xnd-1) -- ++(-.5,0) coordinate (tmp) -- (tmp|-xnd1);
+    \draw[trail, gamma] (tmp|-xnd1) node[below, black] {=} -- (xnd1);
+  \end{tikzpicture}%}}}
+  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}{.32\textwidth}
+  \centering
+  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+    \printstate
+    \draw[trail, beta]
+      (C) -- node[right] {$i$} (lll-1)
+      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
+      (tlll-1) -- (W-20) node[above] {$i$}
+      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
+      (xnd-1) -- (and-1)
+      (W40) node[below] {\phantom{$i$}}
+      ;
+    \draw[trail, gamma]
+      %(W-21) node[above] {$i$}
+      (txor-1) -- (xnd1)
+      (xnd1) -- (xor1) (xor1) -- (lll1)
+      (xnd1) -- (and1)
+      (xor1) -- (txor1)
+      (txor1) -- (W-24) node[above] {$i$}
+      (lll1) -- (W41) node[below] {$i+b_1$}
+      ;
+    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
+    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
+  \end{tikzpicture}%}}}
+  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
+  \end{subfigure}
+  \caption{Weight of $\beta_i \oplus \gamma_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
+  \label{fig:minimorusfragments}
+\end{figure}
+
+TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
+(maybe not relevant due to overall complexity\dots)
 
@@
 
-      \draw (\S{4}{2}) node[above] {\normalsize weight 7 (not 9)};
+      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
     \end{scope}
@@
 
-      \draw (\S{5}{2}) node[above] {\normalsize weight 9 (not 11)};
+      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
       \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
@@
 
-      \draw (\S{4}{2}) node[above] {\normalsize weight 7 (not 9)};
+      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
       \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
@@
 
-      \draw (\S{5}{2}) node[above] {\normalsize weight 9 (not 11)};
+      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
     \end{scope}
@@
 
-\begin{figure}
+\clearpage
 
-  \caption{Suspicious things I'd like to verify with full \cipher{MORUS}}
-\end{figure}
+\section{Linear Approximations for \cipher{MORUS}}
+
+\section{Differential Attack on the Finalization}
+
+\section{Key-Recovery Attack in Nonce-Misuse Setting}
 

### 2018-03-24 minimorus.py

@@ -5,18 +5,32 @@ import math
 import sys
 from functools import reduce
 
-width=64
 approximations = dict( \
-        alpha=dict(w=1, masks=lambda S, C, width: [mask(C[0], [27]),
-                                                   mask(S[1][0], [0])]) if width==32 else \
-              dict(w=1, masks=lambda S, C, width: [mask(C[0], [51]),
+        alpha=dict(w=1, masks=lambda S, C, width: [mask(C[0], [27 if width==32 else 51]),
                                                    mask(S[1][0], [0])]),
+        beta =dict(w=1, masks=lambda S, C, width: [mask(C[0], [0]),
+                                                   mask(S[0][0], [0]),
+                                                   mask(S[0][1], [0])]),
+        gamma=dict(w=1, masks=lambda S, C, width: [mask(S[0][1], [0]),
+                                                   mask(S[0][4], [0]),
+                                                   mask(S[1][1], [31 if width==32 else 46])]),
+        delta=dict(w=1, masks=lambda S, C, width: [mask(S[0][4], [0]),
+                                                   mask(S[1][2], [0]),
+                                                   mask(S[1][4], [13 if width==32 else 4])]),
+        epsil=dict(w=1, masks=lambda S, C, width: [mask(S[0][2], [0]),
+                                                   mask(S[1][0], [0]),
+                                                   mask(S[1][2], [7 if width==32 else 38])]),
+        #
+        #alphabeta=dict(w=2, masks=lambda S, C, width: [mask(S[0][0], [0]),
+        #                                               mask(S[0][1], [0]),
+        #                                               mask(S[1][0], [5 if width==32 else 13])]),
+        #
         appr1=dict(w=7, masks=lambda S, C, width: [mask(C[0], [27]),
                                                    mask(C[1], [0,26,8]),
                                                    mask(C[2], [31,13,7]),
                                                    mask(C[3], [12]),
-                                                   mask(S[2][2], [0])]) if width==32 else \
-              dict(w=7, masks=lambda S, C, width: [mask(C[0], [51]),
+                                                   mask(S[2][2], [0])] if width==32 else \
+                                                  [mask(C[0], [51]),
                                                    mask(C[1], [0,33,55]),
                                                    mask(C[2], [4,37,46]),
                                                    mask(C[3], [50]),
@@ -25,8 +39,8 @@ approximations = dict( \
[…]

### 2018-04-08 morus.tex
started writeup of the attack for sending to Hongjun Wu

@@
-\documentclass[a4paper, 11pt]{article}
+\newif\ifanonymous
+%\anonymoustrue%uncomment for submission version
 
-\usepackage{geometry}
+\documentclass{llncs}
+
+\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
+
 \usepackage{amsmath,amssymb}
@@
 \usepackage[hidelinks]{hyperref}
+
+\usepackage{xcolor}
+\definecolor{blue_link}{RGB}{0,0,200}
+\hypersetup{
+  colorlinks = true,
+  linkcolor = blue_link,
+  citecolor = blue_link,
+  urlcolor = blue_link,
+  bookmarksdepth = 3
+}
 
@@
 
+%new commands
+\newcommand{\bias}{\ensuremath{\operatorname{bias}}}
+\newcommand{\weight}{\ensuremath{\operatorname{weight}}}
+\newcommand\eqdef{:=}%equality by definition
+\newcommand\StateUpdate{\ensuremath{\texttt{StateUpdate}}}
+\newcommand\rotl{\ensuremath{\lll}}
+\newcommand\rotlxy{\ensuremath{\lll_W}}
+
+\ifanonymous
+\author{}
+\institute{}
+\else
+\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Beno\^it Viguier, Yu Sasaki}
+\institute{}
+%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
+\fi
+
 \begin{document}
 
-\section{Description of \cipher{MORUS} and \cipher{MiniMORUS}}
+\mainmatter
+
+\title{Note on Linear Biases in MORUS}
+
+\maketitle
+
+\begin{abstract}
+In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
+More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
+Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
+In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is fixed and known.
+However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
+We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
+\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
+\end{abstract}
+
+\section{Introduction}
+
+Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
+As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
+Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
+Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initalize the inner state of MORUS are irrelevant: the bias will exist regardless of their value.
+
+\section{Notation}
+
+We define the \emph{bias} of an event $E$ as:
+\[
+\bias(E) \eqdef 2\Pr(E)-1.
+\]
+Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
+\[
+\weight(E) \eqdef -\log|\bias(E)|
+\]
+where $\log$ denotes logarithm in base 2.
+By the piling-up lemma, the bias (resp.
+weight) of an XOR of independent variables is equal to the product (resp.
+sum) of their individual biases (resp.
+weights).
+
+We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
+\centerline{
+\def\arraystretch{1.25}
+\begin{tabular}{lcl}
+$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
+$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
+$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
+$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
+\end{tabular}
+}
+In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
+Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
+In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
+
+Finally, let $x$ denote a register.
+Recall that register of \cipher{MORUS} are composed of four words.
+We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
+\centerline{
+\def\arraystretch{1.25}
+\begin{tabular}{lcl}
+$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
+$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
+\end{tabular}
+}
+
+\section{A Linear Bias in \cipher{MORUS}}
+
+To simplify the description of the attack, we assume all plaintext blocks are zero.
+However the same results would hold as long as the plaintext is fixed.
+Let $w$ denote the word size, i.e.
+32 for \cipher{MORUS640}, and 64 for \cipher{MORUS1280}.
+Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
+
+\subsection{Symmetric Linear Combinations}
+
+We begin with a few observations about the \StateUpdate{} function. Beside XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
+\begin{enumerate}
+\item \emph{word-wise} rotations perform a circular shift on each word within a register;
+\item \emph{register-wise} rotations perform a circular shift on a whole register.
+\end{enumerate}
+The second type of rotation always shifts registers by a multiple of the word size (i.e.
+32 for \cipher{MORUS640}).
+This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by 32 bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
+
+Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
+\begin{definition}
+Let $w$ denote the word size in bits, and $r = 4w$ denote the register size. A linear combination of the form:
+\[
+S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
+\]
+is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
+\[
+\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } r : i\leq k\}.
+\]
+\end{definition}
+\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
+\begin{equation}
+S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
+\label{eq:symmetric}
+\end{equation}
+
+This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
+The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
+On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetric property.
+Moreover, the XOR of two symmetric linear combinations is also symmetric; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
+
+This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
+As a result, the effect of register-wise rotations can be ignored.
+Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
+Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
+For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
+
+\subsection{\cipher{MiniMORUS}}
+
+In fact, we can go further an consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
+The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
+We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
+Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
+
+Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS640}.
+We will then extend it to the full \cipher{MORUS640} via the previous symmetry.
+The trail for \cipher{MORUS1280} will be essentially the same.
 
@@
   \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function (1 step = 5 rounds).}
+  \caption{\cipher{MiniMORUS} state update function.}
   \label{fig:minimorus}
@@
 
-\clearpage
+\section{Linear Trail for \cipher{MiniMORUS640}}
 
-\section{Linear Approximations for \cipher{MiniMORUS}}
+\subsection{Overview of the Trail}
 
-Basic idea using the fragments in \autoref{fig:minimorusfragments} and a fixed message $M=0$:
+To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
 \begin{itemize}
-    \item $\alpha_i$ approximates (one bit of) state word $S^0$ with the ciphertext
-    \item $\beta_i$ approximates $S^1$ using $S^0$ and the ciphertext
-    \item $\gamma_i$ approximates $S^4$ using two approximations of $S^1$ in consecutive steps
-    \item $\delta_i$ approximates $S^2$ using two approximations of $S^4$ in consecutive steps
-    \item $\varepsilon_i$ approximates $S^0$ using two approximations of $S^2$ in consecutive steps
+    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
+    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
+    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
+    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
+    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
 \end{itemize}
-Thus, we obtain a ciphertext-only approximation by appropriately connecting
-$1 \times \alpha$ with $1 \times \varepsilon$, $2 \times \delta$, $4 \times \gamma$, $8 \times \beta$, $8 \times \alpha$.
-The attack is essentially independent of the rotation constants $b_i$.
-The resulting biased ciphertext expression is listed for both
-\cipher{MiniMORUS-640} and \cipher{MiniMORUS-1280} in \autoref{tab:miniapproximations}.
-Since each fragment has weight 1, we would expect a total weight of 24;
-however, the approximations are not independent (\autoref{fig:todo}).
-The actual weight is 16, and experiments confirm an even higher correlation of $2^{-15.5}$.
 
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
+The trail fragments are depicted on \autoref{fig:minimorusfragments}, and may be represented as follows (recall that the $b_i$'s are rotation constants):
+\begin{alignat*}{2}
+\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
+\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
+\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
+\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
+\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
+\end{alignat*}
 
-\begin{table}[hb!]
-  \caption{Verification of approximations}
-  \label{tab:miniapproximations}
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS-640}}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS-1280}} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & {??} \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
+In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, so their sum can be written as:
+\[
+x \cdot y \oplus x \cdot z = x \cdot (y \oplus z).
+\]
+As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
+Another way to look at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \autoref{fig:minimorusfragments}.
 
@@
       \intertext{\cipher{MiniMORUS-640}:}
-      \alpha_i:      && C_i &\to S^{0}_{i+5} \\
-      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
-      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+31} \\
-      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+13} \\
-      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+7}
+\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
+\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
+\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
+\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
+\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
       \intertext{\cipher{MiniMORUS-1280}:}
-      \alpha_i:      && C_i &\to S^{0}_{i+13} \\
-      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
-      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+46} \\
-      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+4} \\
-      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+38}
+\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
+\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
+\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
+\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
+\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
     \end{align*}
@@
 \end{figure}
+
+Overall, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
+By combining $\alpha^t_i$ with $\beta^t_i$, we are then able to approximate a bit of $S_1$ using only ciphertext bits (from two consecutive rounds).
+Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
+by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$, we are able to approximate one bit of $S_4$ using only ciphertext bits.
+Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive rounds; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice}.
+In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
+In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
+
+However, the same bit of $S_0$ can also be approximated much more simply by using $\alpha^t_i$ at the corresponding step.
+Thus that bit can be approximated from two different sides: the first one uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps.
+The second only involves using $\alpha^t_i$ at the final step reached by the previous trail.
+By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption rounds.
+
+Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive rounds: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
+Then $\alpha^t_i$ is used one final time to close the trail.
+One may naturally wonder if some components of this trail are in conflict.
+In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
+To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitely.
+
+\subsection{Trail Equation}
+
+The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitely as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
+For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
+\begin{alignat*}{2}
+A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
+B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
+C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
+D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^t_{0,i} \cdot S^t_{1,i}\\
+E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
+\end{alignat*}
+
+From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ on the left-hand side cancel out.
+Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
+By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
+We now set out to do so, and compute the equation of the full trail explicitely.
+
+In order to build the equation for the full trail, we start with $E^3_0$:
+\[
+S^3_{2,0} \oplus S^{4}_{2,b_0} \oplus S^{4}_{0,0} = S^3_{3,0} \cdot S^t_{4,0}.
+\]
+In order to cancel the $S^{4}_{0,0}$ term on the left-hand side, we add in the equation $A^3_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
+This yields:
+\begin{align*}
+&S^3_{2,0} \oplus S^{4}_{2,7} \oplus C^{3}_{0,-b_0}\\
+=\; &S^3_{3,0} \cdot S^t_{4,0} \oplus S^t_{1,-b_0} \oplus S^t_{3,-b_0} \oplus S^3_{1,-b_0} \cdot S^3_{2,-b_0} \oplus S^3_{2,-b_0} \cdot S^t_{3,-b_0}.
+\end{align*}
+We then need to cancel two terms of the form $S^t_{2,i}$.
+To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
+This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
+By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
+By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
+Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
+We do not write the whole process as it is rather tedious.
+Ultimately, for \cipher{MORUS640}, this yields the equation:
+
+\allowdisplaybreaks
+\begin{flalign*}
+&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
+&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
+&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
+&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
+&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
+&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
+&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
+&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
+&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
+&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
+&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
+&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
+&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
+&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
+&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
+&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
+&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
+&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
+\end{flalign*}
+
+\subsection{Bias of the Trail}
+
+In the equation from the previous section, each line on the right-hand side of the equality involves distinct terms, and each line has a weight of 1.
+By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
+Hence the linear combination of ciphertext bits on the left-hand side has a bias of $2^{-16}$.
+
+That the bias is so high may be suprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
+Since each trail fragment has a weight of 1, this would suggest that the total bias should be $1+2+4+8+9 = 24$ rather than 16.
+However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same round between registers $S_2$ and $S_3$.
+In both cases it is approximated by zero.
+When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
+Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
+
+\subsection{Experiments}
+
+To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
+In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
+Results are displayed on \autoref{tab:miniapproximations}.
+While our analysis from the previous section predicts a bias of $2^{-16}$, experiments confirm a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
+The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different rounds are not completely independent.
+
+%\noindent
+%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
+%\begin{align*}
+%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
+%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
+%\intertext{Combined ciphertext-only approximation:}
+%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
+%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
+%\end{align*}
+%\bigskip
+%
+%\noindent
+%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
+%\begin{align*}
+%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
+%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
+%\intertext{Combined ciphertext-only approximation:}
+%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
+%\end{align*}
+
+\begin{table}[h!]
+  \caption{Experimental verification of trail biases.}
+  \label{tab:miniapproximations}
+  \centerline{
+  \begin{tabular}{@{}llSSS@{}}
+    \toprule
+    & & \multicolumn{3}{@{}c@{}}{Weight} \\
+    \cmidrule{3-5}
+    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
+             & {Exp.} & {Bool.} & {Meas.} \\
+    \midrule
+    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
+             & 7    & 7     & 7 \\
+    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
+             & 9    & 9     & 9 \\
+    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
+             & 16   & 16    & 15.5 \\
+    \midrule
+    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
+    \midrule
+    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
+             & 7    & 7     & 7 \\
+    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
+             & 9    & 9     & 9 \\
+    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
+             & 16   & 16    & {??} \\
+    \bottomrule
+  \end{tabular}
+  }
+\end{table}
 
@@
 
-TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
-(maybe not relevant due to overall complexity\dots)
+%TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
+%(maybe not relevant due to overall complexity\dots)
 

### 2018-04-08 morus.tex
more writing

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{cleveref}
+
+\allowdisplaybreaks%this is to allow page breaks within multiline equations (to accomodate full trail equations)
 
@@
 }
+In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
+
 In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
@@
 
-\section{A Linear Bias in \cipher{MORUS}}
+\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
+\label{sec:introminimorus}
 
@@
 Let $w$ denote the word size, i.e.
-32 for \cipher{MORUS640}, and 64 for \cipher{MORUS1280}.
+$w = 32$ in the case of \cipher{MORUS640}, and $w = 64$ for \cipher{MORUS1280}.
 Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
@@
 \end{enumerate}
-The second type of rotation always shifts registers by a multiple of the word size (i.e.
-32 for \cipher{MORUS640}).
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by 32 bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
+The second type of rotation always shifts registers by a multiple of the word size $w$.
+This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
 
@@
 \begin{definition}
-Let $w$ denote the word size in bits, and $r = 4w$ denote the register size. A linear combination of the form:
+Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
 \[
@@
 
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS640}.
-We will then extend it to the full \cipher{MORUS640} via the previous symmetry.
-The trail for \cipher{MORUS1280} will be essentially the same.
+Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
 
@@
 
-\section{Linear Trail for \cipher{MiniMORUS640}}
+\section{Linear Trail for \cipher{MiniMORUS}}
 
@@
 
-The trail fragments are depicted on \autoref{fig:minimorusfragments}, and may be represented as follows (recall that the $b_i$'s are rotation constants):
+The trail fragments are depicted on \Cref{fig:minimorusfragments}, and may be represented as follows (recall that the $b_i$'s are rotation constants):
 \begin{alignat*}{2}
@@
 As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way to look at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \autoref{fig:minimorusfragments}.
+Another way to look at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
 
@@
 \subsection{Trail Equation}
+\label{sec:minitraileq}
 
@@
 Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-We do not write the whole process as it is rather tedious.
 Ultimately, for \cipher{MORUS640}, this yields the equation:
 
-\allowdisplaybreaks
 \begin{flalign*}
@@
 
+The trail for \cipher{MORUS1280} is very similar, and is given in \Cref{sec:traileq}.
+
 \subsection{Bias of the Trail}
+\label{sec:minibias}
 
-In the equation from the previous section, each line on the right-hand side of the equality involves distinct terms, and each line has a weight of 1.
+In the equation for \cipher{MORUS640} from the previous section, each line on the right-hand side of the equality involves distinct terms, and each line has a weight of 1.
 By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
 Hence the linear combination of ciphertext bits on the left-hand side has a bias of $2^{-16}$.
+The same holds for \cipher{MORUS1280} (cf.
+\Cref{sec:traileq}).
 
@@
 In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
-Results are displayed on \autoref{tab:miniapproximations}.
+Results are displayed on \Cref{tab:miniapproximations}.
 While our analysis from the previous section predicts a bias of $2^{-16}$, experiments confirm a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
@@
 \end{table}
+
+\section{Trail for Full \cipher{MORUS}}
+
+\subsection{Symmetrizing the Trail}
+In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
+We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
+As explained in \Cref{sec:introminimorus}, translating a trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher just amounts to making the linear combinations symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
+That is, for every $S^t_{i,j}$ (resp.
+$C^t_j$) term in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
+$C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
+For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
+\[
+S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
+\]
+
+In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
+\begin{align*}
+&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
+\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
+\end{align*}
+yields a biased linear combination on the full \cipher{MORUS640}, namely:
+\begin{align*}
+&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
+\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
+\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
+\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
+\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
+\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
+\end{align*}
+We refer the reader to \Cref{sec:traileq} for the corresponding equation on \cipher{MORUS1280}.
+
+\subsection{Bias of the Full Trail}
+
+Due to the observations from \autoref{sec:introminimorus}, the symmetrized trail on full \cipher{MORUS} more or less consists of four copies of the original trail on \cipher{MiniMORUS}.
+As such, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
+However, reality is a little more complex, as the symmetrized trail does not simply behave as four copies of the original trail.
+
+To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
+Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiple of $w$ bits are ignored.
+However in the real \cipher{MORUS} these terms are of course distinct and do not cancel out.
+
+In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
+Indeed, $\beta^t_i$ involves approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t{3,i-w}$ (cf.
+\Cref{fig:trailcollision}).
+While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
+\begin{align}
+&S^t_{2,i} \cdot S^t_{3,i} \oplus S^t_{3,i} \cdot S^t_{2,i+w}
+\oplus S^t_{2,i+w} \cdot S^t_{3,i+w} \oplus S^t_{3,i+w} \cdot S^t_{2,i+2w}\notag\\
+\oplus\; &S^t_{2,i+2w} \cdot S^t_{3,i+2w} \oplus S^t_{3,i+2w} \cdot S^t_{2,i+3w}
+\oplus S^t_{2,i+3w} \cdot S^t_{3,i+3w} \oplus S^t_{3,i+3w} \cdot S^t_{2,i}.\label{eq:8circle}
+\end{align}
+It may be observed that the products occuring in the equation above involve 8 terms in a circular ring. The weight of this expression can be computed by brute force, and is equal to $3$.
 
@@
   \end{subfigure}
-  \caption{Weight of $\beta_i \oplus \gamma_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
-  \label{fig:minimorusfragments}
+  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
+  \label{fig:trailcollision}
 \end{figure}
+
+In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
+This results in a total weight of $4 \cdot 16 + 12 = 76$ (where 16 is the weight of the trail on \cipher{MiniMORUS1280}).
+Due to the complexity of the trail, we have confirmed that this bias is correct by explicitely computing the full trail equation in \Cref{sec:traileq}, and computing its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
+That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
+In the end, the full trail equation does confirm a weight of 76 for the full trail on \cipher{MORUS1280}.
+
+In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
+More specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t{3,i}$ occurs.
+As explained previously, a quasi-collision with the term $S^t_{2,i} \cdot S^t{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
+However for \cipher{MORUS640}, trail fragment $\alpha^t_{i+d}$ is once used with an offest of $d = b_1+b_4-b_0-b_2 = 31+13-5-7 = 0 \;\text{mod}\; 32$ relative to $\gamma^t_i$.
+This creates another term $S^t_{2,i} \cdot S^t{3,i}$, and destroys one of the four occurences of \Cref{eq:8circle}.
+In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
+\Cref{sec:traileq}).
 
@@
   \end{tikzpicture}
-  \caption{\cipher{MiniMORUS}: Two approximations for $S^{2,2}_0$}
+  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
+  \label{fig:twoapproximations}
 \end{figure}
 
-\clearpage
+\section{Discussion}
 
-\section{Linear Approximations for \cipher{MORUS}}
+%This is just a sketch for now...
 
-\section{Differential Attack on the Finalization}
+We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
+As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
+This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
 
-\section{Key-Recovery Attack in Nonce-Misuse Setting}
+The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
+However, since our attack is independent of the encryption key, and hence immune to rekeying, the limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
+
+With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
+The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
+This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
+
+However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
+Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
+That it is possible to build a trail that ignores rotation constants may be seen as surprising.
+This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation---it seems that this would defeat any trail of the type uncovered in this work.
+
+\appendix
+
+\section{Trail Equations}
+\label{sec:traileq}
+
+In this section, we provide full trail equations for all variants of \cipher{MORUS}.
+In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each connected connected component.
+By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
+
+\subsection{Trail Equation for \cipher{MiniMORUS640}}
+
+This trail equation was already given in \Cref{sec:minitraileq}.
+
+\begin{flalign*}
+&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
+&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
+&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
+&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
+&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
+&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
+&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
+&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
+&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
+&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
+&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
+&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
+&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
+&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
+&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
+&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
+&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
+&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
+\end{flalign*}
+The total weight of the trail is 16.
+
+\subsection{Trail Equation for \cipher{MiniMORUS1280}}
+
+\begin{flalign*}
+&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
+&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
+&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
+\end{flalign*}
+The total weight of the trail is 16.
+
+\subsection{Trail Equation for full \cipher{MORUS640}}
+
+\begin{flalign*}
+&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
+&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
+&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
+&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
+&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
+&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
+%
+&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
+&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
+%
+&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
+&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
+%
+&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
+&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
+%
+&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
+&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
+&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
+&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
+\end{flalign*}
+The total weight of the trail is 73.
+
+\subsection{Trail Equation for full \cipher{MORUS1280}}
+
+\begin{flalign*}
+&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
+&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
+&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
+&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
+&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
+&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
+&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
+%
+&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
+&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
+%
+&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
+&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
+%
+&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
+&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
+%
+&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
+&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
+%
+&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
+&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
+&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
+&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
+\end{flalign*}
+The total weight of the trail is 76.
 

### 2018-04-08 morus.tex
tweaks

@@
 Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is fixed and known.
+In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
 However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
@@
 Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initalize the inner state of MORUS are irrelevant: the bias will exist regardless of their value.
+Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initalize the inner state of MORUS are irrelevant: the bias is independent of their value.
 
@@
 \]
-where $\log$ denotes logarithm in base 2.
+where $\log()$ denotes logarithm in base 2.
 By the piling-up lemma, the bias (resp.
@@
 Finally, let $x$ denote a register.
-Recall that register of \cipher{MORUS} are composed of four words.
+Recall that registers of \cipher{MORUS} are composed of four words.
 We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
@@
 
+The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
+Throughout, $w$ denotes the  word size, i.e.
+$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
+
 \section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
@@
 However the same results would hold as long as the plaintext is fixed.
-Let $w$ denote the word size, i.e.
-$w = 32$ in the case of \cipher{MORUS640}, and $w = 64$ for \cipher{MORUS1280}.
 Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
@@
 \[
-\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } r : i\leq k\}.
+\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
 \]
@@
 The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetric property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
+On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
+Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
 
@@
 
-In fact, we can go further an consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
+In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
 The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
@@
 
-The trail fragments are depicted on \Cref{fig:minimorusfragments}, and may be represented as follows (recall that the $b_i$'s are rotation constants):
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
+The trail fragments are depicted on \Cref{fig:minimorusfragments}.
 In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, so their sum can be written as:
+In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, so their sum may be written in the form:
 \[
@@
 As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way to look at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
+Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
 
@@
   \end{tikzpicture}%}}}
-  \caption*{$\alpha_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
+  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
   \end{subfigure}
@@
   \end{tikzpicture}%}}}
-  \caption*{$\beta_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
+  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
   \end{subfigure}
@@
   \end{tikzpicture}%}}}
-  \caption*{$\gamma_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
+  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
   \end{subfigure}
@@
   \end{tikzpicture}%}}}
-  \caption*{$\delta_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
+  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
   \end{subfigure}
@@
   \end{tikzpicture}%}}}
-  \caption*{$\varepsilon_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
+  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
   \end{subfigure}
@@
   \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear approximation fragments.}
+  \caption{\cipher{MiniMORUS} linear trail fragments.}
   \label{fig:minimorusfragments}
@@
 
-Overall, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^t_i$, we are then able to approximate a bit of $S_1$ using only ciphertext bits (from two consecutive rounds).
+The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
+\begin{alignat*}{2}
+\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
+\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
+\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
+\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
+\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
+\end{alignat*}
+
+In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
+By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive rounds.
 Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive rounds; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice}.
+by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
+Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive rounds; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
 In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
@@
 
-However, the same bit of $S_0$ can also be approximated much more simply by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be approximated from two different sides: the first one uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps.
-The second only involves using $\alpha^t_i$ at the final step reached by the previous trail.
+However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
+Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
+The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
 By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption rounds.
@@
 
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ on the left-hand side cancel out.
+From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
 Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
 By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so, and compute the equation of the full trail explicitely.
+We now set out to do so.
 
@@
 
-The trail for \cipher{MORUS1280} is very similar, and is given in \Cref{sec:traileq}.
+The equation for \cipher{MORUS1280} is very similar, and is given in \Cref{sec:traileq}.
 
@@
 
-In the equation for \cipher{MORUS640} from the previous section, each line on the right-hand side of the equality involves distinct terms, and each line has a weight of 1.
+In the equation for \cipher{MORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
 By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has a bias of $2^{-16}$.
+Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
 The same holds for \cipher{MORUS1280} (cf.
@@
 That the bias is so high may be suprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total bias should be $1+2+4+8+9 = 24$ rather than 16.
+Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
 However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same round between registers $S_2$ and $S_3$.
@@
 
-\subsection{Experiments}
+\subsection{Experimental Verification}
 
@@
 Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments confirm a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
+While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
 The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different rounds are not completely independent.
@@
 We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-As explained in \Cref{sec:introminimorus}, translating a trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher just amounts to making the linear combinations symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
+In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
 That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) term in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
+$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
 $C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
@@
 
+Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
+Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
 In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
@@
 \end{align*}
-yields a biased linear combination on the full \cipher{MORUS640}, namely:
+utlimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
 \begin{align*}
@@
 \end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding equation on \cipher{MORUS1280}.
+We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
 
@@
 
-Due to the observations from \autoref{sec:introminimorus}, the symmetrized trail on full \cipher{MORUS} more or less consists of four copies of the original trail on \cipher{MiniMORUS}.
-As such, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
-However, reality is a little more complex, as the symmetrized trail does not simply behave as four copies of the original trail.
+The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
+Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
+But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
+
+Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
+However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
 
@@
 In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
-Indeed, $\beta^t_i$ involves approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t{3,i-w}$ (cf.
+Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
 \Cref{fig:trailcollision}).
@@
 \end{align}
-It may be observed that the products occuring in the equation above involve 8 terms in a circular ring. The weight of this expression can be computed by brute force, and is equal to $3$.
+It may be observed that the products occuring in the equation above involve 8 terms in a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
 
@@
 In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (where 16 is the weight of the trail on \cipher{MiniMORUS1280}).
+This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
 Due to the complexity of the trail, we have confirmed that this bias is correct by explicitely computing the full trail equation in \Cref{sec:traileq}, and computing its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
 That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation does confirm a weight of 76 for the full trail on \cipher{MORUS1280}.
+In the end, the full trail equation does yield a weight of 76 for the full trail on \cipher{MORUS1280}.
 
 In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
-More specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t{3,i}$ occurs.
-As explained previously, a quasi-collision with the term $S^t_{2,i} \cdot S^t{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However for \cipher{MORUS640}, trail fragment $\alpha^t_{i+d}$ is once used with an offest of $d = b_1+b_4-b_0-b_2 = 31+13-5-7 = 0 \;\text{mod}\; 32$ relative to $\gamma^t_i$.
-This creates another term $S^t_{2,i} \cdot S^t{3,i}$, and destroys one of the four occurences of \Cref{eq:8circle}.
+Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
+As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
+However trail fragment $\alpha^t_{i+d}$ is once used with an offest of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
+This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurences of \Cref{eq:8circle}.
 In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
@@
 The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, the limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
+However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
 
@@
 Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be seen as surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation---it seems that this would defeat any trail of the type uncovered in this work.
+That it is possible to build a trail that ignores rotation constants may be surprising.
+This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
 
@@
 In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each connected connected component.
+In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
 By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.

### 2018-04-09 morus.tex
minor fix

@@
 In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, so their sum may be written in the form:
+In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
 \[
-x \cdot y \oplus x \cdot z = x \cdot (y \oplus z).
+x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
 \]
@@
 In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive rounds.
+By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
 Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
 by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive rounds; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
+Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
 In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
@@
 The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption rounds.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive rounds: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
+By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
+
+Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
 Then $\alpha^t_i$ is used one final time to close the trail.
@@
 C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^t_{0,i} \cdot S^t_{1,i}\\
+D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
 E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
@@
 
-In order to build the equation for the full trail, we start with $E^3_0$:
+In order to build the equation for the full trail, we start with $E^2_0$:
 \[
-S^3_{2,0} \oplus S^{4}_{2,b_0} \oplus S^{4}_{0,0} = S^3_{3,0} \cdot S^t_{4,0}.
+S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
 \]
-In order to cancel the $S^{4}_{0,0}$ term on the left-hand side, we add in the equation $A^3_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
+In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
 This yields:
 \begin{align*}
-&S^3_{2,0} \oplus S^{4}_{2,7} \oplus C^{3}_{0,-b_0}\\
-=\; &S^3_{3,0} \cdot S^t_{4,0} \oplus S^t_{1,-b_0} \oplus S^t_{3,-b_0} \oplus S^3_{1,-b_0} \cdot S^3_{2,-b_0} \oplus S^3_{2,-b_0} \cdot S^t_{3,-b_0}.
+&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
+=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
 \end{align*}
@@
 Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same round between registers $S_2$ and $S_3$.
+However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
 In both cases it is approximated by zero.
@@
 While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different rounds are not completely independent.
+The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
 
@@
 To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiple of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel out.
+Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
+However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
 
@@
 This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-Due to the complexity of the trail, we have confirmed that this bias is correct by explicitely computing the full trail equation in \Cref{sec:traileq}, and computing its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
+We have confirmed this bias by explicitely computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
 That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation does yield a weight of 76 for the full trail on \cipher{MORUS1280}.
+In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
 
@@
 As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used with an offest of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
+However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offest of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
 This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurences of \Cref{eq:8circle}.

### 2018-04-11 morus.tex
typo fix

@@
 Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initalize the inner state of MORUS are irrelevant: the bias is independent of their value.
+Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
 
@@
 In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitely.
+To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
 
@@
 
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitely as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
+The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
 For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
@@
 
-That the bias is so high may be suprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
+That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
 Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
@@
 \end{align*}
-utlimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
+ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
 \begin{align*}
@@
 \end{align}
-It may be observed that the products occuring in the equation above involve 8 terms in a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
+It may be observed that the products occurring in the equation above involve 8 terms in a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
 
@@
 This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitely computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
+We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
 That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
@@
 As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offest of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
-This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurences of \Cref{eq:8circle}.
+However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
+This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
 In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.

### 2018-04-11 morus.tex
fixed mistake in alphabetic ordering of authors (apologies to Yu)

@@
 \else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Beno\^it Viguier, Yu Sasaki}
+\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
 \institute{}
@@
 \end{align}
-It may be observed that the products occurring in the equation above involve 8 terms in a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
+It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
 

### 2018-04-13 src/morus.c
Fix big morus

@@ -196,6 +196,7 @@ void encrypt(state_words* out, state_words message, state st)
   // permute_words(st1,96);
 
   xor_to_word(st1,message);
+  xor_to_word(st1,st[0]);
   and_xor_to_word(st1,st[2],st[3]);
 	iterate(st, mem);
   copy_state_words(*out,st1);

### 2018-04-13 src/medium_morus.c
Fix gen_mask

@@ -70,7 +70,7 @@ void encrypt(state_words* out, state_words message, state st)
 
 inline uint64_t gen_mask(uint64_t* mask, unsigned int i)
 {
-  *mask |= (1 << i);
+  *mask |= ((uint64_t)1 << i);
   return 0;
 }
 

### 2018-04-13 morus.tex
Proofreading

@@
 \usepackage{tikz}
-\usetikzlibrary{calc,math}
+\usetikzlibrary{calc,math,arrows.meta}
 
@@
 Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MORUS640}, this yields the equation:
+Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
 
@@
 
-The equation for \cipher{MORUS1280} is very similar, and is given in \Cref{sec:traileq}.
+The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
 
@@
 
-In the equation for \cipher{MORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
+In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
 By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
 Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MORUS1280} (cf.
+The same holds for \cipher{MiniMORUS1280} (cf.
 \Cref{sec:traileq}).
@@
     $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & {??} \\
+             & 16   & 16    & 15.9 \\
     \bottomrule
@@
     %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
-    \draw[trail, beta]  (xnd-1) -- ++(-.5,0) coordinate (tmp) -- (tmp|-xnd1);
-    \draw[trail, gamma] (tmp|-xnd1) node[below, black] {=} -- (xnd1);
+    \node (eq) at ($(and-1|-and1)+(-1,-2.4pt)$) {=};
+    \draw[beta,{Circle[sep=-1.8pt]}-]  ($(and-1)+(0,-2.4pt)$) -| (eq);
+    \draw[gamma,{Circle[sep=-1.8pt]}-] ($(and1)+(0,-2.4pt)$) -- (eq);
   \end{tikzpicture}%}}}

### 2018-05-01 morusAC.tex
Conference version. Specification was added and merged with a part of Notations.

--- 2018-04-13 morus.tex
@@
 
-\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
+%\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
 
@@
 
-\section{Notation}
+\section{Preliminaries}
+\subsection{Specification of MORUS}
+MORUS is a family of authenticated ciphers with two internal state sizes: 640 and 1280 bits.
+Each member is called \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
+In the recommended parameters, \cipher{MORUS640} supports 128-bit keys and \cipher{MORUS1280} supports both 128-bit and 256-bit keys.
+The tag size is 128 bits or shorter, though the designers strongly recommend using a 128-bit tag.
+
+In MORUS, the internal state is divided in two five registers.
+Let $q$ be the register size, where $q = 128$ for \cipher{MORUS640} and $q=256$ for \cipher{MORUS1280}.
+Let $S_i, i \in\{0,1,2,3,4\}$ be a register.
+Then, the internal state is represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
+In MORUS, the internal state is first initialized by a key and nonce, and then updated by associated data blocks and message blocks step by step.
+Hence, we use a superscript to denote the step number of the state as $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is 0 before the initialization is performed.
+
+The computation of MORUS consists of the four parts; initialization, associated data process, encryption, and finalization.
+All parts share the core design of MORUS called the {\tt StepUpdate} function.
+Hereafter we explain {\tt StepUpdate}.
+
+\subsubsection{Step Update Function.}
+{\tt StepUpdate} takes as input the internal state $S^t = S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$ and additional register-wise value $m^t$.
+$m^t$ is supposed to be associated data block or message block.
+$m^t$ can sometimes be 0 or constant when the additional input is unnecessary.
+
+{\tt StepUpdate} is composed of 5 rounds with similar operations to update
+the state $S^t$, in which the additional input $m^t$ is used in rounds 2 to 5, but not in round 1.
+Each round use the word-wise left circular shift, which divides a $q$-bit register value into 4 words of $q/4$ bits, and performs left circular shift operation for every $q/4$-bit word.
+Throughout the paper, we denote the word size by $w$, i.e.
+$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
+The word-wise circular shift constants $b_i$ in round $i$ are defined in Table~\ref{Tbl:rcon}.
+Besides, the left rotation of a whole $q$-bit register is used.
+The register-wise circular shift constants are also listed in Table~\ref{Tbl:rcon}.
+\begin{table}[!htb]
+\centering
+\caption{Rotation Constants of MORUS} \label{Tbl:rcon}
+\begin{tabular}{c||ccccc|ccccc} \hline
+& \multicolumn{5}{c|}{register-wise circular shift} & \multicolumn{5}{c}{word-wise circular shift} \\
+                  & \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline
+\cipher{MORUS640} &  5 & 31 &  7 & 22 & 13 & 32 &  64 &  96 &  64 & 32 \\
+\cipher{MORUS1280}& 13 & 46 & 38 &  7 &  4 & 64 & 128 & 192 & 128 & 64 \\ \hline
+\end{tabular}
+\end{table}
+We denote those register-wise and word-wise left circular shift by $x \rotl y$ and $x \rotlxy y$, respectively.
+(Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document.)
+
+Let $\cdot$ denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
+\begin{align*}
+\textrm{Round 1:} &&
+S^{t+1}_0 &\leftarrow ( S_t^0 \oplus S^t_1 \cdot S^t_2 \oplus S^t_3 ) \lll_w b_0 &
+S^t_3 \leftarrow S^t_3 \lll b'_0.\\
+\textrm{Round 2:} &&
+S^{t+1}_1 &\leftarrow ( S_t^1 \oplus S^t_2 \cdot S^t_3 \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
+S^t_3 \leftarrow S^t_3 \lll b'_1.\\
+\textrm{Round 3:} &&
+S^{t+1}_2 &\leftarrow ( S_t^2 \oplus S^t_3 \cdot S^t_4 \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
+S^t_3 \leftarrow S^t_3 \lll b'_2.\\
+\textrm{Round 4:} &&
+S^{t+1}_3 &\leftarrow ( S_t^3 \oplus S^t_4 \cdot S^t_0 \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
+S^t_3 \leftarrow S^t_3 \lll b'_3.\\
+\textrm{Round 5:} &&
+S^{t+1}_4 &\leftarrow ( S_t^4 \oplus S^t_0 \cdot S^t_1 \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
+S^t_3 \leftarrow S^t_3 \lll b'_4.
+\end{align*}
+
+\subsubsection{Initialization.}
+The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K$ and a 128-bit nonce IV into the state, and running the {\tt StateUpdate} for 16 steps.
+The key and IV are loaded into the state as follows:
+\begin{align*}
+S^{-16}_0 = \textrm{IV}_{128}, &&
+S^{-16}_1 = K_{128}, &&
+S^{-16}_2 = 1^{128}, &&
+S^{-16}_3 = const_0, &&
+S^{-16}_4 = const_1.
+\end{align*}
+Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\cdots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K$.
+
+The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size. The state is initialized as
+\begin{align*}
+S^{-16}_0 = \textrm{IV}_{128} \| 0^{128}, &&
+S^{-16}_1 = K_{128}\|K_{128}, &&
+S^{-16}_2 = 1^{256}, \\
+S^{-16}_3 = 0^{256}, &&
+S^{-16}_4 = const_1\|const_0.
+\end{align*}
+After updating the state 16 times as \cipher{MORUS640}, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K_{128}\|K_{128}$.
+
+\subsubsection{Associated Data Processing.}
+After the initialization, the associated data $A$ is processed.
+For the padding, if the last associated data block is not a full block, use `0' bits to pad it to 128 bits and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
+Note that if the length of $A$ dented by `$adlen$' is 0, the associated data processing is skipped.
+
+Then, state is updated by $S^{t+1} \leftarrow {\tt StateUpdate}(S^t, A^t)$ for $t=0,1,\cdots,\ell$ where $\ell = \lceil \frac{adlen}{128} \rceil -1$ and $\ell = \lceil \frac{adlen}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
+
+\subsubsection{Encryption.}
+At each step of the encryption, a 16-byte (resp. 32-byte) message block $M_i$ is used to update the state, and $M_t$ is encrypted to $C_t$ in \cipher{MORUS640} (resp. \cipher{MORUS1280}).
+
+If the last message block is not a full block, use `0' bits to pad it to 128 and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
+But only the partial block is encrypted.
+Note that if the message length denoted by `$msglen$' is 0, encryption is skipped.
+
+In \cipher{MORUS640}, Let $u = \lceil \frac{adlen}{128} \rceil$ and $v = \lceil \frac{msglen}{128} \rceil$. The following is performed for $t=0, 1, \cdots, v-1$;
+\begin{align*}
+C^t &\leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 96) \oplus (S^{u+t}_2 \cdot S^{u+t}_3),\\
+S^{u+t+1} &\leftarrow {\tt StateUpdate}(S^{u+t},M^t).
+\end{align*}
+
+For \cipher{MORUS1280}, $u = \lceil \frac{adlen}{256} \rceil$ and $v = \lceil \frac{msglen}{256} \rceil$.
+The equation to compute ciphertext block $C^t$ is $C^t \leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 192) \oplus (S^{u+t}_2 \cdot S^{u+t}_3)$.
+
+\subsubsection{Finalization.}
+The finalization generates the authentication tag $T$ using 10 more steps.
+We only discuss the case that $T$ is not truncated.
+The associated data length and the message length are used to update the state as follows.
+\begin{enumerate}
+\item $tmp \leftarrow adlen\|msglen$ for \cipher{MORUS640} and $tmp \leftarrow adlen\|msglen\|0^{128}$ for \cipher{MORUS1280}, where $adlen$ and $msglen$ are represented as 64-bit integers.
+\item $S^{u+v}_4 \leftarrow S^{u+v}_4 \oplus S^{u+v}_0.$
+\item For $t = u+v, u+v+1, \cdots, u+v+9,$ compute
+%\begin{equation*}
+$S^{t+1} \leftarrow {\tt StateUpdate} (S^t, tmp)$.
+%\end{equation*}
+\item $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ for \cipher{MORUS640} and $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ for \cipher{MORUS1280}.
+\end{enumerate}
+
+\subsubsection{Security.}
+When the tag size is 128 bits, integrity is claimed up to 128-bits and confidentiality is claimed up to the number of bits of the key.
+\begin{table}[!htb]
+\centering
+\caption{Security Goals of MORUS.}
+\begin{tabular}{ccc}\hline
+                        & Confidentiality (bits) & Integrity (bits) \\ \hline
+\cipher{MORUS640}-128   & 128                    & 128              \\
+\cipher{MORUS1280}-128  & 128                    & 128              \\
+\cipher{MORUS1280}-256  & 256                    & 128              \\ \hline
+\end{tabular}
+\end{table}
+
+\subsection{Notation}
 
@@
 
-Finally, let $x$ denote a register.
+%Finally, let $x$ denote a register.
 Recall that registers of \cipher{MORUS} are composed of four words.
 We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-\end{tabular}
-}
-
-The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
+%\centerline{
+%\def\arraystretch{1.25}
+%\begin{tabular}{lcl}
+%$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
+%$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
+%\end{tabular}
+%}
+%
+%The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
 Throughout, $w$ denotes the  word size, i.e.
@@
 This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
+
+\section{Analysis on Initialization and Finalization of Reduced MORUS}
+
+\subsection{Forgery with Reduced Finalization}
+\subsubsection{Overview.}
+\subsubsection{Differential Trails.}
+
+\subsection{Extending State Recovery to Key Recovery}
+\subsubsection{Overview.}
+\subsubsection{Full Diffusion Rounds.}
 

### 2018-05-02 morusAC.tex
update section 1

@@
 In the recommended parameters, \cipher{MORUS640} supports 128-bit keys and \cipher{MORUS1280} supports both 128-bit and 256-bit keys.
-The tag size is 128 bits or shorter, though the designers strongly recommend using a 128-bit tag.
-
-In MORUS, the internal state is divided in two five registers.
+The tag size is 128 bits or shorter.
+The designers strongly recommend using a 128-bit tag.
+
+The state of MORUS is divided into five registers.
 Let $q$ be the register size, where $q = 128$ for \cipher{MORUS640} and $q=256$ for \cipher{MORUS1280}.
@@
 Then, the internal state is represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
-In MORUS, the internal state is first initialized by a key and nonce, and then updated by associated data blocks and message blocks step by step.
-Hence, we use a superscript to denote the step number of the state as $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is 0 before the initialization is performed.
+The state is first initialized by a key and nonce, and then updated by associated data blocks and message blocks step by step.
+Hence, we use a superscript to denote the step number of the state as $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is $-16$ before the initialization and $t=0$ after the initialization.
 
 The computation of MORUS consists of the four parts; initialization, associated data process, encryption, and finalization.
-All parts share the core design of MORUS called the {\tt StepUpdate} function.
+All parts share the core design called the {\tt StepUpdate} function.
 Hereafter we explain {\tt StepUpdate}.
@@
 the state $S^t$, in which the additional input $m^t$ is used in rounds 2 to 5, but not in round 1.
-Each round use the word-wise left circular shift, which divides a $q$-bit register value into 4 words of $q/4$ bits, and performs left circular shift operation for every $q/4$-bit word.
+Each round uses the word-wise left circular shift, which divides a $q$-bit register value into 4 words of $q/4$ bits, and performs left circular shift operation for every $q/4$-bit word.
 Throughout the paper, we denote the word size by $w$, i.e.
@@
 
-Let $\cdot$ denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
+Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
 \begin{align*}
@@
 S^{t+1}_1 &\leftarrow ( S_t^1 \oplus S^t_2 \cdot S^t_3 \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
-S^t_3 \leftarrow S^t_3 \lll b'_1.\\
+S^t_4 \leftarrow S^t_3 \lll b'_1.\\
 \textrm{Round 3:} &&
 S^{t+1}_2 &\leftarrow ( S_t^2 \oplus S^t_3 \cdot S^t_4 \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
-S^t_3 \leftarrow S^t_3 \lll b'_2.\\
+S^t_0 \leftarrow S^t_3 \lll b'_2.\\
 \textrm{Round 4:} &&
 S^{t+1}_3 &\leftarrow ( S_t^3 \oplus S^t_4 \cdot S^t_0 \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
-S^t_3 \leftarrow S^t_3 \lll b'_3.\\
+S^t_1 \leftarrow S^t_3 \lll b'_3.\\
 \textrm{Round 5:} &&
 S^{t+1}_4 &\leftarrow ( S_t^4 \oplus S^t_0 \cdot S^t_1 \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
-S^t_3 \leftarrow S^t_3 \lll b'_4.
+S^t_2 \leftarrow S^t_3 \lll b'_4.
 \end{align*}

### 2018-05-02 morus.tex
add morusAC.log to gitignore, extract the common.sty

@@
 \documentclass{llncs}
-
 \usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-
-\usepackage{amsmath,amssymb}
-\usepackage{ifthen}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage{siunitx}
-\usepackage[section]{placeins}
-\usepackage[hidelinks]{hyperref}
-\usepackage{cleveref}
-
-\allowdisplaybreaks%this is to allow page breaks within multiline equations (to accomodate full trail equations)
-
-\usepackage{xcolor}
-\definecolor{blue_link}{RGB}{0,0,200}
-\hypersetup{
-  colorlinks = true,
-  linkcolor = blue_link,
-  citecolor = blue_link,
-  urlcolor = blue_link,
-  bookmarksdepth = 3
-}
-
-\usepackage{tikz}
-\usetikzlibrary{calc,math,arrows.meta}
-
-\makeatletter
-\pgfdeclareshape{oplus}{%{{{
-  \inheritsavedanchors[from=circle]
-  \inheritanchorborder[from=circle]
-  \foreach \s in {center,mid,base,text, north,south,west,east,
-                  mid west,mid east,base west,base east,
-                  north west,south west,north east,south east} {
-    \inheritanchor[from=circle]{\s}
-  }
-  \backgroundpath{%
-    \pgfutil@tempdima=\radius%
-    \pgfmathsetlength{\pgf@xb}{\pgfkeysvalueof{/pgf/outer xsep}}%  
-    \pgfmathsetlength{\pgf@yb}{\pgfkeysvalueof{/pgf/outer ysep}}%  
-    \ifdim\pgf@xb<\pgf@yb%
-      \advance\pgfutil@tempdima by-\pgf@yb%
-    \else%
-      \advance\pgfutil@tempdima by-\pgf@xb%
-    \fi%
-    \pgfpathcircle{\centerpoint}{\pgfutil@tempdima}%
-    % north-south
-    \centerpoint\advance\pgf@y by\radius  \pgf@xa=\pgf@x \pgf@ya=\pgf@y
-    \centerpoint\advance\pgf@y by-\radius \pgf@xb=\pgf@x \pgf@yb=\pgf@y
-    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
-    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
-    % east-west
-    \centerpoint\advance\pgf@x by\radius  \pgf@xa=\pgf@x \pgf@ya=\pgf@y
-    \centerpoint\advance\pgf@x by-\radius \pgf@xb=\pgf@x \pgf@yb=\pgf@y
-    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
-    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
-  }
-}%}}}
-\makeatother
-\tikzset{op/.style={draw, fill=none, minimum size=1.25ex, inner sep=0pt}}
-\tikzset{opxor/.style={op, shape=oplus}}
-\tikzset{opand/.style={op, rectangle, rounded corners=2pt, minimum height=3ex}}
-\tikzset{oprot/.style={op, rectangle, rounded corners=2pt, inner sep=2pt, minimum width=3ex, font=\scriptsize}}
-\tikzset{optee/.style={shape=circle, fill, draw, inner sep=0pt, minimum size=2pt}}
-\tikzset{next/.style={->, >=latex}}
-\tikzset{trail/.style={line width=2pt, line cap=round}}
-\definecolor{delta}{HTML}{B58900}
-\definecolor{epsil}{HTML}{CB4B16}
-\definecolor{gamma}{HTML}{6C71C4}
-\definecolor{beta}{HTML}{268BD2}
-\definecolor{alpha}{HTML}{859900}
-
-%\definecolor{mybase0}{HTML}{839496}
-%\definecolor{mybase1}{HTML}{93A1A1}
-%\definecolor{mybase2}{HTML}{EEE8D5}
-%\definecolor{mybase3}{HTML}{FDF6E3}
-%\definecolor{myred}{HTML}{DC322F}
-%\definecolor{mymagenta}{HTML}{D33682}
-%\definecolor{mycyan}{HTML}{2AA198}
-
-\newcommand{\cipher}[1]{\textsf{#1}}
-
-\newif\ifsubstates\substatesfalse
-
-\newcommand{\printstate}{
-  \ifsubstates \pgfmathsetmacro{\roundsep}{1.25}
-  \else        \pgfmathsetmacro{\roundsep}{0.80} \fi
-  \pgfmathsetmacro{\opoffset}{.1}
-
-  \ifsubstates
-    \foreach \r in {-2,...,4} {
-      \foreach \w in {0,...,4} {
-        \draw[thick] (\w-.5,-\r*\roundsep) -- ++(0,-.25);
-        \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W\r\w) at (\w,-\r*\roundsep-.125) {};
-      }
-      \draw[thick] (-.5,-\r*\roundsep) node[below left, inner sep=0pt, xshift=-3pt] {$S_{\r,*}^{\r}$} rectangle ++(5,-.25);
-    }
-    \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W-1-1) at (-1,--1*\roundsep-.125) {};
-    \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W-2-1) at (-2,--1*\roundsep-.125) {};
-  \else
-    \foreach \r in {-1,...,4} {
-      \foreach \w in {0,...,4} {
-        \coordinate (W\r\w) at (\w,-\r*\roundsep-.125);
-      }
-    }
-    \foreach \w in {0,...,4} {
-      \coordinate (W-2\w) at (\w,--2.375*\roundsep-.125);
-    }
-    \coordinate (W-1-1) at (-1,--1*\roundsep-.125);
-    \coordinate (W-2-1) at (-2,--1*\roundsep-.125);
-  \fi
-
-  %\foreach \r/\rot in {-1/0,0/5,1/31,2/7,3/22,4/13} { % round
-  \foreach \r/\rot in {-1/0,0/b_0,1/b_1,2/b_2,3/b_3,4/b_4} { % round
-    \ifthenelse{\equal{\r}{-1}}{
-      \pgfmathsetmacro{\txorx}{int(mod(\r+2,5))}
-      \pgfmathsetmacro{\tanAx}{int(mod(\r+4,5))}
-      \pgfmathsetmacro{\tanBx}{int(mod(\r+3,5))}
-      \pgfmathsetmacro{\tlllx}{int(mod(\r+1,5))}
-    }{
-      \pgfmathsetmacro{\txorx}{int(mod(\r+3,5))}
-      \pgfmathsetmacro{\tanAx}{int(mod(\r+2,5))}
-      \pgfmathsetmacro{\tanBx}{int(mod(\r+1,5))}
-    }
-    \pgfmathsetmacro{\rprev}{int(\r-1)}
-    \ifthenelse{\equal{\r}{-1}}%
-    { \node[opxor] (lll\r) at ($(W\r\r) +(0,\opoffset+.25)$)  {}; }%
-    { \node[oprot] (lll\r) at ($(W\r\r) +(0,\opoffset+.25)$)  {$\lll\!\rot$}; }
-      \node[opxor] (xor\r) at ($(W\r\r) +(0,\opoffset+.50)$)  {};
-      \node[opxor] (xnd\r) at ($(W\r\r) +(0,\opoffset+.75)$)  {};
-    \ifthenelse{\equal{\r}{4}}
-    { \node[opand] (and\r) at ($(W\r3) +(.5,\opoffset+.75)$)  {$\cdot$}; }
-    { \node[opand] (and\r) at ($(W\r\r) +(.5,\opoffset+.75)$) {$\cdot$}; }
-    \ifthenelse{\equal{\r}{-1}}%
-    { \node[inner sep=1pt] (M) at ($(xnd\r) +(0,.45)$) {$M$};
-      \node[inner sep=1pt] (C) at ($(lll\r) +(0,-.5)$) {$C$};
-      \draw[next] (M) -- (xnd\r);
-      \draw[next] (lll\r) -- (C);
-      \coordinate[optee] (tlll\r) at ($(W\r\tlllx) +(0,\opoffset+.25)$); }{}
-      \coordinate[optee] (txor\r) at ($(W\r\txorx) +(0,\opoffset+.50)$);
-      \coordinate[optee] (tanA\r) at ($(W\r\tanAx) +(0,\opoffset+.675)$);
-      \coordinate[optee] (tanB\r) at ($(W\r\tanBx) +(0,\opoffset+.825)$);
-
-    \ifthenelse{\equal{\r}{-1}}%
-    { \draw[next] (tlll\r) -- (lll\r); }{}
-      \draw[next] (txor\r) -- (xor\r);
-    \ifthenelse{\tanAx < \r}%
-    { \draw[next] (tanA\r) -- (tanA\r-|and\r.west); }%
-    { \draw[next] (tanA\r) -- (tanA\r-|and\r.east); }
-    \ifthenelse{\tanBx < \r}%
-    { \draw[next] (tanB\r) -- (tanB\r-|and\r.west); }%
-    { \draw[next] (tanB\r) -- (tanB\r-|and\r.east); }
-    \draw[next] (and\r) -- (xnd\r);
-    \draw[    ] (xnd\r) -- (xor\r);
-    \draw[    ] (xor\r) -- (lll\r);
-
-    \foreach \w in {0,...,4} {
-      \ifthenelse{\equal{\w}{\r}}{
-        \draw (W\rprev\w) -- (xnd\r);
-        \draw (lll\r) -- (W\r\w);
-      }{
-        \draw (W\rprev\w) -- (W\r\w);
-      }
-    }
-  }
-}
-
-%new commands
-\newcommand{\bias}{\ensuremath{\operatorname{bias}}}
-\newcommand{\weight}{\ensuremath{\operatorname{weight}}}
-\newcommand\eqdef{:=}%equality by definition
-\newcommand\StateUpdate{\ensuremath{\texttt{StateUpdate}}}
-\newcommand\rotl{\ensuremath{\lll}}
-\newcommand\rotlxy{\ensuremath{\lll_W}}
+\usepackage{common}
 

### 2018-05-02 morusAC.tex
add morusAC.log to gitignore, extract the common.sty

@@
 \documentclass{llncs}
-
-%\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-
-\usepackage{amsmath,amssymb}
-\usepackage{ifthen}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage{siunitx}
-\usepackage[section]{placeins}
-\usepackage[hidelinks]{hyperref}
-\usepackage{cleveref}
-
-\allowdisplaybreaks%this is to allow page breaks within multiline equations (to accomodate full trail equations)
-
-\usepackage{xcolor}
-\definecolor{blue_link}{RGB}{0,0,200}
-\hypersetup{
-  colorlinks = true,
-  linkcolor = blue_link,
-  citecolor = blue_link,
-  urlcolor = blue_link,
-  bookmarksdepth = 3
-}
-
-\usepackage{tikz}
-\usetikzlibrary{calc,math,arrows.meta}
-
-\makeatletter
-\pgfdeclareshape{oplus}{%{{{
-  \inheritsavedanchors[from=circle]
-  \inheritanchorborder[from=circle]
-  \foreach \s in {center,mid,base,text, north,south,west,east,
-                  mid west,mid east,base west,base east,
-                  north west,south west,north east,south east} {
-    \inheritanchor[from=circle]{\s}
-  }
-  \backgroundpath{%
-    \pgfutil@tempdima=\radius%
-    \pgfmathsetlength{\pgf@xb}{\pgfkeysvalueof{/pgf/outer xsep}}%  
-    \pgfmathsetlength{\pgf@yb}{\pgfkeysvalueof{/pgf/outer ysep}}%  
-    \ifdim\pgf@xb<\pgf@yb%
-      \advance\pgfutil@tempdima by-\pgf@yb%
-    \else%
-      \advance\pgfutil@tempdima by-\pgf@xb%
-    \fi%
-    \pgfpathcircle{\centerpoint}{\pgfutil@tempdima}%
-    % north-south
-    \centerpoint\advance\pgf@y by\radius  \pgf@xa=\pgf@x \pgf@ya=\pgf@y
-    \centerpoint\advance\pgf@y by-\radius \pgf@xb=\pgf@x \pgf@yb=\pgf@y
-    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
-    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
-    % east-west
-    \centerpoint\advance\pgf@x by\radius  \pgf@xa=\pgf@x \pgf@ya=\pgf@y
-    \centerpoint\advance\pgf@x by-\radius \pgf@xb=\pgf@x \pgf@yb=\pgf@y
-    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
-    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
-  }
-}%}}}
-\makeatother
-\tikzset{op/.style={draw, fill=none, minimum size=1.25ex, inner sep=0pt}}
-\tikzset{opxor/.style={op, shape=oplus}}
-\tikzset{opand/.style={op, rectangle, rounded corners=2pt, minimum height=3ex}}
-\tikzset{oprot/.style={op, rectangle, rounded corners=2pt, inner sep=2pt, minimum width=3ex, font=\scriptsize}}
-\tikzset{optee/.style={shape=circle, fill, draw, inner sep=0pt, minimum size=2pt}}
-\tikzset{next/.style={->, >=latex}}
-\tikzset{trail/.style={line width=2pt, line cap=round}}
-\definecolor{delta}{HTML}{B58900}
-\definecolor{epsil}{HTML}{CB4B16}
-\definecolor{gamma}{HTML}{6C71C4}
-\definecolor{beta}{HTML}{268BD2}
-\definecolor{alpha}{HTML}{859900}
-
-%\definecolor{mybase0}{HTML}{839496}
-%\definecolor{mybase1}{HTML}{93A1A1}
-%\definecolor{mybase2}{HTML}{EEE8D5}
-%\definecolor{mybase3}{HTML}{FDF6E3}
-%\definecolor{myred}{HTML}{DC322F}
-%\definecolor{mymagenta}{HTML}{D33682}
-%\definecolor{mycyan}{HTML}{2AA198}
-
-\newcommand{\cipher}[1]{\textsf{#1}}
-
-\newif\ifsubstates\substatesfalse
-
-\newcommand{\printstate}{
-  \ifsubstates \pgfmathsetmacro{\roundsep}{1.25}
-  \else        \pgfmathsetmacro{\roundsep}{0.80} \fi
-  \pgfmathsetmacro{\opoffset}{.1}
-
-  \ifsubstates
-    \foreach \r in {-2,...,4} {
-      \foreach \w in {0,...,4} {
-        \draw[thick] (\w-.5,-\r*\roundsep) -- ++(0,-.25);
-        \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W\r\w) at (\w,-\r*\roundsep-.125) {};
-      }
-      \draw[thick] (-.5,-\r*\roundsep) node[below left, inner sep=0pt, xshift=-3pt] {$S_{\r,*}^{\r}$} rectangle ++(5,-.25);
-    }
-    \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W-1-1) at (-1,--1*\roundsep-.125) {};
-    \node[minimum width=1*1.0cm,minimum height=.25*1.5cm, inner sep=0pt] (W-2-1) at (-2,--1*\roundsep-.125) {};
-  \else
-    \foreach \r in {-1,...,4} {
-      \foreach \w in {0,...,4} {
-        \coordinate (W\r\w) at (\w,-\r*\roundsep-.125);
-      }
-    }
-    \foreach \w in {0,...,4} {
-      \coordinate (W-2\w) at (\w,--2.375*\roundsep-.125);
-    }
-    \coordinate (W-1-1) at (-1,--1*\roundsep-.125);
-    \coordinate (W-2-1) at (-2,--1*\roundsep-.125);
-  \fi
-
-  %\foreach \r/\rot in {-1/0,0/5,1/31,2/7,3/22,4/13} { % round
-  \foreach \r/\rot in {-1/0,0/b_0,1/b_1,2/b_2,3/b_3,4/b_4} { % round
-    \ifthenelse{\equal{\r}{-1}}{
-      \pgfmathsetmacro{\txorx}{int(mod(\r+2,5))}
-      \pgfmathsetmacro{\tanAx}{int(mod(\r+4,5))}
-      \pgfmathsetmacro{\tanBx}{int(mod(\r+3,5))}
-      \pgfmathsetmacro{\tlllx}{int(mod(\r+1,5))}
-    }{
-      \pgfmathsetmacro{\txorx}{int(mod(\r+3,5))}
-      \pgfmathsetmacro{\tanAx}{int(mod(\r+2,5))}
-      \pgfmathsetmacro{\tanBx}{int(mod(\r+1,5))}
-    }
-    \pgfmathsetmacro{\rprev}{int(\r-1)}
-    \ifthenelse{\equal{\r}{-1}}%
-    { \node[opxor] (lll\r) at ($(W\r\r) +(0,\opoffset+.25)$)  {}; }%
-    { \node[oprot] (lll\r) at ($(W\r\r) +(0,\opoffset+.25)$)  {$\lll\!\rot$}; }
-      \node[opxor] (xor\r) at ($(W\r\r) +(0,\opoffset+.50)$)  {};
-      \node[opxor] (xnd\r) at ($(W\r\r) +(0,\opoffset+.75)$)  {};
-    \ifthenelse{\equal{\r}{4}}
-    { \node[opand] (and\r) at ($(W\r3) +(.5,\opoffset+.75)$)  {$\cdot$}; }
-    { \node[opand] (and\r) at ($(W\r\r) +(.5,\opoffset+.75)$) {$\cdot$}; }
-    \ifthenelse{\equal{\r}{-1}}%
-    { \node[inner sep=1pt] (M) at ($(xnd\r) +(0,.45)$) {$M$};
-      \node[inner sep=1pt] (C) at ($(lll\r) +(0,-.5)$) {$C$};
-      \draw[next] (M) -- (xnd\r);
-      \draw[next] (lll\r) -- (C);
-      \coordinate[optee] (tlll\r) at ($(W\r\tlllx) +(0,\opoffset+.25)$); }{}
-      \coordinate[optee] (txor\r) at ($(W\r\txorx) +(0,\opoffset+.50)$);
-      \coordinate[optee] (tanA\r) at ($(W\r\tanAx) +(0,\opoffset+.675)$);
-      \coordinate[optee] (tanB\r) at ($(W\r\tanBx) +(0,\opoffset+.825)$);
-
-    \ifthenelse{\equal{\r}{-1}}%
-    { \draw[next] (tlll\r) -- (lll\r); }{}
-      \draw[next] (txor\r) -- (xor\r);
-    \ifthenelse{\tanAx < \r}%
-    { \draw[next] (tanA\r) -- (tanA\r-|and\r.west); }%
-    { \draw[next] (tanA\r) -- (tanA\r-|and\r.east); }
-    \ifthenelse{\tanBx < \r}%
-    { \draw[next] (tanB\r) -- (tanB\r-|and\r.west); }%
-    { \draw[next] (tanB\r) -- (tanB\r-|and\r.east); }
-    \draw[next] (and\r) -- (xnd\r);
-    \draw[    ] (xnd\r) -- (xor\r);
-    \draw[    ] (xor\r) -- (lll\r);
-
-    \foreach \w in {0,...,4} {
-      \ifthenelse{\equal{\w}{\r}}{
-        \draw (W\rprev\w) -- (xnd\r);
-        \draw (lll\r) -- (W\r\w);
-      }{
-        \draw (W\rprev\w) -- (W\r\w);
-      }
-    }
-  }
-}
-
-%new commands
-\newcommand{\bias}{\ensuremath{\operatorname{bias}}}
-\newcommand{\weight}{\ensuremath{\operatorname{weight}}}
-\newcommand\eqdef{:=}%equality by definition
-\newcommand\StateUpdate{\ensuremath{\texttt{StateUpdate}}}
-\newcommand\rotl{\ensuremath{\lll}}
-\newcommand\rotlxy{\ensuremath{\lll_W}}
+\usepackage{common}
 
@@
 
-\title{Note on Linear Biases in MORUS}
+% \title{Note on Linear Biases in MORUS}
+\title{Morus: Biases in the stream}
 

### 2018-05-03 morusAC.tex
fixed mistake in cyclic shift constants

@@
 \centering
-\caption{Rotation Constants of MORUS} \label{Tbl:rcon}
+\caption{Left Circular-Shift Constants of MORUS} \label{Tbl:rcon}
 \begin{tabular}{c||ccccc|ccccc} \hline
-& \multicolumn{5}{c|}{register-wise circular shift} & \multicolumn{5}{c}{word-wise circular shift} \\
+& \multicolumn{5}{c|}{word-wise circular shift} & \multicolumn{5}{c}{register-wise circular shift} \\
                   & \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline

### 2018-05-04 morusAC.tex
smalltypos

@@
 \end{table}
-We denote those register-wise and word-wise left circular shift by $x \rotl y$ and $x \rotlxy y$, respectively.
+We denote those register-wise and word-wise left circular shift by $x \lll y$ and $x \lll_w y$, respectively.
 (Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document.)
@@
 \end{align*}
-Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\cdots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K$.
+Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K$.
 
@@
 For the padding, if the last associated data block is not a full block, use `0' bits to pad it to 128 bits and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
-Note that if the length of $A$ dented by `$adlen$' is 0, the associated data processing is skipped.
-
-Then, state is updated by $S^{t+1} \leftarrow {\tt StateUpdate}(S^t, A^t)$ for $t=0,1,\cdots,\ell$ where $\ell = \lceil \frac{adlen}{128} \rceil -1$ and $\ell = \lceil \frac{adlen}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
+Note that if the length of $A$ denoted by `$adlen$' is 0, the associated data processing is skipped.
+
+Then, state is updated by $S^{t+1} \leftarrow {\tt StateUpdate}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{adlen}{128} \rceil -1$ and $\ell = \lceil \frac{adlen}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
 
@@
 
-In \cipher{MORUS640}, Let $u = \lceil \frac{adlen}{128} \rceil$ and $v = \lceil \frac{msglen}{128} \rceil$. The following is performed for $t=0, 1, \cdots, v-1$;
+In \cipher{MORUS640}, Let $u = \lceil \frac{adlen}{128} \rceil$ and $v = \lceil \frac{msglen}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$;
 \begin{align*}
@@
 \item $S^{u+v}_4 \leftarrow S^{u+v}_4 \oplus S^{u+v}_0.$
-\item For $t = u+v, u+v+1, \cdots, u+v+9,$ compute
+\item For $t = u+v, u+v+1, \ldots, u+v+9,$ compute
 %\begin{equation*}
@@
 
-We begin with a few observations about the \StateUpdate{} function. Beside XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
+We begin with a few observations about the \StateUpdate{} function. Besides XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
 \begin{enumerate}

### 2018-05-07 morusAC.tex
update section 7

@@
 Let $S_i, i \in\{0,1,2,3,4\}$ be a register.
-Then, the internal state is represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
+Then, the state is represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
 The state is first initialized by a key and nonce, and then updated by associated data blocks and message blocks step by step.
@@
 
-The computation of MORUS consists of the four parts; initialization, associated data process, encryption, and finalization.
+The computation of MORUS consists of four parts; initialization, associated data process, encryption, and finalization.
 All parts share the core design called the {\tt StepUpdate} function.
-Hereafter we explain {\tt StepUpdate}.
+%Hereafter we explain {\tt StepUpdate}.
 
@@
 \begin{align*}
-S^{-16}_0 = \textrm{IV}_{128} \| 0^{128}, &&
-S^{-16}_1 = K_{128}\|K_{128}, &&
+S^{-16}_0 &= \textrm{IV}_{128} \| 0^{128}, &
+S^{-16}_1 &= K_{128}\|K_{128}, &
 S^{-16}_2 = 1^{256}, \\
-S^{-16}_3 = 0^{256}, &&
-S^{-16}_4 = const_1\|const_0.
+S^{-16}_3 &= 0^{256}, &
+S^{-16}_4 &= const_1\|const_0.
 \end{align*}
-After updating the state 16 times as \cipher{MORUS640}, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K_{128}\|K_{128}$.
+After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K_{128}\|K_{128}$.
 
@@
 \section{Analysis on Initialization and Finalization of Reduced MORUS}
+The bias discussed in the previous sections analysed the encryption part of the MORUS.
+In this section, for comprehensive security analysis of MORUS, we provide new attacks on reduced version of the initialization and the finalization.
+We emphasize that the results in this section do not threat any security claim by the designers.
+However, we believe that investigating all parts of the design in new approaches provide better understanding of the design and will be useful especially when the design will be tweaked in future.
 
 \subsection{Forgery with Reduced Finalization}
+We present forgery attacks on 3 out of 10 steps of \cipher{MORUS1280} that claims 128-bit security for integrity.
+The attack only works for a limited number of steps, while it works in the nonce-respect setting As far as we know, this is the first attempt to evaluate integrity of MORUS in the nonce-respect setting.
+
 \subsubsection{Overview.} 
+A general strategy of forgery attacks in the nonce-respect setting is to inject some difference in a message block and to propagate the difference so that the difference can be canceled by another difference in another message block.
+However this approach does not work well against MORUS because its large state size prevents the attackers from controlling the difference to be canceled by another message difference.
+
+Here we focus on the property that the padding scheme for the associated data $A$ and the message $M$ is zero-padding, hence $A$ and $A'=A\|0^*$ and $M$ and $M'=M\|0$ will make the result identical for the associated data processing and the encryption parts, as long as $A,A'$ and $M,M'$ fit in the same number of blocks.
+During the finalization, because $A,A'$ (resp.
+$M,M'$) have different lengths, the corresponding $adlen$s (resp.
+$msglen$s) are different, which appears as $\Delta adlen$ (resp.
+$\Delta msglen$), and is injected from the message input.
+Our strategy is propagating this difference to the 128-bit value tag $T$ and $T'$ such that their difference $\Delta T$ appears higher probability than $2^{-128}$.
+All in all, the forgery succeeds as long as the desired $\Delta T$ is obtained, in other words, the attackers do not have to cancel out the state difference, which is the main advantage of attacking the finalization.
+
+Note that if the attacker uses different messages $M,M'$, not only the new tag $T'$ but also new ciphertext $C'$ must be guessed correctly.
+Because the encryption of MORUS is a simple XOR of the key stream, $C'$ can be easily guessed.
+For this purpose, the attacker should first query a longer message $M'=M\|0^*$ to obtain $C'$.
+Then, $C$ can be obtained by truncating $C'$.
+
 \subsubsection{Differential Trails.}

### 2018-05-07 morusAC.tex
Added differential trail

@@
 \subsubsection{Differential Trails.}
+Recall that the message input during the finalization of \cipher{MORUS1280} is $adlen\|msglen\|0^{128}$ where $adlen$ and $msglen$ are 64-bit strings.
+We set $\Delta adlen$ to be low Hamming weight, namely {\tt 0x0000000000000001}.
+Hereafter, we use {\tt typefont} denote hexadecimal numbers.
+This difference propagates during the subsequent 3 steps as specified in Table~\ref{Tbl:fin_diff}.
+
+\begin{table}[!tb]
+\begin{center}
+\caption{Differential Propagation through 3 Steps in the Finalization.}
+\label{Tbl:fin_diff}
+{\tiny
+\begin{tabular}{cccccc}
+Round & \multicolumn{4}{c}{State Difference} & Weight \\ \hline
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+PT    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+1     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000400000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1 \\
+2     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000400000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1 \\
+3     & {\tt 0000004000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\ \hline
+
+%Weight: 0
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1 \\
+4     & {\tt 0000004000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1 \\
+      & {\tt 0020000000000080} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\ \hline
+
+%Weight: 1
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1 \\
+5     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1 \\
+      & {\tt 0020000000000080} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\
+      & {\tt 0000040000000010} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\ \hline
+
+%Weight: 3
+      & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1 \\
+6     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2 \\
+      & {\tt 0000040000000010} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\ \hline
+
+%Weight: 6
+      & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\
+      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4 \\
+7     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2 \\ \hline
+
+%Weight: 10
+      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\
+      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4 \\
+8     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2 \\ \hline
+
+%Weight: 14
+      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4 \\
+9     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4 \\
+      & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7 \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2 \\ \hline
+
+%Weight: 20
+      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2 \\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4 \\
+10    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4 \\
+      & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7 \\
+      & {\tt 4000140000000010} & {\tt 0000000000000000} & {\tt 0000400000000100} & {\tt 0000000000010000} & 7 \\ \hline
+
+%Weight: 28
+      & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9 \\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4 \\
+11    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4 \\
+      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7 \\
+      & {\tt 4000140000000010} & {\tt 0000000000000000} & {\tt 0000400000000100} & {\tt 0000000000010000} & 7 \\ \hline
+
+%Weight: 39
+      & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9 \\
+      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 \\
+12    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4 \\
+      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7 \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7 \\ \hline
+
+%Weight: 53
+      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9 \\
+      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 \\
+13    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 \\
+      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7 \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7 \\ \hline
+
+%Weight: 69
+      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9 \\
+      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 \\
+14    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 \\
+      & {\tt 0228000280020080} & {\tt 0000040000000000} & {\tt 2000008000202008} & {\tt 1000004000004021} & 18 \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7 \\ \hline
+
+%Weight: 69
+      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9 \\
+      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 \\
+15    & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & {\tt 0400114000040000} & 14 \\
+      & {\tt 0228000280020080} & {\tt 0000040000000000} & {\tt 2000008000202008} & {\tt 1000004000004021} & 18 \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7 \\ \hline
+
+%Weight: 98
+\end{tabular}
+}
+\end{center}
+\end{table}
 

### 2018-05-07 morusAC.tex
fit trail picture to page
finishes section 7.1

--- 2018-05-07 morusAC.tex
@@
 %\end{equation*}
-\item $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ for \cipher{MORUS640} and $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ for \cipher{MORUS1280}.
+\item For \cipher{MORUS640}, $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$.
+For \cipher{MORUS1280}, $S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ is computed and the least significant 128 bits are produced as a tag $T$.
 \end{enumerate}
@@
 We emphasize that the results in this section do not threat any security claim by the designers.
-However, we believe that investigating all parts of the design in new approaches provide better understanding of the design and will be useful especially when the design will be tweaked in future.
+However, we believe that investigating all parts of the design with new approaches provides better understanding of the design and will be useful especially when the design will be tweaked in future.
 
@@
 We present forgery attacks on 3 out of 10 steps of \cipher{MORUS1280} that claims 128-bit security for integrity.
-The attack only works for a limited number of steps, while it works in the nonce-respect setting As far as we know, this is the first attempt to evaluate integrity of MORUS in the nonce-respect setting.
+The attack only works for a limited number of steps, while it works in the nonce-respect setting.
+As far as we know, this is the first attempt to evaluate integrity of MORUS in the nonce-respect setting.
 
 \subsubsection{Overview.} 
-A general strategy of forgery attacks in the nonce-respect setting is to inject some difference in a message block and to propagate the difference so that the difference can be canceled by another difference in another message block.
+A general strategy of forgery attacks in the nonce-respect setting is to inject some difference in a message block and to propagate the difference so that it can be canceled by difference in another message block.
 However this approach does not work well against MORUS because its large state size prevents the attackers from controlling the difference to be canceled by another message difference.
 
-Here we focus on the property that the padding scheme for the associated data $A$ and the message $M$ is zero-padding, hence $A$ and $A'=A\|0^*$ and $M$ and $M'=M\|0$ will make the result identical for the associated data processing and the encryption parts, as long as $A,A'$ and $M,M'$ fit in the same number of blocks.
+Here we focus on the property that the padding for an associated data $A$ and a message $M$ is the zero-padding, hence $A$ and $A'=A\|0^*$ and $M$ and $M'=M\|0$ will make the result identical for the associated data processing and the encryption parts, as long as $A,A'$ and $M,M'$ fit in the same number of blocks.
 During the finalization, because $A,A'$ (resp.
@@
 $\Delta msglen$), and is injected from the message input.
-Our strategy is propagating this difference to the 128-bit value tag $T$ and $T'$ such that their difference $\Delta T$ appears higher probability than $2^{-128}$.
-All in all, the forgery succeeds as long as the desired $\Delta T$ is obtained, in other words, the attackers do not have to cancel out the state difference, which is the main advantage of attacking the finalization.
+Our strategy is to propagate this difference to 128-bit tag $T$ and $T'$ such that their difference $\Delta T$ appears higher probability than $2^{-128}$.
+All in all, the forgery succeeds as long as the desired $\Delta T$ is obtained, in other words, the attackers do not have to cancel the state difference, which is the main advantage of attacking the finalization.
 
@@
 \subsubsection{Differential Trails.}
+Recall that the message input during the finalization of \cipher{MORUS1280} is $adlen\|msglen\|0^{128}$ where $adlen$ and $msglen$ are 64-bit strings.
+We set $\Delta adlen$ to be low Hamming weight, e.g.~{\tt 0x0000000000000001}.
+%Hereafter, we use {\tt typefont} denote hexadecimal numbers.
+This difference propagates during 3 steps is specified in Table~\ref{Tbl:fin_diff}.
+
+Recall that each step consists of 5 rounds and the input message is absorbed to the state in rounds 2 to 5.
+The trail in Table~\ref{Tbl:fin_diff} initially does not have any difference and the same continues even after round 1.
+Differences start to appear from round 2 and they will go through the bitwise-AND operation from round 4.
+We need to pay 1-bit to control each active AND gate.
+The probability evaluation for round 15 can be ignored because in this round only $S_4$ is non-linearly updated, while $S_4$ is never used for computing the tag.
+Finally, bitwise-AND in the tag computation is taken into account.
+Note that the tag is only 128 LSBs, thus the number of active AND gates should be counted only in those bits.
+As shown in Table~\ref{Tbl:fin_diff}, we can have a particular tag difference $\Delta T$ with probability $2^{-88}$.
+Thus after observing $A$ and corresponding $T$, $A\|0$ and $(T \oplus \Delta T)$ is a valid pair with probability $2^{-88}$.
+
+\begin{table}[!tb]
+\begin{center}
+\caption{Differential Propagation through Three Steps in the Finalization.
+$i$ Five lines for round $i$ denote the state difference $S_0, \cdots, S_4$ after the the round $i$ transformation.
+Weight is a Hamming weight of the state difference and accumulated probability is a probability to satisfy the trail from the beginning.}
+\label{Tbl:fin_diff}
+{\tiny
+\begin{tabular}{ccccccc} \hline
+Round & \multicolumn{4}{c}{State Difference} & Weight & Accumulated \\
+      & & & &                                &        & probability \\ \hline
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+Ini   & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & $-$ \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+1     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & $1$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000400000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & \\
+2     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & $1$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000400000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & \\
+3     & {\tt 0000004000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & $1$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
+4     & {\tt 0000004000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & $1$\\
+      & {\tt 0020000000000080} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
+5     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-1}$\\
+      & {\tt 0020000000000080} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
+      & {\tt 0000040000000010} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\ \hline
+
+      & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
+6     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-3}$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
+      & {\tt 0000040000000010} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\ \hline
+
+      & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
+      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \\
+7     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-6}$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
+      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \\
+8     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & $2^{-10}$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
+9     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & $2^{-14}$\\
+      & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7  & \\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2  & \\ \hline
+
+      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
+10    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-20}$\\
+      & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7  & \\
+      & {\tt 4000140000000010} & {\tt 0000000000000000} & {\tt 0000400000000100} & {\tt 0000000000010000} & 7  & \\ \hline
+
+      & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9  & \\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
+11    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-28}$\\
+      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
+      & {\tt 4000140000000010} & {\tt 0000000000000000} & {\tt 0000400000000100} & {\tt 0000000000010000} & 7  & \\ \hline
+
+      & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9  & \\
+      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \\
+12    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-39}$\\
+      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline
+
+      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
+      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \\
+13    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 & $2^{-53}$\\
+      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline
+
+      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
+      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 & \\
+14    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 & $2^{-69}$\\
+      & {\tt 0228000280020080} & {\tt 0000040000000000} & {\tt 2000008000202008} & {\tt 1000004000004021} & 18 & \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline
+
+%Weight: 69
+      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
+      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 & \\
+15    & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & {\tt 0400114000040000} & 14 & $-$\\
+      & {\tt 0228000280020080} & {\tt 0000040000000000} & {\tt 2000008000202008} & {\tt 1000004000004021} & 18 & \\
+      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline
+
+$\Delta T$ &                   &                        & {\tt 600080830020f00a} & {\tt 1405414005044421} & & $2^{-88}$ \\ \hline
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\subsubsection{Remarks.} The fact that $S_4$ updated in the last round is not used to compute a tag implies that the MORUS finalization generally includes unnecessary computations.
+It may be interesting to tweak the design such that the tag can also depend on $S_4$.
+Indeed in Table~\ref{Tbl:fin_diff}, we can observe some jump-up of the probability in the tag computation.
+This is because the non-linearly involved terms are $S_2 \cdot S_3$, and $S_3$ that was updated 2-round before has a high Hamming weight.
+In this sense, involving $S_4$ in non-linear terms of the tag computation imposes more difficulties for the attacker.
 

### 2018-05-08 morusAC.tex
Added bib file. Prepared separated tex file for Abstract, Introduction, Preliminaries, Initialization-Finalization analysis

--- 2018-05-07 morusAC.tex
@@
 
-\begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
-More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
-Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
-\end{abstract}
-
-\section{Introduction}
-
-Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
-As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
-Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-
-\section{Preliminaries}
-\subsection{Specification of MORUS}
-MORUS is a family of authenticated ciphers with two internal state sizes: 640 and 1280 bits.
-Each member is called \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
-In the recommended parameters, \cipher{MORUS640} supports 128-bit keys and \cipher{MORUS1280} supports both 128-bit and 256-bit keys.
-The tag size is 128 bits or shorter.
-The designers strongly recommend using a 128-bit tag.
-
-The state of MORUS is divided into five registers.
-Let $q$ be the register size, where $q = 128$ for \cipher{MORUS640} and $q=256$ for \cipher{MORUS1280}.
-Let $S_i, i \in\{0,1,2,3,4\}$ be a register.
-Then, the state is represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
-The state is first initialized by a key and nonce, and then updated by associated data blocks and message blocks step by step.
-Hence, we use a superscript to denote the step number of the state as $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is $-16$ before the initialization and $t=0$ after the initialization.
-
-The computation of MORUS consists of four parts; initialization, associated data process, encryption, and finalization.
-All parts share the core design called the {\tt StepUpdate} function.
-%Hereafter we explain {\tt StepUpdate}.
-
-\subsubsection{Step Update Function.}
-{\tt StepUpdate} takes as input the internal state $S^t = S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$ and additional register-wise value $m^t$.
-$m^t$ is supposed to be associated data block or message block.
-$m^t$ can sometimes be 0 or constant when the additional input is unnecessary.
-
-{\tt StepUpdate} is composed of 5 rounds with similar operations to update
-the state $S^t$, in which the additional input $m^t$ is used in rounds 2 to 5, but not in round 1.
-Each round uses the word-wise left circular shift, which divides a $q$-bit register value into 4 words of $q/4$ bits, and performs left circular shift operation for every $q/4$-bit word.
-Throughout the paper, we denote the word size by $w$, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
-The word-wise circular shift constants $b_i$ in round $i$ are defined in Table~\ref{Tbl:rcon}.
-Besides, the left rotation of a whole $q$-bit register is used.
-The register-wise circular shift constants are also listed in Table~\ref{Tbl:rcon}.
-\begin{table}[!htb]
-\centering
-\caption{Left Circular-Shift Constants of MORUS} \label{Tbl:rcon}
-\begin{tabular}{c||ccccc|ccccc} \hline
-& \multicolumn{5}{c|}{word-wise circular shift} & \multicolumn{5}{c}{register-wise circular shift} \\
-                  & \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline
-\cipher{MORUS640} &  5 & 31 &  7 & 22 & 13 & 32 &  64 &  96 &  64 & 32 \\
-\cipher{MORUS1280}& 13 & 46 & 38 &  7 &  4 & 64 & 128 & 192 & 128 & 64 \\ \hline
-\end{tabular}
-\end{table}
-We denote those register-wise and word-wise left circular shift by $x \lll y$ and $x \lll_w y$, respectively.
-(Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document.)
-
-Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
-\begin{align*}
-\textrm{Round 1:} &&
-S^{t+1}_0 &\leftarrow ( S_t^0 \oplus S^t_1 \cdot S^t_2 \oplus S^t_3 ) \lll_w b_0 &
-S^t_3 \leftarrow S^t_3 \lll b'_0.\\
-\textrm{Round 2:} &&
-S^{t+1}_1 &\leftarrow ( S_t^1 \oplus S^t_2 \cdot S^t_3 \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
-S^t_4 \leftarrow S^t_3 \lll b'_1.\\
-\textrm{Round 3:} &&
-S^{t+1}_2 &\leftarrow ( S_t^2 \oplus S^t_3 \cdot S^t_4 \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
-S^t_0 \leftarrow S^t_3 \lll b'_2.\\
-\textrm{Round 4:} &&
-S^{t+1}_3 &\leftarrow ( S_t^3 \oplus S^t_4 \cdot S^t_0 \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
-S^t_1 \leftarrow S^t_3 \lll b'_3.\\
-\textrm{Round 5:} &&
-S^{t+1}_4 &\leftarrow ( S_t^4 \oplus S^t_0 \cdot S^t_1 \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
-S^t_2 \leftarrow S^t_3 \lll b'_4.
-\end{align*}
-
-\subsubsection{Initialization.}
-The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K$ and a 128-bit nonce IV into the state, and running the {\tt StateUpdate} for 16 steps.
-The key and IV are loaded into the state as follows:
-\begin{align*}
-S^{-16}_0 = \textrm{IV}_{128}, &&
-S^{-16}_1 = K_{128}, &&
-S^{-16}_2 = 1^{128}, &&
-S^{-16}_3 = const_0, &&
-S^{-16}_4 = const_1.
-\end{align*}
-Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K$.
-
-The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size. The state is initialized as
-\begin{align*}
-S^{-16}_0 &= \textrm{IV}_{128} \| 0^{128}, &
-S^{-16}_1 &= K_{128}\|K_{128}, &
-S^{-16}_2 = 1^{256}, \\
-S^{-16}_3 &= 0^{256}, &
-S^{-16}_4 &= const_1\|const_0.
-\end{align*}
-After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K_{128}\|K_{128}$.
-
-\subsubsection{Associated Data Processing.}
-After the initialization, the associated data $A$ is processed.
-For the padding, if the last associated data block is not a full block, use `0' bits to pad it to 128 bits and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
-Note that if the length of $A$ denoted by `$adlen$' is 0, the associated data processing is skipped.
-
-Then, state is updated by $S^{t+1} \leftarrow {\tt StateUpdate}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{adlen}{128} \rceil -1$ and $\ell = \lceil \frac{adlen}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
-
-\subsubsection{Encryption.}
-At each step of the encryption, a 16-byte (resp. 32-byte) message block $M_i$ is used to update the state, and $M_t$ is encrypted to $C_t$ in \cipher{MORUS640} (resp. \cipher{MORUS1280}).
-
-If the last message block is not a full block, use `0' bits to pad it to 128 and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
-But only the partial block is encrypted.
-Note that if the message length denoted by `$msglen$' is 0, encryption is skipped.
-
-In \cipher{MORUS640}, Let $u = \lceil \frac{adlen}{128} \rceil$ and $v = \lceil \frac{msglen}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$;
-\begin{align*}
-C^t &\leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 96) \oplus (S^{u+t}_2 \cdot S^{u+t}_3),\\
-S^{u+t+1} &\leftarrow {\tt StateUpdate}(S^{u+t},M^t).
-\end{align*}
-
-For \cipher{MORUS1280}, $u = \lceil \frac{adlen}{256} \rceil$ and $v = \lceil \frac{msglen}{256} \rceil$.
-The equation to compute ciphertext block $C^t$ is $C^t \leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 192) \oplus (S^{u+t}_2 \cdot S^{u+t}_3)$.
-
-\subsubsection{Finalization.}
-The finalization generates the authentication tag $T$ using 10 more steps.
-We only discuss the case that $T$ is not truncated.
-The associated data length and the message length are used to update the state as follows.
-\begin{enumerate}
-\item $tmp \leftarrow adlen\|msglen$ for \cipher{MORUS640} and $tmp \leftarrow adlen\|msglen\|0^{128}$ for \cipher{MORUS1280}, where $adlen$ and $msglen$ are represented as 64-bit integers.
-\item $S^{u+v}_4 \leftarrow S^{u+v}_4 \oplus S^{u+v}_0.$
-\item For $t = u+v, u+v+1, \ldots, u+v+9,$ compute
-%\begin{equation*}
-$S^{t+1} \leftarrow {\tt StateUpdate} (S^t, tmp)$.
-%\end{equation*}
-\item $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ for \cipher{MORUS640} and $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ for \cipher{MORUS1280}.
-\end{enumerate}
-
-\subsubsection{Security.}
-When the tag size is 128 bits, integrity is claimed up to 128-bits and confidentiality is claimed up to the number of bits of the key.
-\begin{table}[!htb]
-\centering
-\caption{Security Goals of MORUS.}
-\begin{tabular}{ccc}\hline
-                        & Confidentiality (bits) & Integrity (bits) \\ \hline
-\cipher{MORUS640}-128   & 128                    & 128              \\
-\cipher{MORUS1280}-128  & 128                    & 128              \\
-\cipher{MORUS1280}-256  & 256                    & 128              \\ \hline
-\end{tabular}
-\end{table}
-
-\subsection{Notation}
-
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
-weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
-weights).
-
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
-$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
-$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
-$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
-\end{tabular}
-}
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
-
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
-Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
-In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
-
-%Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-%\centerline{
-%\def\arraystretch{1.25}
-%\begin{tabular}{lcl}
-%$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-%$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-%\end{tabular}
-%}
-%
-%The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
+\input{morusAC_00_Abstract}
+
+\input{morusAC_01_Introduction}
+
+\input{morusAC_02_Preliminaries}
 
@@
   \centering
+  \resizebox{!}{0.95\textheight}{
   \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
@@
     \end{scope}
-  \end{tikzpicture}
+  \end{tikzpicture}}
+
   \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
@@
 
-\section{Analysis on Initialization and Finalization of Reduced MORUS}
-The bias discussed in the previous sections analysed the encryption part of the MORUS.
-In this section, for comprehensive security analysis of MORUS, we provide new attacks on reduced version of the initialization and the finalization.
-We emphasize that the results in this section do not threat any security claim by the designers.
-However, we believe that investigating all parts of the design in new approaches provide better understanding of the design and will be useful especially when the design will be tweaked in future.
-
-\subsection{Forgery with Reduced Finalization}
-We present forgery attacks on 3 out of 10 steps of \cipher{MORUS1280} that claims 128-bit security for integrity.
-The attack only works for a limited number of steps, while it works in the nonce-respect setting As far as we know, this is the first attempt to evaluate integrity of MORUS in the nonce-respect setting.
-
-\subsubsection{Overview.} 
-A general strategy of forgery attacks in the nonce-respect setting is to inject some difference in a message block and to propagate the difference so that the difference can be canceled by another difference in another message block.
-However this approach does not work well against MORUS because its large state size prevents the attackers from controlling the difference to be canceled by another message difference.
-
-Here we focus on the property that the padding scheme for the associated data $A$ and the message $M$ is zero-padding, hence $A$ and $A'=A\|0^*$ and $M$ and $M'=M\|0$ will make the result identical for the associated data processing and the encryption parts, as long as $A,A'$ and $M,M'$ fit in the same number of blocks.
-During the finalization, because $A,A'$ (resp.
-$M,M'$) have different lengths, the corresponding $adlen$s (resp.
-$msglen$s) are different, which appears as $\Delta adlen$ (resp.
-$\Delta msglen$), and is injected from the message input.
-Our strategy is propagating this difference to the 128-bit value tag $T$ and $T'$ such that their difference $\Delta T$ appears higher probability than $2^{-128}$.
-All in all, the forgery succeeds as long as the desired $\Delta T$ is obtained, in other words, the attackers do not have to cancel out the state difference, which is the main advantage of attacking the finalization.
-
-Note that if the attacker uses different messages $M,M'$, not only the new tag $T'$ but also new ciphertext $C'$ must be guessed correctly.
-Because the encryption of MORUS is a simple XOR of the key stream, $C'$ can be easily guessed.
-For this purpose, the attacker should first query a longer message $M'=M\|0^*$ to obtain $C'$.
-Then, $C$ can be obtained by truncating $C'$.
-
-\subsubsection{Differential Trails.}
-
-\subsection{Extending State Recovery to Key Recovery}
-\subsubsection{Overview.}
-\subsubsection{Full Diffusion Rounds.}
+\input{morusAC_07_IniFin}
+
+\bibliographystyle{splncs03}
+\bibliography{morusAC}
 

### 2018-05-08 morusAC_00_Abstract.tex
Added bib file. Prepared separated tex file for Abstract, Introduction, Preliminaries, Initialization-Finalization analysis

--- 2018-05-02 morus.tex
@@
-\newif\ifanonymous
-%\anonymoustrue%uncomment for submission version
-
-\documentclass{llncs}
-\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-\usepackage{common}
-
-\ifanonymous
-\author{}
-\institute{}
-\else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
-\institute{}
-%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
-\fi
-
-\begin{document}
-
-\mainmatter
-
-\title{Note on Linear Biases in MORUS}
-
-\maketitle
-
 \begin{abstract}
@@
 \end{abstract}
-
-\section{Introduction}
-
-Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
-As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
-Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-
-\section{Notation}
-
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
-weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
-weights).
-
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
-$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
-$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
-$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
-\end{tabular}
-}
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
-
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
-Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
-In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
-
-Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-\end{tabular}
-}
-
-The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
-
-\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
-\label{sec:introminimorus}
-
-To simplify the description of the attack, we assume all plaintext blocks are zero.
-However the same results would hold as long as the plaintext is fixed.
-Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
-
-\subsection{Symmetric Linear Combinations}
-
-We begin with a few observations about the \StateUpdate{} function. Beside XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
-\begin{enumerate}
-\item \emph{word-wise} rotations perform a circular shift on each word within a register;
-\item \emph{register-wise} rotations perform a circular shift on a whole register.
-\end{enumerate}
-The second type of rotation always shifts registers by a multiple of the word size $w$.
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
-
-Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
-\begin{definition}
-Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
-\[
-S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
-\]
-is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
-\[
-\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
-\]
-\end{definition}
-\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
-\begin{equation}
-S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
-\label{eq:symmetric}
-\end{equation}
-
-This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
-The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
-
-This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
-As a result, the effect of register-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
-For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
-
-\subsection{\cipher{MiniMORUS}}
-
-In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
-We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
-Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
-
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
-
-\begin{figure}[h]
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate
-  \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function.}
-  \label{fig:minimorus}
-\end{figure}
-
-\section{Linear Trail for \cipher{MiniMORUS}}
-
-\subsection{Overview of the Trail}
-
-To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
-\begin{itemize}
-    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
-    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
-    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
-    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
-    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
-\end{itemize}
-
-The trail fragments are depicted on \Cref{fig:minimorusfragments}.
-In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
-\[
-x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
-\]
-As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, alpha]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
-      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
-      ;
-    \draw[trail, alpha, dashed]
-      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-    \small
-    \begin{align*}
-      \intertext{\cipher{MiniMORUS-640}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\cipher{MiniMORUS-1280}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
-    \end{align*}
-    \caption*{\cipher{MiniMORUS} instances}
-  \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear trail fragments.}
-  \label{fig:minimorusfragments}
-\end{figure}
-
-The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
-In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
-Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
-In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
-
-However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
-The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
-Then $\alpha^t_i$ is used one final time to close the trail.
-One may naturally wonder if some components of this trail are in conflict.
-In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
-
-\subsection{Trail Equation}
-\label{sec:minitraileq}
-
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
-\begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
-\end{alignat*}
-
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
-Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so.
-
-In order to build the equation for the full trail, we start with $E^2_0$:
-\[
-S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
-\]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
-This yields:
-\begin{align*}
-&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
-=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
-\end{align*}
-We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
-This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-
-The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
-
-\subsection{Bias of the Trail}
-\label{sec:minibias}
-
-In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS1280} (cf.
-\Cref{sec:traileq}).
-
-That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
-In both cases it is approximated by zero.
-When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
-
-\subsection{Experimental Verification}
-
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
-Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
-
-\begin{table}[h!]
-  \caption{Experimental verification of trail biases.}
-  \label{tab:miniapproximations}
-  \centerline{
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
-    \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
-\section{Trail for Full \cipher{MORUS}}
-
-\subsection{Symmetrizing the Trail}
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
-That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
-$C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
-\[
-S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
-\]
-
-Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
-Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
-\begin{align*}
-&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
-\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
-\end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
-\begin{align*}
-&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
-\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
-\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
-\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
-\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
-\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
-\end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
-
-\subsection{Bias of the Full Trail}
-
-The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
-Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
-
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
-However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
-
-To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
-
-In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
-Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
-\Cref{fig:trailcollision}).
-While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
-\begin{align}
-&S^t_{2,i} \cdot S^t_{3,i} \oplus S^t_{3,i} \cdot S^t_{2,i+w}
-\oplus S^t_{2,i+w} \cdot S^t_{3,i+w} \oplus S^t_{3,i+w} \cdot S^t_{2,i+2w}\notag\\
-\oplus\; &S^t_{2,i+2w} \cdot S^t_{3,i+2w} \oplus S^t_{3,i+2w} \cdot S^t_{2,i+3w}
-\oplus S^t_{2,i+3w} \cdot S^t_{3,i+3w} \oplus S^t_{3,i+3w} \cdot S^t_{2,i}.\label{eq:8circle}
-\end{align}
-It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      %(xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      %(xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
-    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
-    \node (eq) at ($(and-1|-and1)+(-1,-2.4pt)$) {=};
-    \draw[beta,{Circle[sep=-1.8pt]}-]  ($(and-1)+(0,-2.4pt)$) -| (eq);
-    \draw[gamma,{Circle[sep=-1.8pt]}-] ($(and1)+(0,-2.4pt)$) -- (eq);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
-    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
-  \label{fig:trailcollision}
-\end{figure}
-
-In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
-
-In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
-Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
-As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
-This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
-\Cref{sec:traileq}).
-
-%TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
-%(maybe not relevant due to overall complexity\dots)
-
-\begin{figure}
-  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
-  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
-  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
-  \centering
-  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
-    \begin{scope}
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{27}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
-                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
-                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
-                   (\M{2}{1}) node[right] {$\alpha_7$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
-                   (\M{2}{2}) node[right] {$\gamma_{13}$};
-
-      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
-                   (\M{3}{0}) node[right] {$\beta_{12}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
-                   (\M{1}{0}) node[right] {$\alpha_{2}$};
-
-      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
-                   (\M{2}{0}) node[right] {$\beta_7$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
-      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
-                   (\M{2}{2}) node[right] {$\gamma_7$};
-      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
-                   (\M{2}{4}) node[right] {$\delta_7$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
-                   (\M{3}{1}) node[right] {$\alpha_{14}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
-                   (\M{3}{2}) node[right] {$\gamma_{20}$};
-
-      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
-                   (\M{4}{0}) node[right] {$\beta_{19}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
-    \end{scope}
-
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
-
-    \begin{scope}[yshift=-7.0cm]
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{51}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
-                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
-                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
-                   (\M{2}{1}) node[right] {$\alpha_{37}$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
-                   (\M{2}{2}) node[right] {$\gamma_{4}$};
-
-      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
-                   (\M{3}{0}) node[right] {$\beta_{50}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
-                   (\M{1}{0}) node[right] {$\alpha_{25}$};
-
-      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
-                   (\M{2}{0}) node[right] {$\beta_{38}$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
-      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
-                   (\M{2}{2}) node[right] {$\gamma_{38}$};
-      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
-                   (\M{2}{4}) node[right] {$\delta_{38}$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
-                   (\M{3}{1}) node[right] {$\alpha_{11}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
-                   (\M{3}{2}) node[right] {$\gamma_{42}$};
-
-      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
-                   (\M{4}{0}) node[right] {$\beta_{24}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-    \end{scope}
-  \end{tikzpicture}
-  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
-  \label{fig:twoapproximations}
-\end{figure}
-
-\section{Discussion}
-
-%This is just a sketch for now...
-
-We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
-As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
-This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
-
-The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
-
-With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
-The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
-This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
-
-However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
-Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
-
-\appendix
-
-\section{Trail Equations}
-\label{sec:traileq}
-
-In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
-By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
-
-\subsection{Trail Equation for \cipher{MiniMORUS640}}
-
-This trail equation was already given in \Cref{sec:minitraileq}.
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for \cipher{MiniMORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for full \cipher{MORUS640}}
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
-&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
-&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
-&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
-&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
-&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
-%
-&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
-&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
-&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
-&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 73.
-
-\subsection{Trail Equation for full \cipher{MORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
-&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
-&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
-&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
-&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
-&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
-&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
-%
-&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
-&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
-&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
-&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
-&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 76.
-
-\end{document}

### 2018-05-08 morusAC_01_Introduction.tex
Added bib file. Prepared separated tex file for Abstract, Introduction, Preliminaries, Initialization-Finalization analysis

--- 2018-05-02 morus.tex
@@
-\newif\ifanonymous
-%\anonymoustrue%uncomment for submission version
+\section{Introduction}
+\label{sec/Introduction}
 
-\documentclass{llncs}
-\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-\usepackage{common}
+Authenticated encryption is ... \\
 
-\ifanonymous
-\author{}
-\institute{}
-\else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
-\institute{}
-%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
-\fi
+\noindent
+CAESAR \cite{CAESAR} ...\\
 
-\begin{document}
-
-\mainmatter
-
-\title{Note on Linear Biases in MORUS}
-
-\maketitle
-
-\begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
-More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
-Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
-\end{abstract}
-
-\section{Introduction}
-
+\subsubsection*{Our Contributions.}
 Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
@@
 
-\section{Notation}
-
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
-weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
-weights).
-
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
-$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
-$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
-$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
-\end{tabular}
-}
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
-
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
-Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
-In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
-
-Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-\end{tabular}
-}
-
-The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
-
-\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
-\label{sec:introminimorus}
-
-To simplify the description of the attack, we assume all plaintext blocks are zero.
-However the same results would hold as long as the plaintext is fixed.
-Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
-
-\subsection{Symmetric Linear Combinations}
-
-We begin with a few observations about the \StateUpdate{} function. Beside XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
-\begin{enumerate}
-\item \emph{word-wise} rotations perform a circular shift on each word within a register;
-\item \emph{register-wise} rotations perform a circular shift on a whole register.
-\end{enumerate}
-The second type of rotation always shifts registers by a multiple of the word size $w$.
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
-
-Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
-\begin{definition}
-Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
-\[
-S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
-\]
-is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
-\[
-\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
-\]
-\end{definition}
-\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
-\begin{equation}
-S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
-\label{eq:symmetric}
-\end{equation}
-
-This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
-The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
-
-This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
-As a result, the effect of register-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
-For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
-
-\subsection{\cipher{MiniMORUS}}
-
-In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
-We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
-Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
-
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
-
-\begin{figure}[h]
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate
-  \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function.}
-  \label{fig:minimorus}
-\end{figure}
-
-\section{Linear Trail for \cipher{MiniMORUS}}
-
-\subsection{Overview of the Trail}
-
-To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
-\begin{itemize}
-    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
-    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
-    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
-    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
-    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
-\end{itemize}
-
-The trail fragments are depicted on \Cref{fig:minimorusfragments}.
-In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
-\[
-x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
-\]
-As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, alpha]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
-      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
-      ;
-    \draw[trail, alpha, dashed]
-      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-    \small
-    \begin{align*}
-      \intertext{\cipher{MiniMORUS-640}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\cipher{MiniMORUS-1280}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
-    \end{align*}
-    \caption*{\cipher{MiniMORUS} instances}
-  \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear trail fragments.}
-  \label{fig:minimorusfragments}
-\end{figure}
-
-The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
-In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
-Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
-In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
-
-However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
-The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
-Then $\alpha^t_i$ is used one final time to close the trail.
-One may naturally wonder if some components of this trail are in conflict.
-In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
-
-\subsection{Trail Equation}
-\label{sec:minitraileq}
-
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
-\begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
-\end{alignat*}
-
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
-Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so.
-
-In order to build the equation for the full trail, we start with $E^2_0$:
-\[
-S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
-\]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
-This yields:
-\begin{align*}
-&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
-=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
-\end{align*}
-We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
-This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-
-The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
-
-\subsection{Bias of the Trail}
-\label{sec:minibias}
-
-In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS1280} (cf.
-\Cref{sec:traileq}).
-
-That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
-In both cases it is approximated by zero.
-When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
-
-\subsection{Experimental Verification}
-
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
-Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
-
-\begin{table}[h!]
-  \caption{Experimental verification of trail biases.}
-  \label{tab:miniapproximations}
-  \centerline{
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
-    \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
-\section{Trail for Full \cipher{MORUS}}
-
-\subsection{Symmetrizing the Trail}
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
-That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
-$C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
-\[
-S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
-\]
-
-Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
-Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
-\begin{align*}
-&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
-\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
-\end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
-\begin{align*}
-&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
-\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
-\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
-\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
-\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
-\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
-\end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
-
-\subsection{Bias of the Full Trail}
-
-The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
-Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
-
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
-However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
-
-To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
-
-In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
-Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
-\Cref{fig:trailcollision}).
-While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
-\begin{align}
-&S^t_{2,i} \cdot S^t_{3,i} \oplus S^t_{3,i} \cdot S^t_{2,i+w}
-\oplus S^t_{2,i+w} \cdot S^t_{3,i+w} \oplus S^t_{3,i+w} \cdot S^t_{2,i+2w}\notag\\
-\oplus\; &S^t_{2,i+2w} \cdot S^t_{3,i+2w} \oplus S^t_{3,i+2w} \cdot S^t_{2,i+3w}
-\oplus S^t_{2,i+3w} \cdot S^t_{3,i+3w} \oplus S^t_{3,i+3w} \cdot S^t_{2,i}.\label{eq:8circle}
-\end{align}
-It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      %(xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      %(xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
-    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
-    \node (eq) at ($(and-1|-and1)+(-1,-2.4pt)$) {=};
-    \draw[beta,{Circle[sep=-1.8pt]}-]  ($(and-1)+(0,-2.4pt)$) -| (eq);
-    \draw[gamma,{Circle[sep=-1.8pt]}-] ($(and1)+(0,-2.4pt)$) -- (eq);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
-    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
-  \label{fig:trailcollision}
-\end{figure}
-
-In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
-
-In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
-Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
-As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
-This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
-\Cref{sec:traileq}).
-
-%TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
-%(maybe not relevant due to overall complexity\dots)
-
-\begin{figure}
-  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
-  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
-  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
-  \centering
-  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
-    \begin{scope}
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{27}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
-                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
-                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
-                   (\M{2}{1}) node[right] {$\alpha_7$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
-                   (\M{2}{2}) node[right] {$\gamma_{13}$};
-
-      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
-                   (\M{3}{0}) node[right] {$\beta_{12}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
-                   (\M{1}{0}) node[right] {$\alpha_{2}$};
-
-      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
-                   (\M{2}{0}) node[right] {$\beta_7$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
-      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
-                   (\M{2}{2}) node[right] {$\gamma_7$};
-      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
-                   (\M{2}{4}) node[right] {$\delta_7$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
-                   (\M{3}{1}) node[right] {$\alpha_{14}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
-                   (\M{3}{2}) node[right] {$\gamma_{20}$};
-
-      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
-                   (\M{4}{0}) node[right] {$\beta_{19}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
-    \end{scope}
-
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
-
-    \begin{scope}[yshift=-7.0cm]
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{51}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
-                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
-                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
-                   (\M{2}{1}) node[right] {$\alpha_{37}$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
-                   (\M{2}{2}) node[right] {$\gamma_{4}$};
-
-      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
-                   (\M{3}{0}) node[right] {$\beta_{50}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
-                   (\M{1}{0}) node[right] {$\alpha_{25}$};
-
-      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
-                   (\M{2}{0}) node[right] {$\beta_{38}$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
-      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
-                   (\M{2}{2}) node[right] {$\gamma_{38}$};
-      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
-                   (\M{2}{4}) node[right] {$\delta_{38}$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
-                   (\M{3}{1}) node[right] {$\alpha_{11}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
-                   (\M{3}{2}) node[right] {$\gamma_{42}$};
-
-      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
-                   (\M{4}{0}) node[right] {$\beta_{24}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-    \end{scope}
-  \end{tikzpicture}
-  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
-  \label{fig:twoapproximations}
-\end{figure}
-
-\section{Discussion}
-
-%This is just a sketch for now...
-
-We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
-As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
-This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
-
-The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
-
-With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
-The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
-This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
-
-However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
-Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
-
-\appendix
-
-\section{Trail Equations}
-\label{sec:traileq}
-
-In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
-By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
-
-\subsection{Trail Equation for \cipher{MiniMORUS640}}
-
-This trail equation was already given in \Cref{sec:minitraileq}.
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for \cipher{MiniMORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for full \cipher{MORUS640}}
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
-&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
-&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
-&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
-&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
-&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
-%
-&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
-&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
-&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
-&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 73.
-
-\subsection{Trail Equation for full \cipher{MORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
-&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
-&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
-&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
-&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
-&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
-&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
-%
-&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
-&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
-&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
-&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
-&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 76.
-
-\end{document}
+\subsubsection*{Paper Outline.}
+This paper is organized as follows......

### 2018-05-08 morusAC_02_Preliminaries.tex
Added bib file. Prepared separated tex file for Abstract, Introduction, Preliminaries, Initialization-Finalization analysis

\section{Preliminaries}
\label{Sec/Preliminaries}

\subsection{Specification of MORUS}
\label{subsec/Spec}
MORUS \cite{MORUS} is a family of authenticated ciphers with two internal state sizes: 640 and 1280 bits. Each member is called \cipher{MORUS640} and \cipher{MORUS1280}, respectively. In the recommended parameters, \cipher{MORUS640} supports 128-bit keys and \cipher{MORUS1280} supports both 128-bit and 256-bit keys. The tag size is 128 bits or shorter. The designers strongly recommend using a 128-bit tag.

The state of MORUS is divided into five registers. Let $q$ be the register size, where $q = 128$ for \cipher{MORUS640} and $q=256$ for \cipher{MORUS1280}. Let $S_i, i \in\{0,1,2,3,4\}$ be a register. Then, the state is represented as $S_0\|S_1\|S_2\|S_3\|S_4$. The state is first initialized by a key and nonce, and then updated by associated data blocks and message blocks step by step. Hence, we use a superscript to denote the step number of the state as $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is $-16$ before the initialization and $t=0$ after the initialization.

The computation of MORUS consists of four parts; initialization, associated data process, encryption, and finalization. All parts share the core design called the {\tt StepUpdate} function. %Hereafter we explain {\tt StepUpdate}.

\subsubsection{Step Update Function.}
{\tt StepUpdate} takes as input the internal state $S^t = S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$ and additional register-wise value $m^t$. $m^t$ is supposed to be associated data block or message block. $m^t$ can sometimes be 0 or constant when the additional input is unnecessary.

{\tt StepUpdate} is composed of 5 rounds with similar operations to update
the state $S^t$, in which the additional input $m^t$ is used in rounds 2 to 5, but not in round 1. Each round uses the word-wise left circular shift, which divides a $q$-bit register value into 4 words of $q/4$ bits, and performs left circular shift operation for every $q/4$-bit word. Throughout the paper, we denote the word size by $w$, i.e. $w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}. The word-wise circular shift constants $b_i$ in round $i$ are defined in Table~\ref{Tbl:rcon}. Besides, the left rotation of a whole $q$-bit register is used. The register-wise circular shift constants are also listed in Table~\ref{Tbl:rcon}.
\begin{table}[!htb]
\centering
\caption{Left Circular-Shift Constants of MORUS} \label{Tbl:rcon}
\begin{tabular}{c||ccccc|ccccc} \hline
& \multicolumn{5}{c|}{word-wise circular shift} & \multicolumn{5}{c}{register-wise circular shift} \\
                  & \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline
\cipher{MORUS640} &  5 & 31 &  7 & 22 & 13 & 32 &  64 &  96 &  64 & 32 \\
\cipher{MORUS1280}& 13 & 46 & 38 &  7 &  4 & 64 & 128 & 192 & 128 & 64 \\ \hline
\end{tabular}
\end{table}
We denote those register-wise and word-wise left circular shift by $x \lll y$ and $x \lll_w y$, respectively. (Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document.)

Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
\begin{align*}
\textrm{Round 1:} &&
S^{t+1}_0 &\leftarrow ( S_t^0 \oplus S^t_1 \cdot S^t_2 \oplus S^t_3 ) \lll_w b_0 &
S^t_3 \leftarrow S^t_3 \lll b'_0.\\
\textrm{Round 2:} &&
S^{t+1}_1 &\leftarrow ( S_t^1 \oplus S^t_2 \cdot S^t_3 \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
S^t_4 \leftarrow S^t_3 \lll b'_1.\\
\textrm{Round 3:} &&
S^{t+1}_2 &\leftarrow ( S_t^2 \oplus S^t_3 \cdot S^t_4 \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
S^t_0 \leftarrow S^t_3 \lll b'_2.\\
\textrm{Round 4:} &&
S^{t+1}_3 &\leftarrow ( S_t^3 \oplus S^t_4 \cdot S^t_0 \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
S^t_1 \leftarrow S^t_3 \lll b'_3.\\
\textrm{Round 5:} &&
S^{t+1}_4 &\leftarrow ( S_t^4 \oplus S^t_0 \cdot S^t_1 \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
S^t_2 \leftarrow S^t_3 \lll b'_4.
\end{align*}

\subsubsection{Initialization.}
The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K$ and a 128-bit nonce IV into the state, and running the {\tt StateUpdate} for 16 steps. The key and IV are loaded into the state as follows:
\begin{align*}
S^{-16}_0 = \textrm{IV}_{128}, &&
S^{-16}_1 = K_{128}, &&
S^{-16}_2 = 1^{128}, &&
S^{-16}_3 = const_0, &&
S^{-16}_4 = const_1.
\end{align*}
Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K$.

The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size. The state is initialized as
\begin{align*}
S^{-16}_0 &= \textrm{IV}_{128} \| 0^{128}, &
S^{-16}_1 &= K_{128}\|K_{128}, &
S^{-16}_2 = 1^{256}, \\
S^{-16}_3 &= 0^{256}, &
S^{-16}_4 &= const_1\|const_0.
\end{align*}
After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K_{128}\|K_{128}$.

\subsubsection{Associated Data Processing.}
After the initialization, the associated data $A$ is processed. For the padding, if the last associated data block is not a full block, use `0' bits to pad it to 128 bits and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state. Note that if the length of $A$ denoted by `$adlen$' is 0, the associated data processing is skipped.

Then, state is updated by $S^{t+1} \leftarrow {\tt StateUpdate}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{adlen}{128} \rceil -1$ and $\ell = \lceil \frac{adlen}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.

\subsubsection{Encryption.}
At each step of the encryption, a 16-byte (resp. 32-byte) message block $M_i$ is used to update the state, and $M_t$ is encrypted to $C_t$ in \cipher{MORUS640} (resp. \cipher{MORUS1280}).

If the last message block is not a full block, use `0' bits to pad it to 128 and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state. But only the partial block is encrypted. Note that if the message length denoted by `$msglen$' is 0, encryption is skipped.

In \cipher{MORUS640}, Let $u = \lceil \frac{adlen}{128} \rceil$ and $v = \lceil \frac{msglen}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$;
\begin{align*}
C^t &\leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 96) \oplus (S^{u+t}_2 \cdot S^{u+t}_3),\\
S^{u+t+1} &\leftarrow {\tt StateUpdate}(S^{u+t},M^t).
\end{align*}

For \cipher{MORUS1280}, $u = \lceil \frac{adlen}{256} \rceil$ and $v = \lceil \frac{msglen}{256} \rceil$. The equation to compute ciphertext block $C^t$ is $C^t \leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 192) \oplus (S^{u+t}_2 \cdot S^{u+t}_3)$.

\subsubsection{Finalization.}
The finalization generates the authentication tag $T$ using 10 more steps. We only discuss the case that $T$ is not truncated. The associated data length and the message length are used to update the state as follows.
\begin{enumerate}
\item $tmp \leftarrow adlen\|msglen$ for \cipher{MORUS640} and $tmp \leftarrow adlen\|msglen\|0^{128}$ for \cipher{MORUS1280}, where $adlen$ and $msglen$ are represented as 64-bit integers.
\item $S^{u+v}_4 \leftarrow S^{u+v}_4 \oplus S^{u+v}_0.$
\item For $t = u+v, u+v+1, \ldots, u+v+9,$ compute
%\begin{equation*}
$S^{t+1} \leftarrow {\tt StateUpdate} (S^t, tmp)$.
%\end{equation*}
\item For \cipher{MORUS640}, $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$. For \cipher{MORUS1280}, $S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ is computed and the least significant 128 bits are produced as a tag $T$.
\end{enumerate}

\subsubsection{Security.}
When the tag size is 128 bits, integrity is claimed up to 128-bits and confidentiality is claimed up to the number of bits of the key.
\begin{table}[!htb]
\centering
\caption{Security Goals of MORUS.}
\begin{tabular}{ccc}\hline
                        & Confidentiality (bits) & Integrity (bits) \\ \hline
\cipher{MORUS640}-128   & 128                    & 128              \\
\cipher{MORUS1280}-128  & 128                    & 128              \\
\cipher{MORUS1280}-256  & 256                    & 128              \\ \hline
\end{tabular}
\end{table}

\subsection{Notation}
\label{subsec/Notation}

We define the \emph{bias} of an event $E$ as:
\[
\bias(E) \eqdef 2\Pr(E)-1.
\]
Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
\[
\weight(E) \eqdef -\log|\bias(E)|
\]
where $\log()$ denotes logarithm in base 2. By the piling-up lemma, the bias (resp. weight) of an XOR of independent variables is equal to the product (resp. sum) of their individual biases (resp. weights).

We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
\centerline{
\def\arraystretch{1.25}
\begin{tabular}{lcl}
$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
\end{tabular}
}
In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.

In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts. Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow. In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.

%Finally, let $x$ denote a register. Recall that registers of \cipher{MORUS} are composed of four words. We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
%\centerline{
%\def\arraystretch{1.25}
%\begin{tabular}{lcl}
%$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
%$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
%\end{tabular}
%}
%
%The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered. Throughout, $w$ denotes the  word size, i.e. $w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.

### 2018-05-08 morusAC_07_IniFin.tex
Added bib file. Prepared separated tex file for Abstract, Introduction, Preliminaries, Initialization-Finalization analysis

\section{Analysis on Initialization and Finalization of Reduced MORUS}
\label{sec/IniFin}

The bias discussed in the previous sections analysed the encryption part of the MORUS. In this section, for comprehensive security analysis of MORUS, we provide new attacks on reduced version of the initialization and the finalization. We emphasize that the results in this section do not threat any security claim by the designers. However, we believe that investigating all parts of the design with new approaches provides better understanding of the design and will be useful especially when the design will be tweaked in future.

\subsection{Forgery with Reduced Finalization}
We present forgery attacks on 3 out of 10 steps of \cipher{MORUS1280} that claims 128-bit security for integrity. The attack only works for a limited number of steps, while it works in the nonce-respect setting. As far as we know, this is the first attempt to evaluate integrity of MORUS in the nonce-respect setting.

\subsubsection{Overview.}
A general strategy of forgery attacks in the nonce-respect setting is to inject some difference in a message block and to propagate the difference so that it can be canceled by difference in another message block. However this approach does not work well against MORUS because its large state size prevents the attackers from controlling the difference to be canceled by another message difference.

Here we focus on the property that the padding for an associated data $A$ and a message $M$ is the zero-padding, hence $A$ and $A'=A\|0^*$ and $M$ and $M'=M\|0$ will make the result identical for the associated data processing and the encryption parts, as long as $A,A'$ and $M,M'$ fit in the same number of blocks. During the finalization, because $A,A'$ (resp. $M,M'$) have different lengths, the corresponding $adlen$s (resp. $msglen$s) are different, which appears as $\Delta adlen$ (resp. $\Delta msglen$), and is injected from the message input. Our strategy is to propagate this difference to 128-bit tag $T$ and $T'$ such that their difference $\Delta T$ appears higher probability than $2^{-128}$. All in all, the forgery succeeds as long as the desired $\Delta T$ is obtained, in other words, the attackers do not have to cancel the state difference, which is the main advantage of attacking the finalization.

Note that if the attacker uses different messages $M,M'$, not only the new tag $T'$ but also new ciphertext $C'$ must be guessed correctly. Because the encryption of MORUS is a simple XOR of the key stream, $C'$ can be easily guessed. For this purpose, the attacker should first query a longer message $M'=M\|0^*$ to obtain $C'$. Then, $C$ can be obtained by truncating $C'$.

\subsubsection{Differential Trails.}
Recall that the message input during the finalization of \cipher{MORUS1280} is $adlen\|msglen\|0^{128}$ where $adlen$ and $msglen$ are 64-bit strings. We set $\Delta adlen$ to be low Hamming weight, e.g.~{\tt 0x0000000000000001}. 
%Hereafter, we use {\tt typefont} denote hexadecimal numbers.
This difference propagates during 3 steps is specified in Table~\ref{Tbl:fin_diff}.

Recall that each step consists of 5 rounds and the input message is absorbed to the state in rounds 2 to 5. The trail in Table~\ref{Tbl:fin_diff} initially does not have any difference and the same continues even after round 1. Differences start to appear from round 2 and they will go through the bitwise-AND operation from round 4. We need to pay 1-bit to control each active AND gate. The probability evaluation for round 15 can be ignored because in this round only $S_4$ is non-linearly updated, while $S_4$ is never used for computing the tag. Finally, bitwise-AND in the tag computation is taken into account. Note that the tag is only 128 LSBs, thus the number of active AND gates should be counted only in those bits. As shown in Table~\ref{Tbl:fin_diff}, we can have a particular tag difference $\Delta T$ with probability $2^{-88}$. Thus after observing $A$ and corresponding $T$, $A\|0$ and $(T \oplus \Delta T)$ is a valid pair with probability $2^{-88}$.

\begin{table}[!tb]
\begin{center}
\caption{Differential Propagation through Three Steps in the Finalization. $i$ Five lines for round $i$ denote the state difference $S_0, \cdots, S_4$ after the the round $i$ transformation. Weight is a Hamming weight of the state difference and accumulated probability is a probability to satisfy the trail from the beginning.}
\label{Tbl:fin_diff}
{\tiny
\begin{tabular}{ccccccc} \hline
Round & \multicolumn{4}{c}{State Difference} & Weight & Accumulated \\
      & & & &                                &        & probability \\ \hline
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
Ini   & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & $-$ \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
1     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & $1$\\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000400000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & \\
2     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & $1$\\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000400000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & \\
3     & {\tt 0000004000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & $1$\\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
4     & {\tt 0000004000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 1  & $1$\\
      & {\tt 0020000000000080} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
5     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-1}$\\
      & {\tt 0020000000000080} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
      & {\tt 0000040000000010} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\ \hline

      & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
6     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-3}$\\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
      & {\tt 0000040000000010} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\ \hline

      & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \\
7     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-6}$\\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \\
8     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & $2^{-10}$\\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
9     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & $2^{-14}$\\
      & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7  & \\
      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000040000000010} & {\tt 0000000000000000} & 2  & \\ \hline

      & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
10    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-20}$\\
      & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7  & \\
      & {\tt 4000140000000010} & {\tt 0000000000000000} & {\tt 0000400000000100} & {\tt 0000000000010000} & 7  & \\ \hline

      & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9  & \\
      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
11    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-28}$\\
      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
      & {\tt 4000140000000010} & {\tt 0000000000000000} & {\tt 0000400000000100} & {\tt 0000000000010000} & 7  & \\ \hline

      & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9  & \\
      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \\
12    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-39}$\\
      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline

      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \\
13    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 & $2^{-53}$\\
      & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline

      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 & \\
14    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 & $2^{-69}$\\
      & {\tt 0228000280020080} & {\tt 0000040000000000} & {\tt 2000008000202008} & {\tt 1000004000004021} & 18 & \\
      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline

%Weight: 69
      & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 & \\
15    & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & {\tt 0400114000040000} & 14 & $-$\\
      & {\tt 0228000280020080} & {\tt 0000040000000000} & {\tt 2000008000202008} & {\tt 1000004000004021} & 18 & \\
      & {\tt 0000400000000100} & {\tt 0000000000010000} & {\tt 4000140000000010} & {\tt 0000000000000000} & 7  & \\ \hline

$\Delta T$ &                   &                        & {\tt 600080830020f00a} & {\tt 1405414005044421} & & $2^{-88}$ \\ \hline
\end{tabular}
}
\end{center}
\end{table}

\subsubsection{Remarks.} The fact that $S_4$ updated in the last round is not used to compute a tag implies that the MORUS finalization generally includes unnecessary computations. It may be interesting to tweak the design such that the tag can also depend on $S_4$. Indeed in Table~\ref{Tbl:fin_diff}, we can observe some jump-up of the probability in the tag computation. This is because the non-linearly involved terms are $S_2 \cdot S_3$, and $S_3$ that was updated 2-round before has a high Hamming weight. In this sense, involving $S_4$ in non-linear terms of the tag computation imposes more difficulties for the attacker.

\subsection{Extending State Recovery to Key Recovery}
Kales et al. \cite{cryptoeprint:2017:1137} showed that internal state recovery of \cipher{MORUS640} can be recovered under the nonce-misuse scenario only with $2^7$ plaintext-ciphertext pairs. As claimed by \cite{cryptoeprint:2017:1137} the attack is naturally extended to \cipher{MORUS1280} though Kales et al. \cite{cryptoeprint:2017:1137} did not demonstrate specific attacks. The recovered state allows the attacker to mount universal forgery attack under the same nonce, whereas the key still cannot be recovered.

\subsubsection{Overview.}
\subsubsection{Full Diffusion Rounds.}

### 2018-05-08 src-pyth-sse/Ciphers/full_morus_640.c
try to implement full morus with SSE for faster speed

@@ -1,0 +1,255 @@
+#include "full_morus_640.h"
+
+#ifdef _MSC_VER
+#define inline __inline
+// #define _mm_store_si128 _mm_storeu_si128
+// #define _mm_load_si128 _mm_loadu_si128
+#endif
+
+#define n1 5
+#define n2 31
+#define n3 7
+#define n4 22
+#define n5 13
+
+#define ROTL_128(x,n)     XOR_128(_mm_slli_epi32((x), (n)),  _mm_srli_epi32((x),(32-n)))  /*Rotate 4 32-bit unsigned integers in x to the left by n-bit positions*/
+#define ROTL8(x)       		_mm_shuffle_epi8((x), _mm_set_epi8(14,13,12,11,10,9,8,7,6,5,4,3,2,1,0,15) )  /*Rotate 4 32-bit unsigned integers in x to the left by 8-bit positions*/
+
+#define ROTL32(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(2,1,0,3))   /*Rotate x by 32-bit positions to the left*/
+#define ROTL64(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(1,0,3,2))   /*Rotate x by 64-bit positions to the left*/
+#define ROTL96(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(0,3,2,1))   /*Rotate x by 96-bit positions to the left*/
+
+// #define STORE_128(x,p)     _mm_store_si128((__m128i *)(p), (x))         /*store the 128-bit word x into memeory address p, where p is the multile of 16 bytes*/
+// #define LOAD_128(p)        _mm_load_si128((__m128i *)(p))               /*load 16 bytes from the memory address p, return a 128-bit word, where p is the multile of 16 bytes*/
+
+void morus_640_print_state(state state) {
+	// printf("--------------------------------------------\n");
+	// printf("   S0   |   S1   |   S3   |   S3   |   S4   \n");
+	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[3], state[1].f32[3], state[2].f32[3], state[3].f32[3], state[4].f32[3]);
[…]

### 2018-05-08 src-pyth-sse/Ciphers/morus.h

@@ -1,0 +1,6 @@
+#pragma once
+
+#include "full_morus_640.h"
+#include "mini_morus_640.h"
+#include "mini_morus_1280.h"
+#include "../Utilities/utilities.h"

### 2018-05-08 src-pyth-sse/Utilities/aes-rng.c

@@ -1,0 +1,155 @@
+// Use AES-based RNG
+#include "aes-rng.h"
+
+#ifdef __AES__
+
+struct RNG_state *init_aesrand_r(uint32_t seed1, uint32_t seed2) {
+	// Internal state: count + t
+	struct RNG_state *rng = (struct RNG_state *) malloc(sizeof(struct RNG_state));
+	assert(rng);
+	rng->count = _mm_setr_epi32(seed1, seed2, 0, 0);
+	rng->k = _mm_setr_epi32(0, 0, seed1, seed2);
+	rng->next = 0;
+	return rng;
+}
+
+uint32_t aesrand_int32_r(struct RNG_state *rng) {
+	switch (rng->next++) {
+		case 0:
+			// Increment counter
+			rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
+
+			// 6 AES rounds
+			rng->t = rng->count;
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+
+			return (uint32_t) _mm_extract_epi32(rng->t, 0);
+		case 1:
+			return (uint32_t) _mm_extract_epi32(rng->t, 1);
+		case 2:
+			return (uint32_t) _mm_extract_epi32(rng->t, 2);
+		case 3:
+			rng->next = 0;
+			return (uint32_t) _mm_extract_epi32(rng->t, 3);
+		default:
+			assert(0);
+	}
+}
+
+uint64_t aesrand_int64_r(struct RNG_state *rng) {
+	switch (rng->next++) {
+		case 0:
+			// Increment counter
+			rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
+
+			// 6 AES rounds
+			rng->t = rng->count;
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
[…]

### 2018-05-08 src-pyth-sse/Utilities/aes-rng.h

@@ -1,0 +1,20 @@
+#pragma once
+
+#include "types.h"
+
+//struct RNG_state {
+//    __m128i count;
+//    __m128i t;
+//    __m128i k;
+//    int next;
+//};
+
+struct RNG_state *init_aesrand_r(uint32_t seed1, uint32_t seed2);
+
+uint32_t aesrand_int32_r(struct RNG_state *rng);
+
+uint64_t aesrand_int64_r(struct RNG_state *rng);
+
+uint8_t aesrand_int8_r(struct RNG_state *rng);
+
+__m128i aesrand_m128i_r(struct RNG_state *rng);

### 2018-05-08 src-pyth-sse/Utilities/sse.h

@@ -1,0 +1,4 @@
+#include <immintrin.h>
+
+#define XOR_128(x,y)      _mm_xor_si128((x),(y))     /*XOR_128(x,y) = x ^ y, where x and y are two 128-bit word*/
+#define AND_128(x,y)      _mm_and_si128((x),(y))     /*AND_128(x,y) = x & y, where x and y are two 128-bit word*/

### 2018-05-08 src-pyth-sse/Utilities/std.h

@@ -1,0 +1,15 @@
+#pragma once
+
+#include <math.h>
+#include <omp.h>
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <time.h>
+#include <wchar.h>
+
+// for AES
+#include <smmintrin.h>
+#include <wmmintrin.h>
+#include <assert.h>

### 2018-05-08 src-pyth-sse/Utilities/types.h

@@ -1,0 +1,24 @@
+#pragma once
+
+#include "std.h"
+
+union Register {
+	uint32_t m32;
+	uint64_t m64;
+	__m128i f32;
+	// uint32_t f32[4];
+	// uint64_t f64[4];
+};
+
+struct RNG_state {
+	__m128i count;
+	__m128i t;
+	__m128i k;
+	int next;
+};
+
+typedef union Register state[5];
+
+struct Properties {
+	uint8_t num_proc_to_use;       // specify the number of cores to use
+};

### 2018-05-08 src-pyth-sse/Utilities/utilities.h

@@ -1,0 +1,9 @@
+#pragma once
+
+#include "std.h"
+#include "types.h"
+#include "aes-rng.h"
+#include "maths.h"
+#include "printing.h"
+#include "args.h"
+#include "sse.h"

### 2018-05-08 src-pyth-sse/gen_masks.py

@@ -1,0 +1,109 @@
+#!/usr/bin/env python3
+
+from tools import save, gen_name, bcolors
+from masks import masks_list
+
+def header_head():
+    z = ''
+    z += '#pragma once\n'
+    z += '\n'
+    z += '#include "../Utilities/utilities.h"\n'
+    z += '\n'
+    return z
+
+def main_head(name):
+    z = '#include "' + name + '.h"\n'
+    z += '\n'
+    return z
+
+def gen_mask(list_index):
+    mask = 0;
+    for idx in list_index:
+        mask ^= 1 << idx
+    return mask
+
+def gen_union(kind,width):
+    union = ''
+    if kind == 1:
+        union = 'm' + str(width)
+    if kind == 2:
+        union = 'f' + str(width)
+    return union
+
+def gen_function(kind,width,num,states,ciphers, weight):
+
+    union = gen_union(kind,width)
+    name = gen_name(kind, width)
+    name += '_' + str(num)
+
+    z = '\n'
+    z += 'uint64_t ' + name + '(state *saved_states, union Register *saved_ciphers) {\n'
+    z += '\tuint' + str(width) + '_t res = 0;\n'
+    if kind != 1:
+        z += '\tuint' + str(width) + '_t t = 0;\n'
+
+    for K, C in ciphers.items():
+        if kind == 1:
+            z += '\tres ^= saved_ciphers' + K + '.' + union + ' & ' + hex(gen_mask(C)) + ';\n'
+        else:
+            z += '\tt = (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 0);\n'
+            z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 1);\n'
+            z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 2);\n'
+            z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 3);\n'
+            z += '\tres ^= t &' + hex(gen_mask(C)) + ';\n'
+
+    for K, C in states.items():
+        if kind == 1:
+            z += '\tres ^= saved_states' + K + '.' + union + ' & ' + hex(gen_mask(C)) + ';\n'
+        else:
+            z += '\tt = (uint32_t) _mm_extract_epi32(saved_states' + K + '.' + union + ', 0);\n'
[…]

### 2018-05-08 src-pyth-sse/masks.py

@@ -1,0 +1,137 @@
+############################################################
+#                       MORUS 640
+############################################################
+
+ciphers_640 = [
+{   '[0]': [27]},               # alpha
+{   '[0]': [0]},                # beta
+{},                             # gamma
+{},                             # delta
+{},                             # epsil
+{   '[0]': [27],                # appr1
+    '[1]': [0,26,8],
+    '[2]': [31,13,7],
+    '[3]': [12]},
+{   '[1]': [2],                 # appr2
+    '[2]': [1,7,15,27],
+    '[3]': [6,20,14],
+    '[4]': [19]},
+{   '[0]': [27],                # full
+    '[1]': [0,2,8,26],
+    '[2]': [31,27,15,13,1],
+    '[3]': [20,14,12,6],
+    '[4]': [19]}
+]
+
+states_640 = [
+{   '[1][0]': [0]},             # alpha
+{   '[0][0]': [0],              # beta
+    '[0][1]': [0]},
+{   '[0][1]': [0],              # gamma
+    '[0][4]': [0],
+    '[1][1]': [31]},
+{   '[0][4]': [0],              # delta
+    '[1][2]': [0],
+    '[1][4]': [13]},
+{   '[0][2]': [0],              # epsil
+    '[1][0]': [0],
+    '[1][2]': [7]},
+{   '[2][2]': [0]},             # appr1
+{   '[2][2]': [0]},             # appr2
+{}                              # full
+]
+
+weight_mini_640 = [
+1,              # alpha
+1,              # beta
+1,              # gamma
+1,              # delta
+1,              # epsil
+7,              # appr1
+9,              # appr2
+16,             # full
+]
+
+weight_640 = [
+10,              # alpha
+10,              # beta
+10,              # gamma
+10,              # delta
+10,              # epsil
+# 14,              # appr1
+# 18,              # appr2
+# 16,             # full
+]
+
+
+############################################################
+#                       MORUS 1280
+############################################################
+
+ciphers_1280 = [
+{   '[0]': [51]},               # alpha
+{   '[0]': [0]},                # beta
+{},                             # gamma
+{},                             # delta
+{},                             # epsil
+{   '[0]': [51],                # appr1
+    '[1]': [0,33,55],
[…]

### 2018-05-08 src-pyth-sse/masks.py
update weights expected

@@ -53,14 +53,14 @@ weight_mini_640 = [
 ]
 
 weight_640 = [
-10,              # alpha
-10,              # beta
-10,              # gamma
-10,              # delta
-10,              # epsil
-# 14,              # appr1
-# 18,              # appr2
-# 16,             # full
+4,              # alpha
+4,              # beta
+4,              # gamma
+4,              # delta
+4,              # epsil
+28,             # appr1
+36,             # appr2
+73,             # full
 ]
 
 

### 2018-05-08 morusAC_07_IniFin.tex
Updated analysis on initialization function

@@
 \subsection{Forgery with Reduced Finalization}
+\label{subsec/Fin}
 We present forgery attacks on 3 out of 10 steps of \cipher{MORUS1280} that claims 128-bit security for integrity.
@@
 \subsection{Extending State Recovery to Key Recovery}
-Kales et al.
-\cite{cryptoeprint:2017:1137} showed that internal state recovery of \cipher{MORUS640} can be recovered under the nonce-misuse scenario only with $2^7$ plaintext-ciphertext pairs.
+\label{subsec/Ini}
+Kales et al.~\cite{cryptoeprint:2017:1137} showed that internal state of \cipher{MORUS640} can be recovered under the nonce-misuse scenario using $2^5$ plaintext-ciphertext pairs.
 As claimed by \cite{cryptoeprint:2017:1137} the attack is naturally extended to \cipher{MORUS1280} though Kales et al.
 \cite{cryptoeprint:2017:1137} did not demonstrate specific attacks.
-The recovered state allows the attacker to mount universal forgery attack under the same nonce, whereas the key still cannot be recovered.
+The recovered state allows the attacker to mount universal forgery attack under the same nonce.
+However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker to back track the state value to the initial state.
+In this section, we show that meet-in-the-middle attacks can allow the attacker to recover the key faster than the exhaustive search for a relatively large number of steps, i.e.~10 out of 16 steps in \cipher{MORUS1280}.
 
 \subsubsection{Overview.}
+We divide the 10 steps of the initialization computation into two subsequent parts $F_0$ and $F_1$.
+(We later set that $F_0$ is the first 4 steps and $F_1$ is the last 6 steps.) Let $S^{-10}$ be the initial state value before setting the key, i.e.~$S^{-10} = (IV\|0^{128},0^{256},1^{256},0^{256},const_0\|const_1)$.
+Also let $S^0$ be 1280-bit state value after the finalization, which is now assumed to be recovered with the nonce-misuse analysis~\cite{cryptoeprint:2017:1137}.
+We then have the following relation.
+\begin{align*}
+F_1 \circ F_0 \bigl(S^{-10} \oplus (0, K\|K, 0, 0, 0)\bigr) \oplus (0,K\|K,0,0,0) = S^0.
+\end{align*}
+
+Here, our strategy is to recover $K$ by independently processing $F_0$ and $F_1^{-1}$ to find the following match.
+\begin{align*}
+F_0 (S^{-10} \oplus (0, K\|K, 0, 0, 0)) \stackrel{?}{=} F_1^{-1} (S^0 \oplus (0,K\|K,0,0,0)).
+\end{align*}
+
+To evaluate the attack complexity, we consider the following configuration of the attack parameters.
+\begin{itemize}
+\item $G_0$: a set of bits of $K$ that are guessed for computing $F_0$.
+\item $G_1$: a set of bits of $K$ that are guessed for computing $F_1^{-1}$.
+\item $G_2$: a set of bits in the intersection of $G_0$ and $G_1$.
+\item $x$ bits can match after processing $F_0$ and $F_1^{-1}$.
+\end{itemize}
+Suppose that the union of $G_0$ and $G_1$ covers all the bits of $K$. The attack exhaustively guesses $G_2$ and performs the following procedure for each guess.
+\begin{enumerate}
+\item $F_0$ is computed $2^{|G_0|-|G_2|}$ times and the results are stored in a table $T$. (Because $|G_1|-|G_2|$ bits are unknown, only a part of the state is computed.)
+\item $F_1^{-1}$ is computed $2^{|G_1|-|G_2|}$ times and for each result we check the match with any entry in $T$.
+\item There are $2^{|G_0|-|G_2| + |G_1|-|G_2|}$ combinations, and the number of valid matches reduces to $2^{|G_0|-|G_2| + |G_1|-|G_2| - x}$ after matching the $x$ bits.
+\item Check the correctness of the guess by using one plaintext-ciphertext pair.
+\end{enumerate}
+
+In the end,
+$F_0$ is computed $2^{|G_2|} \cdot 2^{|G_0|-|G_2|} = 2^{|G_0|}$ times.
+Similarly, $F_1^{-1}$ is computed $2^{|G_1|}$ times.
+The number of the total candidates after the $x$-bit match is $2^{|G_2|} \cdot 2^{|G_0|-|G_2| + |G_1|-|G_2| - x} = 2^{|G_0| + |G_1| - |G_2| - x}$.
+Hence, the key $K$ is recovered with complexity $$\max( 2^{|G_0|}, 2^{|G_1|}, 2^{|G_0| + |G_1| - |G_2| - x}).$$
+Suppose that we choose $|G_0|$ and $|G_1|$ to be balanced i.e. $|G_0|=|G_1|$. Then, the complexity is $$\max( 2^{|G_0|}, 2^{2|G_0| - |G_2| - x}).$$
+Two terms are balanced when $x = |G_0| - |G_2|$, namely, the number of matched bits in the middle of two functions must be equal to the number of independently guessed bits to compute $F_0$ and $F_1^{-1}$.
+
+In the attack below, we choose $|G_0|=|G_1|=127$ and $|G_2|=126$ (equivalently $|G_2|-|G_0| = |G_2|-|G_1| = 1$) in order to aim $x=1$-bit match in the middle, which maximizes the number of attacked rounds.
+
 \subsubsection{Full Diffusion Rounds.}

### 2018-05-08 morusAC_02_Preliminaries.tex

@@
 \textrm{Round 1:} &&
-S^{t+1}_0 &\leftarrow ( S_t^0 \oplus S^t_1 \cdot S^t_2 \oplus S^t_3 ) \lll_w b_0 &
+S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 ) \lll_w b_0 &
 S^t_3 \leftarrow S^t_3 \lll b'_0.\\
 \textrm{Round 2:} &&
-S^{t+1}_1 &\leftarrow ( S_t^1 \oplus S^t_2 \cdot S^t_3 \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
-S^t_4 \leftarrow S^t_3 \lll b'_1.\\
+S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
+S^t_4 \leftarrow S^t_4 \lll b'_1.\\
 \textrm{Round 3:} &&
-S^{t+1}_2 &\leftarrow ( S_t^2 \oplus S^t_3 \cdot S^t_4 \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
-S^t_0 \leftarrow S^t_3 \lll b'_2.\\
+S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
+S^t_0 \leftarrow S^t_0 \lll b'_2.\\
 \textrm{Round 4:} &&
-S^{t+1}_3 &\leftarrow ( S_t^3 \oplus S^t_4 \cdot S^t_0 \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
-S^t_1 \leftarrow S^t_3 \lll b'_3.\\
+S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
+S^t_1 \leftarrow S^t_1 \lll b'_3.\\
 \textrm{Round 5:} &&
-S^{t+1}_4 &\leftarrow ( S_t^4 \oplus S^t_0 \cdot S^t_1 \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
-S^t_2 \leftarrow S^t_3 \lll b'_4.
+S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
+S^t_2 \leftarrow S^t_2 \lll b'_4.
 \end{align*}
@@
 \subsubsection{Initialization.}
-The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K$ and a 128-bit nonce IV into the state, and running the {\tt StateUpdate} for 16 steps.
-The key and IV are loaded into the state as follows:
+The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K_{128}$ and a 128-bit nonce $IV_{128}$ into the state, and running the {\tt StateUpdate} for 16 steps.
+The key and nonce are loaded into the state as follows:
 \begin{align*}
@@
 \end{align*}
-Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K$.
+Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K_{128}$.
 
-The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size. The state is initialized as
+The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size and the two possible key size:
+\begin{itemize}
+  \item[-] When the key size is 128 bits, $K = K_{128} \|K_{128}$
+  \item[-] When the key size is 256 bits, $K = K_{256}$
+\end{itemize}
+The state is initialized as
 \begin{align*}
 S^{-16}_0 &= \textrm{IV}_{128} \| 0^{128}, &
-S^{-16}_1 &= K_{128}\|K_{128}, &
+S^{-16}_1 &= K, &
 S^{-16}_2 = 1^{256}, \\
@@
 \end{align*}
-After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K_{128}\|K_{128}$.
+After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K$.
 

### 2018-05-08 src-pyth-sse/Ciphers/full_morus_1280.c
add full_morus_1280 with AVX2 instruction set

@@ -1,0 +1,148 @@
+#include "full_morus_1280.h"
+
+#ifdef __AVX2__
+
+/*define data alignment for different C compilers*/
+#if defined(__GNUC__)
+#define DATA_ALIGN32(x) x __attribute__ ((aligned(32)))
+#else
+#define DATA_ALIGN32(x) __declspec(align(32)) x
+#endif
+
+#ifdef _MSC_VER
+#define  inline __inline
+#endif
+
+#define n1 13
+#define n2 46
+#define n3 38
+#define n4 7
+#define n5 4
+
+#define XOR_256(x,y)      _mm256_xor_si256((x),(y))        /*XOR_256(x,y) = x ^ y, where x and y are two 256-bit word*/
+#define AND_256(x,y)      _mm256_and_si256((x),(y))        /*AND(x,y) = x & y, where x and y are two 256-bit word*/
+
+#define ROTL_256(x,n)      XOR_256( _mm256_slli_epi64((x), (n)), _mm256_srli_epi64((x),(64-n)) )   /*Rotate 4 64-bit unsigned integers in x to the left by n-bit positions*/
+
+#define ROTL_256_64(x)     _mm256_permute4x64_epi64((x), _MM_SHUFFLE(2,1,0,3))  /*Rotate x by 64-bit  positions to the left*/
+#define ROTL_256_128(x)    _mm256_permute4x64_epi64((x), _MM_SHUFFLE(1,0,3,2))  /*Rotate x by 128-bit positions to the left*/
+#define ROTL_256_192(x)    _mm256_permute4x64_epi64((x), _MM_SHUFFLE(0,3,2,1))  /*Rotate x by 192-bit positions to the left*/
+
+void morus_1280_print_state(state state) {
+	// printf("--------------------------------------------\n");
+	// printf("   S0   |   S1   |   S3   |   S3   |   S4   \n");
+	// printf("%08x %08x %08x %08x %08x\n", state[0].f64[3], state[1].f64[3], state[2].f64[3], state[3].f64[3], state[4].f64[3]);
+	// printf("%08x %08x %08x %08x %08x\n", state[0].f64[2], state[1].f64[2], state[2].f64[2], state[3].f64[2], state[4].f64[2]);
[…]

### 2018-05-08 src-pyth-sse/Ciphers/full_morus_640.c

@@ -2,8 +2,6 @@
 
 #ifdef _MSC_VER
 #define inline __inline
-// #define _mm_store_si128 _mm_storeu_si128
-// #define _mm_load_si128 _mm_loadu_si128
 #endif
 
 #define n1 5
@@ -12,15 +10,15 @@
 #define n4 22
 #define n5 13
 
+
+#define XOR_128(x,y)      _mm_xor_si128((x),(y))     /*XOR_128(x,y) = x ^ y, where x and y are two 128-bit word*/
+#define AND_128(x,y)      _mm_and_si128((x),(y))     /*AND_128(x,y) = x & y, where x and y are two 128-bit word*/
+
 #define ROTL_128(x,n)     XOR_128(_mm_slli_epi32((x), (n)),  _mm_srli_epi32((x),(32-n)))  /*Rotate 4 32-bit unsigned integers in x to the left by n-bit positions*/
-#define ROTL8(x)       		_mm_shuffle_epi8((x), _mm_set_epi8(14,13,12,11,10,9,8,7,6,5,4,3,2,1,0,15) )  /*Rotate 4 32-bit unsigned integers in x to the left by 8-bit positions*/
 
-#define ROTL32(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(2,1,0,3))   /*Rotate x by 32-bit positions to the left*/
-#define ROTL64(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(1,0,3,2))   /*Rotate x by 64-bit positions to the left*/
-#define ROTL96(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(0,3,2,1))   /*Rotate x by 96-bit positions to the left*/
-
-// #define STORE_128(x,p)     _mm_store_si128((__m128i *)(p), (x))         /*store the 128-bit word x into memeory address p, where p is the multile of 16 bytes*/
-// #define LOAD_128(p)        _mm_load_si128((__m128i *)(p))               /*load 16 bytes from the memory address p, return a 128-bit word, where p is the multile of 16 bytes*/
[…]

### 2018-05-08 src-pyth-sse/Utilities/aes-rng.c

@@ -81,6 +81,37 @@ __m128i aesrand_m128i_r(struct RNG_state *rng) {
 		return rng->t;
 }
 
+#ifdef __AVX2__
+
+__m256i aesrand_m256i_r(struct RNG_state *rng) {
+		__m128i t;
+		// Increment counter
+		rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
+
+		// 6 AES rounds
+		t = rng->count;
+		t = _mm_aesenc_si128(t, rng->k);
+		t = _mm_aesenc_si128(t, rng->k);
+		t = _mm_aesenc_si128(t, rng->k);
+		t = _mm_aesenc_si128(t, rng->k);
+		t = _mm_aesenc_si128(t, rng->k);
+		t = _mm_aesenc_si128(t, rng->k);
+
+		// Increment counter
+		rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
+		// 6 AES rounds
+		rng->t = rng->count;
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+
+		return _mm256_set_i128(t,rng->t);
+}
+
+#endif
 
 uint8_t aesrand_int8_r(struct RNG_state *rng) {
 	switch (rng->next++) {

### 2018-05-08 src-pyth-sse/Utilities/aes-rng.h

@@ -18,3 +18,9 @@ uint64_t aesrand_int64_r(struct RNG_state *rng);
 uint8_t aesrand_int8_r(struct RNG_state *rng);
 
 __m128i aesrand_m128i_r(struct RNG_state *rng);
+
+#ifdef __AVX2__
+
+__m256i aesrand_m256i_r(struct RNG_state *rng);
+
+#endif

### 2018-05-08 src-pyth-sse/Utilities/sse.h

@@ -1,4 +1,0 @@
-#include <immintrin.h>
-
-#define XOR_128(x,y)      _mm_xor_si128((x),(y))     /*XOR_128(x,y) = x ^ y, where x and y are two 128-bit word*/
-#define AND_128(x,y)      _mm_and_si128((x),(y))     /*AND_128(x,y) = x & y, where x and y are two 128-bit word*/

### 2018-05-08 src-pyth-sse/Utilities/std.h

@@ -12,4 +12,5 @@
 // for AES
 #include <smmintrin.h>
 #include <wmmintrin.h>
+#include <immintrin.h>
 #include <assert.h>

### 2018-05-08 src-pyth-sse/Utilities/types.h

@@ -6,8 +6,9 @@ union Register {
 	uint32_t m32;
 	uint64_t m64;
 	__m128i f32;
-	// uint32_t f32[4];
-	// uint64_t f64[4];
+#ifdef __AVX2__
+	__m256i f64;
+#endif
 };
 
 struct RNG_state {

### 2018-05-08 src-pyth-sse/Utilities/utilities.h

@@ -6,4 +6,3 @@
 #include "maths.h"
 #include "printing.h"
 #include "args.h"
-#include "sse.h"

### 2018-05-08 src-pyth-sse/gen_masks.py

@@ -46,21 +46,40 @@ def gen_function(kind,width,num,states,ciphers, weight):
         if kind == 1:
             z += '\tres ^= saved_ciphers' + K + '.' + union + ' & ' + hex(gen_mask(C)) + ';\n'
         else:
-            z += '\tt = (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 0);\n'
-            z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 1);\n'
-            z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 2);\n'
-            z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 3);\n'
-            z += '\tres ^= t &' + hex(gen_mask(C)) + ';\n'
+            if width == 32:
+                z += '\tt = (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 0);\n'
+                z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 1);\n'
+                z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 2);\n'
+                z += '\tt ^= (uint32_t) _mm_extract_epi32(saved_ciphers' + K + '.' + union + ', 3);\n'
+                z += '\tres ^= t &' + hex(gen_mask(C)) + ';\n'
+            else:
+                z += '#ifdef __AVX2__\n'
+                z += '\tt = (uint64_t) _mm256_extract_epi64(saved_ciphers' + K + '.' + union + ', 0);\n'
+                z += '\tt ^= (uint64_t) _mm256_extract_epi64(saved_ciphers' + K + '.' + union + ', 1);\n'
+                z += '\tt ^= (uint64_t) _mm256_extract_epi64(saved_ciphers' + K + '.' + union + ', 2);\n'
+                z += '\tt ^= (uint64_t) _mm256_extract_epi64(saved_ciphers' + K + '.' + union + ', 3);\n'
+                z += '\tres ^= t &' + hex(gen_mask(C)) + ';\n'
+                z += '#endif\n'
+
 
     for K, C in states.items():
         if kind == 1:
[…]

### 2018-05-08 src-pyth-sse/masks.py

@@ -118,20 +118,20 @@ weight_mini_1280 = [
 16,             # full
 ]
 
-# weight_1280 = [
-# 1,              # alpha
-# 1,              # beta
-# 1,              # gamma
-# 1,              # delta
-# 1,              # epsil
-# 7,              # appr1
-# 9,              # appr2
-# 16,             # full
-# ]
+weight_1280 = [
+4,              # alpha
+4,              # beta
+4,              # gamma
+4,              # delta
+4,              # epsil
+28,              # appr1
+36,              # appr2
+76,             # full
+]
 
 masks_list = [
 {'kind': 1, 'width': 32, 'states': states_640, 'ciphers': ciphers_640, 'weight': weight_mini_640},
 {'kind': 1, 'width': 64, 'states': states_1280, 'ciphers': ciphers_1280, 'weight': weight_mini_1280},
 {'kind': 2, 'width': 32, 'states': states_640, 'ciphers': ciphers_640, 'weight': weight_640},
-# {'kind': 2, 'width': 64, 'states': states_1280, 'ciphers': ciphers_1280, 'weight': weight_1280},
+{'kind': 2, 'width': 64, 'states': states_1280, 'ciphers': ciphers_1280, 'weight': weight_1280},
 ]

### 2018-05-08 src-pyth-sse/Ciphers/morus.h
add build target for full Morus 1280

@@ -1,6 +1,7 @@
 #pragma once
 
 #include "full_morus_640.h"
+#include "full_morus_1280.h"
 #include "mini_morus_640.h"
 #include "mini_morus_1280.h"
 #include "../Utilities/utilities.h"

### 2018-05-08 src-pyth-sse/Utilities/aes-rng.c
update AES to AVX2

@@ -84,18 +84,21 @@ __m128i aesrand_m128i_r(struct RNG_state *rng) {
 #ifdef __AVX2__
 
 __m256i aesrand_m256i_r(struct RNG_state *rng) {
-		__m128i t;
+
+		__m256i r;
 		// Increment counter
 		rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
 
 		// 6 AES rounds
-		t = rng->count;
-		t = _mm_aesenc_si128(t, rng->k);
-		t = _mm_aesenc_si128(t, rng->k);
-		t = _mm_aesenc_si128(t, rng->k);
-		t = _mm_aesenc_si128(t, rng->k);
-		t = _mm_aesenc_si128(t, rng->k);
-		t = _mm_aesenc_si128(t, rng->k);
+		rng->t = rng->count;
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+		rng->t = _mm_aesenc_si128(rng->t, rng->k);
+
+		r = _mm256_castsi128_si256(rng->t);
 
 		// Increment counter
 		rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
@@ -108,7 +111,7 @@ __m256i aesrand_m256i_r(struct RNG_state *rng) {
 		rng->t = _mm_aesenc_si128(rng->t, rng->k);
 		rng->t = _mm_aesenc_si128(rng->t, rng->k);
 
-		return _mm256_set_i128(t,rng->t);
+		return _mm256_inserti128_si256(r, rng->t, 1);
 }
 
 #endif

### 2018-05-08 src-pyth-sse/Ciphers/full_morus_640.c
Implement Morus with SSE and AVX2
Implement Morus640 with SSE instructions
Implement Morus1280 with AVX2 instructions
Allows to test trails on full versions.

@@ -1,0 +1,106 @@
+#include "full_morus_640.h"
+
+#ifdef _MSC_VER
+#define inline __inline
+#endif
+
+#define n1 5
+#define n2 31
+#define n3 7
+#define n4 22
+#define n5 13
+
+
+#define XOR_128(x,y)      _mm_xor_si128((x),(y))     /*XOR_128(x,y) = x ^ y, where x and y are two 128-bit word*/
+#define AND_128(x,y)      _mm_and_si128((x),(y))     /*AND_128(x,y) = x & y, where x and y are two 128-bit word*/
+
+#define ROTL_128(x,n)     XOR_128(_mm_slli_epi32((x), (n)),  _mm_srli_epi32((x),(32-n)))  /*Rotate 4 32-bit unsigned integers in x to the left by n-bit positions*/
+
+#define ROTL_128_32(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(2,1,0,3))   /*Rotate x by 32-bit positions to the left*/
+#define ROTL_128_64(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(1,0,3,2))   /*Rotate x by 64-bit positions to the left*/
+#define ROTL_128_96(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(0,3,2,1))   /*Rotate x by 96-bit positions to the left*/
+
+void morus_640_print_state(state state) {
+	// printf("--------------------------------------------\n");
+	// printf("   S0   |   S1   |   S3   |   S3   |   S4   \n");
+	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[3], state[1].f32[3], state[2].f32[3], state[3].f32[3], state[4].f32[3]);
+	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[2], state[1].f32[2], state[2].f32[2], state[3].f32[2], state[4].f32[2]);
+	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[1], state[1].f32[1], state[2].f32[1], state[3].f32[1], state[4].f32[1]);
+	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[0], state[1].f32[0], state[2].f32[0], state[3].f32[0], state[4].f32[0]);
+}
+
+void morus_640_copy_state(state to, state from) {
+	to[0].f32 = from[0].f32;
[…]

### 2018-05-08 src-pyth-sse/Utilities/aes-rng.c

@@ -1,0 +1,189 @@
+// Use AES-based RNG
+#include "aes-rng.h"
+
+#ifdef __AES__
+
+struct RNG_state *init_aesrand_r(uint32_t seed1, uint32_t seed2) {
+	// Internal state: count + t
+	struct RNG_state *rng = (struct RNG_state *) malloc(sizeof(struct RNG_state));
+	assert(rng);
+	rng->count = _mm_setr_epi32(seed1, seed2, 0, 0);
+	rng->k = _mm_setr_epi32(0, 0, seed1, seed2);
+	rng->next = 0;
+	return rng;
+}
+
+uint32_t aesrand_int32_r(struct RNG_state *rng) {
+	switch (rng->next++) {
+		case 0:
+			// Increment counter
+			rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
+
+			// 6 AES rounds
+			rng->t = rng->count;
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+
+			return (uint32_t) _mm_extract_epi32(rng->t, 0);
+		case 1:
+			return (uint32_t) _mm_extract_epi32(rng->t, 1);
+		case 2:
+			return (uint32_t) _mm_extract_epi32(rng->t, 2);
+		case 3:
+			rng->next = 0;
+			return (uint32_t) _mm_extract_epi32(rng->t, 3);
+		default:
+			assert(0);
+	}
+}
+
+uint64_t aesrand_int64_r(struct RNG_state *rng) {
+	switch (rng->next++) {
+		case 0:
+			// Increment counter
+			rng->count = _mm_add_epi64(rng->count, _mm_setr_epi32(0, 0, 1, 0));
+
+			// 6 AES rounds
+			rng->t = rng->count;
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
+			rng->t = _mm_aesenc_si128(rng->t, rng->k);
[…]

### 2018-05-09 morusAC.tex
split text into sections

@@
 \input{morusAC_00_Abstract}
-
 \input{morusAC_01_Introduction}
-
 \input{morusAC_02_Preliminaries}
-
-\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
-\label{sec:introminimorus}
-
-To simplify the description of the attack, we assume all plaintext blocks are zero.
-However the same results would hold as long as the plaintext is fixed.
-Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
-
-\subsection{Symmetric Linear Combinations}
-
-We begin with a few observations about the \StateUpdate{} function. Besides XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
-\begin{enumerate}
-\item \emph{word-wise} rotations perform a circular shift on each word within a register;
-\item \emph{register-wise} rotations perform a circular shift on a whole register.
-\end{enumerate}
-The second type of rotation always shifts registers by a multiple of the word size $w$.
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
-
-Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
-\begin{definition}
-Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
-\[
-S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
-\]
-is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
-\[
-\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
-\]
-\end{definition}
-\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
-\begin{equation}
-S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
-\label{eq:symmetric}
-\end{equation}
-
-This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
-The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
-
-This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
-As a result, the effect of register-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
-For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
-
-\subsection{\cipher{MiniMORUS}}
-
-In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
-We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
-Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
-
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
-
-\begin{figure}[h]
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate
-  \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function.}
-  \label{fig:minimorus}
-\end{figure}
-
-\section{Linear Trail for \cipher{MiniMORUS}}
-
-\subsection{Overview of the Trail}
-
-To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
-\begin{itemize}
-    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
-    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
-    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
-    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
-    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
-\end{itemize}
-
-The trail fragments are depicted on \Cref{fig:minimorusfragments}.
-In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
-\[
-x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
-\]
-As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, alpha]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
-      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
-      ;
-    \draw[trail, alpha, dashed]
-      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-    \small
-    \begin{align*}
-      \intertext{\cipher{MiniMORUS-640}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\cipher{MiniMORUS-1280}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
-    \end{align*}
-    \caption*{\cipher{MiniMORUS} instances}
-  \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear trail fragments.}
-  \label{fig:minimorusfragments}
-\end{figure}
-
-The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
-In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
-Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
-In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
-
-However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
-The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
-Then $\alpha^t_i$ is used one final time to close the trail.
-One may naturally wonder if some components of this trail are in conflict.
-In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
-
-\subsection{Trail Equation}
-\label{sec:minitraileq}
-
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
-\begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
-\end{alignat*}
-
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
-Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so.
-
-In order to build the equation for the full trail, we start with $E^2_0$:
-\[
-S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
-\]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
-This yields:
-\begin{align*}
-&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
-=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
-\end{align*}
-We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
-This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-
-The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
-
-\subsection{Bias of the Trail}
-\label{sec:minibias}
-
-In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS1280} (cf.
-\Cref{sec:traileq}).
-
-That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
-In both cases it is approximated by zero.
-When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
-
-\subsection{Experimental Verification}
-
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
-Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
-
-\begin{table}[h!]
-  \caption{Experimental verification of trail biases.}
-  \label{tab:miniapproximations}
-  \centerline{
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
-    \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
-\section{Trail for Full \cipher{MORUS}}
-
-\subsection{Symmetrizing the Trail}
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
-That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
-$C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
-\[
-S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
-\]
-
-Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
-Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
-\begin{align*}
-&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
-\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
-\end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
-\begin{align*}
-&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
-\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
-\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
-\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
-\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
-\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
-\end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
-
-\subsection{Bias of the Full Trail}
-
-The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
-Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
-
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
-However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
-
-To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
-
-In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
-Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
-\Cref{fig:trailcollision}).
-While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
-\begin{align}
-&S^t_{2,i} \cdot S^t_{3,i} \oplus S^t_{3,i} \cdot S^t_{2,i+w}
-\oplus S^t_{2,i+w} \cdot S^t_{3,i+w} \oplus S^t_{3,i+w} \cdot S^t_{2,i+2w}\notag\\
-\oplus\; &S^t_{2,i+2w} \cdot S^t_{3,i+2w} \oplus S^t_{3,i+2w} \cdot S^t_{2,i+3w}
-\oplus S^t_{2,i+3w} \cdot S^t_{3,i+3w} \oplus S^t_{3,i+3w} \cdot S^t_{2,i}.\label{eq:8circle}
-\end{align}
-It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      %(xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      %(xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
-    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
-    \node (eq) at ($(and-1|-and1)+(-1,-2.4pt)$) {=};
-    \draw[beta,{Circle[sep=-1.8pt]}-]  ($(and-1)+(0,-2.4pt)$) -| (eq);
-    \draw[gamma,{Circle[sep=-1.8pt]}-] ($(and1)+(0,-2.4pt)$) -- (eq);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
-    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
-  \label{fig:trailcollision}
-\end{figure}
-
-In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
-
-In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
-Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
-As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
-This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
-\Cref{sec:traileq}).
-
-%TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
-%(maybe not relevant due to overall complexity\dots)
-
-\begin{figure}
-  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
-  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
-  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
-  \centering
-  \resizebox{!}{0.95\textheight}{
-  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
-    \begin{scope}
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{27}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
-                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
-                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
-                   (\M{2}{1}) node[right] {$\alpha_7$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
-                   (\M{2}{2}) node[right] {$\gamma_{13}$};
-
-      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
-                   (\M{3}{0}) node[right] {$\beta_{12}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
-                   (\M{1}{0}) node[right] {$\alpha_{2}$};
-
-      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
-                   (\M{2}{0}) node[right] {$\beta_7$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
-      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
-                   (\M{2}{2}) node[right] {$\gamma_7$};
-      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
-                   (\M{2}{4}) node[right] {$\delta_7$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
-                   (\M{3}{1}) node[right] {$\alpha_{14}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
-                   (\M{3}{2}) node[right] {$\gamma_{20}$};
-
-      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
-                   (\M{4}{0}) node[right] {$\beta_{19}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
-    \end{scope}
-
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
-
-    \begin{scope}[yshift=-7.0cm]
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{51}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
-                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
-                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
-                   (\M{2}{1}) node[right] {$\alpha_{37}$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
-                   (\M{2}{2}) node[right] {$\gamma_{4}$};
-
-      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
-                   (\M{3}{0}) node[right] {$\beta_{50}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
-                   (\M{1}{0}) node[right] {$\alpha_{25}$};
-
-      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
-                   (\M{2}{0}) node[right] {$\beta_{38}$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
-      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
-                   (\M{2}{2}) node[right] {$\gamma_{38}$};
-      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
-                   (\M{2}{4}) node[right] {$\delta_{38}$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
-                   (\M{3}{1}) node[right] {$\alpha_{11}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
-                   (\M{3}{2}) node[right] {$\gamma_{42}$};
-
-      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
-                   (\M{4}{0}) node[right] {$\beta_{24}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-    \end{scope}
-  \end{tikzpicture}}
-
-  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
-  \label{fig:twoapproximations}
-\end{figure}
-
-\section{Discussion}
-
-%This is just a sketch for now...
-
-We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
-As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
-This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
-
-The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
-
-With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
-The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
-This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
-
-However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
-Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
-
+\input{morusAC_03_MiniMorus}
+\input{morusAC_04_Mini_trails}
+\input{morusAC_05_Full_trails}
+\input{morusAC_06_Discussion}
 \input{morusAC_07_IniFin}
@@
 
-\section{Trail Equations}
-\label{sec:traileq}
-
-In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
-By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
-
-\subsection{Trail Equation for \cipher{MiniMORUS640}}
-
-This trail equation was already given in \Cref{sec:minitraileq}.
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for \cipher{MiniMORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for full \cipher{MORUS640}}
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
-&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
-&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
-&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
-&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
-&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
-%
-&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
-&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
-&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
-&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 73.
-
-\subsection{Trail Equation for full \cipher{MORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
-&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
-&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
-&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
-&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
-&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
-&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
-%
-&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
-&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
-&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
-&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
-&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 76.
+\input{morusAC_App_IniFin}
 

### 2018-05-09 morusAC_03_MiniMorus.tex
split text into sections

--- 2018-05-02 morus.tex
@@
-\newif\ifanonymous
-%\anonymoustrue%uncomment for submission version
-
-\documentclass{llncs}
-\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-\usepackage{common}
-
-\ifanonymous
-\author{}
-\institute{}
-\else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
-\institute{}
-%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
-\fi
-
-\begin{document}
-
-\mainmatter
-
-\title{Note on Linear Biases in MORUS}
-
-\maketitle
-
-\begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
-More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
-Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
-\end{abstract}
-
-\section{Introduction}
-
-Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
-As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
-Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-
-\section{Notation}
-
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
-weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
-weights).
-
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
-$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
-$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
-$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
-\end{tabular}
-}
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
-
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
-Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
-In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
-
-Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-\end{tabular}
-}
-
-The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
-
 \section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
@@
 \end{figure}
-
-\section{Linear Trail for \cipher{MiniMORUS}}
-
-\subsection{Overview of the Trail}
-
-To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
-\begin{itemize}
-    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
-    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
-    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
-    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
-    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
-\end{itemize}
-
-The trail fragments are depicted on \Cref{fig:minimorusfragments}.
-In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
-\[
-x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
-\]
-As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, alpha]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
-      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
-      ;
-    \draw[trail, alpha, dashed]
-      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-    \small
-    \begin{align*}
-      \intertext{\cipher{MiniMORUS-640}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\cipher{MiniMORUS-1280}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
-    \end{align*}
-    \caption*{\cipher{MiniMORUS} instances}
-  \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear trail fragments.}
-  \label{fig:minimorusfragments}
-\end{figure}
-
-The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
-In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
-Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
-In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
-
-However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
-The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
-Then $\alpha^t_i$ is used one final time to close the trail.
-One may naturally wonder if some components of this trail are in conflict.
-In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
-
-\subsection{Trail Equation}
-\label{sec:minitraileq}
-
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
-\begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
-\end{alignat*}
-
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
-Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so.
-
-In order to build the equation for the full trail, we start with $E^2_0$:
-\[
-S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
-\]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
-This yields:
-\begin{align*}
-&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
-=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
-\end{align*}
-We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
-This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-
-The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
-
-\subsection{Bias of the Trail}
-\label{sec:minibias}
-
-In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS1280} (cf.
-\Cref{sec:traileq}).
-
-That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
-In both cases it is approximated by zero.
-When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
-
-\subsection{Experimental Verification}
-
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
-Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
-
-\begin{table}[h!]
-  \caption{Experimental verification of trail biases.}
-  \label{tab:miniapproximations}
-  \centerline{
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
-    \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
-\section{Trail for Full \cipher{MORUS}}
-
-\subsection{Symmetrizing the Trail}
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
-That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
-$C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
-\[
-S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
-\]
-
-Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
-Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
-\begin{align*}
-&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
-\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
-\end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
-\begin{align*}
-&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
-\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
-\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
-\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
-\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
-\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
-\end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
-
-\subsection{Bias of the Full Trail}
-
-The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
-Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
-
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
-However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
-
-To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
-
-In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
-Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
-\Cref{fig:trailcollision}).
-While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
-\begin{align}
-&S^t_{2,i} \cdot S^t_{3,i} \oplus S^t_{3,i} \cdot S^t_{2,i+w}
-\oplus S^t_{2,i+w} \cdot S^t_{3,i+w} \oplus S^t_{3,i+w} \cdot S^t_{2,i+2w}\notag\\
-\oplus\; &S^t_{2,i+2w} \cdot S^t_{3,i+2w} \oplus S^t_{3,i+2w} \cdot S^t_{2,i+3w}
-\oplus S^t_{2,i+3w} \cdot S^t_{3,i+3w} \oplus S^t_{3,i+3w} \cdot S^t_{2,i}.\label{eq:8circle}
-\end{align}
-It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      %(xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      %(xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
-    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
-    \node (eq) at ($(and-1|-and1)+(-1,-2.4pt)$) {=};
-    \draw[beta,{Circle[sep=-1.8pt]}-]  ($(and-1)+(0,-2.4pt)$) -| (eq);
-    \draw[gamma,{Circle[sep=-1.8pt]}-] ($(and1)+(0,-2.4pt)$) -- (eq);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
-    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
-  \label{fig:trailcollision}
-\end{figure}
-
-In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
-
-In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
-Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
-As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
-This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
-\Cref{sec:traileq}).
-
-%TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
-%(maybe not relevant due to overall complexity\dots)
-
-\begin{figure}
-  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
-  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
-  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
-  \centering
-  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
-    \begin{scope}
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{27}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
-                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
-                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
-                   (\M{2}{1}) node[right] {$\alpha_7$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
-                   (\M{2}{2}) node[right] {$\gamma_{13}$};
-
-      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
-                   (\M{3}{0}) node[right] {$\beta_{12}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
-                   (\M{1}{0}) node[right] {$\alpha_{2}$};
-
-      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
-                   (\M{2}{0}) node[right] {$\beta_7$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
-      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
-                   (\M{2}{2}) node[right] {$\gamma_7$};
-      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
-                   (\M{2}{4}) node[right] {$\delta_7$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
-                   (\M{3}{1}) node[right] {$\alpha_{14}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
-                   (\M{3}{2}) node[right] {$\gamma_{20}$};
-
-      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
-                   (\M{4}{0}) node[right] {$\beta_{19}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
-    \end{scope}
-
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
-
-    \begin{scope}[yshift=-7.0cm]
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{51}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
-                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
-                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
-                   (\M{2}{1}) node[right] {$\alpha_{37}$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
-                   (\M{2}{2}) node[right] {$\gamma_{4}$};
-
-      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
-                   (\M{3}{0}) node[right] {$\beta_{50}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
-                   (\M{1}{0}) node[right] {$\alpha_{25}$};
-
-      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
-                   (\M{2}{0}) node[right] {$\beta_{38}$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
-      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
-                   (\M{2}{2}) node[right] {$\gamma_{38}$};
-      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
-                   (\M{2}{4}) node[right] {$\delta_{38}$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
-                   (\M{3}{1}) node[right] {$\alpha_{11}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
-                   (\M{3}{2}) node[right] {$\gamma_{42}$};
-
-      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
-                   (\M{4}{0}) node[right] {$\beta_{24}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-    \end{scope}
-  \end{tikzpicture}
-  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
-  \label{fig:twoapproximations}
-\end{figure}
-
-\section{Discussion}
-
-%This is just a sketch for now...
-
-We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
-As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
-This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
-
-The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
-
-With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
-The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
-This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
-
-However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
-Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
-
-\appendix
-
-\section{Trail Equations}
-\label{sec:traileq}
-
-In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
-By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
-
-\subsection{Trail Equation for \cipher{MiniMORUS640}}
-
-This trail equation was already given in \Cref{sec:minitraileq}.
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for \cipher{MiniMORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for full \cipher{MORUS640}}
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
-&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
-&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
-&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
-&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
-&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
-%
-&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
-&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
-&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
-&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 73.
-
-\subsection{Trail Equation for full \cipher{MORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
-&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
-&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
-&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
-&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
-&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
-&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
-%
-&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
-&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
-&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
-&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
-&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 76.
-
-\end{document}

### 2018-05-09 morusAC_04_Mini_trails.tex
split text into sections

--- 2018-05-02 morus.tex
@@
-\newif\ifanonymous
-%\anonymoustrue%uncomment for submission version
-
-\documentclass{llncs}
-\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-\usepackage{common}
-
-\ifanonymous
-\author{}
-\institute{}
-\else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
-\institute{}
-%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
-\fi
-
-\begin{document}
-
-\mainmatter
-
-\title{Note on Linear Biases in MORUS}
-
-\maketitle
-
-\begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
-More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
-Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
-\end{abstract}
-
-\section{Introduction}
-
-Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
-As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
-Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-
-\section{Notation}
-
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
-weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
-weights).
-
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
-$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
-$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
-$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
-\end{tabular}
-}
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
-
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
-Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
-In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
-
-Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-\end{tabular}
-}
-
-The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
-
-\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
-\label{sec:introminimorus}
-
-To simplify the description of the attack, we assume all plaintext blocks are zero.
-However the same results would hold as long as the plaintext is fixed.
-Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
-
-\subsection{Symmetric Linear Combinations}
-
-We begin with a few observations about the \StateUpdate{} function. Beside XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
-\begin{enumerate}
-\item \emph{word-wise} rotations perform a circular shift on each word within a register;
-\item \emph{register-wise} rotations perform a circular shift on a whole register.
-\end{enumerate}
-The second type of rotation always shifts registers by a multiple of the word size $w$.
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
-
-Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
-\begin{definition}
-Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
-\[
-S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
-\]
-is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
-\[
-\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
-\]
-\end{definition}
-\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
-\begin{equation}
-S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
-\label{eq:symmetric}
-\end{equation}
-
-This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
-The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
-
-This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
-As a result, the effect of register-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
-For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
-
-\subsection{\cipher{MiniMORUS}}
-
-In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
-We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
-Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
-
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
-
-\begin{figure}[h]
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate
-  \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function.}
-  \label{fig:minimorus}
-\end{figure}
-
 \section{Linear Trail for \cipher{MiniMORUS}}
@@
 \end{table}
-
-\section{Trail for Full \cipher{MORUS}}
-
-\subsection{Symmetrizing the Trail}
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
-That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
-$C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
-\[
-S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
-\]
-
-Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
-Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
-\begin{align*}
-&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
-\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
-\end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
-\begin{align*}
-&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
-\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
-\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
-\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
-\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
-\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
-\end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
-
-\subsection{Bias of the Full Trail}
-
-The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
-Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
-
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
-However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
-
-To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
-
-In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
-Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
-\Cref{fig:trailcollision}).
-While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
-\begin{align}
-&S^t_{2,i} \cdot S^t_{3,i} \oplus S^t_{3,i} \cdot S^t_{2,i+w}
-\oplus S^t_{2,i+w} \cdot S^t_{3,i+w} \oplus S^t_{3,i+w} \cdot S^t_{2,i+2w}\notag\\
-\oplus\; &S^t_{2,i+2w} \cdot S^t_{3,i+2w} \oplus S^t_{3,i+2w} \cdot S^t_{2,i+3w}
-\oplus S^t_{2,i+3w} \cdot S^t_{3,i+3w} \oplus S^t_{3,i+3w} \cdot S^t_{2,i}.\label{eq:8circle}
-\end{align}
-It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      %(xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      %(xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
-    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
-    \node (eq) at ($(and-1|-and1)+(-1,-2.4pt)$) {=};
-    \draw[beta,{Circle[sep=-1.8pt]}-]  ($(and-1)+(0,-2.4pt)$) -| (eq);
-    \draw[gamma,{Circle[sep=-1.8pt]}-] ($(and1)+(0,-2.4pt)$) -- (eq);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
-    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
-  \label{fig:trailcollision}
-\end{figure}
-
-In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
-
-In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
-Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
-As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
-This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
-\Cref{sec:traileq}).
-
-%TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
-%(maybe not relevant due to overall complexity\dots)
-
-\begin{figure}
-  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
-  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
-  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
-  \centering
-  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
-    \begin{scope}
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{27}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
-                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
-                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
-                   (\M{2}{1}) node[right] {$\alpha_7$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
-                   (\M{2}{2}) node[right] {$\gamma_{13}$};
-
-      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
-                   (\M{3}{0}) node[right] {$\beta_{12}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
-                   (\M{1}{0}) node[right] {$\alpha_{2}$};
-
-      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
-                   (\M{2}{0}) node[right] {$\beta_7$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
-      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
-                   (\M{2}{2}) node[right] {$\gamma_7$};
-      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
-                   (\M{2}{4}) node[right] {$\delta_7$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
-                   (\M{3}{1}) node[right] {$\alpha_{14}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
-                   (\M{3}{2}) node[right] {$\gamma_{20}$};
-
-      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
-                   (\M{4}{0}) node[right] {$\beta_{19}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
-    \end{scope}
-
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
-
-    \begin{scope}[yshift=-7.0cm]
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{51}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
-                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
-                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
-                   (\M{2}{1}) node[right] {$\alpha_{37}$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
-                   (\M{2}{2}) node[right] {$\gamma_{4}$};
-
-      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
-                   (\M{3}{0}) node[right] {$\beta_{50}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
-                   (\M{1}{0}) node[right] {$\alpha_{25}$};
-
-      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
-                   (\M{2}{0}) node[right] {$\beta_{38}$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
-      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
-                   (\M{2}{2}) node[right] {$\gamma_{38}$};
-      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
-                   (\M{2}{4}) node[right] {$\delta_{38}$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
-                   (\M{3}{1}) node[right] {$\alpha_{11}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
-                   (\M{3}{2}) node[right] {$\gamma_{42}$};
-
-      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
-                   (\M{4}{0}) node[right] {$\beta_{24}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-    \end{scope}
-  \end{tikzpicture}
-  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
-  \label{fig:twoapproximations}
-\end{figure}
-
-\section{Discussion}
-
-%This is just a sketch for now...
-
-We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
-As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
-This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
-
-The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
-
-With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
-The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
-This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
-
-However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
-Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
-
-\appendix
-
-\section{Trail Equations}
-\label{sec:traileq}
-
-In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
-By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
-
-\subsection{Trail Equation for \cipher{MiniMORUS640}}
-
-This trail equation was already given in \Cref{sec:minitraileq}.
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for \cipher{MiniMORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for full \cipher{MORUS640}}
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
-&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
-&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
-&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
-&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
-&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
-%
-&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
-&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
-&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
-&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 73.
-
-\subsection{Trail Equation for full \cipher{MORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
-&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
-&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
-&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
-&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
-&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
-&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
-%
-&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
-&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
-&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
-&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
-&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 76.
-
-\end{document}

### 2018-05-09 morusAC_05_Full_trails.tex
split text into sections

--- 2018-05-02 morus.tex
@@
-\newif\ifanonymous
-%\anonymoustrue%uncomment for submission version
-
-\documentclass{llncs}
-\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-\usepackage{common}
-
-\ifanonymous
-\author{}
-\institute{}
-\else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
-\institute{}
-%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
-\fi
-
-\begin{document}
-
-\mainmatter
-
-\title{Note on Linear Biases in MORUS}
-
-\maketitle
-
-\begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
-More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
-Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
-\end{abstract}
-
-\section{Introduction}
-
-Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
-As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
-Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-
-\section{Notation}
-
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
-weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
-weights).
-
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
-$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
-$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
-$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
-\end{tabular}
-}
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
-
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
-Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
-In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
-
-Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-\end{tabular}
-}
-
-The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
-
-\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
-\label{sec:introminimorus}
-
-To simplify the description of the attack, we assume all plaintext blocks are zero.
-However the same results would hold as long as the plaintext is fixed.
-Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
-
-\subsection{Symmetric Linear Combinations}
-
-We begin with a few observations about the \StateUpdate{} function. Beside XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
-\begin{enumerate}
-\item \emph{word-wise} rotations perform a circular shift on each word within a register;
-\item \emph{register-wise} rotations perform a circular shift on a whole register.
-\end{enumerate}
-The second type of rotation always shifts registers by a multiple of the word size $w$.
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
-
-Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
-\begin{definition}
-Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
-\[
-S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
-\]
-is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
-\[
-\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
-\]
-\end{definition}
-\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
-\begin{equation}
-S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
-\label{eq:symmetric}
-\end{equation}
-
-This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
-The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
-
-This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
-As a result, the effect of register-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
-For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
-
-\subsection{\cipher{MiniMORUS}}
-
-In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
-We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
-Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
-
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
-
-\begin{figure}[h]
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate
-  \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function.}
-  \label{fig:minimorus}
-\end{figure}
-
-\section{Linear Trail for \cipher{MiniMORUS}}
-
-\subsection{Overview of the Trail}
-
-To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
-\begin{itemize}
-    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
-    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
-    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
-    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
-    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
-\end{itemize}
-
-The trail fragments are depicted on \Cref{fig:minimorusfragments}.
-In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
-\[
-x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
-\]
-As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, alpha]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
-      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
-      ;
-    \draw[trail, alpha, dashed]
-      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-    \small
-    \begin{align*}
-      \intertext{\cipher{MiniMORUS-640}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\cipher{MiniMORUS-1280}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
-    \end{align*}
-    \caption*{\cipher{MiniMORUS} instances}
-  \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear trail fragments.}
-  \label{fig:minimorusfragments}
-\end{figure}
-
-The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
-In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
-Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
-In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
-
-However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
-The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
-Then $\alpha^t_i$ is used one final time to close the trail.
-One may naturally wonder if some components of this trail are in conflict.
-In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
-
-\subsection{Trail Equation}
-\label{sec:minitraileq}
-
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
-\begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
-\end{alignat*}
-
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
-Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so.
-
-In order to build the equation for the full trail, we start with $E^2_0$:
-\[
-S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
-\]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
-This yields:
-\begin{align*}
-&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
-=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
-\end{align*}
-We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
-This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-
-The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
-
-\subsection{Bias of the Trail}
-\label{sec:minibias}
-
-In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS1280} (cf.
-\Cref{sec:traileq}).
-
-That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
-In both cases it is approximated by zero.
-When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
-
-\subsection{Experimental Verification}
-
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
-Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
-
-\begin{table}[h!]
-  \caption{Experimental verification of trail biases.}
-  \label{tab:miniapproximations}
-  \centerline{
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
-    \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
 \section{Trail for Full \cipher{MORUS}}
@@
 \end{figure}
-
-\section{Discussion}
-
-%This is just a sketch for now...
-
-We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
-As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
-This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
-
-The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
-
-With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
-The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
-This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
-
-However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
-Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
-
-\appendix
-
-\section{Trail Equations}
-\label{sec:traileq}
-
-In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
-By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
-
-\subsection{Trail Equation for \cipher{MiniMORUS640}}
-
-This trail equation was already given in \Cref{sec:minitraileq}.
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for \cipher{MiniMORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for full \cipher{MORUS640}}
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
-&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
-&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
-&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
-&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
-&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
-%
-&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
-&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
-&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
-&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 73.
-
-\subsection{Trail Equation for full \cipher{MORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
-&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
-&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
-&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
-&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
-&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
-&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
-%
-&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
-&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
-&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
-&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
-&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 76.
-
-\end{document}

### 2018-05-09 morusAC_06_Discussion.tex
split text into sections

--- 2018-05-02 morus.tex
@@
-\newif\ifanonymous
-%\anonymoustrue%uncomment for submission version
-
-\documentclass{llncs}
-\usepackage[hmargin=3cm,vmargin=2.8cm]{geometry}
-\usepackage{common}
-
-\ifanonymous
-\author{}
-\institute{}
-\else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
-\institute{}
-%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
-\fi
-
-\begin{document}
-
-\mainmatter
-
-\title{Note on Linear Biases in MORUS}
-
-\maketitle
-
-\begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
-More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
-Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
-\end{abstract}
-
-\section{Introduction}
-
-Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
-As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
-Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-
-\section{Notation}
-
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
-weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
-weights).
-
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$C^t$ &:& the ciphertext block output during the $t$-th encryption step.\\
-$C^t_j$ &:& the $j$-th bit of $C^t$, with $C^t_0$ being the rightmost bit.\\
-$S^t_i$ &:& the $i$-th register at the beginning of $t$-th encryption step.\\
-$S^t_{i,j}$ &:& the $j$-th bit of $S^t_i$, with $S^t_{i,0}$ being the rightmost bit.
-\end{tabular}
-}
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
-
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
-Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.
-In particular the $0$-th encryption step from the perspective of the trail does not have to be the first encryption step after initialization.
-
-Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
-\centerline{
-\def\arraystretch{1.25}
-\begin{tabular}{lcl}
-$x \rotl y$ &:& circular shift of $x$ by $y$ bits to the left.\\
-$x \rotlxy y$ &:& circular shift of each word within $x$ by $y$ bits to the left.\\
-\end{tabular}
-}
-
-The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
-
-\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
-\label{sec:introminimorus}
-
-To simplify the description of the attack, we assume all plaintext blocks are zero.
-However the same results would hold as long as the plaintext is fixed.
-Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
-
-\subsection{Symmetric Linear Combinations}
-
-We begin with a few observations about the \StateUpdate{} function. Beside XOR and AND operations, the \StateUpdate{} function uses two types of bit rotations:
-\begin{enumerate}
-\item \emph{word-wise} rotations perform a circular shift on each word within a register;
-\item \emph{register-wise} rotations perform a circular shift on a whole register.
-\end{enumerate}
-The second type of rotation always shifts registers by a multiple of the word size $w$.
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
-
-Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
-\begin{definition}
-Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
-\[
-S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
-\]
-is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
-\[
-\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
-\]
-\end{definition}
-\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
-\begin{equation}
-S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
-\label{eq:symmetric}
-\end{equation}
-
-This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
-The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
-
-This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
-As a result, the effect of register-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
-For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
-
-\subsection{\cipher{MiniMORUS}}
-
-In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
-We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
-Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
-
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
-
-\begin{figure}[h]
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
-    \printstate
-  \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function.}
-  \label{fig:minimorus}
-\end{figure}
-
-\section{Linear Trail for \cipher{MiniMORUS}}
-
-\subsection{Overview of the Trail}
-
-To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
-\begin{itemize}
-    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
-    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
-    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
-    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
-    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
-\end{itemize}
-
-The trail fragments are depicted on \Cref{fig:minimorusfragments}.
-In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
-\[
-x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
-\]
-As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, alpha]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
-      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
-      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
-      ;
-    \draw[trail, alpha, dashed]
-      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-    \small
-    \begin{align*}
-      \intertext{\cipher{MiniMORUS-640}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\cipher{MiniMORUS-1280}:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
-    \end{align*}
-    \caption*{\cipher{MiniMORUS} instances}
-  \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear trail fragments.}
-  \label{fig:minimorusfragments}
-\end{figure}
-
-The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
-In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
-Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
-In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
-
-However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
-The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
-Then $\alpha^t_i$ is used one final time to close the trail.
-One may naturally wonder if some components of this trail are in conflict.
-In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
-
-\subsection{Trail Equation}
-\label{sec:minitraileq}
-
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
-\begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
-\end{alignat*}
-
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
-Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so.
-
-In order to build the equation for the full trail, we start with $E^2_0$:
-\[
-S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
-\]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
-This yields:
-\begin{align*}
-&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
-=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
-\end{align*}
-We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
-This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-
-The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
-
-\subsection{Bias of the Trail}
-\label{sec:minibias}
-
-In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS1280} (cf.
-\Cref{sec:traileq}).
-
-That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
-In both cases it is approximated by zero.
-When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
-
-\subsection{Experimental Verification}
-
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
-Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
-
-\begin{table}[h!]
-  \caption{Experimental verification of trail biases.}
-  \label{tab:miniapproximations}
-  \centerline{
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
-    \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
-\section{Trail for Full \cipher{MORUS}}
-
-\subsection{Symmetrizing the Trail}
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
-That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
-$C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
-\[
-S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
-\]
-
-Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
-Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
-\begin{align*}
-&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
-\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
-\end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
-\begin{align*}
-&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
-\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
-\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
-\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
-\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
-\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
-\end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
-
-\subsection{Bias of the Full Trail}
-
-The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
-Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
-
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
-However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
-
-To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
-
-In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
-Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
-\Cref{fig:trailcollision}).
-While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
-\begin{align}
-&S^t_{2,i} \cdot S^t_{3,i} \oplus S^t_{3,i} \cdot S^t_{2,i+w}
-\oplus S^t_{2,i+w} \cdot S^t_{3,i+w} \oplus S^t_{3,i+w} \cdot S^t_{2,i+2w}\notag\\
-\oplus\; &S^t_{2,i+2w} \cdot S^t_{3,i+2w} \oplus S^t_{3,i+2w} \cdot S^t_{2,i+3w}
-\oplus S^t_{2,i+3w} \cdot S^t_{3,i+3w} \oplus S^t_{3,i+3w} \cdot S^t_{2,i}.\label{eq:8circle}
-\end{align}
-It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      %(xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      %(xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
-    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
-    \node (eq) at ($(and-1|-and1)+(-1,-2.4pt)$) {=};
-    \draw[beta,{Circle[sep=-1.8pt]}-]  ($(and-1)+(0,-2.4pt)$) -| (eq);
-    \draw[gamma,{Circle[sep=-1.8pt]}-] ($(and1)+(0,-2.4pt)$) -- (eq);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-    \draw[trail, gamma]
-      %(W-21) node[above] {$i$}
-      (txor-1) -- (xnd1)
-      (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
-    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
-  \end{tikzpicture}%}}}
-  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
-  \label{fig:trailcollision}
-\end{figure}
-
-In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
-
-In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
-Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
-As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
-This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
-\Cref{sec:traileq}).
-
-%TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
-%(maybe not relevant due to overall complexity\dots)
-
-\begin{figure}
-  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
-  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
-  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
-  \centering
-  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
-    \begin{scope}
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{27}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
-                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
-                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
-                   (\M{2}{1}) node[right] {$\alpha_7$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
-                   (\M{2}{2}) node[right] {$\gamma_{13}$};
-
-      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
-                   (\M{3}{0}) node[right] {$\beta_{12}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
-                   (\M{1}{0}) node[right] {$\alpha_{2}$};
-
-      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
-                   (\M{2}{0}) node[right] {$\beta_7$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
-      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
-                   (\M{2}{2}) node[right] {$\gamma_7$};
-      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
-                   (\M{2}{4}) node[right] {$\delta_7$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
-                   (\M{3}{1}) node[right] {$\alpha_{14}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
-                   (\M{3}{2}) node[right] {$\gamma_{20}$};
-
-      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
-                   (\M{4}{0}) node[right] {$\beta_{19}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
-    \end{scope}
-
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
-
-    \begin{scope}[yshift=-7.0cm]
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{51}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
-                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
-                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
-                   (\M{2}{1}) node[right] {$\alpha_{37}$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
-                   (\M{2}{2}) node[right] {$\gamma_{4}$};
-
-      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
-                   (\M{3}{0}) node[right] {$\beta_{50}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
-                   (\M{1}{0}) node[right] {$\alpha_{25}$};
-
-      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
-                   (\M{2}{0}) node[right] {$\beta_{38}$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
-      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
-                   (\M{2}{2}) node[right] {$\gamma_{38}$};
-      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
-                   (\M{2}{4}) node[right] {$\delta_{38}$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
-                   (\M{3}{1}) node[right] {$\alpha_{11}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
-                   (\M{3}{2}) node[right] {$\gamma_{42}$};
-
-      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
-                   (\M{4}{0}) node[right] {$\beta_{24}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-    \end{scope}
-  \end{tikzpicture}
-  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
-  \label{fig:twoapproximations}
-\end{figure}
-
 \section{Discussion}
@@
 This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
-
-\appendix
-
-\section{Trail Equations}
-\label{sec:traileq}
-
-In this section, we provide full trail equations for all variants of \cipher{MORUS}.
-In each case, we decompose the right-hand side of the equality (involving state bits) into connected components, and compute the weight of each of these connected component.
-By the piling-up lemma, if we assume that the state bits are uniformly random and independent, then the weight of the full equation is equal to the sum of the weights of the connected components.
-
-\subsection{Trail Equation for \cipher{MiniMORUS640}}
-
-This trail equation was already given in \Cref{sec:minitraileq}.
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for \cipher{MiniMORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 16.
-
-\subsection{Trail Equation for full \cipher{MORUS640}}
-
-\begin{flalign*}
-&&&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}&&\\
-&&\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}&&\\
-&&\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}&&\\
-&&\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}&&\\
-&&\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}&&\\
-&&\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}&&\\
-%
-&&=\; &S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,0} \cdot S^1_{3,96} \oplus S^1_{2,32} \cdot S^1_{3,0} \oplus S^1_{2,96} \cdot S^1_{3,96}&&\\
-&&&\quad \oplus S^1_{2,96} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,64} \oplus S^1_{2,64} \cdot S^1_{3,32} \oplus S^1_{2,32} \cdot S^1_{3,32}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,13} \cdot S^2_{3,13} \oplus S^2_{2,13} \cdot S^2_{3,109} \oplus S^2_{2,45} \cdot S^2_{3,13} \oplus S^2_{2,109} \cdot S^2_{3,109}&&\\
-&&&\quad \oplus S^2_{2,45} \cdot S^2_{3,45} \oplus S^2_{2,109} \cdot S^2_{3,77} \oplus S^2_{2,77} \cdot S^2_{3,45} \oplus S^2_{2,77} \cdot S^2_{3,77}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20} \oplus S^3_{2,20} \cdot S^3_{3,116} \oplus S^3_{2,52} \cdot S^3_{3,20} \oplus S^3_{2,116} \cdot S^3_{3,116}&&\\
-&&&\quad \oplus S^3_{2,52} \cdot S^3_{3,52} \oplus S^3_{2,116} \cdot S^3_{3,84} \oplus S^3_{2,84} \cdot S^3_{3,52} \oplus S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27} \oplus S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,59} \oplus S^0_{1,59} \cdot S^0_{2,59} \oplus S^0_{2,59} \cdot S^0_{3,59} \oplus S^0_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,91} \oplus S^0_{1,91} \cdot S^0_{2,91} \oplus S^0_{2,91} \cdot S^0_{3,91} \oplus S^0_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,123} \cdot S^0_{2,123} \oplus S^0_{2,123} \cdot S^0_{3,123} \oplus S^0_{1,123} \oplus S^0_{3,123}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2} \oplus S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8} \oplus S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26} \oplus S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,34} \oplus S^1_{1,34} \cdot S^1_{2,34} \oplus S^1_{2,34} \cdot S^1_{3,34} \oplus S^1_{3,34}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,40} \oplus S^1_{1,40} \cdot S^1_{2,40} \oplus S^1_{2,40} \cdot S^1_{3,40} \oplus S^1_{3,40}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,58} \oplus S^1_{1,58} \cdot S^1_{2,58} \oplus S^1_{2,58} \cdot S^1_{3,58} \oplus S^1_{3,58}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,66} \oplus S^1_{1,66} \cdot S^1_{2,66} \oplus S^1_{2,66} \cdot S^1_{3,66} \oplus S^1_{3,66}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,72} \oplus S^1_{1,72} \cdot S^1_{2,72} \oplus S^1_{2,72} \cdot S^1_{3,72} \oplus S^1_{3,72}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,90} \oplus S^1_{1,90} \cdot S^1_{2,90} \oplus S^1_{2,90} \cdot S^1_{3,90} \oplus S^1_{3,90}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,98} \oplus S^1_{1,98} \cdot S^1_{2,98} \oplus S^1_{2,98} \cdot S^1_{3,98} \oplus S^1_{3,98}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,104} \oplus S^1_{1,104} \cdot S^1_{2,104} \oplus S^1_{2,104} \cdot S^1_{3,104} \oplus S^1_{3,104}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,122} \oplus S^1_{3,122} \oplus S^1_{1,122} \cdot S^1_{2,122} \oplus S^1_{2,122} \cdot S^1_{3,122}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1} \oplus S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,103} \oplus S^2_{3,103}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15} \oplus S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27} \oplus S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,33} \oplus S^2_{1,33} \cdot S^2_{2,33} \oplus S^2_{2,33} \cdot S^2_{3,33} \oplus S^2_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,39} \oplus S^2_{1,39} \cdot S^2_{2,39} \oplus S^2_{2,39} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,47} \oplus S^2_{1,47} \cdot S^2_{2,47} \oplus S^2_{2,47} \cdot S^2_{3,47} \oplus S^2_{3,47}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,59} \oplus S^2_{1,59} \cdot S^2_{2,59} \oplus S^2_{2,59} \cdot S^2_{3,59} \oplus S^2_{3,59}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,65} \oplus S^2_{1,65} \cdot S^2_{2,65} \oplus S^2_{2,65} \cdot S^2_{3,65} \oplus S^2_{3,65}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \oplus S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{2,71} \cdot S^2_{3,39} \oplus S^2_{3,39}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,79} \oplus S^2_{1,79} \cdot S^2_{2,79} \oplus S^2_{2,79} \cdot S^2_{3,79} \oplus S^2_{3,79}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,91} \oplus S^2_{1,91} \cdot S^2_{2,91} \oplus S^2_{2,91} \cdot S^2_{3,91} \oplus S^2_{3,91}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,97} \oplus S^2_{1,97} \cdot S^2_{2,97} \oplus S^2_{2,97} \cdot S^2_{3,97} \oplus S^2_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,103} \oplus S^2_{1,103} \cdot S^2_{2,103} \oplus S^2_{2,103} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,111} \oplus S^2_{1,111} \cdot S^2_{2,111} \oplus S^2_{2,111} \cdot S^2_{3,111} \oplus S^2_{3,111}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,123} \oplus S^2_{3,123} \oplus S^2_{2,123} \cdot S^2_{3,123} \oplus S^2_{1,123} \cdot S^2_{2,123}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14} \oplus S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,46} \oplus S^3_{1,46} \cdot S^3_{2,46} \oplus S^3_{2,46} \cdot S^3_{3,46} \oplus S^3_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,78} \oplus S^3_{1,78} \cdot S^3_{2,78} \oplus S^3_{2,78} \cdot S^3_{3,78} \oplus S^3_{3,78}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,110} \oplus S^3_{3,110} \oplus S^3_{1,110} \cdot S^3_{2,110} \oplus S^3_{2,110} \cdot S^3_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,32} \cdot S^2_{1,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,96} \cdot S^2_{1,96}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,63} \cdot S^2_{3,63}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,95} \cdot S^2_{3,95}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,127} \cdot S^2_{3,127}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,32} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,32}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,96} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,96}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,39} \cdot S^3_{1,39}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,71} \cdot S^3_{1,71}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,103} \cdot S^3_{1,103}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,38} \cdot S^3_{3,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,44} \cdot S^3_{3,44}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,70} \cdot S^3_{3,70}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,76} \cdot S^3_{3,76}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,102} \cdot S^3_{3,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,108} \cdot S^3_{3,108}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,19} \cdot S^4_{3,19}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,51} \cdot S^4_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,83} \cdot S^4_{3,83}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,115} \cdot S^4_{3,115}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 73.
-
-\subsection{Trail Equation for full \cipher{MORUS1280}}
-
-\begin{flalign*}
-&&&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}&&\\
-&&\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}&&\\
-&&\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}&&\\
-&&\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}&&\\
-&&\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}&&\\
-&&\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}&&\\
-&&\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}&&\\
-%
-&&=\; & S^1_{2,0} \cdot S^1_{3,192} \oplus S^1_{2,0} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,0} \oplus S^1_{2,64} \cdot S^1_{3,64}&&\\
-&&&\quad \oplus S^1_{2,128} \cdot S^1_{3,64} \oplus S^1_{2,128} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,128} \oplus S^1_{2,192} \cdot S^1_{3,192}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,4} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,4} \oplus S^2_{2,68} \cdot S^2_{3,68} \oplus S^2_{2,132} \cdot S^2_{3,68}&&\\
-&&&\quad \oplus S^2_{2,132} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,132} \oplus S^2_{2,196} \cdot S^2_{3,196} \oplus S^2_{2,4} \cdot S^2_{3,196}&&\text{weight 3}\\
-%
-&&\oplus\; & S^2_{2,102} \cdot S^2_{3,38} \oplus S^2_{2,102} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,102} \oplus S^2_{2,166} \cdot S^2_{3,166}&&\\
-&&&\quad \oplus S^2_{2,230} \cdot S^2_{3,166} \oplus S^2_{2,230} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,230} \oplus S^2_{2,38} \cdot S^2_{3,38}&&\text{weight 3}\\
-%
-&&\oplus\; & S^3_{2,42} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,42} \oplus S^3_{2,106} \cdot S^3_{3,106} \oplus S^3_{2,170} \cdot S^3_{3,106}&&\\
-&&&\quad \oplus S^3_{2,170} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,170} \oplus S^3_{2,234} \cdot S^3_{3,234} \oplus S^3_{2,42} \cdot S^3_{3,234}&&\text{weight 3}\\
-%
-&&\oplus\; & S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{1,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,115} \cdot S^0_{2,115} \oplus S^0_{1,115} \oplus S^0_{2,115} \cdot S^0_{3,115} \oplus S^0_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,179} \cdot S^0_{2,179} \oplus S^0_{1,179} \oplus S^0_{2,179} \cdot S^0_{3,179} \oplus S^0_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^0_{1,243} \cdot S^0_{2,243} \oplus S^0_{1,243} \oplus S^0_{2,243} \cdot S^0_{3,243} \oplus S^0_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{1,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{1,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{1,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,89} \cdot S^1_{2,89} \oplus S^1_{1,89} \oplus S^1_{2,89} \cdot S^1_{3,89} \oplus S^1_{3,89}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,97} \cdot S^1_{2,97} \oplus S^1_{1,97} \oplus S^1_{2,97} \cdot S^1_{3,97} \oplus S^1_{3,97}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,119} \cdot S^1_{2,119} \oplus S^1_{1,119} \oplus S^1_{2,119} \cdot S^1_{3,119} \oplus S^1_{3,119}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,153} \cdot S^1_{2,153} \oplus S^1_{1,153} \oplus S^1_{2,153} \cdot S^1_{3,153} \oplus S^1_{3,153}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,161} \cdot S^1_{2,161} \oplus S^1_{1,161} \oplus S^1_{2,161} \cdot S^1_{3,161} \oplus S^1_{3,161}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,183} \cdot S^1_{2,183} \oplus S^1_{1,183} \oplus S^1_{2,183} \cdot S^1_{3,183} \oplus S^1_{3,183}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,217} \cdot S^1_{2,217} \oplus S^1_{1,217} \oplus S^1_{2,217} \cdot S^1_{3,217} \oplus S^1_{3,217}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,225} \cdot S^1_{2,225} \oplus S^1_{1,225} \oplus S^1_{2,225} \cdot S^1_{3,225} \oplus S^1_{3,225}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,247} \cdot S^1_{2,247} \oplus S^1_{1,247} \oplus S^1_{2,247} \cdot S^1_{3,247} \oplus S^1_{3,247}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{1,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{1,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{1,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{1,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,71} \cdot S^2_{2,71} \oplus S^2_{1,71} \oplus S^2_{2,71} \cdot S^2_{3,71} \oplus S^2_{3,71}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,93} \cdot S^2_{2,93} \oplus S^2_{1,93} \oplus S^2_{2,93} \cdot S^2_{3,93} \oplus S^2_{3,93}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,101} \cdot S^2_{2,101} \oplus S^2_{1,101} \oplus S^2_{2,101} \cdot S^2_{3,101} \oplus S^2_{3,101}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,115} \cdot S^2_{2,115} \oplus S^2_{1,115} \oplus S^2_{2,115} \cdot S^2_{3,115} \oplus S^2_{3,115}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,135} \cdot S^2_{2,135} \oplus S^2_{1,135} \oplus S^2_{2,135} \cdot S^2_{3,135} \oplus S^2_{3,135}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,157} \cdot S^2_{2,157} \oplus S^2_{1,157} \oplus S^2_{2,157} \cdot S^2_{3,157} \oplus S^2_{3,157}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,165} \cdot S^2_{2,165} \oplus S^2_{1,165} \oplus S^2_{2,165} \cdot S^2_{3,165} \oplus S^2_{3,165}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,179} \cdot S^2_{2,179} \oplus S^2_{1,179} \oplus S^2_{2,179} \cdot S^2_{3,179} \oplus S^2_{3,179}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,199} \cdot S^2_{2,199} \oplus S^2_{1,199} \oplus S^2_{2,199} \cdot S^2_{3,199} \oplus S^2_{3,199}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,221} \cdot S^2_{2,221} \oplus S^2_{1,221} \oplus S^2_{2,221} \cdot S^2_{3,221} \oplus S^2_{3,221}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,229} \cdot S^2_{2,229} \oplus S^2_{1,229} \oplus S^2_{2,229} \cdot S^2_{3,229} \oplus S^2_{3,229}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,243} \cdot S^2_{2,243} \oplus S^2_{1,243} \oplus S^2_{2,243} \cdot S^2_{3,243} \oplus S^2_{3,243}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{1,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,75} \cdot S^3_{2,75} \oplus S^3_{1,75} \oplus S^3_{2,75} \cdot S^3_{3,75} \oplus S^3_{3,75}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,139} \cdot S^3_{2,139} \oplus S^3_{1,139} \oplus S^3_{2,139} \cdot S^3_{3,139} \oplus S^3_{3,139}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,203} \cdot S^3_{2,203} \oplus S^3_{1,203} \oplus S^3_{2,203} \cdot S^3_{3,203} \oplus S^3_{3,203}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,64} \cdot S^2_{1,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,128} \cdot S^2_{1,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,192} \cdot S^2_{1,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,230} \cdot S^3_{1,230}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,110} \cdot S^2_{3,110}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,174} \cdot S^2_{3,174}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,238} \cdot S^2_{3,238}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,64} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,128} \cdot S^2_{4,64}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,192} \cdot S^2_{4,128}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,192}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,102} \cdot S^3_{1,102}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,166} \cdot S^3_{1,166}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,84} \cdot S^3_{3,84}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,114} \cdot S^3_{3,114}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,148} \cdot S^3_{3,148}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,178} \cdot S^3_{3,178}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,212} \cdot S^3_{3,212}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,242} \cdot S^3_{3,242}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,88} \cdot S^4_{3,88}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,152} \cdot S^4_{3,152}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,216} \cdot S^4_{3,216}&&\text{weight 1}
-\end{flalign*}
-The total weight of the trail is 76.
-
-\end{document}

### 2018-05-09 morusAC_01_Introduction.tex
propose introduction (poorly written)

@@
 
-Authenticated encryption is ... \\
+Block ciphers provide confidentiality through encryption.
+Hash functions are the building blocks of integrity and authenticity.
+These two operations combine in a single scheme class named Authenticated Encryption.
+From a message; IV and key, it produces a ciphertext and a final authentication tag.
 
 \noindent
-CAESAR \cite{CAESAR} ...\\
+The CAESAR\cite{CAESAR} Competition aims to provide with a portfolio of authenticated encryption schemes.
+The selection is made by comparing the resistances to a wide range of attacks.
+Morus is one of the last 3 finalist.
 
@@
 \subsubsection*{Paper Outline.}
-This paper is organized as follows......
+This paper is organized as follows, we first provide a brief introduction of how Morus operates\ref{sec/Preliminaries}.
+By showing existance of a certain class of rotational symmetries, we provide with a reduction of Morus namely MiniMorus\ref{sec/introminimorus}.
+We analysed this simplified scheme in \ref{sec/minitrails}, building a ciphertext-only trail with a weight of 16.
+We then extend our result to the full scheme, showing a bias in the keystream over 5 rounds. In \ref{sec/IniFin} we provide new attacks on reduced version of the initialization and the finalization.

### 2018-05-09 morusAC_02_Preliminaries.tex
split text into sections
propose introduction (poorly written)

@@
 \section{Preliminaries}
-\label{Sec/Preliminaries}
+\label{sec/Preliminaries}
 
-\subsection{Specification of MORUS}
-\label{subsec/Spec}
 MORUS \cite{MORUS} is a family of authenticated ciphers with two internal state sizes: 640 and 1280 bits.
@@
 
+\subsection{Specification of MORUS}
+\label{subsec/Spec}
 The state of MORUS is divided into five registers.
@@
 S^{-16}_3 &= 0^{256}, &
-S^{-16}_4 &= const_1\|const_0.
+S^{-16}_4 &= const_0\|const_1.
 \end{align*}

### 2018-05-09 morusAC_05_Full_trails.tex
propose introduction (poorly written)

@@
 \section{Trail for Full \cipher{MORUS}}
+\label{sec/fulltrails}
+
+In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
+We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
 
 \subsection{Symmetrizing the Trail}
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec:introminimorus}.
+In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
 That is, for every $S^t_{i,j}$ (resp.
@@
 Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec:introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
+But as observed in \Cref{sec/introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
 

### 2018-05-09 morusAC_01_Introduction.tex
update overview

@@
 \subsubsection*{Paper Outline.}
-This paper is organized as follows, we first provide a brief introduction of how Morus operates\ref{sec/Preliminaries}.
-By showing existance of a certain class of rotational symmetries, we provide with a reduction of Morus namely MiniMorus\ref{sec/introminimorus}.
-We analysed this simplified scheme in \ref{sec/minitrails}, building a ciphertext-only trail with a weight of 16.
-We then extend our result to the full scheme, showing a bias in the keystream over 5 rounds. In \ref{sec/IniFin} we provide new attacks on reduced version of the initialization and the finalization.
+This paper is organized as follows, we first provide a brief introduction of how Morus operates (Sect. \ref{sec/Preliminaries}).
+By showing existance of a certain class of rotational symmetries, we provide a reduction of Morus namely MiniMorus (Sect. \ref{sec/introminimorus}).
+We analysed this simplified scheme in section \ref{sec/minitrails}, building a ciphertext-only trail with a weight of 16.
+We then extend our result to the full scheme, showing a bias in the keystream over 5 rounds (Sec. \ref{sec/fulltrails}).
+In section \ref{sec/Discussion} we discuss about the implication of such weakness.
+In section \ref{sec/IniFin} we provide new attacks on reduced version of the initialization and the finalization.

### 2018-05-09 morusAC_07_IniFin.tex
finished section 7.2 (finished section 7)

@@
 \subsubsection{Full Diffusion Rounds.}
+We found that {\tt StepUpdate} was designed to have good diffusion in the forward direction.
+Thus, once the state is recovered, the attacker can perform the partial computation in the backward direction longer than the forward direction.
+We set $G_0$ and $G_1$ as follows.
+\begin{align*}
+G_0 &= \{1, 2, \cdots, 127\} & \textrm {Bit position 0 is unknown.}\\
+G_1 &= \{0, 1, \cdots, 7, 9, 10, \cdots, 127\} & \textrm {Bit position 8 is unknown.}
+\end{align*}
+Those will lead to 4 matching bits after the 4-step forward computation and the 6-step backward computation.
+The analysis of the diffusion is given in Table~\ref{Tbl:ini_diffuse}.
+In the end, $K$ can be recovered faster than the exhaustive search by 1 bit, i.e.~with complexity $2^{127}$.
+
+\begin{table}[!tb]
+\begin{center}
+\caption{Analysis of the diffusion and matching bits over 10 steps.
+`0' and `1' denote that the state bit can be computed from a partial knowledge of $K$.
+After the partial computations from each direction, 4 bits of $S^{-6}$ can match.}
+\label{Tbl:ini_diffuse}
+{\tiny
+\begin{tabular}{ccccc} \hline
+Round & \multicolumn{4}{c}{State Difference}\\ \hline
+                  & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+                  & {\tt 0000000000000000} & {\tt 0000000000000001} & {\tt 0000000000000000} & {\tt 0000000000000001} \\
+$S^{-10}\oplus K$ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+                  & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+                  & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\ \hline
+
+      & {\tt 0000000000002000} & {\tt 0000000000000000} & {\tt 0000000000002000} & {\tt 0000000000000000} \\
+      & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} \\
+1     & {\tt 0008000000000000} & {\tt 0000000000000000} & {\tt 0008000000000000} & {\tt 0000000000000000} \\
+      & {\tt 0000000000100000} & {\tt 0020000000000000} & {\tt 0000000000100000} & {\tt 0020000000000000} \\
+      & {\tt 0000000000020000} & {\tt 0084000000000000} & {\tt 0000000000020000} & {\tt 0084000000000000} \\ \hline
+
+      & {\tt 0800000000000004} & {\tt 0000000204000001} & {\tt 0800000000000004} & {\tt 0000000204000001} \\
+      & {\tt 8000000a00000000} & {\tt 0000002110000004} & {\tt 8000000a00000000} & {\tt 0000002110000004} \\
+2     & {\tt 0400010221000000} & {\tt 008000400a000081} & {\tt 0400010221000000} & {\tt 008000400a000081} \\
+      & {\tt 1000050001000244} & {\tt 4200118a08000280} & {\tt 1000050001000244} & {\tt 4200118a08000280} \\
+      & {\tt 880004a0a0200858} & {\tt 4840123350000050} & {\tt 880004a0a0200858} & {\tt 4840123350000050} \\ \hline
+
+      & {\tt 023d63c00050a850} & {\tt 00a1442000489380} & {\tt 023d63c00050a850} & {\tt 00a1442000489380} \\
+      & {\tt 02b63380056aaa48} & {\tt 00b5563005dcd6c0} & {\tt 02b63380056aaa48} & {\tt 00b5563005dcd6c0} \\
+3     & {\tt d42ab556bf5dfcd6} & {\tt 5a26f633a8556aaa} & {\tt d42ab556bf5dfcd6} & {\tt 5a26f633a8556aaa} \\
+      & {\tt 5fbbf556bd556c65} & {\tt 7aab99aaee6bea2c} & {\tt 5fbbf556bd556c65} & {\tt 7aab99aaee6bea2c} \\
+      & {\tt abff7f3ad7feafad} & {\tt cfff777ffddffd6d} & {\tt abff7f3ad7feafad} & {\tt cfff777ffddffd6d} \\ \hline
+
+      & {\tt fff77dfffffdcf57} & {\tt fefad7efffdffbf7} & {\tt fff77dfffffdcf57} & {\tt fefad7efffdffbf7} \\
+      & {\tt ffffffffffffbfff} & {\tt fffbf7fffddfff77} & {\tt ffffffffffffbfff} & {\tt fffbf7fffddfff77} \\
+4     & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} \\
+      & {\tt ffffffffffffffff} & {\tt fffbffffefffffff} & {\tt ffffffffffffffff} & {\tt fffbffffefffffff} \\
+      & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} \\ \hline 
+%
+%      
+      &                        & 2-bits match           & {\tt                 } & 2-bits match            \\ \hline
+%
+%
+      & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} \\
+      & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} \\
+5     & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} & {\tt ffffffffffffffff} \\
+      & {\tt fffff7ffedfffff7} & {\tt ffffffffedffffff} & {\tt fffff7ffedfffff7} & {\tt ffffffffedffffff} \\
+      & {\tt ffffffffedffffff} & {\tt fffff7ffedfffff7} & {\tt ffffffffedffffff} & {\tt fffff7ffedfffff7} \\ \hline
+
+      & {\tt ffffffffedffffff} & {\tt fffff7ffedfffff7} & {\tt ffffffffedffffff} & {\tt fffff7ffedfffff7} \\
+      & {\tt fffbf5e7cdfffbf7} & {\tt fffff7bfcdfff757} & {\tt fffbf5e7cdfffbf7} & {\tt fffff7bfcdfff757} \\
+6     & {\tt fffbf5e7cdfffbf7} & {\tt fffff7bfcdfff757} & {\tt fffbf5e7cdfffbf7} & {\tt fffff7bfcdfff757} \\
+      & {\tt 7ffd75b6cdfff357} & {\tt fffbf5a6ccfcfb73} & {\tt 7ffd75b6cdfff357} & {\tt fffbf5a6ccfcfb73} \\
+      & {\tt 7ffbf5a6ccfcf373} & {\tt 7ff975b6ccfff353} & {\tt 7ffbf5a6ccfcf373} & {\tt 7ff975b6ccfff353} \\ \hline
+
+      & {\tt 7efbf5a6cc7cf353} & {\tt 7fd975a6cceff353} & {\tt 7efbf5a6cc7cf353} & {\tt 7fd975a6cceff353} \\
+      & {\tt 7eb950a4cc78e353} & {\tt 7dd07184cced7153} & {\tt 7eb950a4cc78e353} & {\tt 7dd07184cced7153} \\
+7     & {\tt 7eb950a4cc78e353} & {\tt 7dd07184ccec7153} & {\tt 7eb950a4cc78e353} & {\tt 7dd07184ccec7153} \\
+      & {\tt 7cd051044c6c3153} & {\tt 3e985024cc48a313} & {\tt 7cd051044c6c3153} & {\tt 3e985024cc48a313} \\
+      & {\tt 3c905004cc482313} & {\tt 7c9051044c6c2113} & {\tt 3c905004cc482313} & {\tt 7c9051044c6c2113} \\ \hline
+
+      & {\tt 2c905004c4482113} & {\tt 7c9050040c682113} & {\tt 2c905004c4482113} & {\tt 7c9050040c682113} \\
+      & {\tt 2810100444082112} & {\tt 5c1010040c402113} & {\tt 2810100444082112} & {\tt 5c1010040c402113} \\
+8     & {\tt 2810100444082112} & {\tt 1c1010040c402113} & {\tt 2810100444082112} & {\tt 1c1010040c402113} \\
+      & {\tt 0c00100404400113} & {\tt 2800000404082112} & {\tt 0c00100404400113} & {\tt 2800000404082112} \\
+      & {\tt 0800000404002112} & {\tt 0800100404400113} & {\tt 0800000404002112} & {\tt 0800100404400113} \\ \hline
+
+      & {\tt 0800000404002112} & {\tt 0800100004000112} & {\tt 0800000404002112} & {\tt 0800100004000112} \\
+      & {\tt 0000000404000102} & {\tt 0000100004000110} & {\tt 0000000404000102} & {\tt 0000100004000110} \\
+9     & {\tt 0000000404000102} & {\tt 0000000004000110} & {\tt 0000000404000102} & {\tt 0000000004000110} \\
+      & {\tt 0000000004000110} & {\tt 0000000000000102} & {\tt 0000000004000110} & {\tt 0000000000000102} \\
+      & {\tt 0000000000000100} & {\tt 0000000004000110} & {\tt 0000000000000100} & {\tt 0000000004000110} \\ \hline
+
+      & {\tt 0000000000000100} & {\tt 0000000004000100} & {\tt 0000000000000100} & {\tt 0000000004000100} \\
+      & {\tt 0000000000000000} & {\tt 0000000004000100} & {\tt 0000000000000000} & {\tt 0000000004000100} \\
+10    & {\tt 0000000000000000} & {\tt 0000000000000100} & {\tt 0000000000000000} & {\tt 0000000000000100} \\
+      & {\tt 0000000000000100} & {\tt 0000000000000000} & {\tt 0000000000000100} & {\tt 0000000000000000} \\
+      & {\tt 0000000000000000} & {\tt 0000000000000100} & {\tt 0000000000000000} & {\tt 0000000000000100} \\ \hline
+
+              & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+              & {\tt 0000000000000000} & {\tt 0000000000000100} & {\tt 0000000000000000} & {\tt 0000000000000100} \\
+$S^0\oplus K$ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+              & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+              & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\ \hline
+
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\subsubsection{Remarks.}
+The matching state does not have te be a border of a step.
+It can be defined on a border of a round (or even some complicated way).
+We did not find the extension of the number of attacked steps even with this way.
+
+As can be seen in Table~\ref{Tbl:ini_diffuse}, the updated register in step $i$ is independent of the update function in step $i+1$ in the forward direction, and starts to impact from step $i+2$.
+By modifying this point, the diffusion speed can increase faster, which makes this attack harder.

### 2018-05-09 morusAC.tex
citations, common.sty
morus analysis references

@@
 
-\documentclass{llncs}
+\newif\ifshrink\shrinkfalse
+%\shrinktrue for screen-friendly original LNCS paper size
+
+\documentclass[a4paper]{llncs}
 \usepackage{common}
@@
 \else
-\author{Tomer Ashur, Maria Eichlseder, Ga\"etan Leurent,\\Brice Minaud, Yann Rotella, Yu Sasaki, Beno\^it Viguier}
+\author{Tomer Ashur \and Maria Eichlseder \and Ga\"etan Leurent \and Brice Minaud \and Yann Rotella \and Yu Sasaki \and Beno\^it Viguier}
 \institute{}
@@
 
-\mainmatter
-
-% \title{Note on Linear Biases in MORUS}
-\title{Morus: Biases in the stream}
-
+\title{\morus: Biases in the Stream}
 \maketitle
+\todo{The title should reflect the full analysis. How about something more generic?}
 
@@
 
+\ifanonymous
+\else
+  %\subsubsection{Acknowledgments.}
+  %...
+  %\horizonack{The research leading to these results has received funding from
+  %the European Union's Horizon 2020 research and innovation programme under
+  %grant agreement No 644052 (HECTOR) and \dots.}
+\fi
+
 \bibliographystyle{splncs03}

### 2018-05-09 morusAC_04_Mini_trails.tex
propose introduction (poorly written)
Describe attack on MORUS1280 primarly

@@
 \section{Linear Trail for \cipher{MiniMORUS}}
+\label{sec/minitrails}
+
+In this section, we describe how we build a trail for \cipher{MiniMORUS} before computing its bias and experimentally verify it.
 
@@
   \centering
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
     \printstate
@@
   \begin{subfigure}{.32\textwidth}
-  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
+  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
     \printstate
@@
 Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS640}, this yields the equation:
-
+Ultimately, for \cipher{MiniMORUS1280}, this yields the equation:
 \begin{flalign*}
-&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
-&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
-&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
-&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
-&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
-&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
-&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
-&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
-&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
-&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
+&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
+&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
+&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
+&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
+&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
+&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
+&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
+&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
+&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
+&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
+&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
+&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
 \end{flalign*}
 
-The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.
+The equation for \cipher{MiniMORUS640} is very similar, and is given in \Cref{sec:traileq}.
 
@@
 
-In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
+In the equation for \cipher{MiniMORUS1280} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
 By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
 Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS1280} (cf.
+The same holds for \cipher{MiniMORUS640} (cf.
 \Cref{sec:traileq}).
@@
 
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
+To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS1280} and \cipher{MiniMORUS640}.
 In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
@@
 \end{table}
+
+%%% Local Variables:
+%%% TeX-master: "morusAC"
+%%% End:

### 2018-05-09 morusAC_05_Full_trails.tex
Describe attack on MORUS1280 primarly

@@
 
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}.
-We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.
+In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS1280} and \cipher{MiniMORUS640}.
+We now turn to the full ciphers \cipher{MORUS1280} and \cipher{MORUS640}.
 
@@
 $C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
+For example, if $w=64$ (for \cipher{MORUS1280}), the term $S^3_{2,0}$ is symmetrized into:
 \[
-S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
+S^3_{2,0} \oplus S^3_{2,64} \oplus S^3_{2,128} \oplus S^3_{2,192}.
 \]
@@
 Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
+In this way, the biased linear combination on \cipher{MiniMORUS1280} given in \Cref{sec:minitraileq}, namely:
 \begin{align*}
-&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
-\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
+&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}\\
+\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}
 \end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
+ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS1280}:
 \begin{align*}
-&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
-\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
-\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
-\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
-\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
-\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
+&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}\\
+\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}\\
+\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}\\
+\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}\\
+\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}\\
+\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}\\
+\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}
 \end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.
+We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS640}.
 
@@
 
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}.
+Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS1280} and \cipher{MORUS640}.
 However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
@@
 \end{figure}
+
+%%% Local Variables:
+%%% TeX-master: "morusAC"
+%%% End:

### 2018-05-09 morusAC_01_Introduction.tex
citations, common.sty
morus analysis references
working on introduction

@@
 
-Block ciphers provide confidentiality through encryption.
-Hash functions are the building blocks of integrity and authenticity.
-These two operations combine in a single scheme class named Authenticated Encryption.
-From a message; IV and key, it produces a ciphertext and a final authentication tag.
+Authenticated Encryption (AE) schemes combine the functionality of symmetric encryption schemes and message authentication codes.
+Based on a shared secret key $K$, they encrypt a plaintext message $M$ to a ciphertext $C$ and authentication tag $T$ in order to protect both the confidentiality and the authenticity of $M$.
+Most modern authenticated encryption algorithms are nonce-based schemes with associated data (AEAD),
+where $(C, T)$ additionally depends on a unique nonce $N$ (or initialization value IV) and optional associated metadata $A$.
+One of the most prominent standardized AEAD designs is \cipher{AES-GCM} \cite{TODOMV05,TODONIST},
+which is widely deployed in protocols such as \cipher{TLS v1.3}.
 
-\noindent
-The CAESAR\cite{CAESAR} Competition aims to provide with a portfolio of authenticated encryption schemes.
-The selection is made by comparing the resistances to a wide range of attacks.
-Morus is one of the last 3 finalist.
+To address the growing need for modern authenticated encryption designs for different application scenarios, 
+the CAESAR competition was launched in 2013 \cite{CAESAR}.
+The goal of this competition is to select a final portfolio of AEAD designs for three different use-cases:
+(1) lightweight hardware characteristics,
+(2) high-speed software performance, and
+(3) robustness.
+The competition attracted 57 first-round submissions, 7 of which were recently selected as finalists in the fourth selection round.
+
+\morus is one of the three finalists for use-case (2), together with \cipher{OCB} and \cipher{AEGIS}.
+\todo[inline]{Advertise MORUS}
+
+\subsubsection*{Related Work.}
+
+\todo[inline]{Analysis of MORUS: Designers and \cite{secryptDwivediMW17,trustcomSalamSBDPW17,balkancryptsecMilevaDV15}, and nonce-misuse analysis}
+
+\todo[inline]{Related designs; Analysis of related designs; discuss \cite{sacryptMinaud14} on AEGIS \cite{AEGIS}}
 
 \subsubsection*{Our Contributions.}
+% Paragraph: We have 3 results, the main one is on full MORUS
+
+% Paragraph: Details on keystream bias
 Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
@@
 Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
+\todo{somewhere: linear cryptanalysis \cite{eurocryptMatsui93,eurocryptMatsuiY92}}
+
+% Paragraph: Details on Section 7
 

### 2018-05-09 morusAC_01_Introduction.tex
working on introduction

@@
 \subsubsection*{Our Contributions.}
-% Paragraph: We have 3 results, the main one is on full MORUS
+We analyze the security of \morus and present a keystream distinguisher based on
+linear approximations for full \morus[1280].
+In addition, we provide results for round-reduced \morus with reduced initialization or finalization.
 
 % Paragraph: Details on keystream bias
-Our main result is the existence of a linear trail linking ciphertext bits spanning five consecutive encryption steps.
-As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
-Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
-Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-\todo{somewhere: linear cryptanalysis \cite{eurocryptMatsui93,eurocryptMatsuiY92}}
+Our main result is a linear approximation \cite{eurocryptMatsui93,eurocryptMatsuiY92} linking ciphertext bits of five consecutive encryption blocks.
+%As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
+%Technically, this assumes that the inner state of the cipher at the start of the trail is uniformly random, but this is a fair assumption since the inner state should be indistinguishable from uniform randomness by design.
+%Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
+More precisely, we show that when encrypting a fixed plaintext block multiple times, a linear correlation exists between some output bits of the cipher.
+Moreover, the bias depends purely on the plaintext, and not on the secret key of the cipher.
+In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
+For \morus[1280], the \morus family member with 256-bit security, the linear bias is $2^{-73}$
+\todo{double-check bias vs. correlation everywhere}
+and can be exploited using about $2^{152}$ encrypted blocks.
+\todo{double-check}
+We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext block is encrypted sufficiently often, regardless of whether each encryption uses a different key.
 
-% Paragraph: Details on Section 7
+To obtain this result, we propose a simplified abstraction of \morus, called \minimorus, which takes advantage of certain rotational symmetries in \morus.
+We observe that the actual bias of the approximation is better than suggested by the construction of the characteristics.
+To confirm the validity of our analysis,
+we practically verified the bias of the linear approximation for \minimorus and of partial approximations for \morus.
+In addition, we symbolically evaluated the relevant equations.
+
+In addition to this attack on full \morus, we also analyze the security of \morus with round-reduced initialization (in a nonce-misuse setting) or finalization.
+While this analysis does not threaten full \morus, it complements the main result to provide a better understanding of the \morus design approach.
+We present a forgery attack for round-reduced \morus[1280] with success probability $2^{-88}$ for a 128-bit tag if the finalization is reduced to 3 out of 10 rounds.
+This nonce-respecting attack is based on a differential analysis of the padding rule.
+The second result targets a round-reduced initialization with 10 out of 16 steps, and extends a state-recovery attack (e.g., in a nonce-misuse setting) into a key-recovery attack.
 
 \subsubsection*{Paper Outline.}
-This paper is organized as follows, we first provide a brief introduction of how Morus operates (Sect. \ref{sec/Preliminaries}).
-By showing existance of a certain class of rotational symmetries, we provide a reduction of Morus namely MiniMorus (Sect. \ref{sec/introminimorus}).
-We analysed this simplified scheme in section \ref{sec/minitrails}, building a ciphertext-only trail with a weight of 16.
-We then extend our result to the full scheme, showing a bias in the keystream over 5 rounds (Sec. \ref{sec/fulltrails}).
-In section \ref{sec/Discussion} we discuss about the implication of such weakness.
-In section \ref{sec/IniFin} we provide new attacks on reduced version of the initialization and the finalization.
+This paper is organized as follows.
+We first provide a brief description of \morus in \autoref{sec/Preliminaries}.
+In \autoref{sec/introminimorus}, we introduce \minimorus, an abstraction of \morus based on a certain class of rotational symmetries.
+We analyze this simplified scheme in \autoref{sec/minitrails} and provide a ciphertext-only linear approximation with a weight of 16.
+We then extend our result to the full scheme in \autoref{sec/fulltrails}, showing a bias in the keystream over 5 rounds,
+and discuss the implications of our observation for the security of \morus in \autoref{sec/Discussion}.
+In \autoref{sec/IniFin}, we present our results on the security of \morus with round-reduced initialization (in a nonce-misuse setting) or finalization.
+Finally, we conclude in \autoref{TODO}.

### 2018-05-09 morusAC_06_Discussion.tex
propose introduction (poorly written)
working on introduction

@@
 \section{Discussion}
+\label{sec/Discussion}
+
+\todo{ME: If this doesn't grow much longer, I'd suggest to make this a subsection of Section 5, and repeat a summary in the Conclusion.}
 
@@
 
-We emphasize that the biases we uncover between ciphertext bits are \emph{absolute}, in the sense that they do not depend on the encryption key, or the nonce.
-As such, they could be leveraged to mount an attack in a known-prefix, unknown-suffix scenario, similar to biases on the RC4 stream cipher.
-This is especially effective in a setting where the attacker is able to force multiple encryptions of the same plaintext, while controlling the length of the known prefix, such as in the BEAST attack on TLS.
+We now discuss the impact of our attacks for the security of \morus.
 
-The design document of \cipher{MORUS} imposes a limit of $2^{64}$ encrypted blocks for a given key.
-However, since our attack is independent of the encryption key, and hence immune to rekeying, this limitation does not apply: all that matters for our attack is that the same plaintext be encrypted enough times.
+\subsubsection{Keystream biases.}
 
-With the trail presented in this work, the data complexity is still out of reach, since exploiting the bias would require $2^{146}$ encrypted blocks for \cipher{MORUS640}, and $2^{152}$ encrypted blocks for \cipher{MORUS1280}.
-The data complexity could be slightly lowered by leveraging multilinear cryptanalysis; indeed, the trail holds for any bit shift, and if we assume independence, we could run $w$ copies of the trail in parallel on the same encrypted blocks (recall that $w$ is the word size, and the trail is invariant by rotation by $w$ bits).
-This would save a factor $2^5$ on the data complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but the resulting complexity is still out of reach.
+We emphasize that the biases we uncover between ciphertext bits are
+\emph{absolute}, in the sense that they do not depend on the encryption
+key, or the nonce.  This is the same situation as the keystream biases
+in AEGIS~\cite{sacryptMinaud14}.  As such, they can be leveraged to
+mount an attack in the broacast setting, where the same message is
+encrypted multiple times with different IVs and potentially different
+keys~\cite{DBLP:conf/fse/MantinS01}.  In particular, the broadcast setting appears
+in practice in man-in-the-browser attacks against HTTPS connections
+following the BEAST model~\cite{duong2011here}.  In this scenario, an
+attacker uses Javascript code running in the victim's browser (by
+tricking the victim to visit a malicious website) to generate a large
+number of request to a secure website.  Because of details of the HTTP
+protocol, each request includes an authentication token to identify the
+user, and the attacker can target this token as a repeated plaintext.
+Concretely, biases in the RC4 keystream have been exploited in this
+setting, leading to the recovery of authentication cookies in
+practice~\cite{DBLP:conf/uss/AlFardanBPPS13}.
 
-However, the existence of this trail does hint at some weakness in the design of \cipher{MORUS}.
-Indeed, a notable feature of the trail is that the values of rotation constants are mostly irrelevant: a similar trail would exist for most choices of the constants.
-That it is possible to build a trail that ignores rotation constants may be surprising.
-This would have been prevented by adding a word-wise rotation to one of the state registers at the input of the ciphertext equation.
+\subsubsection{Data complexity.}
+
+The design document of \cipher{MORUS} imposes a limit of $2^{64}$
+encrypted blocks for a given key. However, since our attack is
+independent of the encryption key, and hence immune to rekeying, this
+limitation does not apply: all that matters for our attack is that the
+same plaintext be encrypted enough times.
+
+With the trail presented in this work, the data complexity is clearly
+out of reach in practice, since exploiting the bias would require
+$2^{152}$ encrypted blocks for \cipher{MORUS1280}, and $2^{146}$
+encrypted blocks for \cipher{MORUS640}. The data complexity could be
+slightly lowered by leveraging multilinear cryptanalysis; indeed, the
+trail holds for any bit shift, and if we assume independence, we could
+run $w$ copies of the trail in parallel on the same encrypted blocks
+(recall that $w$ is the word size, and the trail is invariant by
+rotation by $w$ bits). This would save a factor $2^5$ on the data
+complexity for \cipher{MORUS640}, and $2^6$ for \cipher{MORUS1280}; but
+the resulting complexity is still out of reach.
+
+However, \cipher{MORUS1280} with a 256-bit key claims a security level
+of 256 bits for confidentiality, and an attack with complexity $2^{152}$
+should violate this claim, even if it is not practical.  The
+\cipher{MORUS} specification limits the amount of plaintext encrypted
+with a given key to $2^{64}$, but this does not affect our attack
+because we can use keystream generated with different keys.
+
+Moreover, the existence of this trail does hint at some weakness in the
+design of \cipher{MORUS}. Indeed, a notable feature of the trail is that
+the values of rotation constants are mostly irrelevant: a similar trail
+would exist for most choices of the constants. That it is possible to
+build a trail that ignores rotation constants may be surprising. This
+would have been prevented by adding a word-wise rotation to one of the
+state registers at the input of the ciphertext equation.
+
+%%% Local Variables:
+%%% TeX-master: "morusAC"
+%%% End:

### 2018-05-09 morusAC_02_Preliminaries.tex
minor rewrites in section 2

@@
+%!TEX root = morusAC.tex
+
 \section{Preliminaries}
@@
 Each member is called \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
+
+An instance of MORUS is parametrized by a secret key.
+During encryption, it takes as input a plaintext message and possibly some associated data, and outputs a ciphertext together with an authentication tag.
 In the recommended parameters, \cipher{MORUS640} supports 128-bit keys and \cipher{MORUS1280} supports both 128-bit and 256-bit keys.
@@
 \label{subsec/Spec}
-The state of MORUS is divided into five registers.
+
+The internal state of MORUS is composed of five registers.
 Let $q$ be the register size, where $q = 128$ for \cipher{MORUS640} and $q=256$ for \cipher{MORUS1280}.
 Let $S_i, i \in\{0,1,2,3,4\}$ be a register.
-Then, the state is represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
-The state is first initialized by a key and nonce, and then updated by associated data blocks and message blocks step by step.
-Hence, we use a superscript to denote the step number of the state as $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is $-16$ before the initialization and $t=0$ after the initialization.
+Then the internal state of MORUS may be represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
+Registers are themselves divided into four $q/4$-bit \emph{words}.
+Throughout the paper, we denote the word size by $w = q/4$, i.e. $w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
 
-The computation of MORUS consists of four parts; initialization, associated data process, encryption, and finalization.
-All parts share the core design called the {\tt StepUpdate} function.
-%Hereafter we explain {\tt StepUpdate}.
+The encryption process of MORUS consists of four parts: initialization, associated data processing, encryption, and finalization. 
+During the initialization phase, the value of the state is initialized using a key and nonce.
+Associated data block and plaintext blocks are then processed block by block.
+Then the internal state undergoes the finalization phase, which outputs the authentication tag.
 
-\subsubsection{Step Update Function.}
-{\tt StepUpdate} takes as input the internal state $S^t = S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$ and additional register-wise value $m^t$.
-$m^t$ is supposed to be associated data block or message block.
-$m^t$ can sometimes be 0 or constant when the additional input is unnecessary.
+Every part of this process relies on iterating the \StateUpdate{} function at the core of MORUS. Each call to the \StateUpdate{} function is called a step.
+In order to denote the internal state of the cipher at the $t$-th step, we use a superscript to denote the step number: the internal state at step $t$ is denoted by $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is $-16$ before the initialization and $t=0$ after the initialization.
 
-{\tt StepUpdate} is composed of 5 rounds with similar operations to update
-the state $S^t$, in which the additional input $m^t$ is used in rounds 2 to 5, but not in round 1.
-Each round uses the word-wise left circular shift, which divides a $q$-bit register value into 4 words of $q/4$ bits, and performs left circular shift operation for every $q/4$-bit word.
-Throughout the paper, we denote the word size by $w$, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
+%Hereafter we explain \StateUpdate{}.
+
+\subsubsection{The \StateUpdate{} Function.}
+\StateUpdate{} takes as input the internal state $S^t = S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$ and an additional $q$-bit value $m^t$ (recall that $q$ is the size of a register), and outputs an updated internal state.
+%During the associated data processing phase and encryption phase, $m^t$ is is an associated data block or message block. Otherwise $m^t$ can be 0 or constant when the additional input is unnecessary.
+
+\StateUpdate{} is composed of 5 rounds with similar operations.
+The additional input $m^t$ is used in rounds 2 to 5, but not in round 1.
+Each round uses the word-wise left circular shift operation, which divides a $q$-bit register value into 4 words of $w = q/4$ bits, and performs a left circular shift operation on each $w$-bit word.
 The word-wise circular shift constants $b_i$ in round $i$ are defined in Table~\ref{Tbl:rcon}.
-Besides, the left rotation of a whole $q$-bit register is used.
+Left circular shifts on a whole $q$-bit register are also used.
 The register-wise circular shift constants are also listed in Table~\ref{Tbl:rcon}.
@@
 \centering
-\caption{Left Circular-Shift Constants of MORUS} \label{Tbl:rcon}
-\begin{tabular}{c||ccccc|ccccc} \hline
-& \multicolumn{5}{c|}{word-wise circular shift} & \multicolumn{5}{c}{register-wise circular shift} \\
+\caption{Left Circular Shift Constants of MORUS} \label{Tbl:rcon}
+\begin{tabular}{r||ccccc|ccccc} \hline
+& \multicolumn{5}{c|}{Word-wise circular shift} & \multicolumn{5}{c}{Register-wise circular shift} \\
                   & \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline
@@
 \end{table}
-We denote those register-wise and word-wise left circular shift by $x \lll y$ and $x \lll_w y$, respectively.
+
+We denote those register-wise and word-wise left circular shift by $x \lll y$ and $x \lll_W y$, respectively.
 (Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document.)
 
-Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow {\tt StateUpdate}(S^t,m^t)$ is defined as follows.
+Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow \StateUpdate{}(S^t,m^t)$ is defined as follows.
 \begin{align*}
 \textrm{Round 1:} &&
-S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 ) \lll_w b_0 &
+S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 ) \lll_W b_0 &
 S^t_3 \leftarrow S^t_3 \lll b'_0.\\
 \textrm{Round 2:} &&
-S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_w b_1 &
+S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_W b_1 &
 S^t_4 \leftarrow S^t_4 \lll b'_1.\\
 \textrm{Round 3:} &&
-S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_w b_2 &
+S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_W b_2 &
 S^t_0 \leftarrow S^t_0 \lll b'_2.\\
 \textrm{Round 4:} &&
-S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_w b_3 &
+S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_W b_3 &
 S^t_1 \leftarrow S^t_1 \lll b'_3.\\
 \textrm{Round 5:} &&
-S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_w b_4 &
+S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_W b_4 &
 S^t_2 \leftarrow S^t_2 \lll b'_4.
@@
 \subsubsection{Initialization.}
-The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K_{128}$ and a 128-bit nonce $IV_{128}$ into the state, and running the {\tt StateUpdate} for 16 steps.
+The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K_{128}$ and a 128-bit nonce $IV_{128}$ into the state, and running \StateUpdate{} for 16 steps.
 The key and nonce are loaded into the state as follows:
@@
 \end{align*}
-Then the internal state is updated 16 times by  ${\tt StateUpdate}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is xored to the state again as $S^0_1 \leftarrow S^0_1 \oplus K_{128}$.
+The internal state is then updated 16 times with  $\StateUpdate{}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is XORed into the state again with $S^0_1 \leftarrow S^0_1 \oplus K_{128}$.
 
-The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the different register size and the two possible key size:
+The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the difference in register size and the two possible key sizes:
 \begin{itemize}
-  \item[-] When the key size is 128 bits, $K = K_{128} \|K_{128}$
-  \item[-] When the key size is 256 bits, $K = K_{256}$
+  \item[-] when the key size is 128 bits, $K = K_{128} \|K_{128}$.
+  \item[-] when the key size is 256 bits, $K = K_{256}$.
 \end{itemize}
-The state is initialized as
+The state is initialized as:
 \begin{align*}
@@
 \end{align*}
-After updating the state 16 times, the state is updated by $S^0_1 \leftarrow S^0_1 \oplus K$.
+After updating the state 16 times with \StateUpdate{}, the state is updated with $S^0_1 \leftarrow S^0_1 \oplus K$.
 
 \subsubsection{Associated Data Processing.}
-After the initialization, the associated data $A$ is processed.
-For the padding, if the last associated data block is not a full block, use `0' bits to pad it to 128 bits and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
-Note that if the length of $A$ denoted by `$adlen$' is 0, the associated data processing is skipped.
+After initialization, the associated data $A$ is processed.
+For the padding, if the last associated data block is not a full block,  it is padded to 128 bits and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280} respectively, using 0 bits.
+The padded full block is then used to update the state.
+If the length of $A$, denoted by `${\it adlen}$' is 0, the associated data processing phase is skipped.
 
-Then, state is updated by $S^{t+1} \leftarrow {\tt StateUpdate}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{adlen}{128} \rceil -1$ and $\ell = \lceil \frac{adlen}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
+The state is then updated with $S^{t+1} \leftarrow \StateUpdate{}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{{\it adlen}}{128} \rceil -1$ and $\ell = \lceil \frac{{\it adlen}}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
 
 \subsubsection{Encryption.}
-At each step of the encryption, a 16-byte (resp. 32-byte) message block $M_i$ is used to update the state, and $M_t$ is encrypted to $C_t$ in \cipher{MORUS640} (resp. \cipher{MORUS1280}).
+At each encryption step, a 16-byte (resp. 32-byte) message block $M_t$ is used to update the state in \cipher{MORUS640} (resp. \cipher{MORUS1280}), and a ciphertext $C_t$ is produced.
 
-If the last message block is not a full block, use `0' bits to pad it to 128 and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
+If the last message block is not a full block, 0 bits are used to pad it to 128 and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
 But only the partial block is encrypted.
-Note that if the message length denoted by `$msglen$' is 0, encryption is skipped.
+Note that if the message length denoted by `${\it msglen}$' is 0, encryption is skipped.
 
-In \cipher{MORUS640}, Let $u = \lceil \frac{adlen}{128} \rceil$ and $v = \lceil \frac{msglen}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$;
+In \cipher{MORUS640}, Let $u = \lceil \frac{{\it adlen}}{128} \rceil$ and $v = \lceil \frac{{\it msglen}}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$:
 \begin{align*}
 C^t &\leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 96) \oplus (S^{u+t}_2 \cdot S^{u+t}_3),\\
-S^{u+t+1} &\leftarrow {\tt StateUpdate}(S^{u+t},M^t).
+S^{u+t+1} &\leftarrow \StateUpdate{}(S^{u+t},M^t).
 \end{align*}
 
-For \cipher{MORUS1280}, $u = \lceil \frac{adlen}{256} \rceil$ and $v = \lceil \frac{msglen}{256} \rceil$.
-The equation to compute ciphertext block $C^t$ is $C^t \leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 192) \oplus (S^{u+t}_2 \cdot S^{u+t}_3)$.
+For \cipher{MORUS1280}, $u = \lceil \frac{{\it adlen}}{256} \rceil$ and $v = \lceil \frac{{\it msglen}}{256} \rceil$.
+The ciphertext block $C^t$ is computed as $C^t \leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 192) \oplus (S^{u+t}_2 \cdot S^{u+t}_3)$.
 
 \subsubsection{Finalization.}
-The finalization generates the authentication tag $T$ using 10 more steps.
-We only discuss the case that $T$ is not truncated.
+The finalization phase generates the authentication tag $T$ using 10 more \StateUpdate{} steps.
+We only discuss the case where $T$ is not truncated.
 The associated data length and the message length are used to update the state as follows.
 \begin{enumerate}
-\item $tmp \leftarrow adlen\|msglen$ for \cipher{MORUS640} and $tmp \leftarrow adlen\|msglen\|0^{128}$ for \cipher{MORUS1280}, where $adlen$ and $msglen$ are represented as 64-bit integers.
+\item ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}$ for \cipher{MORUS640} and ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}\|0^{128}$ for \cipher{MORUS1280}, where ${\it adlen}$ and ${\it msglen}$ are represented as 64-bit integers.
 \item $S^{u+v}_4 \leftarrow S^{u+v}_4 \oplus S^{u+v}_0.$
@@
 %\begin{equation*}
-$S^{t+1} \leftarrow {\tt StateUpdate} (S^t, tmp)$.
+$S^{t+1} \leftarrow \StateUpdate{} (S^t, {\it tmp})$.
 %\end{equation*}
@@
 \subsubsection{Security.}
-When the tag size is 128 bits, integrity is claimed up to 128-bits and confidentiality is claimed up to the number of bits of the key.
+With a 128-bit tag, integrity is claimed up to 128 bits and confidentiality is claimed up to the number of bits of the key.
 \begin{table}[!htb]
@@
 sum) of their individual biases (resp.
-weights).
+weights) \cite{eurocryptMatsui93}.
 
-We will also use the following notation, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
+We also recall the following notation from the previous section, where an \emph{encryption step} refers to one call to the \StateUpdate{} function:\\
 \centerline{
@@
 
-In the remainder, the $0$-th encryption step will denote the encryption step where our linear trail starts.
+For simplicity, in the remainder, the $0$-th encryption step will often denote the encryption step where our linear trail starts.
 Any encryption step could be chosen for that purpose, as long as at least four more encryption steps follow.

### 2018-05-09 morusAC_00_Abstract.tex

@@
 \begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
+This paper presents linear biases in the output of an authenticated encryption scheme \cipher{MORUS} that is one of the seven finalists in CAESAR.
+The bias allows an attacker to distinguish it from an ideal cipher, thus breaks the designer's confidentiality claim.
+%
+More precisely, we show that a linear correlation exists between some bits of the plaintext and ciphertext.
+Moreover the bias depends purely on the plaintext, and not on the secret key nor the nonce, hence restricting the single use of nonce or even rekeying does not prevent the attacks.
+%
+From the nature of the linear cryptanalysis, the attack works in the known plaintext attacks.
+Moreover, if a fixed plaintext is encrypted multiple times, the correlation can be used to recover a 1-bit information of the plaintext.
+The attack principally has larger impact because it can be used to recover more unknown bits of the plaintext, provided an initial segment of the plaintext is known.
+%
+The bias for \cipher{MORUS1280}, claiming a security level of 256 bits, is $2^{-76}$, would thus be identified by around $2^{152}$ encryptions.
+%
+To identify the bias, the self-similarity of the \cipher{MORUS1280} operation is exploited.
+More precisely, \cipher{MORUS1280} operates on 256-bit registers that are further divided into four 64-bit words.
+Many operations of \cipher{MORUS1280} work in a symmetric-way.
+This motivates us to introduce word-oriented version of \cipher{MORUS} called \cipher{MiniMORUS}, which helps the analysis.
+The attack is implemented and verified on \cipher{MiniMORUS} having a bias of $2^{-16}$.
+
+%The bias for \cipher{MORUS1280} is slightly worse than the 4 copies of \cipher{MiniMORUS}, and the mechanism behind is explained.
+%
+%Towards the better understanding of \cipher{MORUS}, several other properties are investigated though these do not break any security claim by the designers.
+Those include the bias for \cipher{MORUS640} requiring $2^{146}$ encryption against 128-bit security, and analyses on initialization and finalization of the step-reduced version of \cipher{MORUS1280}.
+
+%In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
 More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
@@
 We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
+\keywords{MORUS, CAESAR, Authenticated Encryption, Nonce Respect, Linear Cryptanalysis, Confidentiality.}
 \end{abstract}

### 2018-05-09 morusAC_01_Introduction.tex

@@
 \morus is one of the three finalists for use-case (2), together with \cipher{OCB} and \cipher{AEGIS}.
-\todo[inline]{Advertise MORUS}
+This family of authenticated ciphers by Wu and Huang \cite{MORUS} provides three main variants:
+\morus[640] with a 128-bit key and \morus[1280] with either a 128-bit or a 256-bit key.
+The design approach is reminiscent of classical stream cipher designs and
+continuously updates a relatively large state with a few fast operations.
+\morus can be efficiently implemented in both software and hardware;
+in particular, the designers claim that the software performance even surpasses
+\cipher{AES-GCM} implementations using Intel's \cipher{AES-NI} instructions,
+and that \morus is the fastest authenticated cipher not using \cipher{AES-NI} \cite{MORUS}.
 
 \subsubsection*{Related Work.}
-
-\todo[inline]{Analysis of MORUS: Designers and \cite{secryptDwivediMW17,trustcomSalamSBDPW17,balkancryptsecMilevaDV15}, and nonce-misuse analysis}
+In the submission document, the designers discuss the security of \morus against
+several attacks, including algebraic, differential, and guess-and-determine attacks,
+as well as the general design rationale of the initialization, encryption process, and finalization \cite{MORUS}.
+The main focus is on differential properties, whereas not many details are given for the other attack vectors.
+In third-party analysis,
+Mileva et~al.~\cite{balkancryptsecMilevaDV15} propose a distinguisher for the nonce-reuse setting and practically evaluate the differential behaviour of toy variants of \morus.
+Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discussed the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \morus[640] is well beyond the security claim.
+Dwivedi et~al.~\cite{secryptDwivediMW17} later proposed key-recovery attacks for \morus[1280] if the initialization is reduced to 3.6 out of 16 rounds and discussed the security of \morus against internal differentials and rotational cryptanalysis.
+Salam et~al.~\cite{trustcomSalamSBDPW17} applied cube attacks to obtain distinguishers for up to 5 out of 16 rounds of the initialization of \morus[1280] with negligible complexity.
+Additionally, 
+Kales et~al.~\cite{cryptoeprint:2017:1137} and
+Vaudenay and Viz\'{a}r \cite{cryptoeprint:2017:1147}
+independently proposed state-recovery and forgery attacks on \morus in a nonce-misuse setting with negligible data and time complexities.
 

### 2018-05-09 morusAC_01_Introduction.tex
bib, macros

@@
 where $(C, T)$ additionally depends on a unique nonce $N$ (or initialization value IV) and optional associated metadata $A$.
-One of the most prominent standardized AEAD designs is \cipher{AES-GCM} \cite{TODOMV05,TODONIST},
+One of the most prominent standardized AEAD designs is \cipher{AES-GCM} \cite{indocryptMcGrewV04,nistDworkin07},
 which is widely deployed in protocols such as \cipher{TLS v1.3}.
@@
 
-\todo[inline]{Related designs; Analysis of related designs; discuss \cite{sacryptMinaud14} on AEGIS \cite{AEGIS}}
+\todo[inline]{Related designs. Brice, do you want to add a sentence to discuss \cite{sacryptMinaud14} on AEGIS \cite{AEGIS}?}
 

### 2018-05-09 morusAC_01_Introduction.tex
tense consistency

@@
 Mileva et~al.~\cite{balkancryptsecMilevaDV15} propose a distinguisher for the nonce-reuse setting and practically evaluate the differential behaviour of toy variants of \morus.
-Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discussed the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \morus[640] is well beyond the security claim.
-Dwivedi et~al.~\cite{secryptDwivediMW17} later proposed key-recovery attacks for \morus[1280] if the initialization is reduced to 3.6 out of 16 rounds and discussed the security of \morus against internal differentials and rotational cryptanalysis.
-Salam et~al.~\cite{trustcomSalamSBDPW17} applied cube attacks to obtain distinguishers for up to 5 out of 16 rounds of the initialization of \morus[1280] with negligible complexity.
+Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discuss the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \morus[640] is well beyond the security claim.
+Dwivedi et~al.~\cite{secryptDwivediMW17} also propose key-recovery attacks for \morus[1280] if the initialization is reduced to 3.6 out of 16 rounds and discuss the security of \morus against internal differentials and rotational cryptanalysis.
+Salam et~al.~\cite{trustcomSalamSBDPW17} apply cube attacks to obtain distinguishers for up to 5 out of 16 rounds of the initialization of \morus[1280] with negligible complexity.
 Additionally, 
@@
 Vaudenay and Viz\'{a}r \cite{cryptoeprint:2017:1147}
-independently proposed state-recovery and forgery attacks on \morus in a nonce-misuse setting with negligible data and time complexities.
+independently propose state-recovery and forgery attacks on \morus in a nonce-misuse setting with negligible data and time complexities.
 

### 2018-05-09 morusAC_00_Abstract.tex
Reorganising abstract

--- 2018-05-08 morusAC_00_Abstract.tex
@@
 \begin{abstract}
-In this note, we show the existence of linear biases in the output of the authenticated encryption scheme MORUS.
-More precisely, we show that when encrypting a fixed plaintext multiple times, a linear correlation exists between some bits at the output of the cipher.
-Moreover the bias depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-However, the biases are quite small: for the smaller MORUS640 variant, claiming a security level of 128 bits, the bias is $2^{-73}$, and would thus require around $2^{146}$ encryptions to be exploited directly; for MORUS1280, claiming a security level of 256 bits, the bias is $2^{-76}$, corresponding to a data requirement of $2^{152}$ encryptions.
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext is encrypted enough times, regardless of whether each encryption uses a different key.
-\keywords{Linear Cryptanalysis, MORUS, CAESAR.}
+  \morus is a high-performance authenticated encryption algorithm
+  submitted to the CAESAR competition, and recently selected as a
+  finalist.  There are three version of \morus: \morus[640] with a
+  640-bit state and a 128-bit key, and \morus[1280] with a 1280-bit
+  state and a 128-bit or 256-bit key; for all versions the security
+  claim for confidentiality matches the key size.  In this paper, we
+  analyze the components of this algorithm (initialization, state
+  update and tag generation), and report several results.
+
+  Our main result is on the full state update of \morus: we found a
+  linear bias in the keystream that can be used to distinguish the
+  output from random, and to recover some plaintext bits in the
+  broadcast setting.  For \morus[1280], the bias is $2^{-76}$,
+  which can be identified after around $2^{152}$ encryptions, less than
+  would be expected for a 256-bit secure cipher.  On the other hand, for
+  \morus[640], the bias is $2^{-73}$, which does not threaten
+  the security of the cipher.
+
+% More precisely, we show that a linear correlation exists between some bits of the plaintext and ciphertext.
+Moreover the bias depends purely on the plaintext, and not on the secret key nor the nonce, hence restricting the single use of nonce or even rekeying does not prevent the attacks.
+%
+% From the nature of the linear cryptanalysis, the attack works in the known plaintext attacks.
+Moreover, if a fixed plaintext is encrypted multiple times, the correlation can be used to recover a 1-bit information of the plaintext.
+The attack principally has larger impact because it can be used to recover more unknown bits of the plaintext, provided an initial segment of the plaintext is known.
+%
+  To identify the bias, we exploit rotational symmetries in \morus
+  using linear masks that are invariants by word-rotations of the
+  state.  This motivates us to introduce word-oriented versions of
+  \morus called \minimorus, which help the
+  analysis. The attack has been implemented and verified on
+  \minimorus, having a bias of $2^{-16}$.
+
+%The bias for \cipher{MORUS1280} is slightly worse than the 4 copies of \cipher{MiniMORUS}, and the mechanism behind is explained.
+%
+%Towards the better understanding of \cipher{MORUS}, several other properties are investigated though these do not break any security claim by the designers.
+Those include the bias for \cipher{MORUS640} requiring $2^{146}$ encryption against 128-bit security, and analyses on initialization and finalization of the step-reduced version of \cipher{MORUS1280}.
+
+%In this note, we show the existence of linear biases in the output of
+%the authenticated encryption scheme MORUS. More precisely, we show
+%that when encrypting a fixed plaintext multiple times, a linear
+%correlation exists between some bits at the output of the
+%cipher. Moreover the bias depends purely on the plaintext, and not on
+%the secret key of the cipher. In principle, this property could be
+%used to recover unknown bits of a plaintext encrypted a large number
+%of times, provided an initial segment of the plaintext is
+%known. However, the biases are quite small: for the smaller MORUS640
+%variant, claiming a security level of 128 bits, the bias is
+%$2^{-73}$, and would thus require around $2^{146}$ encryptions to be
+%exploited directly; for MORUS1280, claiming a security level of 256
+%bits, the bias is $2^{-76}$, corresponding to a data requirement of
+%$2^{152}$ encryptions. We note that rekeying does not prevent the
+%attack: the biases are independent of the secret encryption key, and
+%can be exploited as long as a given plaintext is encrypted enough
+%times, regardless of whether each encryption uses a different key.
+
+  We also study reduced versions of the initialization and
+  finalization of \morus to evaluate the security margin of these
+  components.  We show a forgery attack when the finalization is
+  reduced from 10 rounds to 3, and a key-recovery attack in the
+  nonce-misue setting when the initialization is reduced from 16 to 10
+  steps.  These additional results do not threaten the full \morus but
+  studying all aspects of the design is useful to understand its
+  strengths and weaknesses.
+
+\keywords{MORUS, CAESAR, Authenticated Encryption, Nonce Respect, Linear Cryptanalysis, Confidentiality.}
 \end{abstract}
+
+%%% Local Variables:
+%%% TeX-master: "morusAC"
+%%% End:

### 2018-05-09 morusAC_01_Introduction.tex
...
Conclusion

@@
 One of the most prominent standardized AEAD designs is \cipher{AES-GCM} \cite{indocryptMcGrewV04,nistDworkin07},
-which is widely deployed in protocols such as \cipher{TLS v1.3}.
+which is widely deployed in protocols such as \cipher{TLS} (since v1.2).
 
@@
 In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-For \morus[1280], the \morus family member with 256-bit security, the linear bias is $2^{-73}$
+For \morus[1280] with 256-bit keys, the linear bias is $2^{-73}$
 \todo{double-check bias vs. correlation everywhere}
@@
 In \autoref{sec/IniFin}, we present our results on the security of \morus with round-reduced initialization (in a nonce-misuse setting) or finalization.
-Finally, we conclude in \autoref{TODO}.
+Finally, we conclude in \autoref{sec/Conclusion}.

### 2018-05-09 morusAC_08_Conclusion.tex
Conclusion

\section{Conclusion}
\label{sec/Conclusion}

This work provides a comprehensive analysis of the components of
\morus.  In particular we show that the state update has important
weaknesses with a linear bias of $2^{-76}$ between ciphertext bits and
plaintext bits in \morus[1280].  This enables a plaintext recovery
attack in the broadcast setting, using about $2^{152}$ blocs of data.
While the amount of data required is impractical, this seems to
violate the security claims of \morus[1280] because the attack works
even if the key is refreshed regularly.  Moreover, the broadcast
setting is practically relevant, as was shown with attacks aginst RC4
as used in TLS~\cite{DBLP:conf/uss/AlFardanBPPS13}.

We have shared an earlier version of this paper with the authors of
\morus and they agree with the technical details of the keystream
bias.  However they consider that it is not a significant weakness in
practice because it requires more than $2^{64}$ ciphertexts bits.  In
the context of the CAESAR competition, we believe that certificational
attacks such as this one should be taken into account, in order to
select a portfolio of candidates that reflect the state of the art in
terms of cryptographic design.

%%% Local Variables:
%%% TeX-master: "morusAC"
%%% End:

### 2018-05-09 morusAC_07_IniFin.tex
minor fix in section 7

@@
 
-The bias discussed in the previous sections analysed the encryption part of the MORUS.
+The bias in the previous sections analysed the encryption part of the MORUS.
 In this section, for comprehensive security analysis of MORUS, we provide new attacks on reduced version of the initialization and the finalization.
 We emphasize that the results in this section do not threat any security claim by the designers.
-However, we believe that investigating all parts of the design with new approaches provides better understanding of the design and will be useful especially when the design will be tweaked in future.
+However, we believe that investigating all parts of the design with different approaches from the existing work on MORUS provides better understanding and will be useful especially when the design will be tweaked in future.
 
@@
 %Hereafter, we use {\tt typefont} denote hexadecimal numbers.
-This difference propagates during 3 steps is specified in Table~\ref{Tbl:fin_diff}.
+This difference propagates during 3 steps as specified in Table~\ref{Tbl:fin_diff}.
 
@@
 \begin{center}
-\caption{Differential Propagation through Three Steps in the Finalization.
-$i$ Five lines for round $i$ denote the state difference $S_0, \cdots, S_4$ after the the round $i$ transformation.
-Weight is a Hamming weight of the state difference and accumulated probability is a probability to satisfy the trail from the beginning.}
+\caption{Differential propagation through 3 Steps.
+Five lines for round $i$ denote the difference of $S_0, \cdots, S_4$ after the round $i$ transformation.} %Weight is a Hamming weight of the state difference and accumulated probability is a probability to satisfy the trail from the beginning.}
 \label{Tbl:fin_diff}
@@
 \begin{tabular}{ccccccc} \hline
-Round & \multicolumn{4}{c}{State Difference} & Weight & Accumulated \\
-      & & & &                                &        & probability \\ \hline
+Round & \multicolumn{4}{c}{State difference} & Weight & Accumulated probability \\ \hline
+%      & & & &                                &        & 
       & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
@@
 
-\subsubsection{Remarks.} The fact that $S_4$ updated in the last round is not used to compute a tag implies that the MORUS finalization generally includes unnecessary computations.
+\subsubsection{Remarks.} The fact that $S_4$ updated in the last round is not used to compute a tag implies that the MORUS finalization generally includes unnecessary computations with respect to security.
 It may be interesting to tweak the design such that the tag can also depend on $S_4$.
@@
 \cite{cryptoeprint:2017:1137} did not demonstrate specific attacks.
-The recovered state allows the attacker to mount universal forgery attack under the same nonce.
+The recovered state allows the attacker to mount a universal forgery attack under the same nonce.
 However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker to back track the state value to the initial state.
@@
 
-To evaluate the attack complexity, we consider the following configuration of the attack parameters.
+To evaluate the attack complexity, we consider the following parameters.
 \begin{itemize}
@@
 Suppose that we choose $|G_0|$ and $|G_1|$ to be balanced i.e. $|G_0|=|G_1|$. Then, the complexity is $$\max( 2^{|G_0|}, 2^{2|G_0| - |G_2| - x}).$$
-Two terms are balanced when $x = |G_0| - |G_2|$, namely, the number of matched bits in the middle of two functions must be equal to the number of independently guessed bits to compute $F_0$ and $F_1^{-1}$.
+Two terms are balanced when $x = |G_0| - |G_2|$.
+Hence, the number of matched bits in the middle of two functions must be greater than or equal to the number of independently guessed bits to compute $F_0$ and $F_1^{-1}$.
 
@@
 \caption{Analysis of the diffusion and matching bits over 10 steps.
-`0' and `1' denote that the state bit can be computed from a partial knowledge of $K$.
+`0' and `1' denote that the state bit can and cannot be computed from a partial knowledge of $K$, respectively.
 After the partial computations from each direction, 4 bits of $S^{-6}$ can match.}

### 2018-05-10 morusAC_01_Introduction.tex
minor

@@
+%!TEX root = morusAC.tex
+
 \section{Introduction}
@@
 \subsubsection*{Related Work.}
-In the submission document, the designers discuss the security of \morus against
-several attacks, including algebraic, differential, and guess-and-determine attacks,
-as well as the general design rationale of the initialization, encryption process, and finalization \cite{MORUS}.
-The main focus is on differential properties, whereas not many details are given for the other attack vectors.
+In the \morus submission document, the designers discuss the security of \morus against
+several attacks, including algebraic, differential, and guess-and-determine attacks.
+The main focus is on differential properties, and not many details are given for other attack vectors.
 In third-party analysis,
-Mileva et~al.~\cite{balkancryptsecMilevaDV15} propose a distinguisher for the nonce-reuse setting and practically evaluate the differential behaviour of toy variants of \morus.
+Mileva et~al.~\cite{balkancryptsecMilevaDV15} propose a distinguisher in the nonce-reuse setting and practically evaluate the differential behaviour of toy variants of \morus.
 Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discuss the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \morus[640] is well beyond the security claim.
-Dwivedi et~al.~\cite{secryptDwivediMW17} also propose key-recovery attacks for \morus[1280] if the initialization is reduced to 3.6 out of 16 rounds and discuss the security of \morus against internal differentials and rotational cryptanalysis.
+Dwivedi et~al.~\cite{secryptDwivediMW17} also propose key-recovery attacks for \morus[1280] if initialization is reduced to 3.6 out of 16 rounds, and discuss the security of \morus against internal differentials and rotational cryptanalysis.
 Salam et~al.~\cite{trustcomSalamSBDPW17} apply cube attacks to obtain distinguishers for up to 5 out of 16 rounds of the initialization of \morus[1280] with negligible complexity.
@@
 independently propose state-recovery and forgery attacks on \morus in a nonce-misuse setting with negligible data and time complexities.
-
-\todo[inline]{Related designs. Brice, do you want to add a sentence to discuss \cite{sacryptMinaud14} on AEGIS \cite{AEGIS}?}
+Finally, a keystream bias of the same nature as our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher AEGIS \cite{AEGIS}, which shares the same overall structure as \morus.
 

### 2018-05-10 morusAC_05_Full_trails.tex
minor update

@@
 \subsection{Symmetrizing the Trail}
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
+In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following exactly the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
 That is, for every $S^t_{i,j}$ (resp.

### 2018-05-10 morusAC_01_Introduction.tex
homogenized references

@@
 
-\subsubsection*{Paper Outline.}
+\subsubsection*{Article Outline.}
 This paper is organized as follows.
-We first provide a brief description of \morus in \autoref{sec/Preliminaries}.
-In \autoref{sec/introminimorus}, we introduce \minimorus, an abstraction of \morus based on a certain class of rotational symmetries.
-We analyze this simplified scheme in \autoref{sec/minitrails} and provide a ciphertext-only linear approximation with a weight of 16.
-We then extend our result to the full scheme in \autoref{sec/fulltrails}, showing a bias in the keystream over 5 rounds,
-and discuss the implications of our observation for the security of \morus in \autoref{sec/Discussion}.
-In \autoref{sec/IniFin}, we present our results on the security of \morus with round-reduced initialization (in a nonce-misuse setting) or finalization.
-Finally, we conclude in \autoref{sec/Conclusion}.
+We first provide a brief description of \morus in \Cref{sec/Preliminaries}.
+In \Cref{sec/introminimorus}, we introduce \minimorus, an abstraction of \morus based on a certain class of rotational symmetries.
+We analyze this simplified scheme in \Cref{sec/minitrails} and provide a ciphertext-only linear approximation with a weight of 16.
+We then extend our result to the full scheme in \Cref{sec/fulltrails}, showing a bias in the keystream over 5 rounds,
+and discuss the implications of our observation for the security of \morus in \Cref{sec/Discussion}.
+In \Cref{sec/IniFin}, we present our results on the security of \morus with round-reduced initialization (in a nonce-misuse setting) or finalization.
+Finally, we conclude in \Cref{sec/Conclusion}.

### 2018-05-10 morusAC_02_Preliminaries.tex
minor update
homogenized references

@@
 Each round uses the word-wise left circular shift operation, which divides a $q$-bit register value into 4 words of $w = q/4$ bits, and performs a left circular shift operation on each $w$-bit word.
-The word-wise circular shift constants $b_i$ in round $i$ are defined in Table~\ref{Tbl:rcon}.
+The word-wise circular shift constants $b_i$ in round $i$ are defined in \Cref{Tbl:rcon}.
 Left circular shifts on a whole $q$-bit register are also used.
-The register-wise circular shift constants are also listed in Table~\ref{Tbl:rcon}.
+The register-wise circular shift constants are also listed in \Cref{Tbl:rcon}.
 \begin{table}[!htb]

### 2018-05-10 morusAC_07_IniFin.tex
homogenized references

@@
+%!TEX root = morusAC.tex
+
 \section{Analysis on Initialization and Finalization of Reduced MORUS}
@@
 %Hereafter, we use {\tt typefont} denote hexadecimal numbers.
-This difference propagates during 3 steps as specified in Table~\ref{Tbl:fin_diff}.
+This difference propagates during 3 steps as specified in \Cref{Tbl:fin_diff}.
 
 Recall that each step consists of 5 rounds and the input message is absorbed to the state in rounds 2 to 5.
-The trail in Table~\ref{Tbl:fin_diff} initially does not have any difference and the same continues even after round 1.
+The trail in \Cref{Tbl:fin_diff} initially does not have any difference and the same continues even after round 1.
 Differences start to appear from round 2 and they will go through the bitwise-AND operation from round 4.
@@
 Note that the tag is only 128 LSBs, thus the number of active AND gates should be counted only in those bits.
-As shown in Table~\ref{Tbl:fin_diff}, we can have a particular tag difference $\Delta T$ with probability $2^{-88}$.
+As shown in \Cref{Tbl:fin_diff}, we can have a particular tag difference $\Delta T$ with probability $2^{-88}$.
 Thus after observing $A$ and corresponding $T$, $A\|0$ and $(T \oplus \Delta T)$ is a valid pair with probability $2^{-88}$.
@@
 It may be interesting to tweak the design such that the tag can also depend on $S_4$.
-Indeed in Table~\ref{Tbl:fin_diff}, we can observe some jump-up of the probability in the tag computation.
+Indeed in \Cref{Tbl:fin_diff}, we can observe some jump-up of the probability in the tag computation.
 This is because the non-linearly involved terms are $S_2 \cdot S_3$, and $S_3$ that was updated 2-round before has a high Hamming weight.
@@
 Those will lead to 4 matching bits after the 4-step forward computation and the 6-step backward computation.
-The analysis of the diffusion is given in Table~\ref{Tbl:ini_diffuse}.
+The analysis of the diffusion is given in \Cref{Tbl:ini_diffuse}.
 In the end, $K$ can be recovered faster than the exhaustive search by 1 bit, i.e.~with complexity $2^{127}$.
@@
 
-As can be seen in Table~\ref{Tbl:ini_diffuse}, the updated register in step $i$ is independent of the update function in step $i+1$ in the forward direction, and starts to impact from step $i+2$.
+As can be seen in \Cref{Tbl:ini_diffuse}, the updated register in step $i$ is independent of the update function in step $i+1$ in the forward direction, and starts to impact from step $i+2$.
 By modifying this point, the diffusion speed can increase faster, which makes this attack harder.

### 2018-05-10 morusAC_08_Conclusion.tex
homogenized references

@@
+%!TEX root = morusAC.tex
+
 \section{Conclusion}
@@
 attacks such as this one should be taken into account, in order to
-select a portfolio of candidates that reflect the state of the art in
+select a portfolio of candidates that reflects the state of the art in
 terms of cryptographic design.

### 2018-05-10 morusAC_00_Abstract.tex
homogenized references
minor

@@
+%!TEX root = morusAC.tex
+
 \begin{abstract}
@@
   submitted to the CAESAR competition, and recently selected as a
-  finalist.  There are three version of \morus: \morus[640] with a
+  finalist.  There are three versions of \morus: \morus[640] with a
   640-bit state and a 128-bit key, and \morus[1280] with a 1280-bit
@@
   To identify the bias, we exploit rotational symmetries in \morus
-  using linear masks that are invariants by word-rotations of the
+  using linear masks that are invariant by word-rotations of the
   state.  This motivates us to introduce word-oriented versions of
@@
   analysis. The attack has been implemented and verified on
-  \minimorus, having a bias of $2^{-16}$.
+  \minimorus, where it yields a bias of $2^{-16}$.
 
@@
   nonce-misue setting when the initialization is reduced from 16 to 10
-  steps.  These additional results do not threaten the full \morus but
+  steps.  These additional results do not threaten full \morus, but
   studying all aspects of the design is useful to understand its

### 2018-05-10 src-pyth-sse/Ciphers/full_morus_1280.c
add README in the src-pyth and src-pyth-sse implementations

@@ -2,13 +2,6 @@
 
 #ifdef __AVX2__
 
-/*define data alignment for different C compilers*/
-#if defined(__GNUC__)
-#define DATA_ALIGN32(x) x __attribute__ ((aligned(32)))
-#else
-#define DATA_ALIGN32(x) __declspec(align(32)) x
-#endif
-
 #ifdef _MSC_VER
 #define  inline __inline
 #endif
@@ -29,12 +22,6 @@
 #define ROTL_256_192(x)    _mm256_permute4x64_epi64((x), _MM_SHUFFLE(0,3,2,1))  /*Rotate x by 192-bit positions to the left*/
 
 void morus_1280_print_state(state state) {
-	// printf("--------------------------------------------\n");
-	// printf("   S0   |   S1   |   S3   |   S3   |   S4   \n");
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f64[3], state[1].f64[3], state[2].f64[3], state[3].f64[3], state[4].f64[3]);
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f64[2], state[1].f64[2], state[2].f64[2], state[3].f64[2], state[4].f64[2]);
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f64[1], state[1].f64[1], state[2].f64[1], state[3].f64[1], state[4].f64[1]);
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f64[0], state[1].f64[0], state[2].f64[0], state[3].f64[0], state[4].f64[0]);
 }
 
 void morus_1280_copy_state(state to, state from) {
@@ -46,7 +33,6 @@ void morus_1280_copy_state(state to, state from) {
 }
 
 inline void print_word(union Register *word) {
-	// printf("%08x %08x %08x %08x\n", word->f64[0], word->f64[1], word->f64[2], word->f64[3]);
 }
 
 void morus_1280_print(state *saved_state, union Register *saved_cipher) {

### 2018-05-10 src-pyth-sse/Ciphers/full_morus_640.c

@@ -10,7 +10,6 @@
 #define n4 22
 #define n5 13
 
-
 #define XOR_128(x,y)      _mm_xor_si128((x),(y))     /*XOR_128(x,y) = x ^ y, where x and y are two 128-bit word*/
 #define AND_128(x,y)      _mm_and_si128((x),(y))     /*AND_128(x,y) = x & y, where x and y are two 128-bit word*/
 
@@ -21,12 +20,6 @@
 #define ROTL_128_96(x)      		_mm_shuffle_epi32((x),_MM_SHUFFLE(0,3,2,1))   /*Rotate x by 96-bit positions to the left*/
 
 void morus_640_print_state(state state) {
-	// printf("--------------------------------------------\n");
-	// printf("   S0   |   S1   |   S3   |   S3   |   S4   \n");
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[3], state[1].f32[3], state[2].f32[3], state[3].f32[3], state[4].f32[3]);
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[2], state[1].f32[2], state[2].f32[2], state[3].f32[2], state[4].f32[2]);
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[1], state[1].f32[1], state[2].f32[1], state[3].f32[1], state[4].f32[1]);
-	// printf("%08x %08x %08x %08x %08x\n", state[0].f32[0], state[1].f32[0], state[2].f32[0], state[3].f32[0], state[4].f32[0]);
 }
 
 void morus_640_copy_state(state to, state from) {
@@ -38,7 +31,6 @@ void morus_640_copy_state(state to, state from) {
 }
 
 inline void print_word(union Register *word) {
-	// printf("%08x %08x %08x %08x\n", word->f32[0], word->f32[1], word->f32[2], word->f32[3]);
 }
 
 void morus_640_print(state *saved_state, union Register *saved_cipher) {

### 2018-05-10 src-pyth-sse/Utilities/aes-rng.c

@@ -186,4 +186,17 @@ uint8_t aesrand_int8_r(struct RNG_state* rng) {
   die ("AES-NI support not enabled!\n");
 }
 
+__m128i aesrand_m128i_r(struct RNG_state *rng) {
+	die ("AES-NI support not enabled!\n");
+}
+
+#ifdef __AVX2__
+
+__m256i aesrand_m256i_r(struct RNG_state *rng) {
+	die ("AES-NI support not enabled!\n");
+}
+
+#endif
+
+
 #endif

### 2018-05-10 morusAC_01_Introduction.tex
minor Section 1

@@
 independently propose state-recovery and forgery attacks on \morus in a nonce-misuse setting with negligible data and time complexities.
-Finally, a keystream bias of the same nature as our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher AEGIS \cite{AEGIS}, which shares the same overall structure as \morus.
+Finally, a keystream bias of a similar nature as our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist which shares the same overall structure as \morus, but uses a very different, \cipher{AES}-based state update function.
 
 \subsubsection*{Our Contributions.}
-We analyze the security of \morus and present a keystream distinguisher based on
+We analyze the security of \morus and present a key\-stream distinguisher based on
 linear approximations for full \morus[1280].
@@
 
-% Paragraph: Details on keystream bias
 Our main result is a linear approximation \cite{eurocryptMatsui93,eurocryptMatsuiY92} linking ciphertext bits of five consecutive encryption blocks.
@@
 This nonce-respecting attack is based on a differential analysis of the padding rule.
-The second result targets a round-reduced initialization with 10 out of 16 steps, and extends a state-recovery attack (e.g., in a nonce-misuse setting) into a key-recovery attack.
+The second result targets a round-reduced initialization with 10 out of 16 rounds, and extends a state-recovery attack (e.g., in a nonce-misuse setting) into a key-recovery attack.
 
-\subsubsection*{Article Outline.}
+\subsubsection*{Outline.}
 This paper is organized as follows.

### 2018-05-10 morusAC_00_Abstract.tex
Proofreading

@@
 %
-  To identify the bias, we exploit rotational symmetries in \morus
+  To identify this bias, we exploit rotational symmetries in \morus
   using linear masks that are invariant by word-rotations of the
-  state.  This motivates us to introduce word-oriented versions of
+  state.  This motivates us to introduce single word versions of
   \morus called \minimorus, which help the
@@
   nonce-misue setting when the initialization is reduced from 16 to 10
-  steps.  These additional results do not threaten full \morus, but
+  steps.  These additional results do not threaten the full \morus, but
   studying all aspects of the design is useful to understand its

### 2018-05-10 morusAC_01_Introduction.tex
Proofreading

@@
 \todo{double-check}
-We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key, and can be exploited as long as a given plaintext block is encrypted sufficiently often, regardless of whether each encryption uses a different key.
+We note that our attack is nonce-respecting, and that rekeying does not prevent it: the biases are independent of the secret encryption key and nonce, and can be exploited as long as a given plaintext block is encrypted sufficiently often, regardless of whether each encryption uses a different key.
 
@@
 Finally, we conclude in \Cref{sec/Conclusion}.
+
+%%% Local Variables:
+%%% TeX-master: "morusAC"
+%%% End:

### 2018-05-10 morusAC_01_Introduction.tex
minor consistency fix: MORUS

@@
 
-\morus is one of the three finalists for use-case (2), together with \cipher{OCB} and \cipher{AEGIS}.
+\MORUS is one of the three finalists for use-case (2), together with \cipher{OCB} and \cipher{AEGIS}.
 This family of authenticated ciphers by Wu and Huang \cite{MORUS} provides three main variants:
-\morus[640] with a 128-bit key and \morus[1280] with either a 128-bit or a 256-bit key.
+\MORUS[640] with a 128-bit key and \MORUS[1280] with either a 128-bit or a 256-bit key.
 The design approach is reminiscent of classical stream cipher designs and
 continuously updates a relatively large state with a few fast operations.
-\morus can be efficiently implemented in both software and hardware;
+\MORUS can be efficiently implemented in both software and hardware;
 in particular, the designers claim that the software performance even surpasses
 \cipher{AES-GCM} implementations using Intel's \cipher{AES-NI} instructions,
-and that \morus is the fastest authenticated cipher not using \cipher{AES-NI} \cite{MORUS}.
+and that \MORUS is the fastest authenticated cipher not using \cipher{AES-NI} \cite{MORUS}.
 
 \subsubsection*{Related Work.}
-In the \morus submission document, the designers discuss the security of \morus against
+In the \MORUS submission document, the designers discuss the security of \MORUS against
 several attacks, including algebraic, differential, and guess-and-determine attacks.
@@
 In third-party analysis,
-Mileva et~al.~\cite{balkancryptsecMilevaDV15} propose a distinguisher in the nonce-reuse setting and practically evaluate the differential behaviour of toy variants of \morus.
-Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discuss the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \morus[640] is well beyond the security claim.
-Dwivedi et~al.~\cite{secryptDwivediMW17} also propose key-recovery attacks for \morus[1280] if initialization is reduced to 3.6 out of 16 rounds, and discuss the security of \morus against internal differentials and rotational cryptanalysis.
-Salam et~al.~\cite{trustcomSalamSBDPW17} apply cube attacks to obtain distinguishers for up to 5 out of 16 rounds of the initialization of \morus[1280] with negligible complexity.
+Mileva et~al.~\cite{balkancryptsecMilevaDV15} propose a distinguisher in the nonce-reuse setting and practically evaluate the differential behaviour of toy variants of \MORUS.
+Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discuss the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \MORUS[640] is well beyond the security claim.
+Dwivedi et~al.~\cite{secryptDwivediMW17} also propose key-recovery attacks for \MORUS[1280] if initialization is reduced to 3.6 out of 16 rounds, and discuss the security of \MORUS against internal differentials and rotational cryptanalysis.
+Salam et~al.~\cite{trustcomSalamSBDPW17} apply cube attacks to obtain distinguishers for up to 5 out of 16 rounds of the initialization of \MORUS[1280] with negligible complexity.
 Additionally, 
@@
 Vaudenay and Viz\'{a}r \cite{cryptoeprint:2017:1147}
-independently propose state-recovery and forgery attacks on \morus in a nonce-misuse setting with negligible data and time complexities.
-Finally, a keystream bias of a similar nature as our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist which shares the same overall structure as \morus, but uses a very different, \cipher{AES}-based state update function.
+independently propose state-recovery and forgery attacks on \MORUS in a nonce-misuse setting with negligible data and time complexities.
+Finally, a keystream bias of a similar nature as our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist which shares the same overall structure as \MORUS, but uses a very different, \cipher{AES}-based state update function.
 
 \subsubsection*{Our Contributions.}
-We analyze the security of \morus and present a key\-stream distinguisher based on
-linear approximations for full \morus[1280].
-In addition, we provide results for round-reduced \morus with reduced initialization or finalization.
+We analyze the security of \MORUS and present a key\-stream distinguisher based on
+linear approximations for full \MORUS[1280].
+In addition, we provide results for round-reduced \MORUS with reduced initialization or finalization.
 
@@
 In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-For \morus[1280] with 256-bit keys, the linear bias is $2^{-73}$
+For \MORUS[1280] with 256-bit keys, the linear bias is $2^{-73}$
 \todo{double-check bias vs. correlation everywhere}
@@
 
-To obtain this result, we propose a simplified abstraction of \morus, called \minimorus, which takes advantage of certain rotational symmetries in \morus.
+To obtain this result, we propose a simplified abstraction of \MORUS, called \MiniMORUS, which takes advantage of certain rotational symmetries in \MORUS.
 We observe that the actual bias of the approximation is better than suggested by the construction of the characteristics.
 To confirm the validity of our analysis,
-we practically verified the bias of the linear approximation for \minimorus and of partial approximations for \morus.
+we practically verified the bias of the linear approximation for \MiniMORUS and of partial approximations for \MORUS.
 In addition, we symbolically evaluated the relevant equations.
 
-In addition to this attack on full \morus, we also analyze the security of \morus with round-reduced initialization (in a nonce-misuse setting) or finalization.
-While this analysis does not threaten full \morus, it complements the main result to provide a better understanding of the \morus design approach.
-We present a forgery attack for round-reduced \morus[1280] with success probability $2^{-88}$ for a 128-bit tag if the finalization is reduced to 3 out of 10 rounds.
+In addition to this attack on full \MORUS, we also analyze the security of \MORUS with round-reduced initialization (in a nonce-misuse setting) or finalization.
+While this analysis does not threaten full \MORUS, it complements the main result to provide a better understanding of the \MORUS design approach.
+We present a forgery attack for round-reduced \MORUS[1280] with success probability $2^{-88}$ for a 128-bit tag if the finalization is reduced to 3 out of 10 rounds.
 This nonce-respecting attack is based on a differential analysis of the padding rule.
@@
 This paper is organized as follows.
-We first provide a brief description of \morus in \Cref{sec/Preliminaries}.
-In \Cref{sec/introminimorus}, we introduce \minimorus, an abstraction of \morus based on a certain class of rotational symmetries.
+We first provide a brief description of \MORUS in \Cref{sec/Preliminaries}.
+In \Cref{sec/introminimorus}, we introduce \MiniMORUS, an abstraction of \MORUS based on a certain class of rotational symmetries.
 We analyze this simplified scheme in \Cref{sec/minitrails} and provide a ciphertext-only linear approximation with a weight of 16.
 We then extend our result to the full scheme in \Cref{sec/fulltrails}, showing a bias in the keystream over 5 rounds,
-and discuss the implications of our observation for the security of \morus in \Cref{sec/Discussion}.
-In \Cref{sec/IniFin}, we present our results on the security of \morus with round-reduced initialization (in a nonce-misuse setting) or finalization.
+and discuss the implications of our observation for the security of \MORUS in \Cref{sec/Discussion}.
+In \Cref{sec/IniFin}, we present our results on the security of \MORUS with round-reduced initialization (in a nonce-misuse setting) or finalization.
 Finally, we conclude in \Cref{sec/Conclusion}.

### 2018-05-10 morusAC_02_Preliminaries.tex
minor consistency fix: MORUS

@@
 
-MORUS \cite{MORUS} is a family of authenticated ciphers with two internal state sizes: 640 and 1280 bits.
-Each member is called \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
+\MORUS \cite{MORUS} is a family of authenticated ciphers with two internal state sizes: 640 and 1280 bits.
+Each member is called \MORUS[640] and \MORUS[1280], respectively.
 
-An instance of MORUS is parametrized by a secret key.
+An instance of \MORUS is parametrized by a secret key.
 During encryption, it takes as input a plaintext message and possibly some associated data, and outputs a ciphertext together with an authentication tag.
-In the recommended parameters, \cipher{MORUS640} supports 128-bit keys and \cipher{MORUS1280} supports both 128-bit and 256-bit keys.
+In the recommended parameters, \MORUS[640] supports 128-bit keys and \MORUS[1280] supports both 128-bit and 256-bit keys.
 The tag size is 128 bits or shorter.
@@
 
-\subsection{Specification of MORUS}
+\subsection{Specification of \MORUS}
 \label{subsec/Spec}
 
-The internal state of MORUS is composed of five registers.
-Let $q$ be the register size, where $q = 128$ for \cipher{MORUS640} and $q=256$ for \cipher{MORUS1280}.
+The internal state of \MORUS is composed of five registers.
+Let $q$ be the register size, where $q = 128$ for \MORUS[640] and $q=256$ for \MORUS[1280].
 Let $S_i, i \in\{0,1,2,3,4\}$ be a register.
-Then the internal state of MORUS may be represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
+Then the internal state of \MORUS may be represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
 Registers are themselves divided into four $q/4$-bit \emph{words}.
-Throughout the paper, we denote the word size by $w = q/4$, i.e. $w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
+Throughout the paper, we denote the word size by $w = q/4$, i.e. $w=32$ for \MORUS[640] and $w=64$ for \MORUS[1280].
 
-The encryption process of MORUS consists of four parts: initialization, associated data processing, encryption, and finalization. 
+The encryption process of \MORUS consists of four parts: initialization, associated data processing, encryption, and finalization. 
 During the initialization phase, the value of the state is initialized using a key and nonce.
@@
 
-Every part of this process relies on iterating the \StateUpdate{} function at the core of MORUS. Each call to the \StateUpdate{} function is called a step.
+Every part of this process relies on iterating the \StateUpdate{} function at the core of \MORUS. Each call to the \StateUpdate{} function is called a step.
 In order to denote the internal state of the cipher at the $t$-th step, we use a superscript to denote the step number: the internal state at step $t$ is denoted by $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is $-16$ before the initialization and $t=0$ after the initialization.
@@
 \centering
-\caption{Left Circular Shift Constants of MORUS} \label{Tbl:rcon}
+\caption{Left Circular Shift Constants of \texorpdfstring{\MORUS}{MORUS}} \label{Tbl:rcon}
 \begin{tabular}{r||ccccc|ccccc} \hline
@@
                   & \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline
-\cipher{MORUS640} &  5 & 31 &  7 & 22 & 13 & 32 &  64 &  96 &  64 & 32 \\
-\cipher{MORUS1280}& 13 & 46 & 38 &  7 &  4 & 64 & 128 & 192 & 128 & 64 \\ \hline
+\MORUS[640]  &  5 & 31 &  7 & 22 & 13 & 32 &  64 &  96 &  64 & 32 \\
+\MORUS[1280] & 13 & 46 & 38 &  7 &  4 & 64 & 128 & 192 & 128 & 64 \\ \hline
 \end{tabular}
@@
 We denote those register-wise and word-wise left circular shift by $x \lll y$ and $x \lll_W y$, respectively.
-(Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document.)
+(Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \MORUS design document.)
 
@@
 \subsubsection{Initialization.}
-The initialization of \cipher{MORUS640} consists of loading a 128-bit key $K_{128}$ and a 128-bit nonce $IV_{128}$ into the state, and running \StateUpdate{} for 16 steps.
+The initialization of \MORUS[640] consists of loading a 128-bit key $K_{128}$ and a 128-bit nonce $IV_{128}$ into the state, and running \StateUpdate{} for 16 steps.
 The key and nonce are loaded into the state as follows:
@@
 
-The initialization of \cipher{MORUS1280} is slightly different from that of \cipher{MORUS640} due to the difference in register size and the two possible key sizes:
+The initialization of \MORUS[1280] is slightly different from that of \MORUS[640] due to the difference in register size and the two possible key sizes:
 \begin{itemize}
@@
 After initialization, the associated data $A$ is processed.
-For the padding, if the last associated data block is not a full block,  it is padded to 128 bits and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280} respectively, using 0 bits.
+For the padding, if the last associated data block is not a full block,  it is padded to 128 bits and 256 bits for \MORUS[640] and \MORUS[1280] respectively, using 0 bits.
 The padded full block is then used to update the state.
@@
 
-The state is then updated with $S^{t+1} \leftarrow \StateUpdate{}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{{\it adlen}}{128} \rceil -1$ and $\ell = \lceil \frac{{\it adlen}}{256} \rceil -1$ for \cipher{MORUS640} and \cipher{MORUS1280}, respectively.
+The state is then updated with $S^{t+1} \leftarrow \StateUpdate{}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{{\it adlen}}{128} \rceil -1$ and $\ell = \lceil \frac{{\it adlen}}{256} \rceil -1$ for \MORUS[640] and \MORUS[1280], respectively.
 
 \subsubsection{Encryption.}
-At each encryption step, a 16-byte (resp. 32-byte) message block $M_t$ is used to update the state in \cipher{MORUS640} (resp. \cipher{MORUS1280}), and a ciphertext $C_t$ is produced.
+At each encryption step, a 16-byte (resp. 32-byte) message block $M_t$ is used to update the state in \MORUS[640] (resp. \MORUS[1280]), and a ciphertext $C_t$ is produced.
 
-If the last message block is not a full block, 0 bits are used to pad it to 128 and 256 bits for \cipher{MORUS640} and \cipher{MORUS1280}, respectively, and the padded full block is used to update the state.
+If the last message block is not a full block, 0 bits are used to pad it to 128 and 256 bits for \MORUS[640] and \MORUS[1280], respectively, and the padded full block is used to update the state.
 But only the partial block is encrypted.
@@
 
-In \cipher{MORUS640}, Let $u = \lceil \frac{{\it adlen}}{128} \rceil$ and $v = \lceil \frac{{\it msglen}}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$:
+In \MORUS[640], Let $u = \lceil \frac{{\it adlen}}{128} \rceil$ and $v = \lceil \frac{{\it msglen}}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$:
 \begin{align*}
@@
 
-For \cipher{MORUS1280}, $u = \lceil \frac{{\it adlen}}{256} \rceil$ and $v = \lceil \frac{{\it msglen}}{256} \rceil$.
+For \MORUS[1280], $u = \lceil \frac{{\it adlen}}{256} \rceil$ and $v = \lceil \frac{{\it msglen}}{256} \rceil$.
 The ciphertext block $C^t$ is computed as $C^t \leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 192) \oplus (S^{u+t}_2 \cdot S^{u+t}_3)$.
@@
 \begin{enumerate}
-\item ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}$ for \cipher{MORUS640} and ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}\|0^{128}$ for \cipher{MORUS1280}, where ${\it adlen}$ and ${\it msglen}$ are represented as 64-bit integers.
+\item ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}$ for \MORUS[640] and ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}\|0^{128}$ for \MORUS[1280], where ${\it adlen}$ and ${\it msglen}$ are represented as 64-bit integers.
 \item $S^{u+v}_4 \leftarrow S^{u+v}_4 \oplus S^{u+v}_0.$
@@
 %\end{equation*}
-\item For \cipher{MORUS640}, $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$.
-For \cipher{MORUS1280}, $S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ is computed and the least significant 128 bits are produced as a tag $T$.
+\item For \MORUS[640], $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$.
+For \MORUS[1280], $S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ is computed and the least significant 128 bits are produced as a tag $T$.
 \end{enumerate}
@@
 \subsubsection{Security.}
-With a 128-bit tag, integrity is claimed up to 128 bits and confidentiality is claimed up to the number of bits of the key.
+With a 128-bit tag, integrity is claimed up to 128 bits and confidentiality is claimed up to the number of bits of the key. See Table~\ref{Tbl/security} for summary.
 \begin{table}[!htb]
 \centering
-\caption{Security Goals of MORUS.}
+\caption{Security goals of \MORUS.} \label{Tbl/security}
 \begin{tabular}{ccc}\hline
                         & Confidentiality (bits) & Integrity (bits) \\ \hline
-\cipher{MORUS640}-128   & 128                    & 128              \\
-\cipher{MORUS1280}-128  & 128                    & 128              \\
-\cipher{MORUS1280}-256  & 256                    & 128              \\ \hline
+\MORUS[640-128]   & 128                    & 128              \\
+\MORUS[1280-128]  & 128                    & 128              \\
+\MORUS[1280-256]  & 256                    & 128              \\ \hline
 \end{tabular}
@@
 }
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \cipher{MORUS640} and 256 for \cipher{MORUS1280}.
+In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \MORUS[640] and 256 for \MORUS[1280].
 
@@
 %Finally, let $x$ denote a register.
-Recall that registers of \cipher{MORUS} are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \cipher{MORUS} design document):\\
+Recall that registers of \MORUS are composed of four words.
+We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \MORUS design document):\\
 %\centerline{
@@
 %
-%The $b_i$'s denote rotation constants, whose value depends on which version of \cipher{MORUS} is considered.
-Throughout, $w$ denotes the  word size, i.e.
-$w=32$ for \cipher{MORUS640} and $w=64$ for \cipher{MORUS1280}.
+%The $b_i$'s denote rotation constants, whose value depends on which version of \MORUS is considered. Throughout, $w$ denotes the  word size, i.e. $w=32$ for \MORUS[640] and $w=64$ for \MORUS[1280].

### 2018-05-10 morusAC_03_MiniMorus.tex
propose introduction (poorly written)
homogenized references
minor consistency fix: MORUS

@@
-\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
-\label{sec:introminimorus}
+%!TEX root = morusAC.tex
+
+\section{Symmetric Linear Combinations and \MiniMORUS}
+\label{sec/introminimorus}
 
@@
 \end{definition}
-\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
+\emph{Example.} The following linear combination is symmetric for \MORUS[640], i.e. $w = 32$:
 \begin{equation}
@@
 
-\subsection{\cipher{MiniMORUS}}
+\subsection{\MiniMORUS}
 
-In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}.
-We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively.
-Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
+In fact, we can go further and consider a reduced version of \MORUS where each register contains a single word instead of four.
+The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see \Cref{fig:minimorus}.
+We call these reduced versions \MiniMORUS[640] and \MiniMORUS[1280], for \MORUS[640] and \MORUS[1280] respectively.
+Since registers in \MiniMORUS contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
 
-Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.
+Because the trail we are going to build is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous symmetry.
 
@@
   \centering
+  \tikzsetnextfilename{minimorus_plain}
   \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
@@
   \end{tikzpicture}%}}}
-  \caption{\cipher{MiniMORUS} state update function.}
+  \caption{\MiniMORUS state update function.}
   \label{fig:minimorus}

### 2018-05-10 morusAC_04_Mini_trails.tex
working on introduction
homogenized references
minor consistency fix: MORUS

@@
-\section{Linear Trail for \cipher{MiniMORUS}}
+%!TEX root = morusAC.tex
+
+\section{Linear Trail for \MiniMORUS}
 \label{sec/minitrails}
 
-In this section, we describe how we build a trail for \cipher{MiniMORUS} before computing its bias and experimentally verify it.
+In this section, we describe how we build a trail for \MiniMORUS before computing its bias and experimentally verify it.
 
@@
 
-To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
+To build a linear trail on \MiniMORUS, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
 \begin{itemize}
@@
   \centering
+  \tikzsetnextfilename{minimorus_part_alpha}
   \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
@@
   \centering
+  \tikzsetnextfilename{minimorus_part_beta}
   \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
@@
   \begin{subfigure}{.32\textwidth}
+  \tikzsetnextfilename{minimorus_part_gamma}
   \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
@@
   \begin{subfigure}{.32\textwidth}
+  \tikzsetnextfilename{minimorus_part_delta}
   \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
@@
   \begin{subfigure}{.32\textwidth}
+  \tikzsetnextfilename{minimorus_part_epsil}
   \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
@@
     \begin{align*}
-      \intertext{\cipher{MiniMORUS-640}:}
+      \intertext{\MiniMORUS[640]:}
 \alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
@@
 \varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\cipher{MiniMORUS-1280}:}
+      \intertext{\MiniMORUS[1280]:}
 \alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
@@
     \end{align*}
-    \caption*{\cipher{MiniMORUS} instances}
+    \caption*{\MiniMORUS instances}
   \end{subfigure}
-  \caption{\cipher{MiniMORUS} linear trail fragments.}
+  \caption{\MiniMORUS linear trail fragments.}
   \label{fig:minimorusfragments}
@@
 Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \cipher{MiniMORUS1280}, this yields the equation:
+Ultimately, for \MiniMORUS[1280], this yields the equation:
 \begin{flalign*}
@@
 
-The equation for \cipher{MiniMORUS640} is very similar, and is given in \Cref{sec:traileq}.
+The equation for \MiniMORUS[640] is very similar, and is given in \Cref{sec:traileq}.
 
@@
 
-In the equation for \cipher{MiniMORUS1280} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
+In the equation for \MiniMORUS[1280] from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
 By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
 Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
-The same holds for \cipher{MiniMORUS640} (cf.
+The same holds for \MiniMORUS[640] (cf.
 \Cref{sec:traileq}).
@@
 
-To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS1280} and \cipher{MiniMORUS640}.
+To confirm that our analysis is correct, we ran experiments on an implementation of \MiniMORUS[1280] and \MiniMORUS[640].
 In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
 Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}.
+While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \MORUS[640].
 The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
@@
 %\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
+%2 approximations for $S^{2,2}_{0}$ of \MiniMORUS[640]:
 %\begin{align*}
@@
 %\noindent
-%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
+%2 approximations for $S^{2,2}_{0}$ of \MiniMORUS[1280]:
 %\begin{align*}
@@
     \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
+    \multicolumn{2}{@{}l}{Approximations for \MiniMORUS[640]}
              & {Exp.} & {Bool.} & {Meas.} \\
@@
     \midrule
-    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
+    \multicolumn{2}{@{}l}{Approximations for \MiniMORUS[1280]} \\
     \midrule

### 2018-05-10 morusAC_05_Full_trails.tex
homogenized references
Proofreading
minor consistency fix: MORUS

--- 2018-05-09 morusAC_05_Full_trails.tex
@@
-\section{Trail for Full \cipher{MORUS}}
+%!TEX root = morusAC.tex
+
+\section{Trail for Full \MORUS}
 \label{sec/fulltrails}
 
-In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS1280} and \cipher{MiniMORUS640}.
-We now turn to the full ciphers \cipher{MORUS1280} and \cipher{MORUS640}.
+In the previous section, we have exhibited a linear trail on the reduced ciphers \MiniMORUS[1280] and \MiniMORUS[640]. We now turn to the full ciphers \MORUS[1280] and \MORUS[640].
 
 \subsection{Symmetrizing the Trail}
-In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
+In order to build a trail on the full \MORUS, we proceed exactly like we did for \MiniMORUS, following exactly the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
 That is, for every $S^t_{i,j}$ (resp.
@@
 $C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=64$ (for \cipher{MORUS1280}), the term $S^3_{2,0}$ is symmetrized into:
+For example, if $w=64$ (for \MORUS[1280]), the term $S^3_{2,0}$ is symmetrized into:
 \[
@@
 
-Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
-Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \cipher{MiniMORUS1280} given in \Cref{sec:minitraileq}, namely:
+Thus, translating the trail from one of the \MiniMORUS ciphers to the corresponding full \MORUS cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \MiniMORUS in the first place.
+Concretely, in order to build the full trail equation for \MORUS, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
+In this way, the biased linear combination on \MiniMORUS[1280] given in \Cref{sec:minitraileq}, namely:
 \begin{align*}
@@
 \end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS1280}:
+ultimately yields the following biased symmetrized linear combination on the full \MORUS[1280]:
 \begin{align*}
@@
 \end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS640}.
+We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \MORUS[640].
 
@@
 
-The symmetrized trail on full \cipher{MORUS} may be intuitively understood as consisting of four copies of the original trail on \cipher{MiniMORUS}.
-Indeed, the only difference between full \cipher{MORUS} (for either version of \cipher{MORUS}) and four independent copies of \cipher{MiniMORUS} comes from register-wise rotations, which permute words within a register.
+The symmetrized trail on full \MORUS may be intuitively understood as consisting of four copies of the original trail on \MiniMORUS.
+Indeed, the only difference between full \MORUS (for either version of \MORUS) and four independent copies of \MiniMORUS comes from register-wise rotations, which permute words within a register.
 But as observed in \Cref{sec/introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
 
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS1280} and \cipher{MORUS640}.
+Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \MiniMORUS trail, namely 64 for both \MORUS[1280] and \MORUS[640].
 However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
@@
 To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \cipher{MiniMORUS} these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
-However in the real \cipher{MORUS} these terms are of course distinct and do not cancel each other.
+Then when we XOR the various trail fragments together, in \MiniMORUS these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
+However in the real \MORUS these terms are of course distinct and do not cancel each other.
 
-In the actual trail for (either version of) full \cipher{MORUS}, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
+In the actual trail for (either version of) full \MORUS, this exact situation occurs when combining trail fragments $\beta^t_i$ and $\gamma^t_i$.
 Indeed, $\beta^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i}$, while $\gamma^t_i$ requires approximating the term $S^t_{2,i} \cdot S^t_{3,i-w}$ (cf.
 \Cref{fig:trailcollision}).
-While in \cipher{MiniMORUS}, these terms cancel out, in the full \cipher{MORUS}, when adding all symmetric copies of the trail, we end up with the sum:
+While in \MiniMORUS, these terms cancel out, in the full \MORUS, when adding all symmetric copies of the trail, we end up with the sum:
 \begin{align}
@@
   \centering
-  \begin{subfigure}{.32\textwidth}
+  \begin{subfigure}{.4\textwidth}
   \centering
@@
   \end{tikzpicture}%}}}
-  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
+  \caption*{\MiniMORUS: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
   \end{subfigure}
   \qquad
-  \begin{subfigure}{.32\textwidth}
+  \begin{subfigure}{.4\textwidth}
   \centering
@@
   \end{tikzpicture}%}}}
-  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
+  \caption*{\MORUS: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
   \end{subfigure}
-  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
+  \caption{Weight of $\beta^t_i \oplus \gamma^t_i$ for \MiniMORUS and \MORUS.}
   \label{fig:trailcollision}
@@
 
-In the end, for \cipher{MORUS1280}, since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
-This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \cipher{MiniMORUS1280} is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \cipher{MiniMORUS} in \Cref{sec:minibias}.
+In the end, for \MORUS[1280], since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
+This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \MiniMORUS[1280] is 16).
+We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \MiniMORUS in \Cref{sec:minibias}.
 That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \cipher{MORUS1280}.
+In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \MORUS[1280].
 
-In the case of \cipher{MORUS640}, collisions between rotation constants further complicate the analysis.
+In the case of \MORUS[640], collisions between rotation constants further complicate the analysis.
 Specifically, when using trail fragment $\beta^t_i$, the term $S^t_{2,i} \cdot S^t_{3,i}$ occurs.
 As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \cipher{MORUS640} is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
+However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \MORUS[640] is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
 This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \cipher{MORUS640}, we get that the weight of the trail is 73 (cf.
+In the end, when computing the full trail equation on \MORUS[640], we get that the weight of the trail is 73 (cf.
 \Cref{sec:traileq}).
@@
       \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
+      \draw (\S{0}{2}) node[below] {\normalsize \MiniMORUS[640]};
     \end{scope}
@@
       \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
+      \draw (\S{5}{2}) node[above] {\normalsize \MiniMORUS[1280]};
     \end{scope}
@@
 
-  \caption{\cipher{MiniMORUS}: two approximations for $S^{2,2}_0$.}
+  \caption{\MiniMORUS: two approximations for $S^{2,2}_0$.}
   \label{fig:twoapproximations}

### 2018-05-10 morusAC_07_IniFin.tex
minor consistency fix: MORUS

@@
 
-\section{Analysis on Initialization and Finalization of Reduced MORUS}
+\section{Analysis on Initialization and Finalization of Reduced \MORUS}
 \label{sec/IniFin}
 
-The bias in the previous sections analysed the encryption part of the MORUS.
-In this section, for comprehensive security analysis of MORUS, we provide new attacks on reduced version of the initialization and the finalization.
+The bias in the previous sections analysed the encryption part of the \MORUS.
+In this section, for comprehensive security analysis of \MORUS, we provide new attacks on reduced version of the initialization and the finalization.
 We emphasize that the results in this section do not threat any security claim by the designers.
-However, we believe that investigating all parts of the design with different approaches from the existing work on MORUS provides better understanding and will be useful especially when the design will be tweaked in future.
+However, we believe that investigating all parts of the design with different approaches from the existing work on \MORUS provides better understanding and will be useful especially when the design will be tweaked in future.
 
@@
 \label{subsec/Fin}
-We present forgery attacks on 3 out of 10 steps of \cipher{MORUS1280} that claims 128-bit security for integrity.
+We present forgery attacks on 3 out of 10 steps of \MORUS[1280] that claims 128-bit security for integrity.
 The attack only works for a limited number of steps, while it works in the nonce-respect setting.
-As far as we know, this is the first attempt to evaluate integrity of MORUS in the nonce-respect setting.
+As far as we know, this is the first attempt to evaluate integrity of \MORUS in the nonce-respect setting.
 
@@
 A general strategy of forgery attacks in the nonce-respect setting is to inject some difference in a message block and to propagate the difference so that it can be canceled by difference in another message block.
-However this approach does not work well against MORUS because its large state size prevents the attackers from controlling the difference to be canceled by another message difference.
+However this approach does not work well against \MORUS because its large state size prevents the attackers from controlling the difference to be canceled by another message difference.
 
@@
 Note that if the attacker uses different messages $M,M'$, not only the new tag $T'$ but also new ciphertext $C'$ must be guessed correctly.
-Because the encryption of MORUS is a simple XOR of the key stream, $C'$ can be easily guessed.
+Because the encryption of \MORUS is a simple XOR of the key stream, $C'$ can be easily guessed.
 For this purpose, the attacker should first query a longer message $M'=M\|0^*$ to obtain $C'$.
@@
 \subsubsection{Differential Trails.}
-Recall that the message input during the finalization of \cipher{MORUS1280} is $adlen\|msglen\|0^{128}$ where $adlen$ and $msglen$ are 64-bit strings.
+Recall that the message input during the finalization of \MORUS[1280] is $adlen\|msglen\|0^{128}$ where $adlen$ and $msglen$ are 64-bit strings.
 We set $\Delta adlen$ to be low Hamming weight, e.g.~{\tt 0x0000000000000001}.
@@
 
-\subsubsection{Remarks.} The fact that $S_4$ updated in the last round is not used to compute a tag implies that the MORUS finalization generally includes unnecessary computations with respect to security.
+\subsubsection{Remarks.} The fact that $S_4$ updated in the last round is not used to compute a tag implies that the \MORUS finalization generally includes unnecessary computations with respect to security.
 It may be interesting to tweak the design such that the tag can also depend on $S_4$.
@@
 \label{subsec/Ini}
-Kales et al.~\cite{cryptoeprint:2017:1137} showed that internal state of \cipher{MORUS640} can be recovered under the nonce-misuse scenario using $2^5$ plaintext-ciphertext pairs.
-As claimed by \cite{cryptoeprint:2017:1137} the attack is naturally extended to \cipher{MORUS1280} though Kales et al.
+Kales et al.~\cite{cryptoeprint:2017:1137} showed that internal state of \MORUS[640] can be recovered under the nonce-misuse scenario using $2^5$ plaintext-ciphertext pairs.
+As claimed by \cite{cryptoeprint:2017:1137} the attack is naturally extended to \MORUS[1280] though Kales et al.
 \cite{cryptoeprint:2017:1137} did not demonstrate specific attacks.
@@
 However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker to back track the state value to the initial state.
-In this section, we show that meet-in-the-middle attacks can allow the attacker to recover the key faster than the exhaustive search for a relatively large number of steps, i.e.~10 out of 16 steps in \cipher{MORUS1280}.
+In this section, we show that meet-in-the-middle attacks can allow the attacker to recover the key faster than the exhaustive search for a relatively large number of steps, i.e.~10 out of 16 steps in \MORUS[1280].
 

### 2018-05-10 morusAC_08_Conclusion.tex
minor consistency fix: MORUS

@@
 This work provides a comprehensive analysis of the components of
-\morus.  In particular we show that the state update has important
+\MORUS.  In particular we show that the state update has important
 weaknesses with a linear bias of $2^{-76}$ between ciphertext bits and
-plaintext bits in \morus[1280].  This enables a plaintext recovery
+plaintext bits in \MORUS[1280].  This enables a plaintext recovery
 attack in the broadcast setting, using about $2^{152}$ blocs of data.
 While the amount of data required is impractical, this seems to
-violate the security claims of \morus[1280] because the attack works
+violate the security claims of \MORUS[1280] because the attack works
 even if the key is refreshed regularly.  Moreover, the broadcast
@@
 We have shared an earlier version of this paper with the authors of
-\morus and they agree with the technical details of the keystream
+\MORUS and they agree with the technical details of the keystream
 bias.  However they consider that it is not a significant weakness in

### 2018-05-10 morusAC_01_Introduction.tex
minor Section 2 and rounds vs. steps

@@
 Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discuss the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \MORUS[640] is well beyond the security claim.
-Dwivedi et~al.~\cite{secryptDwivediMW17} also propose key-recovery attacks for \MORUS[1280] if initialization is reduced to 3.6 out of 16 rounds, and discuss the security of \MORUS against internal differentials and rotational cryptanalysis.
-Salam et~al.~\cite{trustcomSalamSBDPW17} apply cube attacks to obtain distinguishers for up to 5 out of 16 rounds of the initialization of \MORUS[1280] with negligible complexity.
+Dwivedi et~al.~\cite{secryptDwivediMW17} also propose key-recovery attacks for \MORUS[1280] if initialization is reduced to 3.6 out of 16 steps, and discuss the security of \MORUS against internal differentials and rotational cryptanalysis.
+Salam et~al.~\cite{trustcomSalamSBDPW17} apply cube attacks to obtain distinguishers for up to 5 out of 16 steps of the initialization of \MORUS[1280] with negligible complexity.
 Additionally, 
@@
 While this analysis does not threaten full \MORUS, it complements the main result to provide a better understanding of the \MORUS design approach.
-We present a forgery attack for round-reduced \MORUS[1280] with success probability $2^{-88}$ for a 128-bit tag if the finalization is reduced to 3 out of 10 rounds.
+We present a forgery attack for round-reduced \MORUS[1280] with success probability $2^{-88}$ for a 128-bit tag if the finalization is reduced to 3 out of 10 steps.
 This nonce-respecting attack is based on a differential analysis of the padding rule.
-The second result targets a round-reduced initialization with 10 out of 16 rounds, and extends a state-recovery attack (e.g., in a nonce-misuse setting) into a key-recovery attack.
+The second result targets a round-reduced initialization with 10 out of 16 steps, and extends a state-recovery attack (e.g., in a nonce-misuse setting) into a key-recovery attack.
 
@@
 We analyze this simplified scheme in \Cref{sec/minitrails} and provide a ciphertext-only linear approximation with a weight of 16.
-We then extend our result to the full scheme in \Cref{sec/fulltrails}, showing a bias in the keystream over 5 rounds,
+We then extend our result to the full scheme in \Cref{sec/fulltrails}, showing a bias in the keystream over 5 steps,
 and discuss the implications of our observation for the security of \MORUS in \Cref{sec/Discussion}.

### 2018-05-10 morusAC_02_Preliminaries.tex
minor Section 2 and rounds vs. steps

@@
 
-\MORUS \cite{MORUS} is a family of authenticated ciphers with two internal state sizes: 640 and 1280 bits.
-Each member is called \MORUS[640] and \MORUS[1280], respectively.
-
-An instance of \MORUS is parametrized by a secret key.
-During encryption, it takes as input a plaintext message and possibly some associated data, and outputs a ciphertext together with an authentication tag.
-In the recommended parameters, \MORUS[640] supports 128-bit keys and \MORUS[1280] supports both 128-bit and 256-bit keys.
-The tag size is 128 bits or shorter.
-The designers strongly recommend using a 128-bit tag.
+\MORUS is a family of authenticated ciphers designed by Wu and Huang \cite{MORUS}.
+In this section, we provide a brief description of \MORUS and introduce the
+notation for linear approximations.
 
@@
 
-The internal state of \MORUS is composed of five registers.
-Let $q$ be the register size, where $q = 128$ for \MORUS[640] and $q=256$ for \MORUS[1280].
-Let $S_i, i \in\{0,1,2,3,4\}$ be a register.
-Then the internal state of \MORUS may be represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
+An instance of \MORUS is parametrized by a secret key $K$.
+During encryption, it takes as input a plaintext message $M$, a nonce $N$, and possibly some associated data $A$, and outputs a ciphertext $C$ together with an authentication tag $T$.
+
+The \MORUS family supports two internal state sizes: 640 and 1280 bits,
+referred to as \MORUS[640] and \MORUS[1280], respectively.
+Three parameter sets are recommended: \MORUS[640] supports 128-bit keys and \MORUS[1280] supports either 128-bit or 256-bit keys.
+The tag size is 128 bits or shorter.
+The designers strongly recommend using a 128-bit tag.
+
+The internal state of \MORUS is composed of five $q$-bit \emph{registers} $S_i$, $i \in\{0,1,2,3,4\}$, where $q = 128$ for \MORUS[640] and $q=256$ for \MORUS[1280].
+The internal state of \MORUS may be represented as $S_0\|S_1\|S_2\|S_3\|S_4$.
 Registers are themselves divided into four $q/4$-bit \emph{words}.
-Throughout the paper, we denote the word size by $w = q/4$, i.e. $w=32$ for \MORUS[640] and $w=64$ for \MORUS[1280].
+Throughout the paper, we denote the word size by $w = q/4$, i.e., $w=32$ for \MORUS[640] and $w=64$ for \MORUS[1280].
 

### 2018-05-10 morusAC_02_Preliminaries.tex
Proofreading

@@
 During the initialization phase, the value of the state is initialized using a key and nonce.
-Associated data block and plaintext blocks are then processed block by block.
+The associated data and the plaintext are then processed block by block.
 Then the internal state undergoes the finalization phase, which outputs the authentication tag.
@@
 
-The state is then updated with $S^{t+1} \leftarrow \StateUpdate{}(S^t, A^t)$ for $t=0,1,\ldots,\ell$ where $\ell = \lceil \frac{{\it adlen}}{128} \rceil -1$ and $\ell = \lceil \frac{{\it adlen}}{256} \rceil -1$ for \MORUS[640] and \MORUS[1280], respectively.
+The state is then updated with $S^{t+1} \leftarrow \StateUpdate{}(S^t, A^t)$ for $t=0,1,\ldots,u-1$ where $u = \lceil \frac{{\it adlen}}{128} \rceil$ and $u = \lceil \frac{{\it adlen}}{256} \rceil$ for \MORUS[640] and \MORUS[1280], respectively.
 
@@
 %The $b_i$'s denote rotation constants, whose value depends on which version of \MORUS is considered. Throughout, $w$ denotes the  word size, i.e. $w=32$ for \MORUS[640] and $w=64$ for \MORUS[1280].
+
+%%% Local Variables:
+%%% TeX-master: "morusAC"
+%%% End:

### 2018-05-10 src/main_mini_morus.c

@@ -1,11 +1,11 @@
 #include "main_mini_morus.h"
 
-inline void copy_word(state_words* to, const state_words* from)
+void copy_word(state_words* to, const state_words* from)
 {
   *to = *from;
 }
 
-inline void save_state(state* state_saved, state state, int i)
+void save_state(state* state_saved, state state, int i)
 {
   copy_word(&state_saved[i][0], &state[0]);
   copy_word(&state_saved[i][1], &state[1]);
@@ -14,7 +14,7 @@ inline void save_state(state* state_saved, state state, int i)
   copy_word(&state_saved[i][4], &state[4]);
 }
 
-inline void sample(state* saved_state, state_words* saved_cipher, int num, struct RNG_state* seed)
+void sample(state* saved_state, state_words* saved_cipher, int num, struct RNG_state* seed)
 {
   state state;
   state_words null_words;
@@ -29,7 +29,7 @@ inline void sample(state* saved_state, state_words* saved_cipher, int num, struc
   save_state(saved_state, state, i);
 }
 
-inline int linear_sample(struct RNG_state* seed)
+int linear_sample(struct RNG_state* seed)
 {
   state saved_state[6];
   state_words saved_cipher[5];
@@ -58,7 +58,7 @@ void linear_stats(unsigned long long num) {
       // printf ( "  %6d  %12d\n", tid, seed);
 
       // try using openmp to speed things up
-      #pragma omp parallel for reduction(+:bias,inbalance)
+      // #pragma omp parallel for reduction(+:bias,inbalance)
       for(i = 0 ; i < num; ++i)
       {
         res = linear_sample(seed);
@@ -74,13 +74,14 @@ void linear_stats(unsigned long long num) {
     printf("--------------------------\n");
     printf("num:  %llu\n", (num* omp_get_max_threads()));
     printf("inba: %lld\n", inbalance);
-    printf("bias: %llu\n", bias);
+    // printf("bias: %llu\n", bias);
+    printf("bias: %f\n", log2(fabs(inbalance)/(num* omp_get_max_threads())));
     printf("--------------------------\n");
 }
 
 int main(int argc, char const *argv[]) {
   long long int num = 1;
-  num <<= 33;
+  num <<= 20;
   srand(time(NULL));
   linear_stats(num);
   return 0;

### 2018-05-10 src/main_mini_morus.h

@@ -2,5 +2,6 @@
 #include <stdio.h>
 #include <time.h>
 #include <omp.h>
+#include <math.h>
 #include "mini_morus.h"
 #include "aes-rng.h"

### 2018-05-10 src/main_morus.c

@@ -1,6 +1,6 @@
 #include "main_morus.h"
 
-inline void copy_word(state_words* to, const state_words* from)
+void copy_word(state_words* to, const state_words* from)
 {
   (*to)[0] = (*from)[0];
   (*to)[1] = (*from)[1];
@@ -8,7 +8,7 @@ inline void copy_word(state_words* to, const state_words* from)
   (*to)[3] = (*from)[3];
 }
 
-inline void save_state(state* state_saved, state state, int i)
+void save_state(state* state_saved, state state, int i)
 {
   copy_word(&(state_saved[i][0]), &(state[0]));
   copy_word(&(state_saved[i][1]), &(state[1]));
@@ -17,7 +17,7 @@ inline void save_state(state* state_saved, state state, int i)
   copy_word(&(state_saved[i][4]), &(state[4]));
 }
 
-inline void sample(state* saved_state, state_words* saved_cipher, int num, struct RNG_state* seed)
+void sample(state* saved_state, state_words* saved_cipher, int num, struct RNG_state* seed)
 {
   state state;
   state_words null_words;
@@ -32,7 +32,7 @@ inline void sample(state* saved_state, state_words* saved_cipher, int num, struc
   save_state(saved_state, state, i);
 }
 
-inline int linear_sample(struct RNG_state* seed)
+int linear_sample(struct RNG_state* seed)
 {
   state saved_state[6];
   state_words saved_cipher[5];
@@ -61,7 +61,7 @@ void linear_stats(unsigned long long num) {
       // printf ( "  %6d  %12d\n", tid, seed);
 
       // try using openmp to speed things up
-      #pragma omp parallel for reduction(+:bias,inbalance)
+      // #pragma omp parallel for reduction(+:bias,inbalance)
       for(i = 0 ; i < num; ++i)
       {
         res = linear_sample(seed);
@@ -76,13 +76,14 @@ void linear_stats(unsigned long long num) {
     printf("--------------------------\n");
     printf("num:  %llu\n", (num* omp_get_max_threads()));
     printf("inba: %lld\n", inbalance);
-    printf("bias: %llu\n", bias);
+    // printf("bias: %llu\n", bias);
+    printf("bias: %f\n", log2(fabs(inbalance)/(num* omp_get_max_threads())));
     printf("--------------------------\n");
 }
 
 int main(int argc, char const *argv[]) {
   long long int num = 1;
-  num <<= 20;
+  num <<= 24;
   srand(time(NULL));
   linear_stats(num);
   return 0;

### 2018-05-10 src/main_morus.h

@@ -1,6 +1,12 @@
 #include <stdint.h>
 #include <stdio.h>
 #include <time.h>
+#ifdef _OPENMP
 #include <omp.h>
+#else
+#define omp_get_max_threads(x) 1
+#define omp_get_thread_num(x) 1
+#endif
+#include <math.h>
 #include "morus.h"
 #include "aes-rng.h"

### 2018-05-10 src/mini_morus.c

@@ -4,11 +4,11 @@ void null_word(state_words* w) {
    *w = 0;
 }
 
-inline void print_state(state state) {
+void print_state(state state) {
   printf("%08x %08x %08x %08x %08x\n", state[0], state[1], state[2], state[3], state[4]);
 }
 
-inline uint32_t rotate_left(uint32_t x, int bits)
+uint32_t rotate_left(uint32_t x, int bits)
 {
   if (bits == 0) return x;
   return (x << bits) | (x >> (32 - bits));
@@ -58,14 +58,14 @@ void encrypt(state_words* out, state_words message, state st)
   *out = message;
 }
 
-inline uint32_t gen_mask(uint32_t* mask, unsigned int i)
+uint32_t gen_mask(uint32_t* mask, unsigned int i)
 {
   *mask |= (1 << i);
   return 0;
 }
 
 // Compute the Hamming Weight
-inline uint32_t HW(uint32_t x)
+uint32_t HW(uint32_t x)
 {
     x = x - ((x >> 1) & 0x55555555);
     x = (x & 0x33333333) + ((x >> 2) & 0x33333333);
@@ -81,6 +81,7 @@ int linear(state* saved_state, state_words* saved_cipher) {
   uint32_t mask = 0;
   uint32_t res = 0;
 
+#if 0
   gen_mask(&mask,27);
   res ^= saved_cipher[0] & mask;
 
@@ -109,6 +110,13 @@ int linear(state* saved_state, state_words* saved_cipher) {
   mask = 0;
   gen_mask(&mask,19);
   res ^= saved_cipher[4] & mask;
+
+#else
+
+  res ^= saved_state[1][0] & (1<<5);
+  res ^= saved_cipher[0] & (1<<0);
+  
+#endif
 
   return 1 & HW(res);
 }

### 2018-05-10 src/morus.c

@@ -1,6 +1,6 @@
 #include "morus.h"
 
-inline void print_state(state state) {
+void print_state(state state) {
   printf("--------------------------------------------\n");
   printf("   S0   |   S1   |   S3   |   S3   |   S4   \n");
   printf("%08x %08x %08x %08x %08x\n", state[0][3], state[1][3], state[2][3], state[3][3], state[4][3]);
@@ -9,11 +9,11 @@ inline void print_state(state state) {
   printf("%08x %08x %08x %08x %08x\n", state[0][0], state[1][0], state[2][0], state[3][0], state[4][0]);
 }
 
-inline void print_word(state_words word) {
+void print_word(state_words word) {
   printf("%08x %08x %08x %08x\n", word[0], word[1], word[2], word[3]);
 }
 
-inline void print(state* saved_state, state_words* saved_cipher)
+void print(state* saved_state, state_words* saved_cipher)
 {
   int i;
   printf("-------------------------------------------- Ciphers\n");
@@ -25,13 +25,13 @@ inline void print(state* saved_state, state_words* saved_cipher)
     print_state(saved_state[i]);
   }
 }
-inline uint32_t rotate_left(uint32_t x, int bits)
+uint32_t rotate_left(uint32_t x, int bits)
 {
   if (bits == 0) return x;
   return (x << bits) | (x >> (32 - bits));
 }
 
-inline void copy_state_words(state_words out, const state_words word)
+void copy_state_words(state_words out, const state_words word)
 {
   out[0] = word[0];
   out[1] = word[1];
@@ -39,7 +39,7 @@ inline void copy_state_words(state_words out, const state_words word)
   out[3] = word[3];
 }
 
-inline void rotate_left_word(state_words word, int bits)
+void rotate_left_word(state_words word, int bits)
 {
   word[0] = rotate_left(word[0], bits);
   word[1] = rotate_left(word[1], bits);
@@ -47,7 +47,7 @@ inline void rotate_left_word(state_words word, int bits)
[…]

### 2018-05-10 morusAC_02_Preliminaries.tex
Section 2

@@
 \MORUS is a family of authenticated ciphers designed by Wu and Huang \cite{MORUS}.
+An instance of \MORUS is parametrized by a secret key $K$.
+During encryption, it takes as input a plaintext message $M$, a nonce $N$, and possibly some associated data $A$, and outputs a ciphertext $C$ together with an authentication tag $T$.
 In this section, we provide a brief description of \MORUS and introduce the
@@
 \label{subsec/Spec}
-
-An instance of \MORUS is parametrized by a secret key $K$.
-During encryption, it takes as input a plaintext message $M$, a nonce $N$, and possibly some associated data $A$, and outputs a ciphertext $C$ together with an authentication tag $T$.
 
@@
 Every part of this process relies on iterating the \StateUpdate{} function at the core of \MORUS. Each call to the \StateUpdate{} function is called a step.
-In order to denote the internal state of the cipher at the $t$-th step, we use a superscript to denote the step number: the internal state at step $t$ is denoted by $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t$ is $-16$ before the initialization and $t=0$ after the initialization.
+%In order to denote the internal state of the cipher at the $t$-th step, we use a superscript to denote the step number:
+The internal state at step $t$ is denoted by $S^t_0\|S^t_1\|S^t_2\|S^t_3\|S^t_4$, where $t = -16$ before the initialization and $t=0$ after the initialization.
 
@@
 The additional input $m^t$ is used in rounds 2 to 5, but not in round 1.
-Each round uses the word-wise left circular shift operation, which divides a $q$-bit register value into 4 words of $w = q/4$ bits, and performs a left circular shift operation on each $w$-bit word.
-The word-wise circular shift constants $b_i$ in round $i$ are defined in \Cref{Tbl:rcon}.
-Left circular shifts on a whole $q$-bit register are also used.
-The register-wise circular shift constants are also listed in \Cref{Tbl:rcon}.
-\begin{table}[!htb]
+Each round uses the word-wise rotation (left circular shift) operation,
+denoted $\lll_w$ in the following and \texttt{Rotl\_xxx\_yy} in the design document.
+It divides a $q$-bit register value into 4 words of $w = q/4$ bits, and performs a rotation on each $w$-bit word.
+The word-wise rotation constants $b_i$ for round $i$ are defined in \Cref{Tbl:rcon}.
+Additionally, each round uses rotations on a whole $q$-bit register by a multiple of the word size,
+denoted $\lll$ in the following and \texttt{<<<} in the design document.
+The register-wise rotation constants $b_i'$ are also listed in \Cref{Tbl:rcon}.
+%\todo{rounds $0\dots4$ or $1\dots5$?}
+\begin{table}%[!bht]
   \centering
-\caption{Left Circular Shift Constants of \texorpdfstring{\MORUS}{MORUS}} \label{Tbl:rcon}
-\begin{tabular}{r||ccccc|ccccc} \hline
-& \multicolumn{5}{c|}{Word-wise circular shift} & \multicolumn{5}{c}{Register-wise circular shift} \\
-                  & \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline
+  \caption{Rotation constants $b_i$ for $\lll_w$ and $b_i'$ for $\lll$ in round $i$ of \texorpdfstring{\MORUS}{MORUS}.} \label{Tbl:rcon}
+  \setlength{\tabcolsep}{6pt}
+  \begin{tabular}{@{}l@{\qquad}rrrrr rrrrr@{}}
+    \toprule
+    & \multicolumn{5}{@{}c@{~}}{Word-wise rotation $\lll_w$} & \multicolumn{5}{@{}c@{}}{Register-wise rotation $\lll$} \\
+    \cmidrule(r){2-6}
+    \cmidrule(l){7-11}
+    & $b_0$  & $b_1$  & $b_2$  & $b_3$  & $b_4$  
+    & $b_0'$ & $b_1'$ & $b_2'$ & $b_3'$ & $b_4'$  \\
+    %& \makebox[2em]{$b_0$} & \makebox[2em]{$b_1$} & \makebox[2em]{$b_2$} & \makebox[2em]{$b_3$} & \makebox[2em]{$b_4$} & \makebox[2em]{$b'_0$} & \makebox[2em]{$b'_1$} & \makebox[2em]{$b'_2$} & \makebox[2em]{$b'_3$} & \makebox[2em]{$b'_4$} \\ \hline
+    \midrule
     \MORUS[640]  &  5 & 31 &  7 & 22 & 13 & 32 &  64 &  96 &  64 & 32 \\
-\MORUS[1280] & 13 & 46 & 38 &  7 &  4 & 64 & 128 & 192 & 128 & 64 \\ \hline
+    \MORUS[1280] & 13 & 46 & 38 &  7 &  4 & 64 & 128 & 192 & 128 & 64 \\
+    \bottomrule
   \end{tabular}
@@
 
-We denote those register-wise and word-wise left circular shift by $x \lll y$ and $x \lll_W y$, respectively.
-(Those are denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \MORUS design document.)
-
-Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow \StateUpdate{}(S^t,m^t)$ is defined as follows.
+Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow \StateUpdate{}(S^t,m^t)$ is defined as follows:
 \begin{align*}
 \textrm{Round 1:} &&
-S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 ) \lll_W b_0 &
+S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 )            \lll_w b_0, &
 S^t_3 \leftarrow S^t_3 \lll b'_0.\\
 \textrm{Round 2:} &&
-S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_W b_1 &
+S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_w b_1, &
 S^t_4 \leftarrow S^t_4 \lll b'_1.\\
 \textrm{Round 3:} &&
-S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_W b_2 &
+S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_w b_2, &
 S^t_0 \leftarrow S^t_0 \lll b'_2.\\
 \textrm{Round 4:} &&
-S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_W b_3 &
+S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_w b_3, &
 S^t_1 \leftarrow S^t_1 \lll b'_3.\\
 \textrm{Round 5:} &&
-S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_W b_4 &
+S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_w b_4, &
 S^t_2 \leftarrow S^t_2 \lll b'_4.
@@
 \subsubsection{Initialization.}
-The initialization of \MORUS[640] consists of loading a 128-bit key $K_{128}$ and a 128-bit nonce $IV_{128}$ into the state, and running \StateUpdate{} for 16 steps.
-The key and nonce are loaded into the state as follows:
+The initialization of \MORUS[640] starts by loading the 128-bit key $K_{128}$ and the 128-bit nonce $N_{128}$ into the state together with constants $c_0, c_1$:
+%and running \StateUpdate{} for 16 steps. The key and nonce are loaded into the state as follows:
 \begin{align*}
-S^{-16}_0 = \textrm{IV}_{128}, &&
-S^{-16}_1 = K_{128}, &&
-S^{-16}_2 = 1^{128}, &&
-S^{-16}_3 = const_0, &&
-S^{-16}_4 = const_1.
+S^{-16}_0 &= N_{128}, &
+S^{-16}_1 &= K_{128}, &
+S^{-16}_2 &= 1^{128}, &
+S^{-16}_3 &= c_0, &
+S^{-16}_4 &= c_1.
 \end{align*}
-The internal state is then updated 16 times with  $\StateUpdate{}(S^t,0)$ for $t=-16,-15,\ldots,-1$. Finally, the key is XORed into the state again with $S^0_1 \leftarrow S^0_1 \oplus K_{128}$.
+Then, $\StateUpdate{}(S^t,0)$ is iterated 16 times for $t=-16,-15,\ldots,-1$.
+Finally, the key is XORed into the state again with $S^0_1 \leftarrow S^0_1 \oplus K_{128}$.
 
-The initialization of \MORUS[1280] is slightly different from that of \MORUS[640] due to the difference in register size and the two possible key sizes:
-\begin{itemize}
-  \item[-] when the key size is 128 bits, $K = K_{128} \|K_{128}$.
-  \item[-] when the key size is 256 bits, $K = K_{256}$.
-\end{itemize}
-The state is initialized as:
+The initialization of \MORUS[1280] differs slightly due to the difference in register size and the two possible key sizes,
+and uses either
+$K = K_{128} \|K_{128}$ (for \MORUS[1280-128]) or
+$K = K_{256}$ (for \MORUS[1280-256])
+to initialize the state:
 \begin{align*}
-S^{-16}_0 &= \textrm{IV}_{128} \| 0^{128}, &
+S^{-16}_0 &= N_{128} \mathrel\| 0^{128}, &
 S^{-16}_1 &= K, &
-S^{-16}_2 = 1^{256}, \\
+S^{-16}_2 &= 1^{256}, &
 S^{-16}_3 &= 0^{256}, &
-S^{-16}_4 &= const_0\|const_1.
+S^{-16}_4 &= c_0\mathrel\|c_1.
 \end{align*}
-After updating the state 16 times with \StateUpdate{}, the state is updated with $S^0_1 \leftarrow S^0_1 \oplus K$.
+After iterating \StateUpdate{} 16 times, the state is updated with $S^0_1 \leftarrow S^0_1 \oplus K$.
 
 \subsubsection{Associated Data Processing.}
-After initialization, the associated data $A$ is processed.
-For the padding, if the last associated data block is not a full block,  it is padded to 128 bits and 256 bits for \MORUS[640] and \MORUS[1280] respectively, using 0 bits.
-The padded full block is then used to update the state.
-If the length of $A$, denoted by `${\it adlen}$' is 0, the associated data processing phase is skipped.
-
-The state is then updated with $S^{t+1} \leftarrow \StateUpdate{}(S^t, A^t)$ for $t=0,1,\ldots,u-1$ where $u = \lceil \frac{{\it adlen}}{128} \rceil$ and $u = \lceil \frac{{\it adlen}}{256} \rceil$ for \MORUS[640] and \MORUS[1280], respectively.
+After initialization, the associated data $A$ is processed in blocks of $q \in \{128, 256\}$ bits.
+For the padding, if the last associated data block is not a full block, it is padded to $q$ bits with zeroes.
+If the length of $A$, denoted by $|A|$, is 0, then the associated data processing phase is skipped;
+else, the state is updated as
+\[S^{t+1} \leftarrow \StateUpdate{}(S^t, A^t) \qquad \text{for } t=0,1,\ldots,\lceil |A|/q\rceil-1.\]
 

### 2018-05-10 morusAC_02_Preliminaries.tex
Section 2

@@
 The designers strongly recommend using a 128-bit tag.
+%
+%\subsubsection{Security.}
+With a 128-bit tag, integrity is claimed up to 128 bits and confidentiality is claimed up to the number of key bits (\Cref{Tbl/security}).
 
+\begin{table}[b]
+\caption{Security goals of \MORUS.} \label{Tbl/security}
+\centering
+\begin{tabular}{@{}l@{\quad}c@{\qquad}c@{}}\toprule
+                        & Confidentiality (bits) & Integrity (bits) \\ \midrule
+\MORUS[640-128]   & 128                    & 128              \\
+\MORUS[1280-128]  & 128                    & 128              \\
+\MORUS[1280-256]  & 256                    & 128              \\ \bottomrule
+\end{tabular}
+\end{table}
+
+\subsubsection{State.}
 The internal state of \MORUS is composed of five $q$-bit \emph{registers} $S_i$, $i \in\{0,1,2,3,4\}$, where $q = 128$ for \MORUS[640] and $q=256$ for \MORUS[1280].
@@
 %\todo{rounds $0\dots4$ or $1\dots5$?}
-\begin{table}%[!bht]
+
+$S^{t+1} \leftarrow \StateUpdate{}(S^t,m^t)$ is defined as follows, where $\cdot$ denotes bit-wise AND, $\oplus$ is bit-wise XOR, and $m_i$ is defined depending on the context:
+\begin{align*}
+\textrm{Round 1:} &&
+S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 )            \lll_w b_0, &
+S^t_3 \leftarrow S^t_3 \lll b'_0.\\
+\textrm{Round 2:} &&
+S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_w b_1, &
+S^t_4 \leftarrow S^t_4 \lll b'_1.\\
+\textrm{Round 3:} &&
+S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_w b_2, &
+S^t_0 \leftarrow S^t_0 \lll b'_2.\\
+\textrm{Round 4:} &&
+S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_w b_3, &
+S^t_1 \leftarrow S^t_1 \lll b'_3.\\
+\textrm{Round 5:} &&
+S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_w b_4, &
+S^t_2 \leftarrow S^t_2 \lll b'_4.
+\end{align*}
+
+\begin{table}[t]%[!bht]
   \centering
@@
 \end{table}
-
-Let `$\cdot$' denote the bit-wise AND operation. $S^{t+1} \leftarrow \StateUpdate{}(S^t,m^t)$ is defined as follows:
-\begin{align*}
-\textrm{Round 1:} &&
-S^{t+1}_0 &\leftarrow ( S^t_0 \oplus (S^t_1 \cdot S^t_2) \oplus S^t_3 )            \lll_w b_0, &
-S^t_3 \leftarrow S^t_3 \lll b'_0.\\
-\textrm{Round 2:} &&
-S^{t+1}_1 &\leftarrow ( S^t_1 \oplus (S^t_2 \cdot S^t_3) \oplus S^t_4 \oplus m_i ) \lll_w b_1, &
-S^t_4 \leftarrow S^t_4 \lll b'_1.\\
-\textrm{Round 3:} &&
-S^{t+1}_2 &\leftarrow ( S^t_2 \oplus (S^t_3 \cdot S^t_4) \oplus S^t_0 \oplus m_i ) \lll_w b_2, &
-S^t_0 \leftarrow S^t_0 \lll b'_2.\\
-\textrm{Round 4:} &&
-S^{t+1}_3 &\leftarrow ( S^t_3 \oplus (S^t_4 \cdot S^t_0) \oplus S^t_1 \oplus m_i ) \lll_w b_3, &
-S^t_1 \leftarrow S^t_1 \lll b'_3.\\
-\textrm{Round 5:} &&
-S^{t+1}_4 &\leftarrow ( S^t_4 \oplus (S^t_0 \cdot S^t_1) \oplus S^t_2 \oplus m_i ) \lll_w b_4, &
-S^t_2 \leftarrow S^t_2 \lll b'_4.
-\end{align*}
 
@@
 \subsubsection{Encryption.}
-At each encryption step, a 16-byte (resp. 32-byte) message block $M_t$ is used to update the state in \MORUS[640] (resp. \MORUS[1280]), and a ciphertext $C_t$ is produced.
-
-If the last message block is not a full block, 0 bits are used to pad it to 128 and 256 bits for \MORUS[640] and \MORUS[1280], respectively, and the padded full block is used to update the state.
+Next, the message is processed in blocks $M_t$ of  $q \in \{128, 256\}$ bits to update the state and produce the ciphertext blocks $C_t$.
+If the last message block is not a full block, zero-bits are used to pad it to 128 or 256 bits for \MORUS[640] and \MORUS[1280], respectively, and the padded full block is used to update the state.
 But only the partial block is encrypted.
-Note that if the message length denoted by `${\it msglen}$' is 0, encryption is skipped.
-
-In \MORUS[640], Let $u = \lceil \frac{{\it adlen}}{128} \rceil$ and $v = \lceil \frac{{\it msglen}}{128} \rceil$. The following is performed for $t=0, 1, \ldots, v-1$:
+Note that if the message length denoted by $|M|$ is 0, encryption is skipped.
+Let $u = \lceil |A|/q \rceil$ and $v = \lceil |M|/q \rceil$. The following is performed for $t=0, 1, \ldots, v-1$:
 \begin{align*}
-C^t &\leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 96) \oplus (S^{u+t}_2 \cdot S^{u+t}_3),\\
+C^t &\leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll b_2') \oplus (S^{u+t}_2 \cdot S^{u+t}_3),\\
 S^{u+t+1} &\leftarrow \StateUpdate{}(S^{u+t},M^t).
 \end{align*}
-
-For \MORUS[1280], $u = \lceil \frac{{\it adlen}}{256} \rceil$ and $v = \lceil \frac{{\it msglen}}{256} \rceil$.
-The ciphertext block $C^t$ is computed as $C^t \leftarrow M^t \oplus S^{u+t}_0 \oplus (S^{u+t}_1 \lll 192) \oplus (S^{u+t}_2 \cdot S^{u+t}_3)$.
 
@@
 We only discuss the case where $T$ is not truncated.
-The associated data length and the message length are used to update the state as follows.
+The associated data length and the message length are used to update the state:
 \begin{enumerate}
-\item ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}$ for \MORUS[640] and ${\it tmp} \leftarrow {\it adlen}\|{\it msglen}\|0^{128}$ for \MORUS[1280], where ${\it adlen}$ and ${\it msglen}$ are represented as 64-bit integers.
+\item $L \leftarrow |A| \mathrel\| |M|$ for \MORUS[640] or $L \leftarrow |A| \mathrel\| |M| \mathrel\|0^{128}$ for \MORUS[1280], where $|A|, |M|$ are represented as 64-bit integers.
 \item $S^{u+v}_4 \leftarrow S^{u+v}_4 \oplus S^{u+v}_0.$
@@
 %\end{equation*}
-\item For \MORUS[640], $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$.
+%\item For \MORUS[640], $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 96) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$.
 For \MORUS[1280], $S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll 192) \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$ is computed and the least significant 128 bits are produced as a tag $T$.
+\item $T = S^{u+v+10}_0 \oplus (S^{u+v+10}_1 \lll b_2') \oplus ( S^{u+v+10}_2 \cdot S^{u+v+10}_3)$, or the least significant 128 bits of this value in case of \MORUS[1280].
 \end{enumerate}
-
-\subsubsection{Security.}
-With a 128-bit tag, integrity is claimed up to 128 bits and confidentiality is claimed up to the number of bits of the key. See Table~\ref{Tbl/security} for summary.
-\begin{table}[!htb]
-\centering
-\caption{Security goals of \MORUS.} \label{Tbl/security}
-\begin{tabular}{ccc}\hline
-                        & Confidentiality (bits) & Integrity (bits) \\ \hline
-\MORUS[640-128]   & 128                    & 128              \\
-\MORUS[1280-128]  & 128                    & 128              \\
-\MORUS[1280-256]  & 256                    & 128              \\ \hline
-\end{tabular}
-\end{table}
 
@@
 
-We define the \emph{bias} of an event $E$ as:
-\[
-\bias(E) \eqdef 2\Pr(E)-1.
-\]
-Note that the bias is sometimes defined as half the above value in the literature. We also define the \emph{weight} of an event as:
-\[
-\weight(E) \eqdef -\log|\bias(E)|
-\]
-where $\log()$ denotes logarithm in base 2.
-By the piling-up lemma, the bias (resp.
+In the following, we use linear approximations \cite{eurocryptMatsui93}
+that hold with probability $\Pr(E) = \frac{1}{2} + \varepsilon$, i.e., they are biased with bias $\varepsilon$.
+The \emph{correlation} $\corr(E)$ of the approximation and its \emph{weight} $\weight(E)$ are defined as
+\begin{align*}
+\corr(E) &\eqdef 2\Pr(E)-1 = 2 \varepsilon\,, \\
+%We also use the \emph{weight} of the approximation, defined as
+\weight(E) &\eqdef -\log_2|\corr(E)|\,,
+\end{align*}
+where $\log_2()$ denotes logarithm in base 2.
+By the piling-up lemma, the correlation (resp.
 weight) of an XOR of independent variables is equal to the product (resp.
-sum) of their individual biases (resp.
+sum) of their individual correlations (resp.
 weights) \cite{eurocryptMatsui93}.
@@
 }
-In the above notation, bit positions are always taken modulo the register size, i.e. 128 for \MORUS[640] and 256 for \MORUS[1280].
+In the above notation, bit positions are always taken modulo the register size $q$, i.e., $q=128$ for \MORUS[640] and $q=256$ for \MORUS[1280].
 

### 2018-05-10 morusAC_00_Abstract.tex
minor rewrites in abstract and contributions

--- 2018-05-10 morusAC_00_Abstract.tex
@@
 \begin{abstract}
-  \morus is a high-performance authenticated encryption algorithm
+  \MORUS is a high-performance authenticated encryption algorithm
   submitted to the CAESAR competition, and recently selected as a
-  finalist.  There are three versions of \morus: \morus[640] with a
-  640-bit state and a 128-bit key, and \morus[1280] with a 1280-bit
-  state and a 128-bit or 256-bit key; for all versions the security
+  finalist.  There are three versions of \MORUS: \MORUS[640] with
+  %a 640-bit state and
+  a 128-bit key, and \MORUS[1280] with
+  % a 1280-bit state and
+  a 128-bit or 256-bit key. For all versions the security
   claim for confidentiality matches the key size.  In this paper, we
@@
 
-  Our main result is on the full state update of \morus: we found a
-  linear bias in the keystream that can be used to distinguish the
-  output from random, and to recover some plaintext bits in the
-  broadcast setting.  For \morus[1280], the bias is $2^{-76}$,
-  which can be identified after around $2^{152}$ encryptions, less than
-  would be expected for a 256-bit secure cipher.  On the other hand, for
-  \morus[640], the bias is $2^{-73}$, which does not threaten
+  As our main result, we exhibit a
+  linear bias in the keystream of full \MORUS, which can be used to distinguish its output
+  from random, and to recover some plaintext bits in the
+  broadcast setting.  For \MORUS[1280], the bias is $2^{-76}$,
+  which can be exploited after around $2^{152}$ encryptions, less than
+  would be expected for a 256-bit secure cipher.  For
+  \MORUS[640], the same attack results in a bias of $2^{-73}$, which does not threaten
   the security of the cipher.
@@
 %
-  To identify the bias, we exploit rotational symmetries in \morus
+  To identify this bias, we make use of rotational symmetries in \MORUS
   using linear masks that are invariant by word-rotations of the
-  state.  This motivates us to introduce word-oriented versions of
-  \morus called \minimorus, which help the
+  state.  This motivates us to introduce single-word versions of
+  \MORUS called \MiniMORUS, which aid in the
   analysis. The attack has been implemented and verified on
-  \minimorus, where it yields a bias of $2^{-16}$.
+  \MiniMORUS, where it yields a bias of $2^{-16}$.
 
@@
   We also study reduced versions of the initialization and
-  finalization of \morus to evaluate the security margin of these
-  components.  We show a forgery attack when the finalization is
+  finalization of \MORUS, aiming to evaluate the security margin of these
+  components.  We show a forgery attack when finalization is
   reduced from 10 rounds to 3, and a key-recovery attack in the
-  nonce-misue setting when the initialization is reduced from 16 to 10
-  steps.  These additional results do not threaten full \morus, but
+  nonce-misue setting when initialization is reduced from 16 to 10
+  steps.  These additional results do not threaten full \MORUS, but
   studying all aspects of the design is useful to understand its

### 2018-05-10 morusAC_01_Introduction.tex
minor rewrites in abstract and contributions

@@
 independently propose state-recovery and forgery attacks on \MORUS in a nonce-misuse setting with negligible data and time complexities.
-Finally, a keystream bias of a similar nature as our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist which shares the same overall structure as \MORUS, but uses a very different, \cipher{AES}-based state update function.
+Finally, a keystream bias similar in nature to our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist, which shares the same overall structure as \MORUS, but uses a very different, \cipher{AES}-based state update function.
 
 \subsubsection*{Our Contributions.}
-We analyze the security of \MORUS and present a key\-stream distinguisher based on
-linear approximations for full \MORUS[1280].
-In addition, we provide results for round-reduced \MORUS with reduced initialization or finalization.
+Our main contribution is a keystream distinguisher on full \MORUS[1280], built from linear approximations of its core \StateUpdate{} function.
+In addition, we provide results for round-reduced \MORUS, targeting both the initialization or finalization phases of the cipher.
 
-Our main result is a linear approximation \cite{eurocryptMatsui93,eurocryptMatsuiY92} linking ciphertext bits of five consecutive encryption blocks.
+In more detail, our main result is a linear approximation \cite{eurocryptMatsui93,eurocryptMatsuiY92} linking ciphertext bits across five consecutive encryption blocks.
 %As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
@@
 %Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-More precisely, we show that when encrypting a fixed plaintext block multiple times, a linear correlation exists between some output bits of the cipher.
+As a consequence, when encrypting a fixed plaintext segment multiple times, a linear correlation exists between some output bits of the cipher.
 Moreover, the bias depends purely on the plaintext, and not on the secret key of the cipher.
 In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-For \MORUS[1280] with 256-bit keys, the linear bias is $2^{-73}$
+For \MORUS[1280] with 256-bit keys, the linear bias is $2^{-76}$
 \todo{double-check bias vs. correlation everywhere}
@@
 \todo{double-check}
-We note that our attack is nonce-respecting, and that rekeying does not prevent it: the biases are independent of the secret encryption key and nonce, and can be exploited as long as a given plaintext block is encrypted sufficiently often, regardless of whether each encryption uses a different key.
 
-To obtain this result, we propose a simplified abstraction of \MORUS, called \MiniMORUS, which takes advantage of certain rotational symmetries in \MORUS.
-We observe that the actual bias of the approximation is better than suggested by the construction of the characteristics.
+To the best of our knowledge, this is the first attack on full \MORUS in the nonce-respecting setting.
+We also note that rekeying does not prevent the attack: the biases are independent of the secret encryption key and nonce, and can be exploited as long as a given plaintext segment is encrypted sufficiently often, regardless of whether each encryption uses a different key.
+
+To obtain this result, we propose a simplified abstraction of \MORUS, called \MiniMORUS.
+\MiniMORUS takes advantage of certain rotational symmetries in \MORUS and simplifies the description and analysis of the attack.
+We then show how the attack can be extended from \MiniMORUS to the real \MORUS.
+%We observe that the actual bias of the approximation is better than suggested by the construction of the characteristics.
 To confirm the validity of our analysis,
-we practically verified the bias of the linear approximation for \MiniMORUS and of partial approximations for \MORUS.
-In addition, we symbolically evaluated the relevant equations.
+we practically verified the bias of the full linear trail for \MiniMORUS, as well as the bias of trail fragments for the full \MORUS.
+Our analysis is also backed by a symbolic evaluation of the full trail equation and its bias on all variants of \MORUS.
 
-In addition to this attack on full \MORUS, we also analyze the security of \MORUS with round-reduced initialization (in a nonce-misuse setting) or finalization.
-While this analysis does not threaten full \MORUS, it complements the main result to provide a better understanding of the \MORUS design approach.
-We present a forgery attack for round-reduced \MORUS[1280] with success probability $2^{-88}$ for a 128-bit tag if the finalization is reduced to 3 out of 10 steps.
+In addition to the previous attack on full \MORUS, we provide two secondary results: (1) we analyze the security of \MORUS against forgery attacks with round-reduced finalization; and (2) we analyze its security against key recovery in a nonce-misuse setting, with round-reduced initialization.
+While this extra analysis does not threaten full \MORUS, it complements the main result to provide a better overall understanding of the security of \MORUS.
+More precisely, we present a forgery attack for round-reduced \MORUS[1280] with success probability $2^{-88}$ for a 128-bit tag if the finalization is reduced to 3 out of 10 steps.
 This nonce-respecting attack is based on a differential analysis of the padding rule.
-The second result targets a round-reduced initialization with 10 out of 16 steps, and extends a state-recovery attack (e.g., in a nonce-misuse setting) into a key-recovery attack.
+The second result targets round-reduced initialization with 10 out of 16 steps, and extends a state-recovery attack (which can be mounted e.g. in a nonce-misuse setting) into a key-recovery attack.
 

### 2018-05-10 morusAC_02_Preliminaries.tex

@@
 Next, the message is processed in blocks $M_t$ of  $q \in \{128, 256\}$ bits to update the state and produce the ciphertext blocks $C_t$.
-If the last message block is not a full block, zero-bits are used to pad it to 128 or 256 bits for \MORUS[640] and \MORUS[1280], respectively, and the padded full block is used to update the state.
-But only the partial block is encrypted.
+If the last message block is not a full block, a string of 0's is used to pad it to 128 or 256 bits for \MORUS[640] and \MORUS[1280], respectively, and the padded full block is used to update the state.
+However, only the partial block is encrypted.
 Note that if the message length denoted by $|M|$ is 0, encryption is skipped.

### 2018-05-10 morusAC_06_Discussion.tex
homogenized references
minor consistency fix: MORUS
copyedit

@@
+%!TEX root = morusAC.tex
+
 \section{Discussion}
@@
 
-\todo{ME: If this doesn't grow much longer, I'd suggest to make this a subsection of Section 5, and repeat a summary in the Conclusion.}
-
 %This is just a sketch for now...
 
-We now discuss the impact of our attacks for the security of \morus.
+We now discuss the impact of these attacks on the security of \MORUS.
 
@@
 of 256 bits for confidentiality, and an attack with complexity $2^{152}$
-should violate this claim, even if it is not practical.  The
-\cipher{MORUS} specification limits the amount of plaintext encrypted
-with a given key to $2^{64}$, but this does not affect our attack
-because we can use keystream generated with different keys.
+should violate this claim, even if it is not practical.
 

### 2018-05-10 morusAC_06_Discussion.tex
Less apologetic

@@
 of 256 bits for confidentiality, and an attack with complexity $2^{152}$
-should violate this claim, even if it is not practical.
+violates this claim, even if it is not practical.
 

### 2018-05-10 morusAC_03_MiniMorus.tex

@@
 
-Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
+To build our linear trail, we start in a linear combinations of bits within a single register.
 \begin{definition}
@@
 
-Because the trail we are going to build is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous symmetry.
+Since the trail we are building is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous symmetry.
 

### 2018-05-10 morusAC_04_Mini_trails.tex
title vote winner: cryptanalysis of morus

@@
 
-In this section, we describe how we build a trail for \MiniMORUS before computing its bias and experimentally verify it.
+In this section, we describe how we build a trail for \MiniMORUS before computing its correlation and experimentally verify it.
 
@@
 
-To build a linear trail on \MiniMORUS, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
+To build a linear trail for \MiniMORUS, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
 \begin{itemize}
@@
 In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form:
+In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form (see \cite[Sec.~3.3]{DBLP:conf/indocrypt/AshurR16})
 \[
@@
 In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-In the end, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
+Eventually, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
 
@@
 The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero bias (in all cases the bias is $2^{-1}$).
+For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero correlation (in all cases the correlation is $2^{-1}$).
 \begin{alignat*}{2}
@@
 Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the bias of the right-hand side expression, we will then be able to deduce the same bias on the linear combination of ciphertext bits on the left-hand side.
+By measuring the correlation of the right-hand side expression, we are then able to determine the correlation of the linear combination of ciphertext bits on the left-hand side.
 We now set out to do so.
@@
 \]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
+In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add to the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
 This yields:
@@
 We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$.
+To do this, we add to the equations $D^t_i$ for appropriate choices of $t$ and $i$.
 This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
@@
 By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
+Finally, applying $A^t_i$ eight times allows to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
 Ultimately, for \MiniMORUS[1280], this yields the equation:
@@
 
-\subsection{Bias of the Trail}
+\subsection{Correlation of the Trail}
 \label{sec:minibias}
@@
 By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$.
+Hence the linear combination of ciphertext bits on the left-hand side has a correlation of $2^{-16}$.
 The same holds for \MiniMORUS[640] (cf.
@@
 
-That the bias is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
+That the correlation is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
 Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
@@
 When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final bias is $2^{-16}$ rather than $2^{-24}$.
+Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final correlation is $2^{-16}$ rather than $2^{-24}$.
 
@@
 To confirm that our analysis is correct, we ran experiments on an implementation of \MiniMORUS[1280] and \MiniMORUS[640].
-In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$).
+In both cases, we measured the correlation of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the correlation of the full trail itself (denoted by $\chi$).
 Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \MORUS[640].
+While our analysis from the previous section predicts a correlation of $2^{-16}$, experiments indicate a slightly better empirical correlation of $2^{-15.5}$ for \MORUS[640].
 The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
+
+In order to preserve anonymity, our implementation is attached to the submission as auxiliary material. In the full version of this paper, a link will be provided instead.
 
@@
 \begin{table}[h!]
-  \caption{Experimental verification of trail biases.}
+  \caption{Experimental verification of trail correlations.}
   \label{tab:miniapproximations}

### 2018-05-10 morusAC_05_Full_trails.tex

@@
 
-In the previous section, we have exhibited a linear trail on the reduced ciphers \MiniMORUS[1280] and \MiniMORUS[640]. We now turn to the full ciphers \MORUS[1280] and \MORUS[640].
+In the previous section, we presented a linear trail for the reduced ciphers \MiniMORUS[1280] and \MiniMORUS[640]. We now turn to the full ciphers \MORUS[1280] and \MORUS[640].
 
 \subsection{Symmetrizing the Trail}
-In order to build a trail on the full \MORUS, we proceed exactly like we did for \MiniMORUS, following exactly the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
+In order to build a trail for the full \MORUS, we proceed exactly as we did for \MiniMORUS, following exactly the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
 That is, for every $S^t_{i,j}$ (resp.
@@
 Thus, translating the trail from one of the \MiniMORUS ciphers to the corresponding full \MORUS cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \MiniMORUS in the first place.
-Concretely, in order to build the full trail equation for \MORUS, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
-In this way, the biased linear combination on \MiniMORUS[1280] given in \Cref{sec:minitraileq}, namely:
+Concretely, in order to build the full trail equation for \MORUS, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same manner as before.
+This way, the biased linear combination on \MiniMORUS[1280] given in \Cref{sec:minitraileq}, namely:
 \begin{align*}
@@
 
-\subsection{Bias of the Full Trail}
+\subsection{Correlation of the Full Trail}
 
@@
 
-To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arises from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arises from a different part of the trail (where $w$ denotes the word size).
+To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arising from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arising from a different part of the trail (where $w$ denotes the word size).
 Then when we XOR the various trail fragments together, in \MiniMORUS these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
@@
 
-In the end, for \MORUS[1280], since the trail fragment $\gamma^t_i$ is used four times, the phenomenon we just described adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
+For \MORUS[1280], since the trail fragment $\gamma^t_i$ is used four times, this phenomenon adds a contribution of $4 \cdot 3 = 12$ to the overall weight of the full trail.
 This results in a total weight of $4 \cdot 16 + 12 = 76$ (recall that the weight of the trail on \MiniMORUS[1280] is 16).
-We have confirmed this bias by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \MiniMORUS in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the bias of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
-In the end, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \MORUS[1280].
+We have confirmed this by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \MiniMORUS in \Cref{sec:minibias}.
+That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the weight of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
+Overall, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \MORUS[1280].
 
@@
 This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
-In the end, when computing the full trail equation on \MORUS[640], we get that the weight of the trail is 73 (cf.
+Therefore, when computing the full trail equation on \MORUS[640], we get that the weight of the trail is 73 (cf.
 \Cref{sec:traileq}).

### 2018-05-10 morusAC_04_Mini_trails.tex
minor fix

@@
 Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a correlation of $2^{-16}$, experiments indicate a slightly better empirical correlation of $2^{-15.5}$ for \MORUS[640].
+While our analysis from the previous section predicts a correlation of $2^{-16}$, experiments indicate a slightly better empirical correlation of $2^{-15.5}$ for \MiniMORUS[640].
 The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
 
-In order to preserve anonymity, our implementation is attached to the submission as auxiliary material. In the full version of this paper, a link will be provided instead.
+In order to preserve anonymity, our implementation is attached to the submission as auxiliary material. In the full version of this paper, a link will be provided.
 

### 2018-05-10 morusAC_00_Abstract.tex
Some more copyediting

--- 2018-05-10 morusAC_00_Abstract.tex
@@
 \begin{abstract}
-  \morus is a high-performance authenticated encryption algorithm
+  \MORUS is a high-performance authenticated encryption algorithm
   submitted to the CAESAR competition, and recently selected as a
-  finalist.  There are three versions of \morus: \morus[640] with a
-  640-bit state and a 128-bit key, and \morus[1280] with a 1280-bit
+  finalist.  There are three versions of \MORUS: \MORUS[640] with a
+  640-bit state and a 128-bit key, and \MORUS[1280] with a 1280-bit
   state and a 128-bit or 256-bit key; for all versions the security
@@
 
-  Our main result is on the full state update of \morus: we found a
+  Our main result is on the full state update of \MORUS: we found a
   linear bias in the keystream that can be used to distinguish the
   output from random, and to recover some plaintext bits in the
-  broadcast setting.  For \morus[1280], the bias is $2^{-76}$,
+  broadcast setting.  For \MORUS[1280], the bias is $2^{-76}$,
   which can be identified after around $2^{152}$ encryptions, less than
   would be expected for a 256-bit secure cipher.  On the other hand, for
-  \morus[640], the bias is $2^{-73}$, which does not threaten
+  \MORUS[640], the bias is $2^{-73}$, which does not threaten
   the security of the cipher.
@@
 %
-  To identify this bias, we exploit rotational symmetries in \morus
+  To identify this bias, we exploit rotational symmetries in \MORUS
   using linear masks that are invariant by word-rotations of the
   state.  This motivates us to introduce single word versions of
-  \morus called \minimorus, which help the
+  \MORUS called \MiniMORUS, which help the
   analysis. The attack has been implemented and verified on
-  \minimorus, where it yields a bias of $2^{-16}$.
+  \MiniMORUS, where it yields a bias of $2^{-16}$.
 
@@
   We also study reduced versions of the initialization and
-  finalization of \morus to evaluate the security margin of these
+  finalization of \MORUS to evaluate the security margin of these
   components.  We show a forgery attack when the finalization is
@@
   nonce-misue setting when the initialization is reduced from 16 to 10
-  steps.  These additional results do not threaten the full \morus, but
+  steps.  These additional results do not threaten the full \MORUS, but
   studying all aspects of the design is useful to understand its

### 2018-05-10 morusAC_05_Full_trails.tex
Some more copyediting

@@
 \subsection{Symmetrizing the Trail}
-In order to build a trail for the full \MORUS, we proceed exactly as we did for \MiniMORUS, following exactly the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
+In order to build a trail for the full \MORUS, we proceed exactly as we did for \MiniMORUS, following the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
 That is, for every $S^t_{i,j}$ (resp.

### 2018-05-10 morusAC_00_Abstract.tex
That's it

--- 2018-05-10 morusAC_00_Abstract.tex
@@
   % a 1280-bit state and
-  a 128-bit or 256-bit key. For all versions the security
+  128-bit or 256-bit keys. For all versions the security
   claim for confidentiality matches the key size.  In this paper, we
@@
 
-  As our main result, we exhibit a
-  linear bias in the keystream of full \MORUS, which can be used to distinguish its output
-  from random, and to recover some plaintext bits in the
-  broadcast setting.  For \MORUS[1280], the bias is $2^{-76}$,
+  As our main result, we present a
+  linear correlation in the keystream of full \MORUS, which can be used to distinguish its output
+  from random and to recover some plaintext bits in the
+  broadcast setting.  For \MORUS[1280], the correlation is $2^{-76}$,
   which can be exploited after around $2^{152}$ encryptions, less than
   would be expected for a 256-bit secure cipher.  For
-  \MORUS[640], the same attack results in a bias of $2^{-73}$, which does not threaten
-  the security of the cipher.
+  \MORUS[640], the same attack results in a correlation of $2^{-73}$, which does not threaten
+  violate the security claims of the cipher.
 
@@
 %
-  To identify this bias, we make use of rotational symmetries in \MORUS
+  To identify this correlation, we make use of rotational symmetries in \MORUS
   using linear masks that are invariant by word-rotations of the
   state.  This motivates us to introduce single-word versions of
-  \MORUS called \MiniMORUS, which aid in the
+  \MORUS called \MiniMORUS, which simplifies the
   analysis. The attack has been implemented and verified on
-  \MiniMORUS, where it yields a bias of $2^{-16}$.
+  \MiniMORUS, where it yields a correlation of $2^{-16}$.
 
@@
   components.  We show a forgery attack when finalization is
-  reduced from 10 rounds to 3, and a key-recovery attack in the
-  nonce-misue setting when initialization is reduced from 16 to 10
-  steps.  These additional results do not threaten full \MORUS, but
+  reduced from 10 steps to 3, and a key-recovery attack in the
+  nonce-misuse setting when initialization is reduced from 16 steps to 10.  These additional results do not threaten the full \MORUS, but
   studying all aspects of the design is useful to understand its

### 2018-05-10 morusAC_01_Introduction.tex
Some more copyediting
That's it

@@
 independently propose state-recovery and forgery attacks on \MORUS in a nonce-misuse setting with negligible data and time complexities.
-Finally, a keystream bias similar in nature to our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist, which shares the same overall structure as \MORUS, but uses a very different, \cipher{AES}-based state update function.
+Finally, a keystream correlation similar in nature to our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist, which shares the same overall structure as \MORUS, but uses a very different, \cipher{AES}-based state update function.
 
@@
 As a consequence, when encrypting a fixed plaintext segment multiple times, a linear correlation exists between some output bits of the cipher.
-Moreover, the bias depends purely on the plaintext, and not on the secret key of the cipher.
+Moreover, the correlation depends purely on the plaintext, and not on the secret key of the cipher.
 In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
-For \MORUS[1280] with 256-bit keys, the linear bias is $2^{-76}$
-\todo{double-check bias vs. correlation everywhere}
+For \MORUS[1280] with 256-bit keys, the linear correlation is $2^{-76}$
 and can be exploited using about $2^{152}$ encrypted blocks.
-\todo{double-check}
 
@@
 To confirm the validity of our analysis,
-we practically verified the bias of the full linear trail for \MiniMORUS, as well as the bias of trail fragments for the full \MORUS.
-Our analysis is also backed by a symbolic evaluation of the full trail equation and its bias on all variants of \MORUS.
+we practically verified the correlation of the full linear trail for \MiniMORUS, as well as the correlation of trail fragments for the full \MORUS.
+Our analysis is also backed by a symbolic evaluation of the full trail equation and its correlation on all variants of \MORUS.
 
@@
 We analyze this simplified scheme in \Cref{sec/minitrails} and provide a ciphertext-only linear approximation with a weight of 16.
-We then extend our result to the full scheme in \Cref{sec/fulltrails}, showing a bias in the keystream over 5 steps,
+We then extend our result to the full scheme in \Cref{sec/fulltrails}, showing a correlation in the keystream over 5 steps,
 and discuss the implications of our observation for the security of \MORUS in \Cref{sec/Discussion}.

### 2018-05-10 morusAC_04_Mini_trails.tex
Some more copyediting
That's it

--- 2018-05-10 morusAC_04_Mini_trails.tex
@@
 In the equation for \MiniMORUS[1280] from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
+By the Piling-Up Lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
 Hence the linear combination of ciphertext bits on the left-hand side has a correlation of $2^{-16}$.
@@
 The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-In order to preserve anonymity, our implementation is attached to the submission as auxiliary material. In the full version of this paper, a link will be provided instead.
 

### 2018-05-10 morusAC_05_Full_trails.tex
That's it

@@
 We have confirmed this by explicitly computing the full trail equation in \Cref{sec:traileq}, and evaluating its exact weight like we did for \MiniMORUS in \Cref{sec:minibias}.
-That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the weight of each connected component separately by brute force, and then add up the weights of all components per the piling-up lemma.
+That is, since the equation is quadratic, we may view it as a graph, which we split into connected components; we then compute the weight of each connected component separately by brute force, and then add up the weights of all components per the Piling-Up Lemma.
 Overall, the full trail equation given in \Cref{sec:traileq} yields a weight of 76 for the full trail on \MORUS[1280].

### 2018-05-10 morusAC_06_Discussion.tex
That's it

@@
 
-\subsubsection{Keystream biases.}
+\subsubsection{Keystream correlation.}
 
-We emphasize that the biases we uncover between ciphertext bits are
-\emph{absolute}, in the sense that they do not depend on the encryption
-key, or the nonce.  This is the same situation as the keystream biases
+We emphasize that the correlation we uncover between ciphertext bits is
+\emph{absolute}, in the sense that it does not depend on the encryption
+key, or on the nonce.  This is the same situation as the keystream correlations
 in AEGIS~\cite{sacryptMinaud14}.  As such, they can be leveraged to
@@
 user, and the attacker can target this token as a repeated plaintext.
-Concretely, biases in the RC4 keystream have been exploited in this
+Concretely, correlations in the RC4 keystream have been exploited in this
 setting, leading to the recovery of authentication cookies in
@@
 With the trail presented in this work, the data complexity is clearly
-out of reach in practice, since exploiting the bias would require
+out of reach in practice, since exploiting the correlation would require
 $2^{152}$ encrypted blocks for \cipher{MORUS1280}, and $2^{146}$

### 2018-05-10 morusAC_07_IniFin.tex
That's it

@@
 We present forgery attacks on 3 out of 10 steps of \MORUS[1280] that claims 128-bit security for integrity.
-The attack only works for a limited number of steps, while it works in the nonce-respect setting.
-As far as we know, this is the first attempt to evaluate integrity of \MORUS in the nonce-respect setting.
+The attack only works for a limited number of steps, while it works in the nonce-respecting setting.
+As far as we know, this is the first attempt to evaluate integrity of \MORUS in the nonce-respecting setting.
 
 \subsubsection{Overview.}
-A general strategy of forgery attacks in the nonce-respect setting is to inject some difference in a message block and to propagate the difference so that it can be canceled by difference in another message block.
-However this approach does not work well against \MORUS because its large state size prevents the attackers from controlling the difference to be canceled by another message difference.
+A general strategy for forgery attacks in the nonce-respecting setting is to inject some difference in a message block and propagate the it so that it can be canceled by a difference in another message block.
+However this approach does not work well against \MORUS due to its large state size which prevents an attacker from easily controlling the differences in different registers.
 
-Here we focus on the property that the padding for an associated data $A$ and a message $M$ is the zero-padding, hence $A$ and $A'=A\|0^*$ and $M$ and $M'=M\|0$ will make the result identical for the associated data processing and the encryption parts, as long as $A,A'$ and $M,M'$ fit in the same number of blocks.
-During the finalization, because $A,A'$ (resp.
+Here we focus on the property that the padding for an associated data $A$ and a message $M$ is the zero-padding, hence $A$ and $A'=A\|0^*$ and $M$ and $M'=M\|0$ result in identical states after the associated data processing and the encryption parts, as long as $A,A'$ and $M,M'$ fit in the same number of blocks.
+During the finalization, since $A,A'$ (resp.
 $M,M'$) have different lengths, the corresponding $adlen$s (resp.
 $msglen$s) are different, which appears as $\Delta adlen$ (resp.
-$\Delta msglen$), and is injected from the message input.
-Our strategy is to propagate this difference to 128-bit tag $T$ and $T'$ such that their difference $\Delta T$ appears higher probability than $2^{-128}$.
-All in all, the forgery succeeds as long as the desired $\Delta T$ is obtained, in other words, the attackers do not have to cancel the state difference, which is the main advantage of attacking the finalization.
+$\Delta msglen$), and is injected through the message input interface.
+Our strategy is to propagate this difference to the 128-bit tags $T$ and $T'$ such that their difference $\Delta T$ appears with higher probability than $2^{-128}$.
+All in all, the forgery succeeds as long as the desired $\Delta T$ is obtained or in other words, the attacker does not have to cancel the state difference, which is the main advantage of attacking the finalization part of the scheme.
 
@@
 Recall that the message input during the finalization of \MORUS[1280] is $adlen\|msglen\|0^{128}$ where $adlen$ and $msglen$ are 64-bit strings.
-We set $\Delta adlen$ to be low Hamming weight, e.g.~{\tt 0x0000000000000001}.
+We set $\Delta adlen$ to be of low Hamming weight, e.g.~{\tt 0x0000000000000001}.
 %Hereafter, we use {\tt typefont} denote hexadecimal numbers.
-This difference propagates during 3 steps as specified in \Cref{Tbl:fin_diff}.
+This difference propagates through 3 steps as specified in \Cref{Tbl:fin_diff}.
 
@@
 We need to pay 1-bit to control each active AND gate.
-The probability evaluation for round 15 can be ignored because in this round only $S_4$ is non-linearly updated, while $S_4$ is never used for computing the tag.
+The probability evaluation for round 15 can be ignored since in this round only $S_4$ is non-linearly updated, while $S_4$ is never used for computing the tag.
 Finally, bitwise-AND in the tag computation is taken into account.
-Note that the tag is only 128 LSBs, thus the number of active AND gates should be counted only in those bits.
+Note that the tag is only 128 LSBs, thus the number of active AND gates should be counted only for those bits.
 As shown in \Cref{Tbl:fin_diff}, we can have a particular tag difference $\Delta T$ with probability $2^{-88}$.
@@
 
-\subsubsection{Remarks.} The fact that $S_4$ updated in the last round is not used to compute a tag implies that the \MORUS finalization generally includes unnecessary computations with respect to security.
+\subsubsection{Remarks.} The fact that the $S_4$ is updated in the last round but is not used in the tag generation implies that the \MORUS finalization generally includes unnecessary computations with respect to security.
 It may be interesting to tweak the design such that the tag can also depend on $S_4$.
@@
 \label{subsec/Ini}
-Kales et al.~\cite{cryptoeprint:2017:1137} showed that internal state of \MORUS[640] can be recovered under the nonce-misuse scenario using $2^5$ plaintext-ciphertext pairs.
+Kales et al.~\cite{cryptoeprint:2017:1137} showed that the internal state of \MORUS[640] can be recovered under the nonce-misuse scenario using $2^5$ plaintext-ciphertext pairs.
 As claimed by \cite{cryptoeprint:2017:1137} the attack is naturally extended to \MORUS[1280] though Kales et al.
@@
 However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker to back track the state value to the initial state.
-In this section, we show that meet-in-the-middle attacks can allow the attacker to recover the key faster than the exhaustive search for a relatively large number of steps, i.e.~10 out of 16 steps in \MORUS[1280].
+In this section, we show that meet-in-the-middle attacks allow the attacker to recover the key faster than exhaustive search for a relatively large number of steps, i.e.~10 out of 16 steps in \MORUS[1280].
 

### 2018-05-10 morusAC_08_Conclusion.tex
That's it

@@
 This work provides a comprehensive analysis of the components of
-\MORUS.  In particular we show that the state update has important
-weaknesses with a linear bias of $2^{-76}$ between ciphertext bits and
-plaintext bits in \MORUS[1280].  This enables a plaintext recovery
-attack in the broadcast setting, using about $2^{152}$ blocs of data.
+\MORUS.  In particular, we show that \MORUS[1280]'s keystream exhibits a correlation of $2^{-76}$ between certain ciphertext bits. This enables a plaintext recovery
+attack in the broadcast setting, using about $2^{152}$ blocks of data.
 While the amount of data required is impractical, this seems to
@@
 even if the key is refreshed regularly.  Moreover, the broadcast
-setting is practically relevant, as was shown with attacks aginst RC4
+setting is practically relevant, as was shown with attacks against RC4
 as used in TLS~\cite{DBLP:conf/uss/AlFardanBPPS13}.

### 2018-05-10 morusAC_05_Full_trails.tex
move trail figure and adapt notation

@@
 
-\begin{figure}
-  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
-  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
-  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
-  \centering
-  \resizebox{!}{0.95\textheight}{
-  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
-    \begin{scope}
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{27}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
-                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
-                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
-                   (\M{2}{1}) node[right] {$\alpha_7$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
-                   (\M{2}{2}) node[right] {$\gamma_{13}$};
-
-      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
-                   (\M{3}{0}) node[right] {$\beta_{12}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
-                   (\M{1}{0}) node[right] {$\alpha_{2}$};
-
-      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
-                   (\M{2}{0}) node[right] {$\beta_7$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
-      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
-                   (\M{2}{2}) node[right] {$\gamma_7$};
-      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
-                   (\M{2}{4}) node[right] {$\delta_7$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
-                   (\M{3}{1}) node[right] {$\alpha_{14}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
-                   (\M{3}{2}) node[right] {$\gamma_{20}$};
-
-      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
-                   (\M{4}{0}) node[right] {$\beta_{19}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-      \draw (\S{0}{2}) node[below] {\normalsize \MiniMORUS[640]};
-    \end{scope}
-
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
-
-    \begin{scope}[yshift=-7.0cm]
-      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
-                   (\M{0}{0}) node[right] {$\alpha_{51}$};
-
-      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
-                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
-                   (\M{1}{0}) node[right] {$\beta_0$};
-      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
-                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
-      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
-                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
-                   (\M{1}{2}) node[right] {$\gamma_0$};
-      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
-                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
-                   (\M{1}{3}) node[right] {$\delta_0$};
-
-      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
-                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
-                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
-                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
-                   (\M{2}{1}) node[right] {$\alpha_{37}$};
-      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
-                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
-                   (\M{2}{2}) node[right] {$\gamma_{4}$};
-
-      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
-                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
-                   (\M{3}{0}) node[right] {$\beta_{50}$};
-
-      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
-      \draw (\S{5}{2}) node[above] {\normalsize \MiniMORUS[1280]};
-    \end{scope}
-
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
-      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
-      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S^{\r}$}; }
-      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
-
-      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
-                   (\M{1}{0}) node[right] {$\alpha_{25}$};
-
-      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
-                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
-                   (\M{2}{0}) node[right] {$\beta_{38}$};
-      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
-                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
-      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
-      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
-      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
-                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
-                   (\M{2}{2}) node[right] {$\gamma_{38}$};
-      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
-                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
-                   (\M{2}{4}) node[right] {$\delta_{38}$};
-      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
-                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
-                   (\M{2}{3}) node[right] {$\varepsilon_0$};
-
-      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
-                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
-                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
-      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
-                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
-      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
-                   (\M{3}{1}) node[right] {$\alpha_{11}$};
-      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
-                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
-                   (\M{3}{2}) node[right] {$\gamma_{42}$};
-
-      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
-                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
-                   (\M{4}{0}) node[right] {$\beta_{24}$};
-
-      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
-    \end{scope}
-  \end{tikzpicture}}
-
-  \caption{\MiniMORUS: two approximations for $S^{2,2}_0$.}
-  \label{fig:twoapproximations}
-\end{figure}
-
 %%% Local Variables:

### 2018-05-10 morusAC_01_Introduction.tex
added sentence about ignoring rotation constants in contributions

@@
 To the best of our knowledge, this is the first attack on full \MORUS in the nonce-respecting setting.
-We also note that rekeying does not prevent the attack: the biases are independent of the secret encryption key and nonce, and can be exploited as long as a given plaintext segment is encrypted sufficiently often, regardless of whether each encryption uses a different key.
+We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key and nonce, and can be exploited as long as a given plaintext segment is encrypted sufficiently often, regardless of whether each encryption uses a different key.
+A surprising feature of the linear trail underpinning our attack is also that it does not depend on the values of rotation constants: a very similar trail would exist for most choices of round constants.
 

### 2018-05-10 morusAC_04_Mini_trails.tex

@@
 Then $\alpha^t_i$ is used one final time to close the trail.
+The full construction with the exact bit indices for \MiniMORUS[640] and
+\MiniMORUS[1280] is illustrated in \Cref{fig:twoapproximations}, where the left
+and right half each show half of the full trail.
 One may naturally wonder if some components of this trail are in conflict.
@@
 To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
+
+\begin{figure}
+  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
+  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
+  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
+  \centering
+  \resizebox{!}{0.95\textheight}{
+  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
+    \begin{scope}
+      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S_{\r}$}; }
+      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
+
+      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
+                   (\M{0}{0}) node[right] {$\alpha_{27}$};
+
+      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
+                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
+                   (\M{1}{0}) node[right] {$\beta_0$};
+      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
+                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
+      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
+      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
+                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
+                   (\M{1}{2}) node[right] {$\gamma_0$};
+      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
+                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
+                   (\M{1}{3}) node[right] {$\delta_0$};
+
+      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
+                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
+                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
+      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
+                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
+      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
+                   (\M{2}{1}) node[right] {$\alpha_7$};
+      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
+                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
+                   (\M{2}{2}) node[right] {$\gamma_{13}$};
+
+      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
+                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
+                   (\M{3}{0}) node[right] {$\beta_{12}$};
+
+      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
+    \end{scope}
+
+    \begin{scope}[xshift=2.5cm]
+      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S_{\r}$}; }
+      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
+
+      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
+                   (\M{1}{0}) node[right] {$\alpha_{2}$};
+
+      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
+                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
+                   (\M{2}{0}) node[right] {$\beta_7$};
+      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
+                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
+      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
+      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
+      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
+                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
+                   (\M{2}{2}) node[right] {$\gamma_7$};
+      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
+                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
+                   (\M{2}{4}) node[right] {$\delta_7$};
+      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
+                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
+                   (\M{2}{3}) node[right] {$\varepsilon_0$};
+
+      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
+                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
+                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
+      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
+                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
+      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
+                   (\M{3}{1}) node[right] {$\alpha_{14}$};
+      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
+                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
+                   (\M{3}{2}) node[right] {$\gamma_{20}$};
+
+      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
+                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
+                   (\M{4}{0}) node[right] {$\beta_{19}$};
+
+      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
+      \draw (\S{0}{2}) node[below] {\normalsize \MiniMORUS[640]};
+    \end{scope}
+
+    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
+
+    \begin{scope}[yshift=-7.0cm]
+      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S_{\r}$}; }
+      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};
+
+      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
+                   (\M{0}{0}) node[right] {$\alpha_{51}$};
+
+      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
+                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
+                   (\M{1}{0}) node[right] {$\beta_0$};
+      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
+                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
+      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
+      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
+                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
+                   (\M{1}{2}) node[right] {$\gamma_0$};
+      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
+                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
+                   (\M{1}{3}) node[right] {$\delta_0$};
+
+      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
+                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
+                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
+      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
+                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
+      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
+                   (\M{2}{1}) node[right] {$\alpha_{37}$};
+      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
+                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
+                   (\M{2}{2}) node[right] {$\gamma_{4}$};
+
+      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
+                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
+                   (\M{3}{0}) node[right] {$\beta_{50}$};
+
+      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
+      \draw (\S{5}{2}) node[above] {\normalsize \MiniMORUS[1280]};
+    \end{scope}
+
+    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
+      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
+      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S_{\r}$}; }
+      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};
+
+      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
+                   (\M{1}{0}) node[right] {$\alpha_{25}$};
+
+      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
+                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
+                   (\M{2}{0}) node[right] {$\beta_{38}$};
+      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
+                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
+      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
+      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
+      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
+                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
+                   (\M{2}{2}) node[right] {$\gamma_{38}$};
+      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
+                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
+                   (\M{2}{4}) node[right] {$\delta_{38}$};
+      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
+                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
+                   (\M{2}{3}) node[right] {$\varepsilon_0$};
+
+      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
+                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
+                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
+      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
+                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
+      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
+                   (\M{3}{1}) node[right] {$\alpha_{11}$};
+      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
+                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
+                   (\M{3}{2}) node[right] {$\gamma_{42}$};
+
+      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
+                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
+                   (\M{4}{0}) node[right] {$\beta_{24}$};
+
+      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
+    \end{scope}
+  \end{tikzpicture}}
+
+  \caption{\MiniMORUS: two approximations for $S_{2,0}^2$.}
+  \label{fig:twoapproximations}
+\end{figure}
 

### 2018-05-10 morusAC_07_IniFin.tex
proofread Section 7

@@
 In this section, for comprehensive security analysis of \MORUS, we provide new attacks on reduced version of the initialization and the finalization.
-We emphasize that the results in this section do not threat any security claim by the designers.
-However, we believe that investigating all parts of the design with different approaches from the existing work on \MORUS provides better understanding and will be useful especially when the design will be tweaked in future.
+We emphasize that the results in this section do not threaten any security claim by the designers.
+However, we believe that investigating all parts of the design with different approaches from the existing work on \MORUS provides a better understanding and will be useful especially when the design will be tweaked in future.
 
@@
 \subsubsection{Overview.}
-A general strategy for forgery attacks in the nonce-respecting setting is to inject some difference in a message block and propagate the it so that it can be canceled by a difference in another message block.
+A general strategy for forgery attacks in the nonce-respecting setting is to inject some difference in a message block and propagate it so that it can be canceled by a difference in another message block.
 However this approach does not work well against \MORUS due to its large state size which prevents an attacker from easily controlling the differences in different registers.
@@
 During the finalization, since $A,A'$ (resp.
-$M,M'$) have different lengths, the corresponding $adlen$s (resp.
-$msglen$s) are different, which appears as $\Delta adlen$ (resp.
-$\Delta msglen$), and is injected through the message input interface.
+$M,M'$) have different lengths, the corresponding 64-bit values $\adlen$ (resp.
+$\msglen$) are different, which appears as $\Delta \adlen$ (resp.
+$\Delta \msglen$) during the finalization, and is injected through the message input interface.
 Our strategy is to propagate this difference to the 128-bit tags $T$ and $T'$ such that their difference $\Delta T$ appears with higher probability than $2^{-128}$.
@@
 \subsubsection{Differential Trails.}
-Recall that the message input during the finalization of \MORUS[1280] is $adlen\|msglen\|0^{128}$ where $adlen$ and $msglen$ are 64-bit strings.
-We set $\Delta adlen$ to be of low Hamming weight, e.g.~{\tt 0x0000000000000001}.
+Recall that the message input during the finalization of \MORUS[1280] is $\adlen\mathrel\|\msglen\mathrel\|0^{128}$ where $\adlen$ and $\msglen$ are 64-bit strings.
+We set $\Delta \adlen$ to be of low Hamming weight, e.g.,~{\tt 0x0000000000000001}.
 %Hereafter, we use {\tt typefont} denote hexadecimal numbers.
@@
 Differences start to appear from round 2 and they will go through the bitwise-AND operation from round 4.
-We need to pay 1-bit to control each active AND gate.
+We need to pay 1 bit to control each active AND gate.
 The probability evaluation for round 15 can be ignored since in this round only $S_4$ is non-linearly updated, while $S_4$ is never used for computing the tag.
@@
 Indeed in \Cref{Tbl:fin_diff}, we can observe some jump-up of the probability in the tag computation.
-This is because the non-linearly involved terms are $S_2 \cdot S_3$, and $S_3$ that was updated 2-round before has a high Hamming weight.
+This is because the non-linearly involved terms are $S_2 \cdot S_3$, and $S_3$ that was updated 2 rounds before has a high Hamming weight.
 In this sense, involving $S_4$ in non-linear terms of the tag computation imposes more difficulties for the attacker.
@@
 The recovered state allows the attacker to mount a universal forgery attack under the same nonce.
-However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker to back track the state value to the initial state.
-In this section, we show that meet-in-the-middle attacks allow the attacker to recover the key faster than exhaustive search for a relatively large number of steps, i.e.~10 out of 16 steps in \MORUS[1280].
+However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker from backtracking the state value to the initial state.
+In this section, we show that meet-in-the-middle attacks allow the attacker to recover the key faster than exhaustive search for a relatively large number of steps, i.e.,~10 out of 16 steps in \MORUS[1280].
 
@@
 We divide the 10 steps of the initialization computation into two subsequent parts $F_0$ and $F_1$.
-(We later set that $F_0$ is the first 4 steps and $F_1$ is the last 6 steps.) Let $S^{-10}$ be the initial state value before setting the key, i.e.~$S^{-10} = (IV\|0^{128},0^{256},1^{256},0^{256},const_0\|const_1)$.
-Also let $S^0$ be 1280-bit state value after the finalization, which is now assumed to be recovered with the nonce-misuse analysis~\cite{cryptoeprint:2017:1137}.
+(We later set that $F_0$ is the first 4 steps and $F_1$ is the last 6 steps.) Let $S^{-10}$ be the initial state value before setting the key, i.e.,~$S^{-10} = (\IV\mathrel\|0^{128},0^{256},1^{256},0^{256},\const_0\mathrel\|\const_1)$.
+Also let $S^0$ be 1280-bit state value after the initialization, which is now assumed to be recovered with the nonce-misuse analysis~\cite{cryptoeprint:2017:1137}.
 We then have the following relation.
 \begin{align*}
-F_1 \circ F_0 \bigl(S^{-10} \oplus (0, K\|K, 0, 0, 0)\bigr) \oplus (0,K\|K,0,0,0) = S^0.
+%F_1 \circ F_0 \bigl(S^{-10} \oplus (0, K\|K, 0, 0, 0)\bigr) \oplus (0,K\|K,0,0,0) = S^0.
+F_1 \circ F_0 \bigl(S^{-10} \oplus (0, K, 0, 0, 0)\bigr) \oplus (0,K,0,0,0) = S^0.
 \end{align*}
+We target the variant \MORUS[1280-128], where $K = K_{128} \mathrel\| K_{128}$.
 
-Here, our strategy is to recover $K$ by independently processing $F_0$ and $F_1^{-1}$ to find the following match.
+Here, our strategy is to recover $K_{128}$ by independently processing $F_0$ and $F_1^{-1}$ to find the following match.
 \begin{align*}
-F_0 (S^{-10} \oplus (0, K\|K, 0, 0, 0)) \stackrel{?}{=} F_1^{-1} (S^0 \oplus (0,K\|K,0,0,0)).
+  F_0 (S^{-10} \oplus (0, K_{128}\|K_{128}, 0, 0, 0)) \stackrel{?}{=} F_1^{-1} (S^0 \oplus (0,K_{128}\|K_{128},0,0,0)).
 \end{align*}
@@
 \begin{itemize}
-\item $G_0$: a set of bits of $K$ that are guessed for computing $F_0$.
-\item $G_1$: a set of bits of $K$ that are guessed for computing $F_1^{-1}$.
+\item $G_0$: a set of bits of $K_{128}$ that are guessed for computing $F_0$.
+\item $G_1$: a set of bits of $K_{128}$ that are guessed for computing $F_1^{-1}$.
 \item $G_2$: a set of bits in the intersection of $G_0$ and $G_1$.
@@
 \end{itemize}
-Suppose that the union of $G_0$ and $G_1$ covers all the bits of $K$. The attack exhaustively guesses $G_2$ and performs the following procedure for each guess.
+Suppose that the union of $G_0$ and $G_1$ covers all the bits of $K_{128}$. The attack exhaustively guesses $G_2$ and performs the following procedure for each guess.
 \begin{enumerate}
@@
 The number of the total candidates after the $x$-bit match is $2^{|G_2|} \cdot 2^{|G_0|-|G_2| + |G_1|-|G_2| - x} = 2^{|G_0| + |G_1| - |G_2| - x}$.
-Hence, the key $K$ is recovered with complexity $$\max( 2^{|G_0|}, 2^{|G_1|}, 2^{|G_0| + |G_1| - |G_2| - x}).$$
+Hence, the key $K_{128}$ is recovered with complexity $$\max( 2^{|G_0|}, 2^{|G_1|}, 2^{|G_0| + |G_1| - |G_2| - x}).$$
 Suppose that we choose $|G_0|$ and $|G_1|$ to be balanced i.e. $|G_0|=|G_1|$. Then, the complexity is $$\max( 2^{|G_0|}, 2^{2|G_0| - |G_2| - x}).$$
@@
 The analysis of the diffusion is given in \Cref{Tbl:ini_diffuse}.
-In the end, $K$ can be recovered faster than the exhaustive search by 1 bit, i.e.~with complexity $2^{127}$.
+In the end, $K_{128}$ can be recovered faster than the exhaustive search by 1 bit, i.e.~with complexity $2^{127}$.
 
@@
 \caption{Analysis of the diffusion and matching bits over 10 steps.
-`0' and `1' denote that the state bit can and cannot be computed from a partial knowledge of $K$, respectively.
+`0' and `1' denote that the state bit can and cannot be computed from a partial knowledge of $K_{128}$, respectively.
 After the partial computations from each direction, 4 bits of $S^{-6}$ can match.}
@@
                          & {\tt 0000000000000000} & {\tt 0000000000000001} & {\tt 0000000000000000} & {\tt 0000000000000001} \\
-$S^{-10}\oplus K$ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+$S^{-10}\oplus K_{128} $ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
                          & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
@@
                      & {\tt 0000000000000000} & {\tt 0000000000000100} & {\tt 0000000000000000} & {\tt 0000000000000100} \\
-$S^0\oplus K$ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+$S^0\oplus K_{128} $ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
                      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\

### 2018-05-10 morusAC_07_IniFin.tex
one more bib reference

@@
 However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker from backtracking the state value to the initial state.
-In this section, we show that meet-in-the-middle attacks allow the attacker to recover the key faster than exhaustive search for a relatively large number of steps, i.e.,~10 out of 16 steps in \MORUS[1280].
+In this section, we show that meet-in-the-middle attacks allow the attacker to recover the key faster than exhaustive search for a relatively large number of steps, i.e., 10 out of 16 steps in \MORUS[1280].
 
@@
 We divide the 10 steps of the initialization computation into two subsequent parts $F_0$ and $F_1$.
-(We later set that $F_0$ is the first 4 steps and $F_1$ is the last 6 steps.) Let $S^{-10}$ be the initial state value before setting the key, i.e.,~$S^{-10} = (\IV\mathrel\|0^{128},0^{256},1^{256},0^{256},\const_0\mathrel\|\const_1)$.
+(We later set that $F_0$ is the first 4 steps and $F_1$ is the last 6 steps.) Let $S^{-10}$ be the initial state value before setting the key, i.e., $S^{-10} = (\IV\mathrel\|0^{128},0^{256},1^{256},0^{256},\const_0\mathrel\|\const_1)$.
 Also let $S^0$ be 1280-bit state value after the initialization, which is now assumed to be recovered with the nonce-misuse analysis~\cite{cryptoeprint:2017:1137}.
@@
 Hence, the key $K_{128}$ is recovered with complexity $$\max( 2^{|G_0|}, 2^{|G_1|}, 2^{|G_0| + |G_1| - |G_2| - x}).$$
-Suppose that we choose $|G_0|$ and $|G_1|$ to be balanced i.e. $|G_0|=|G_1|$. Then, the complexity is $$\max( 2^{|G_0|}, 2^{2|G_0| - |G_2| - x}).$$
+Suppose that we choose $|G_0|$ and $|G_1|$ to be balanced i.e., $|G_0|=|G_1|$. Then, the complexity is $$\max( 2^{|G_0|}, 2^{2|G_0| - |G_2| - x}).$$
 Two terms are balanced when $x = |G_0| - |G_2|$.
@@
 The analysis of the diffusion is given in \Cref{Tbl:ini_diffuse}.
-In the end, $K_{128}$ can be recovered faster than the exhaustive search by 1 bit, i.e.~with complexity $2^{127}$.
+In the end, $K_{128}$ can be recovered faster than the exhaustive search by 1 bit, i.e., with complexity $2^{127}$.
 

### 2018-05-10 morusAC_07_IniFin.tex
last typo

@@
 \subsubsection{Remarks.}
-The matching state does not have te be a border of a step.
-It can be defined on a border of a round (or even some complicated way).
+The matching state does not have to be a border of a step.
+It can be defined on a border of a round, or even in some more complicated way.
 We did not find the extension of the number of attacked steps even with this way.

### 2018-05-10 morusAC_05_Full_trails.tex
page break optimization
added subsection 5.3 about variable plaintext

@@
 
-\begin{figure}
+\begin{figure}[t!]
   \substatesfalse
@@
 
+\subsection{Taking Variable Plaintext into Account}
+\label{subsec:variable}
+
+In our analysis so far, for the sake of simplicity, we have assumed that all plaintext blocks are zero.
+We now examine what happens if we remove that assumption, and integrate plaintext variables into our analysis.
+What we show is that plaintext variables only contribute linearly to the trail.
+In other words, the full trail equation with plaintext variables is equal to the full trail equation with all-zero plaintext XORed with a linear combination of plaintext variables.
+
+To see this, recall that plaintext bits contribute to the encryption process in two ways (cf. \Cref{subsec/Spec}):
+\begin{enumerate}
+\item They are added to some bits derived from the state to form the ciphertext.\label{item:1}
+\item During each encryption step, the \StateUpdate function adds a plaintext block to every register except $S_0$.\label{item:2}
+\end{enumerate}
+
+The effect of \Cref{item:1} is that whenever we use a ciphertext bit in our full trail equation, the corresponding plaintext bit also needs to be XORed in.
+Because ciphertext bits only contribute linearly to the trail equation, this only adds a linear combination of plaintext bits to the equation.
+
+Regarding \Cref{item:2}, recall that the full trail equation is a linear combination of (the symmetrized version of) equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ in \Cref{sec:minitraileq}.
+Also observe that in each equation, state bits that are shifted by a word-wise rotation only contribute linearly.
+Because plaintext bits are XORed into each register at the same time word-wise rotation is performed, this implies that plaintext bits resulting from \Cref{item:2} also only contribute linearly.
+In fact in all cases, it so happens that updating the equation to take plaintext variables into account simply involves XORing in the plaintext bit $M^t_i$.
+
+It may be observed that message blocks in the \StateUpdate{} function only contribute linearly to the state, and in that regard play a role similar to key bits in an SPN cipher; and indeed in SPN ciphers, key bits contribute linearly to linear trails \cite{eurocryptMatsui93}.
+
+In the end, with variable plaintext, our trail yields a biased linear combination of ciphertext bits and plaintext bits.
+In regards to attacks, this means the situation is effectively the same as with a biased stream cipher: in particular if the plaintext is known we obtain a distinguisher; and if a fixed unknown plaintext is encrypted multiple times (possibly also with some known variable part) then our trail yields a plaintext recovery attack.
+
 %%% Local Variables:

### 2018-05-10 morusAC_01_Introduction.tex
updated contributions and discussion to reflect the fact that the correlation is between plaintext and ciphertext bits, rather than ciphertext bits with a fixed plaintext

--- 2018-05-10 morusAC_01_Introduction.tex
@@
 Mileva et~al.~\cite{balkancryptsecMilevaDV15} propose a distinguisher in the nonce-reuse setting and practically evaluate the differential behaviour of toy variants of \MORUS.
+Shi et~al.~\cite{aiieShiGLZ16} analyze the differential properties of the finalization reduced to 2 out of 10 steps, but find no attacks.
 Dwivedi et~al.~\cite{cryptoeprint:2016:1053} discuss the applicability of SAT solvers for state recovery, but the resulting complexity of $2^{370}$ for \MORUS[640] is well beyond the security claim.
@@
 
-In more detail, our main result is a linear approximation \cite{eurocryptMatsui93,eurocryptMatsuiY92} linking ciphertext bits across five consecutive encryption blocks.
+In more detail, our main result is a linear approximation \cite{eurocryptMatsui93,eurocryptMatsuiY92} linking plaintext and ciphertext bits spanning five consecutive encryption blocks.
 %As a consequence of this trail, we exhibit a biased linear combination of ciphertext bits, which holds across any given sequence of five consecutive encryption steps.
@@
 %Moreover, that is the only assumption required; in particular, the secret key or the nonce used to initialize the inner state of MORUS are irrelevant: the bias is independent of their value.
-As a consequence, when encrypting a fixed plaintext segment multiple times, a linear correlation exists between some output bits of the cipher.
-Moreover, the correlation depends purely on the plaintext, and not on the secret key of the cipher.
-In principle, this property could be used to recover unknown bits of a plaintext encrypted a large number of times, provided an initial segment of the plaintext is known.
+%As a consequence, when encrypting a fixed plaintext segment multiple times, a linear correlation exists between some output bits of the cipher.
+Moreover, the correlation does not depend on the secret key of the cipher.
+In principle, this property could be used as a known-plaintext distinguisher, or to recover unknown bits of a plaintext encrypted a large number of times.
 For \MORUS[1280] with 256-bit keys, the linear correlation is $2^{-76}$
@@
 To the best of our knowledge, this is the first attack on full \MORUS in the nonce-respecting setting.
-We also note that rekeying does not prevent the attack: the biases are independent of the secret encryption key and nonce, and can be exploited as long as a given plaintext segment is encrypted sufficiently often, regardless of whether each encryption uses a different key.
+We note that rekeying does not prevent the attack: the biases are independent of the secret encryption key and nonce, and can be exploited for plaintext recovery as long as a given plaintext segment is encrypted sufficiently often, regardless of whether each encryption uses a different key.
+A notable feature of the linear trail underpinning our attack is also that it does not depend on the values of rotation constants: a very similar trail would exist for most choices of round constants.
 
@@
 In \Cref{sec/IniFin}, we present our results on the security of \MORUS with round-reduced initialization (in a nonce-misuse setting) or finalization.
-Finally, we conclude in \Cref{sec/Conclusion}.
+We conclude in \Cref{sec/Conclusion}.
 

### 2018-05-10 morusAC_03_MiniMorus.tex
Some more copyediting
updated contributions and discussion to reflect the fact that the correlation is between plaintext and ciphertext bits, rather than ciphertext bits with a fixed plaintext

@@
 To simplify the description of the attack, we assume all plaintext blocks are zero.
-However the same results would hold as long as the plaintext is fixed.
+This assumption will be removed in \Cref{subsec:variable}, where we will show that plaintext bits only contribute linearly to the trail.
 Recall that the inner state of the cipher consists of five $4w$-bit registers $S_0,\dots,S_4$, each containing four $w$-bit words.
@@
 
-To build our linear trail, we start in a linear combinations of bits within a single register.
+To build our linear trail, we start with a linear combinations of bits within a single register.
 \begin{definition}

### 2018-05-10 morusAC_05_Full_trails.tex
updated contributions and discussion to reflect the fact that the correlation is between plaintext and ciphertext bits, rather than ciphertext bits with a fixed plaintext

@@
 
-It may be observed that message blocks in the \StateUpdate{} function only contribute linearly to the state, and in that regard play a role similar to key bits in an SPN cipher; and indeed in SPN ciphers, key bits contribute linearly to linear trails \cite{eurocryptMatsui93}.
+It may be observed that message blocks in the \StateUpdate{} function only contribute linearly to the state, and in that regard play a role similar to key bits in an SPN cipher; and indeed in SPN ciphers, it is the case that key bits contribute linearly to linear trails \cite{eurocryptMatsui93}.
+In this light the previous result may not be surprising.
 

### 2018-05-10 morusAC_06_Discussion.tex
page break optimization
added subsection 5.3 about variable plaintext
updated contributions and discussion to reflect the fact that the correlation is between plaintext and ciphertext bits, rather than ciphertext bits with a fixed plaintext

@@
 %!TEX root = morusAC.tex
+
+%\pagebreak[4]
 
@@
 
-We emphasize that the correlation we uncover between ciphertext bits is
+We emphasize that the correlation we uncover between plaintext and ciphertext bits is
 \emph{absolute}, in the sense that it does not depend on the encryption
@@
 
-Moreover, the existence of this trail does hint at some weakness in the
+\subsubsection{Design considerations.}
+The existence of this trail does hint at some weakness in the
 design of \cipher{MORUS}. Indeed, a notable feature of the trail is that

### 2018-05-11 morusAC.tex
Page numbers
working on introduction
Conclusion
Some more copyediting
rollback for Tomer
ePrint version: author names + acknowledgments + increased margins

@@
 \newif\ifanonymous
+\newif\iflncsmargins
 %\anonymoustrue%uncomment for submission version
 
-\newif\ifshrink\shrinkfalse
-%\shrinktrue for screen-friendly original LNCS paper size
+\newif\ifshrink
+%\shrinktrue % for screen-friendly original LNCS paper size
 
@@
 \usepackage{common}
+\pagestyle{plain}
+\usepackage{multirow}
 
 \ifanonymous
+\lncsmarginstrue
 \author{}
@@
 \else
-\author{Tomer Ashur \and Maria Eichlseder \and Ga\"etan Leurent \and Brice Minaud \and Yann Rotella \and Yu Sasaki \and Beno\^it Viguier}
-\institute{}
-%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
+\author{Tomer Ashur\inst{1} \and Maria Eichlseder\inst{2} \and Martin M.
+Lauridsen \and Ga\"etan Leurent\inst{3} \and Brice Minaud\inst{4} \and\\ Yann Rotella\inst{3} \and Yu Sasaki\inst{5} \and Beno\^it Viguier\inst{6}}
+\institute{
+imec-COSIC KULeuven, Belgium
+\and
+Graz University of Technology, Austria
+\and
+Inria, France
+\and
+Royal Holloway, University of London, United Kingdom
+\and
+NTT, Japan
+\and
+Radboud University, Netherlands\\
+\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|mail@martinlauridsen.info|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
+}
+\fi
+
+\iflncsmargins
+\else
+\usepackage[hmargin=2.5cm,vmargin=3cm]{geometry}
 \fi
@@
 
-\title{\morus: Biases in the Stream}
+\title{Cryptanalysis of \MORUS}
 \maketitle
-\todo{The title should reflect the full analysis. How about something more generic?}
+\thispagestyle{plain}
 
@@
 \input{morusAC_07_IniFin}
+\input{morusAC_08_Conclusion}
 
@@
 
+%\section*{Auxiliary Material}
+\clearpage
+
 \appendix

### 2018-05-11 morusAC_08_Conclusion.tex
ePrint version: author names + acknowledgments + increased margins
Acknowledgement

@@
 
+\ifanonymous
+\else
+\section{Acknowledgments}
+
+The results presented here were originally found during the Flexible
+Symmetric Cryptography workshop held at the Lorentz Center in Leiden,
+Netherlands.  The authors would like to thank Meltem Sonmez Turan who
+participated to the initial discussion.  The fifth author was
+supported by EPSRC Grant EP/M013472/1.  The fouth and sixth authors
+are partially supported by the French Agence Nationale de la Recherche
+through the BRUTUS project under Contract ANR-14-CE28-0015.
+
+\fi
+
+\FloatBarrier
+
 %%% Local Variables:

### 2018-05-12 morusAC_08_Conclusion.tex
added link to implementation

@@
 Netherlands.  The authors would like to thank Meltem Sonmez Turan who
-participated to the initial discussion.  The fifth author was
-supported by EPSRC Grant EP/M013472/1.  The fouth and sixth authors
+participated in the initial discussion.
+The fourth and sixth authors
 are partially supported by the French Agence Nationale de la Recherche
 through the BRUTUS project under Contract ANR-14-CE28-0015.
+The fifth author was supported by EPSRC Grant EP/M013472/1.
 

### 2018-05-15 morusAC_04_Mini_trails.tex
Mention src code as supportial material.
ePrint version: author names + acknowledgments + increased margins
added link to implementation
minor typos/clarifications

@@
 
-In this section, we describe how we build a trail for \MiniMORUS before computing its correlation and experimentally verify it.
+In this section, we describe how we build a trail for \MiniMORUS, then compute its correlation and validate the correlation experimentally.
 
@@
   \begin{subfigure}{.32\textwidth}
+  \centering
   \tikzsetnextfilename{minimorus_part_delta}
@@
   \begin{subfigure}{.32\textwidth}
+  \centering
   \tikzsetnextfilename{minimorus_part_epsil}
@@
 
-In the equation for \MiniMORUS[1280] from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
+In the equation for \MiniMORUS[1280] from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms (in the sense that no two lines share a common term), and each line has a weight of 1.
 By the Piling-Up Lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
@@
 The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
+
+\ifanonymous
+The programs used to verify the bias experimentally have been
+submitted as supporting material.
+\else
+The programs we used to verify the bias experimentally are available at:\\
+\centerline{
+\url{https://github.com/ildyria/MorusBias}
+}
+\fi
 

### 2018-05-15 morusAC_05_Full_trails.tex
typo
minor typos/clarifications

@@
 While in \MiniMORUS, these terms cancel out, in the full \MORUS, when adding all symmetric copies of the trail, we end up with the sum:
+{
+\allowdisplaybreaks[0]
 \begin{align}
@@
 \end{align}
+}
 It may be observed that the products occurring in the equation above involve eight terms forming a ring. The weight of this expression can be computed by brute force, and is equal to $3$.
@@
 As explained previously, a partial collision with the term $S^t_{2,i} \cdot S^t_{3,i-w}$ from trail fragment $\gamma^t_i$ results in \Cref{eq:8circle}.
-However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case \MORUS[640] is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
+However trail fragment $\alpha^t_{i+d}$ is once used in the course of the full trail with an offset of $d = b_1+b_4-b_0-b_2$ (relative to $\gamma^t_i$), which in the case of \MORUS[640] is equal to $31+13-5-7 = 0 \;\text{mod}\; 32$.
 This creates another term $S^t_{2,i} \cdot S^t_{3,i}$, which ultimately destroys one of the four occurrences of \Cref{eq:8circle}.
@@
 \item They are added to some bits derived from the state to form the ciphertext.\label{item:1}
-\item During each encryption step, the \StateUpdate function adds a plaintext block to every register except $S_0$.\label{item:2}
+\item During each encryption step, the \StateUpdate{} function adds a plaintext block to every register except $S_0$.\label{item:2}
 \end{enumerate}

### 2018-05-17 morusAC_00_Abstract.tex
removed extra word in abstract
typo

@@
   would be expected for a 256-bit secure cipher.  For
-  \MORUS[640], the same attack results in a correlation of $2^{-73}$, which does not threaten
+  \MORUS[640], the same attack results in a correlation of $2^{-73}$, which does not
   violate the security claims of the cipher.
@@
 
-\keywords{MORUS, CAESAR, Authenticated Encryption, Nonce Respect, Linear Cryptanalysis, Confidentiality.}
+\keywords{MORUS, CAESAR, Authenticated Encryption, Nonce Respecting, Linear Cryptanalysis, Confidentiality.}
 \end{abstract}

### 2018-06-21 slides/tikz/trail-alpha.tex
Benoit Lunch Talk

--- 2018-05-10 morusAC_04_Mini_trails.tex
@@
-%!TEX root = morusAC.tex
-
-\section{Linear Trail for \MiniMORUS}
-\label{sec/minitrails}
-
-In this section, we describe how we build a trail for \MiniMORUS before computing its correlation and experimentally verify it.
-
-\subsection{Overview of the Trail}
-
-To build a linear trail for \MiniMORUS, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
-\begin{itemize}
-    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
-    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
-    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
-    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
-    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
-\end{itemize}
-
-The trail fragments are depicted on \Cref{fig:minimorusfragments}.
-In all cases except $\alpha^t_i$, the trail fragment approximates a single AND gate by zero, which holds with probability 3/4, and hence the trail fragment has weight 1.
-In the case of $\alpha^t_i$, two AND gates are involved; however the two gates share an entry in common, and in both cases the other entry also has a linear contribution to the trail, which results in an overall contribution of the form (see \cite[Sec.~3.3]{DBLP:conf/indocrypt/AshurR16})
-\[
-x \cdot y \oplus x \cdot z \oplus y \oplus z= (x \oplus 1) \cdot (y \oplus z).
-\]
-As a result, the trail fragment $\alpha^t_i$ also has a weight of 1.
-Another way of looking at this phenomenon is that the trail holds for two different approximations of the AND gates: the alternative approximation is depicted by a dashed line on \Cref{fig:minimorusfragments}.
-
-\begin{figure}
-  \substatesfalse
-  % \substatesfalse to label state words and/or masks
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \tikzsetnextfilename{minimorus_part_alpha}
 \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
+  \rotwfalse
+  \messagefalse
   \printstate
@@
   \draw[trail, alpha, dashed]
-      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
-      ;
+    (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east);
+  % \draw[alpha, very thick] (W40) + (0,-0.2) ellipse (0.6cm and 0.175cm);
+
+  \draw (W42) + (0,-0.5) node[below] {weight($\alpha^t_i$) = 1 (not 2)};
+
 \end{tikzpicture}%}}}
-  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \centering
-  \tikzsetnextfilename{minimorus_part_beta}
-  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, beta]
-      (C) -- node[right] {$i$} (lll-1)
-      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
-      (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \tikzsetnextfilename{minimorus_part_gamma}
-  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
-      (xnd1) -- (and1)
-      (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \bigskip
-
-  \begin{subfigure}{.32\textwidth}
-  \tikzsetnextfilename{minimorus_part_delta}
-  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-  \tikzsetnextfilename{minimorus_part_epsil}
-  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
-    \printstate
-    \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
-      ;
-  \end{tikzpicture}%}}}
-  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{.32\textwidth}
-    \small
-    \begin{align*}
-      \intertext{\MiniMORUS[640]:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
-      \intertext{\MiniMORUS[1280]:}
-\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
-\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
-\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
-    \end{align*}
-    \caption*{\MiniMORUS instances}
-  \end{subfigure}
-  \caption{\MiniMORUS linear trail fragments.}
-  \label{fig:minimorusfragments}
-\end{figure}
-
-The way we are going to use each trail fragment may be summarized as follows, where in each case, elements to the left of the arrow $\to$ are used to approximate the element on the right of the arrow:
-\begin{alignat*}{2}
-\alpha^t_i:\quad&& C^t_i &\to S^{t+1}_{0,i+b_0} \\
-\beta^t_i:\quad&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
-\gamma^t_i:\quad&& S^t_{1,i}, S^{t+1}_{1,i+b_1} &\to S^t_{4,i} \\
-\delta^t_i:\quad&& S^t_{4,i}, S^{t+1}_{4,i+b_4} &\to S^{t+1}_{2,i} \\
-\varepsilon^t_i:\quad&& S^t_{2,i}, S^{t+1}_{2,i+b_2} &\to S^{t+1}_{0,i}.
-\end{alignat*}
-
-In more detail, the idea is that by using $\alpha^t_i$, we are able to approximate a bit of $S_0$ using only a ciphertext bit.
-By combining $\alpha^t_i$ with $\beta^{t+1}_{i+b_0}$, we are then able to approximate a bit of $S_1$ (at step $t+1$) using only ciphertext bits from two consecutive steps.
-Likewise, $\gamma^t_i$ allows us to ``jump'' from $S_1$ to $S_4$, i.e.
-by combining $\alpha^t_i$ with $\beta^t_i$ and $\gamma^t_i$ with appropriate choices of parameters $t$ and $i$ for each, we are able to approximate one bit of $S_4$ using only ciphertext bits.
-Notice however that $\gamma^t_i$ requires approximating $S_1$ in two consecutive steps; and so the previous combination requires using $\alpha^t_i$ and $\beta^t_i$ \emph{twice} at different steps.
-In the same way, $\delta^t_i$ allows us to jump from $S_4$ to $S_2$; and $\varepsilon^t_i$ allows jumping from $S_2$ back to $S_0$.
-Eventually, we are able to approximate a bit of $S_0$ using only ciphertext bits via the combination of all trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$.
-
-However, the same bit of $S_0$ can also be approximated directly by using $\alpha^t_i$ at the corresponding step.
-Thus that bit can be linearly approximated from two different sides: the first approximation uses a combination of all trail fragments, and involves successive approximations of all state registers (except $S_3$) spanning several encryption steps, as explained in the previous paragraph.
-The second approximation only involves using $\alpha^t_i$ at the final step reached by the previous trail.
-By XORing up these two approximations, we are left with only ciphertext bits, spanning five consecutive encryption steps.
-
-Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice.
-Then $\alpha^t_i$ is used one final time to close the trail.
-One may naturally wonder if some components of this trail are in conflict.
-In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$.
-To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.
-
-\subsection{Trail Equation}
-\label{sec:minitraileq}
-
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
-For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero correlation (in all cases the correlation is $2^{-1}$).
-\begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
-\end{alignat*}
-
-From an algebraic point of view, building the full trail amounts to adding up copies of the previous equations for various choices of $t$ and $i$, so that eventually all $S^x_{y,z}$ terms on the left-hand side cancel out.
-Then we are left with only ciphertext terms on the left-hand side, while the right-hand side consists of a sum of biased expressions.
-By measuring the correlation of the right-hand side expression, we are then able to determine the correlation of the linear combination of ciphertext bits on the left-hand side.
-We now set out to do so.
-
-In order to build the equation for the full trail, we start with $E^2_0$:
-\[
-S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
-\]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add to the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
-This yields:
-\begin{align*}
-&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
-=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
-\end{align*}
-We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add to the equations $D^t_i$ for appropriate choices of $t$ and $i$.
-This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
-Ultimately, for \MiniMORUS[1280], this yields the equation:
-\begin{flalign*}
-&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
-&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
-&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
-&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
-&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
-&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
-&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
-&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
-&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
-&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
-&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
-&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
-\end{flalign*}
-
-The equation for \MiniMORUS[640] is very similar, and is given in \Cref{sec:traileq}.
-
-\subsection{Correlation of the Trail}
-\label{sec:minibias}
-
-In the equation for \MiniMORUS[1280] from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1.
-By the Piling-Up Lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16.
-Hence the linear combination of ciphertext bits on the left-hand side has a correlation of $2^{-16}$.
-The same holds for \MiniMORUS[640] (cf.
-\Cref{sec:traileq}).
-
-That the correlation is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
-Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
-In both cases it is approximated by zero.
-When XORing the corresponding equations, these two ANDs cancel each other, which saves two AND gates.
-Since $\gamma^t_i$ is used four times in the course of the full trail, this results in saving 8 AND gates overall, which explains why the final correlation is $2^{-16}$ rather than $2^{-24}$.
-
-\subsection{Experimental Verification}
-
-To confirm that our analysis is correct, we ran experiments on an implementation of \MiniMORUS[1280] and \MiniMORUS[640].
-In both cases, we measured the correlation of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the correlation of the full trail itself (denoted by $\chi$).
-Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a correlation of $2^{-16}$, experiments indicate a slightly better empirical correlation of $2^{-15.5}$ for \MORUS[640].
-The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
-
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \MiniMORUS[640]:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
-%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
-%\end{align*}
-%\bigskip
-%
-%\noindent
-%2 approximations for $S^{2,2}_{0}$ of \MiniMORUS[1280]:
-%\begin{align*}
-%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
-%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
-%\intertext{Combined ciphertext-only approximation:}
-%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
-%\end{align*}
-
-\begin{table}[h!]
-  \caption{Experimental verification of trail correlations.}
-  \label{tab:miniapproximations}
-  \centerline{
-  \begin{tabular}{@{}llSSS@{}}
-    \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
-    \multicolumn{2}{@{}l}{Approximations for \MiniMORUS[640]}
-             & {Exp.} & {Bool.} & {Meas.} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
-    \midrule
-    \multicolumn{2}{@{}l}{Approximations for \MiniMORUS[1280]} \\
-    \midrule
-    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
-    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
-    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
-    \bottomrule
-  \end{tabular}
-  }
-\end{table}
-
-%%% Local Variables:
-%%% TeX-master: "morusAC"
-%%% End:

### 2018-07-06 AC18_rebuttal.txt
Asiacrypt rebuttal draft

Dear Tomer Ashur; Maria Eichlseder; Martin M. Lauridsen; Gaëtan Leurent; Brice Minaud; Yann Rotella; Yu Sasaki; Benoît Viguier,

Below please find the reviewer comments on your paper

  "Cryptanalysis of MORUS"

that was submitted to ASIACRYPT 2018.

https://secure.iacr.org/websubrev/ac2018/submit/rebuttal.php

Remember that the deadline for the rebuttals is July 13th, 11:45 UTC.

************************************************************************
* REBUTTAL *************************************************************
************************************************************************

We thank all reviewers for taking the time to review this submission.
We will take all editorial comments into account to prepare an updated version.
Below, we discuss some specific comments raised by the reviewers:

> A small drawback that I would add is that this paper would benefit from a
> concrete comparison with similar attacks, as the previous one that was found
> on AEGIS. While it does not low down the impact of the attack on a new scheme
> it could help to understand the differences for further cryptanalysis.

... TODO ...

> Moreover, the provided attack works either if the key is fixed or if the same
> plaintext is repeated a large number of times. While the former case is not a
> big limitation in practice even considering design recommendations, the later
> seems to be less probable in practice. Maybe examples of concrete scenarii
> should be emphasized to strengthen the proposal.

... TODO ...

************************************************************************
* REVIEWS **************************************************************
************************************************************************

This paper presents three attacks on MORUS: a linear attack on the
confidentiality of full MORUS, a
differential attack on the integrity of MORUS with reduced finalisation, and a
meet-in-the-middle
key-recovery attack on MORUS with reduced initialisation.

The main result, i.e. the linear attack, is quite interesting in itself, and
shows a potential
weakness in the parameters of MORUS-1280. In particular, the attack shows that
increasing the
internal state size alone from MORUS-640 to MORUS-1280 has little influence on
the linear
correlation. Additionally, it is very interesting that the correlation is
completely independent of
the key and nonce. While that attack requires repeated encryption of the same
plaintext with
different nonces, the authors argue that previous attacks on RC4 have been
carried out in this
setting.

Presentation wise, the authors introduce a small version of MORUS which, along
with excellent
figures, makes it significantly easier to grasp the details of the attack. It
is also nice to see
that the authors explicitly derive formulas for the linear expressions, and
additionally perform
experiments to back up these derivations.

For the two other attacks, while maybe not as interesting, the presentation is
equally clear and
easy to follow. All in all, the paper is well written as present valuable
cryptanalysis of a CAESAR
finalist. I recommend acceptance.

Some editorial comments:

- Page 6: Under finalization, tmp isn't defined. 
- The word "Symmetric" is a bit confusing. Maybe change to "Rotation
invariant"?
- It might be good to add the LAT for AND somewhere, as a help to the reader.

- Figure 3: Good, but needs a much more descriptive caption. Explain what the
numbers are, what the
x's are, annotate with round-numbers.
- I could use a bit more detail about the experimental verification in section
4.4.
- Page 21: "beginning and end of finalization" -> initialization
- Table 4 and 5 should be in the main body.

========================================================================

This paper presents new cryptanalysis against MORUS, one of the three
finalists in the CAESAR authenticated encryption competition in the group for
use-case high-speed software performance.

The main result of this paper is a keystream distinguisher on full MORUS-1280
based on linear approximations of its StateUpdate-function. 
It does not depend the secret key and thus does not allow key recovery, so in
principle it can only recover unknown bits of plaintext.
Against full MORUS-1280 with 256 bits of claimed security, one can recover
plaintext bits using about 2^152 encrypted blocks *under various keys*, which
is clearly impractical.
While MORUS claims 256-bit security and restricts each key to at most 2^64
encryptions, this does not hamper this attack.
Even though the data complexity of 2^152 is impractically large, as this is
the same scenario as the real-world decrypting plaintext attacks on RC4 in
TLS, this attack should be taken as a certificational break of full
MORUS-1280.

The linear approximation on the StateUpdate function can be seen as rotational
linear cryptanalysis, where every used linear mask on a 'register' (256-bit
for MORUS-1280) uses the same bits of each 'word' within a register (a
'register' consists of 4 'words').
The authors call this property 'symmetric', and this property is maintained
under various operations on linear masks: XORing two linear masks, performing
the same bit-wise rotation on each word in a register, and performing
word-wise rotations on a register.
This simplifies analysis and naturally gives rise to analysis where one just
considers the linear mask on the first word of each register (as the other
words have the same linear mask).
The authors introduce minimorus where each register just contains a single
word and remove all word-wise rotations.

The authors introduce a simplified analysis on simplified minimorus which
showcases their results and allowed experimental verification by the authors.
The full attack is then an extension of this analysis, where the authors
analyze the interaction of the non-linear components.

The paper is well written and provides new cryptanalysis, besides the main
result on full MORUS-1280 the paper includes attacks against round-reduce
initialization and finalization.

- consider renaming rotations similar to the existing terminology of 'bit-wise
rotation of machine registers'. i.e., X-wise rotation does rotation in steps
of unit X. This means that 'word-wise' rotation becomes 'bit-wise' rotation,
and 'register-wise' rotation becomes 'word-wise' rotation in this paper.
- consider renaming 'symmetric property' to 'rotational property' to make the
link to rotational differential cryptanalysis and to make it more specific
that it is a rotational symmetry instead of e.g. reflection symmetry.

========================================================================

Morus is a stream cipher based authenticated encryption scheme, typically
designed for high-performance applications. It is one of the finalist of the
CAESAR competition. As the major contribution of the paper, the authors have
presented a linear correlation in the keystream of full MORUS. For MORUS-1280
(-640), the authors have showed a correlation of $2^{-76}$ ($2^{-73}$ resp.),
which can be used to mount a distinguishing attack after $2^{152}$ ($2^{146}$
resp.) encryptions. This attack depicts that MORUS-1280-256 doesn't offer 
$256$-bit security for confidentiality, as claimed by the designers. The
authors have also presented one integrity attack and one key-recovery attacks
on reduced round MORUS.

The overall presentation of the paper is really good. The concept of using
symmetric linear combinations to deal with the bit rotations is novel. I also
appreciate the idea of building the linear trail for a reduced version (called
MiniMORUS) first, and then extending it to full MORUS using symmetrization.

The editorial quality of the paper is also good. Some minor issues of the
papers are:
1. abstract: “less than would be” -> less than what would be.
2. Same notation $C_i^t$ is used to denote a ciphertext as well as an equation
(page 11). 
3. “That the correlation ... 9 times” -> change the sentence.
4. Capitalization is not consistent (Keystream correlation, Data complexity,
Differential Trails, Full Diffusion Rounds etc).

Overall, the reported results of this paper are very interesting, especially
MORUS being one of the finalist of the ongoing CAESAR competition. Although
the reported attacks are not practical, it certainly breaks the claimed
security of MORUS-1280-256 and depicts some major weaknesses in the inherent
design of MORUS such as:
(i) It is possible to build trails irrespective of the rotation constants,
	(i)	(ii) Finalization phase involves some unnecessary non-linear computations
which are ultimately not used during the tag generation.
	(ii)	
With the light of the ongoing final round of the CAESAR competition, these
results are very relevant and might play a crucial role in the
selection/rejection of MORUS in the final portfolio; and hence I think the
work should be accepted.

========================================================================

This paper proposes several attacks on a family of authenticated encryption
schemes MORUS, which is one of finalists of the CAESAR competition. The
authors found linear correlations in the keystreams with 2^-73 and 2^-76 of on
MORUS-640 and MORUS-1280, respectively. These biases enable to develop
distinguishing attacks with 2^146 and 2^152 encryptions, respectively. Since
these linear correlations are independent from values of key, a typical
countermeasure such as key updating does not make sense, and broadcast setting
where the same plaintext is encrypted with different keys is directly applied.

Also, they analyze the reduced version of MORUS, and present forgery and key
recovery attacks on several reduced variants of MORUS.

The strongest result in this paper is the distinguishing attack on MORUS-1280,
because this attack successfully breaks the claimed security of 256-bit key
version of MORUS-1280 for the first time, while the required time and data
complexity is still very high and impractical.
Linear correlation attacks to stream ciphers are not new techniques but
finding good trails are non-trivial task. Furthermore, these attacks are
nicely backed up by experimental verifications of toy models. On the other
hands, the security analysis of the reduced variant of MORUS is less
interesting.

In the plaintext recovery attack in the broad setting, it requires more
samples than distinguishing attacks. So, the attack does not match the
practical attack scenario such as BEAST.

Overall, since MORUS is one of finalists of CAESAR, third party cryptanalysis
is valuable, I believe that the result should be published from this
conference

-The fact that correlations are independent from the key is not uncommon in
stream ciphers when the weakness of the key generation function is exploited,
e.g. RC4.

-In the plaintext recovery setting, it requires more samples than
distinguishing attacks. So, the attack does not match practical attack
scenario of BEAST.

========================================================================

This paper provides a cryptanalysis of the authenticated encryption MORUS,
finalist of the CAESAR competition. Its main contribution relies on a linear
correlation in the keystream of the full version of MORUS, which provides a
distinguisher on its output. The attack complexity for MORUS-1280 is lower
than what we would expect for a 256-bit security scheme. Two other smaller
contributions of this paper are forgery attack and key-recovery attack in the
nonce-misuse setting both in reduced versions of the authenticated encryption
scheme.

The editorial quality of this paper makes it very easy to follow. The authors
start with a concrete description of MORUS. Then they introduce a lighter
scheme they refer to as MiniMORUS to describe a simple version of their main
attack. Afterwards, the concrete attack of full MORUS is given, followed by
two extra contributions on round reduced versions of the scheme.

Strengths of this paper are listed below.

1) The main strength of this paper is the impact of the linear correlation
attack. Namely, the provided attack provides a way to distinguish MORUS'
output from random with a lower complexity than expected for such security.

2) The preliminary description of the attack on MiniMORUS is very appreciable
to understand the full attack. The trail fragments depicted in Figure 2 are
very helpful to get the procedure. Basically, the idea is to write equations
between ciphertext and registers in the state at different steps, using a zero
plaintext. Then, dependencies and equations' weights allow to conclude the
approximation.

3) The source code of MiniMORUS' attack is provided as a supplementary
material.

4) Going from MiniMORUS to full-MORUS is done through an elegant use of
symmetries. Furthermore, changing the zero plaintext to a known one is pretty
straightforward as the considered dependencies are linear.

5) The two extra attacks on initialization and finalization are not concrete
attacks in the sense that they only apply on reduced versions of MORUS.
Nevertheless, they nicely complete the analysis of MORUS.

A small drawback that I would add is that this paper would benefit from a
concrete comparison with similar attacks, as the previous one that was found
on AEGIS. While it does not low down the impact of the attack on a new scheme
it could help to understand the differences for further cryptanalysis.

Moreover, the provided attack works either if the key is fixed or if the same
plaintext is repeated a large number of times. While the former case is not a
big limitation in practice even considering design recommendations, the later
seems to be less probable in practice. Maybe examples of concrete scenarii
should be emphasized to strengthen the proposal. Nevertheless, as soon as it
is possible to mount this main attack under certain circumstances and because
this scheme is currently under evaluation, I think such work is mandatory to
determine the concrete security of proposals.

All in one, this paper presents huge contributions that totally fit the scope
of Asiacrypt conference. While I would appreciate more feedback on previous
related works, the paper is very well written and advantageously describes
every attack precisely enough for a 'non-symmetric' audience. For these
reasons, I lean towards acceptance.

Minor comments:
- few typos in README.nb

### 2018-07-10 AC18_rebuttal.txt
added email discussion to rebuttal

@@
 
-We thank all reviewers for taking the time to review this submission.
+We thank all reviewers for their positive comments.
 We will take all editorial comments into account to prepare an updated version.
@@
 
-... TODO ...
+... TODO (Brice?)...
 
@@
 ... TODO ...
+
+> In the plaintext recovery attack in the broad setting, it requires more
+> samples than distinguishing attacks. So, the attack does not match the
+> practical attack scenario such as BEAST.
+
+... TODO ...
+
+  Brice wrote:
+  "We are not aware of a distinguishing attack on MORUS with lower complexity than our plaintext recovery attack.
+Note that MORUS is not a block cipher, and distinguishing attacks on modes of operation (such as the CBC mode targeted in the original BEAST attack) do not apply."
+  A small possibility is that in the broadcast setting of BEAST the same nonce would be used with different keys.
+If that's true the reviewer could be thinking of a generic multi-target key-guessing attack using the initial known plaintext segment, and they would be kind of right, except it's not a distinguishing attack.
+
+  Gaëtan wrote:
+  "Another option is that he might be talking about the small constant factors that are needed to get a high probability of success (with 1/epsilon^2 data we only have a small advantage)..
+There is also a factor in the complexity depending on the number of bits/bytes that we want to recover (that we didn't consider in the paper, but are needed in order to recover a password or a web cookie)."
 

### 2018-07-10 AC18_rebuttal.txt
updated rebuttal

@@
 
-... TODO ...
+As discussed in Section 6, repeated plaintexts actually appear in many practical
+applications due to fixed formats (headers, metadata, file formats) and often
+include some repeated secret parts (password, authentication cookie, ...).
+This scenario has been exploited in several practical attacks, including BEAST
+(TLS), RC4 attacks (TLS, WPA), and Sweet32 (TLS, OpenVPN).
 
@@
 
-... TODO ...
+Clearly, the complexity of the attack we propose is not practical, unlike the
+attacks on legacy cryptosystems in other BEAST-style attacks.
+However, the attack complexity is less than suggested by the security claim of
+MORUS1280. We are not aware of a distinguishing attack on MORUS with lower
+complexity than our plaintext recovery attack. Note that MORUS is not a block
+cipher, and distinguishing attacks on modes of operation (such as the CBC mode
+targeted in the original BEAST attack) do not apply.
 
   Brice wrote:
-  "We are not aware of a distinguishing attack on MORUS with lower complexity than our plaintext recovery attack.
-Note that MORUS is not a block cipher, and distinguishing attacks on modes of operation (such as the CBC mode targeted in the original BEAST attack) do not apply."
   A small possibility is that in the broadcast setting of BEAST the same nonce would be used with different keys.

### 2018-07-10 morusAC_06_Discussion.tex
updated rebuttal

@@
 in AEGIS~\cite{sacryptMinaud14}.  As such, they can be leveraged to
-mount an attack in the broacast setting, where the same message is
+mount an attack in the broadcast setting, where the same message is
 encrypted multiple times with different IVs and potentially different

### 2018-07-12 AC18_rebuttal.txt
added comparison to AEGIS to rebuttal

@@
 
-... TODO (Brice?)...
+Broadly speaking, the attack on MORUS is similar to traditional linear attacks
+on stream ciphers. It is closest to the attack on AEGIS. In both cases, the
+approach is to build a linear trail linking ciphertext bits, while canceling the
+contribution of inner state bits. How the trail is built depends on the state
+update function, and differs significantly between AEGIS and MORUS, as the AEGIS
+state update function relies on the parallel application of AES rounds, rather
+than shift/AND/XOR operations.
 
@@
 Clearly, the complexity of the attack we propose is not practical, unlike the
-attacks on legacy cryptosystems in other BEAST-style attacks.
-However, the attack complexity is less than suggested by the security claim of
-MORUS1280. We are not aware of a distinguishing attack on MORUS with lower
-complexity than our plaintext recovery attack. Note that MORUS is not a block
-cipher, and distinguishing attacks on modes of operation (such as the CBC mode
-targeted in the original BEAST attack) do not apply.
+attacks on legacy cryptosystems in other BEAST-style attacks.  However, the
+attack complexity is less than suggested by the security claim of MORUS1280. We
+are not aware of a distinguishing attack on MORUS with lower complexity than our
+plaintext recovery attack. Note that MORUS is not a block cipher, and
+distinguishing attacks on modes of operation (such as the CBC mode targeted in
+the original BEAST attack) do not apply.
 

### 2018-07-12 AC18_rebuttal.txt
Benoît's suggestion
rebuttal tweak following Yu's comment

@@
 We will take all editorial comments into account to prepare an updated version.
-Below, we discuss some specific comments raised by the reviewers:
+Below, we discuss some specific comments raised by the reviewers. We agree with
+those comments and will update the paper with the relevant clarifications.
 
@@
 attacks on legacy cryptosystems in other BEAST-style attacks.  However, the
-attack complexity is less than suggested by the security claim of MORUS1280. We
-are not aware of a distinguishing attack on MORUS with lower complexity than our
-plaintext recovery attack. Note that MORUS is not a block cipher, and
+attack complexity is less than suggested by the security claim of MORUS1280-256.
+We are not aware of a distinguishing attack on MORUS with lower complexity than
+our plaintext recovery attack. Note that MORUS is not a block cipher, and
 distinguishing attacks on modes of operation (such as the CBC mode targeted in

### 2018-09-06 AC2018_camera_ready/morusAC.tex
Camera-ready version

--- 2018-05-11 morusAC.tex
@@
 \author{Tomer Ashur\inst{1} \and Maria Eichlseder\inst{2} \and Martin M.
-Lauridsen \and Ga\"etan Leurent\inst{3} \and Brice Minaud\inst{4} \and\\ Yann Rotella\inst{3} \and Yu Sasaki\inst{5} \and Beno\^it Viguier\inst{6}}
+Lauridsen \and Ga\"etan Leurent\inst{3} \and Brice Minaud\inst{4} \and Yann Rotella\inst{3} \and Yu Sasaki\inst{5} \and Beno\^it Viguier\inst{6}}
 \institute{
-imec-COSIC KULeuven, Belgium
+imec-COSIC, KU Leuven, Belgium
 \and
@@
 \and
-Royal Holloway, University of London, United Kingdom
+Royal Holloway University of London, United Kingdom
 \and
@@
 \else
-\usepackage[hmargin=2.5cm,vmargin=3cm]{geometry}
+%\usepackage[hmargin=2.5cm,vmargin=3cm]{geometry}
 \fi
+
+\pagestyle{empty}
 
@@
 \maketitle
-\thispagestyle{plain}
+%\thispagestyle{empty}
 
@@
 
-\ifanonymous
-\else
-  %\subsubsection{Acknowledgments.}
-  %...
-  %\horizonack{The research leading to these results has received funding from
-  %the European Union's Horizon 2020 research and innovation programme under
-  %grant agreement No 644052 (HECTOR) and \dots.}
-\fi
-
 \bibliographystyle{splncs03}
@@
 
-%\section*{Auxiliary Material}
-\clearpage
+\ifanonymous
+\section*{Auxiliary Material}
+\else
+%\clearpage
+\fi
 

### 2018-09-06 AC2018_camera_ready/morusAC_00_Abstract.tex
Camera-ready version

--- 2018-05-17 morusAC_00_Abstract.tex
@@
   which can be exploited after around $2^{152}$ encryptions, less than
-  would be expected for a 256-bit secure cipher.  For
+  what would be expected for a 256-bit secure cipher.  For
   \MORUS[640], the same attack results in a correlation of $2^{-73}$, which does not
@@
 %
-  To identify this correlation, we make use of rotational symmetries in \MORUS
+  To identify this correlation, we make use of rotational invariants in \MORUS
   using linear masks that are invariant by word-rotations of the

### 2018-09-06 AC2018_camera_ready/morusAC_01_Introduction.tex
Camera-ready version

--- 2018-05-10 morusAC_01_Introduction.tex
@@
 To obtain this result, we propose a simplified abstraction of \MORUS, called \MiniMORUS.
-\MiniMORUS takes advantage of certain rotational symmetries in \MORUS and simplifies the description and analysis of the attack.
+\MiniMORUS takes advantage of certain rotational invariants in \MORUS and simplifies the description and analysis of the attack.
 We then show how the attack can be extended from \MiniMORUS to the real \MORUS.
@@
 We first provide a brief description of \MORUS in \Cref{sec/Preliminaries}.
-In \Cref{sec/introminimorus}, we introduce \MiniMORUS, an abstraction of \MORUS based on a certain class of rotational symmetries.
+In \Cref{sec/introminimorus}, we introduce \MiniMORUS, an abstraction of \MORUS based on a certain class of rotational invariants.
 We analyze this simplified scheme in \Cref{sec/minitrails} and provide a ciphertext-only linear approximation with a weight of 16.

### 2018-09-06 AC2018_camera_ready/morusAC_02_Preliminaries.tex
Camera-ready version

--- 2018-05-10 morusAC_02_Preliminaries.tex
@@
 The additional input $m^t$ is used in rounds 2 to 5, but not in round 1.
-Each round uses the word-wise rotation (left circular shift) operation,
+Each round uses the bit-wise rotation (left circular shift) operation inside word,
 denoted $\lll_w$ in the following and \texttt{Rotl\_xxx\_yy} in the design document.
 It divides a $q$-bit register value into 4 words of $w = q/4$ bits, and performs a rotation on each $w$-bit word.
-The word-wise rotation constants $b_i$ for round $i$ are defined in \Cref{Tbl:rcon}.
+The bit-wise rotation constants $b_i$ for round $i$ are defined in \Cref{Tbl:rcon}.
 Additionally, each round uses rotations on a whole $q$-bit register by a multiple of the word size,
 denoted $\lll$ in the following and \texttt{<<<} in the design document.
-The register-wise rotation constants $b_i'$ are also listed in \Cref{Tbl:rcon}.
+The word-wise rotation constants $b_i'$ are also listed in \Cref{Tbl:rcon}.
 %\todo{rounds $0\dots4$ or $1\dots5$?}
@@
     \toprule
-    & \multicolumn{5}{@{}c@{~}}{Word-wise rotation $\lll_w$} & \multicolumn{5}{@{}c@{}}{Register-wise rotation $\lll$} \\
+    & \multicolumn{5}{@{}c@{~}}{Bit-wise rotation $\lll_w$} & \multicolumn{5}{@{}c@{}}{Word-wise rotation $\lll$} \\
     \cmidrule(r){2-6}
@@
 %\begin{equation*}
-$S^{t+1} \leftarrow \StateUpdate{} (S^t, {\it tmp})$.
+$S^{t+1} \leftarrow \StateUpdate{} (S^t, L)$.
 %\end{equation*}
@@
 where $\log_2()$ denotes logarithm in base 2.
-By the piling-up lemma, the correlation (resp.
+By the Piling-Up Lemma, the correlation (resp.
 weight) of an XOR of independent variables is equal to the product (resp.
@@
 Recall that registers of \MORUS are composed of four words.
-We now define register-wise and word-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \MORUS design document):\\
+We now define word-wise and bit-wise rotations (denoted by \texttt{<<<} and \texttt{Rotl\_xxx\_yy} respectively in the \MORUS design document):\\
 %\centerline{

### 2018-09-06 AC2018_camera_ready/morusAC_03_MiniMorus.tex
Camera-ready version

--- 2018-05-10 morusAC_03_MiniMorus.tex
@@
 
-\section{Symmetric Linear Combinations and \MiniMORUS}
+\section{Rotational Invariant Linear Combinations and \MiniMORUS}
 \label{sec/introminimorus}
@@
 
-\subsection{Symmetric Linear Combinations}
+\subsection{Rotational Invariant Linear Combinations}
 
@@
 \begin{enumerate}
-\item \emph{word-wise} rotations perform a circular shift on each word within a register;
-\item \emph{register-wise} rotations perform a circular shift on a whole register.
+\item \emph{bit-wise} rotations perform a circular shift on each word within a register;
+\item \emph{word-wise} rotations perform a circular shift on a whole register.
 \end{enumerate}
 The second type of rotation always shifts registers by a multiple of the word size $w$.
-This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a register-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
+This amounts to a (circular) permutation of the words within the register: for example, if a register contains the words $(A, B, C, D)$, and a word-wise rotation by $w$ bits to the left is performed, then the register now contains the words $(B, C, D, A)$.
 
@@
 \]
-is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
+is said to be \emph{rotational invariant} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
 \[
@@
 \end{definition}
-\emph{Example.} The following linear combination is symmetric for \MORUS[640], i.e. $w = 32$:
+\emph{Example.} The following linear combination is rotational invariant for \MORUS[640], i.e. $w = 32$:
 \begin{equation}
@@
 This definition naturally extends to a linear combination across multiple registers, and also across ciphertext blocks.
-The value of such a linear combination is unaffected by register-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since word-wise rotations always shift all four words within a register by the same amount, word-wise rotations preserve the symmetry property.
-Moreover, the XOR of two symmetric linear combinations is also symmetric.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
+The value of such a linear combination is unaffected by word-wise rotations, since those rotations always shift registers by a multiple of the word size.
+On the other hand, since bit-wise rotations always shift all four words within a register by the same amount, bit-wise rotations preserve the rotational invariant property.
+Moreover, the XOR of two rotational invariant linear combinations is also rotational invariant.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
 
-This naturally leads to the idea of building a linear trail using only symmetric linear combinations, which is what we are going to do.
-As a result, the effect of register-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be symmetric, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is symmetric, is that it involves the same bits in each word within a register.
+This naturally leads to the idea of building a linear trail using only rotational invariant linear combinations, which is what we are going to do.
+As a result, the effect of word-wise rotations can be ignored.
+Moreover, since all linear combinations we consider are going to be rotational invariant, they can be described by truncating the linear combination to the first word of a register.
+Indeed, an equivalent way of saying a linear combination is rotational invariant, is that it involves the same bits in each word within a register.
 For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
@@
 In fact, we can go further and consider a reduced version of \MORUS where each register contains a single word instead of four.
-The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see \Cref{fig:minimorus}.
+The \StateUpdate{} function is unchanged, except for the fact that word-wise rotations are removed: see \Cref{fig:minimorus}.
 We call these reduced versions \MiniMORUS[640] and \MiniMORUS[1280], for \MORUS[640] and \MORUS[1280] respectively.
-Since registers in \MiniMORUS contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.
+Since registers in \MiniMORUS contain a single word, bit-wise and word-wise rotations are the same operation; for simplicity we write $\lll$ for bit-wise rotations.
 
-Since the trail we are building is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous symmetry.
+Since the trail we are building is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous rotational invariant property.
 

### 2018-09-06 AC2018_camera_ready/morusAC_04_Mini_trails.tex
Camera-ready version

--- 2018-05-15 morusAC_04_Mini_trails.tex
@@
 
-  \caption{\MiniMORUS: two approximations for $S_{2,0}^2$.}
+  \caption{\MiniMORUS: two approximations for $S_{2,0}^2$.
+Numbers in each diagram denote bit positions used in the linear approximation, i.e.~subscripts of $\alpha,\beta,\gamma,\delta$ and $\varepsilon$.
+$\chi_1$ and $\chi_2$ are two halves of the full trail which we experimentally verify.}
   \label{fig:twoapproximations}
@@
 
-The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ as follows.
+The equation corresponding to each of the five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$ may be written explicitly as $\mathbf{A}^t_i$, $\mathbf{B}^t_i$, $\mathbf{C}^t_i$, $\mathbf{D}^t_i$, $\mathbf{E}^t_i$ as follows.
 For each equation, we write on the left-hand side of the equality the biased linear combination used in the trail; and on the right-hand side, the remainder of the equation, which must have non-zero correlation (in all cases the correlation is $2^{-1}$).
 \begin{alignat*}{2}
-A^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
-B^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
-C^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
-D^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
-E^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
+\mathbf{A}^t_i:\quad&& C^t_i \oplus S^{t+1}_{0,i+b_0} &= S^t_{1,i} \oplus S^t_{3,i} \oplus S^t_{1,i} \cdot S^t_{2,i} \oplus S^t_{2,i} \cdot S^t_{3,i}\\
+\mathbf{B}^t_i:\quad&& C^t_i \oplus S^t_{0,i} \oplus S^t_{1,i} &= S^t_{2,i} \cdot S^t_{3,i} \\
+\mathbf{C}^t_i:\quad&& S^t_{1,i} \oplus S^{t+1}_{1,i+b_1} \oplus S^t_{4,i} &= S^t_{2,i} \cdot S^t_{3,i}\\
+\mathbf{D}^t_i:\quad&& S^t_{4,i} \oplus S^{t+1}_{4,i+b_4} \oplus S^{t+1}_{2,i} &= S^{t+1}_{0,i} \cdot S^{t+1}_{1,i}\\
+\mathbf{E}^t_i:\quad&& S^t_{2,i} \oplus S^{t+1}_{2,i+b_2} \oplus S^{t+1}_{0,i} &= S^t_{3,i} \cdot S^t_{4,i}
 \end{alignat*}
@@
 
-In order to build the equation for the full trail, we start with $E^2_0$:
+In order to build the equation for the full trail, we start with $\mathbf{E}^2_0$:
 \[
@@
 \]
-In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add to the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
+In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add to the equation $\mathbf{A}^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$).
 This yields:
@@
 We then need to cancel two terms of the form $S^t_{2,i}$.
-To do this, we add to the equations $D^t_i$ for appropriate choices of $t$ and $i$.
+To do this, we add to the equations $\mathbf{D}^t_i$ for appropriate choices of $t$ and $i$.
 This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms.
-By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
-By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
-Finally, applying $A^t_i$ eight times allows to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
+By using equation $\mathbf{B}^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms.
+By applying equation $\mathbf{B}^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms).
+Finally, applying $\mathbf{A}^t_i$ eight times allows to replace these eight $S^t_{0,i}$ terms by only ciphertext bits.
 Ultimately, for \MiniMORUS[1280], this yields the equation:
@@
 
-That the correlation is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
+%That the correlation is so high may be surprising: indeed, the full trail uses trail fragment $\varepsilon^t_i$ 1 time, $\delta^t_i$ 2 times, $\gamma^t_i$ 4 times, $\beta^t_i$ 8 times, and $\alpha^t_i$ 9 times.
+The correlation is surprising high.
+The full trail uses trail fragments $\varepsilon^t_i$, $\delta^t_i$, $\gamma^t_i$, $\beta^t_i$, and $\alpha^t_i$, once, twice, 4 times, 8 times, and 9 times, respectively.
 Since each trail fragment has a weight of 1, this would suggest that the total weight should be $1+2+4+8+9 = 24$ rather than 16.
-However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $B^t_i$ and $C^t_i$ is equal).
+However, when combining trail fragments $\beta_i$ and $\gamma_i$, notice that the same AND is computed at the same step between registers $S_2$ and $S_3$ (equivalently, notice that the right-hand side of equations $\mathbf{B}^t_i$ and $\mathbf{C}^t_i$ is equal).
 In both cases it is approximated by zero.

### 2018-09-06 AC2018_camera_ready/morusAC_05_Full_trails.tex
Camera-ready version

--- 2018-05-15 morusAC_05_Full_trails.tex
@@
 
-\subsection{Symmetrizing the Trail}
-In order to build a trail for the full \MORUS, we proceed exactly as we did for \MiniMORUS, following the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
-That is, for every $S^t_{i,j}$ (resp.
-$C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
+%\subsection{Symmetrizing the Trail}
+\subsection{Making the Trail Rotational Invariant}
+In order to build a trail for the full \MORUS, we proceed exactly as we did for \MiniMORUS, following the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ rotational invariant, in the sense of \Cref{sec/introminimorus}.
+That is, for every $S^t_{i,j}$ (resp. $C^t_j$) component in every trail fragment and every equation, we
+%\emph{symmetrize}
+expand the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp.
 $C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size.
-For example, if $w=64$ (for \MORUS[1280]), the term $S^3_{2,0}$ is symmetrized into:
+For example, if $w=64$ (for \MORUS[1280]), the term $S^3_{2,0}$ is expanded into:
+%symmetrized into:
 \[
@@
 
-Thus, translating the trail from one of the \MiniMORUS ciphers to the corresponding full \MORUS cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \MiniMORUS in the first place.
-Concretely, in order to build the full trail equation for \MORUS, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same manner as before.
+Thus, translating the trail from one of the \MiniMORUS ciphers to the corresponding full \MORUS cipher amounts to making every linear combination rotational invariant---indeed, that was the point of introducing \MiniMORUS in the first place.
+Concretely, in order to build the full trail equation for \MORUS, we write rotational invariant versions of equations $\mathbf{A}^t_i$, $\mathbf{B}^t_i$, $\mathbf{C}^t_i$, $\mathbf{D}^t_i$, $\mathbf{E}^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same manner as before.
 This way, the biased linear combination on \MiniMORUS[1280] given in \Cref{sec:minitraileq}, namely:
@@
 \end{align*}
-ultimately yields the following biased symmetrized linear combination on the full \MORUS[1280]:
+ultimately yields the following biased rotational invariant linear combination on the full \MORUS[1280]:
 \begin{align*}
@@
 
-The symmetrized trail on full \MORUS may be intuitively understood as consisting of four copies of the original trail on \MiniMORUS.
-Indeed, the only difference between full \MORUS (for either version of \MORUS) and four independent copies of \MiniMORUS comes from register-wise rotations, which permute words within a register.
-But as observed in \Cref{sec/introminimorus}, register-wise rotations leave symmetric linear combinations invariant; and so, insofar as we only ever use symmetric linear combinations on all registers along the trail, register-wise rotations have no effect.
+The rotational invariant trail on full \MORUS may be intuitively understood as consisting of four copies of the original trail on \MiniMORUS.
+Indeed, the only difference between full \MORUS (for either version of \MORUS) and four independent copies of \MiniMORUS comes from word-wise rotations, which permute words within a register.
+But as observed in \Cref{sec/introminimorus}, word-wise rotations preserves rotational invariant property; and so, insofar as we only ever use rotational invariant linear combinations on all registers along the trail, word-wise rotations have no effect.
 
-Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \MiniMORUS trail, namely 64 for both \MORUS[1280] and \MORUS[640].
-However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
+Following the previous intuition, one may expect that the weight of the full trail should simply be four times the weight of the corresponding \MiniMORUS trail, namely 64 for both \MORUS[1280] and \MORUS[640].
+However, reality is a little more complex, as the full trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
 
 To understand why that might be the case, assume a nonlinear term $S^0_{2,0} \cdot S^0_{3,0}$ arising from some part of the trail, and another term $S^0_{2,0} \cdot S^0_{3,w}$ arising from a different part of the trail (where $w$ denotes the word size).
-Then when we XOR the various trail fragments together, in \MiniMORUS these two terms are actually equal and will cancel out, since register-wise rotations by multiples of $w$ bits are ignored.
+Then when we XOR the various trail fragments together, in \MiniMORUS these two terms are actually equal and will cancel out, since word-wise rotations by multiples of $w$ bits are ignored.
 However in the real \MORUS these terms are of course distinct and do not cancel each other.
@@
 \Cref{fig:trailcollision}).
-While in \MiniMORUS, these terms cancel out, in the full \MORUS, when adding all symmetric copies of the trail, we end up with the sum:
+While in \MiniMORUS, these terms cancel out, in the full \MORUS, when adding all rotational invariant copies of the trail, we end up with the sum:
 {
@@
 
-Regarding \Cref{item:2}, recall that the full trail equation is a linear combination of (the symmetrized version of) equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ in \Cref{sec:minitraileq}.
-Also observe that in each equation, state bits that are shifted by a word-wise rotation only contribute linearly.
-Because plaintext bits are XORed into each register at the same time word-wise rotation is performed, this implies that plaintext bits resulting from \Cref{item:2} also only contribute linearly.
+Regarding \Cref{item:2}, recall that the full trail equation is a linear combination of (the rotational invariant version of) equations $\mathbf{A}^t_i$, $\mathbf{B}^t_i$, $\mathbf{C}^t_i$, $\mathbf{D}^t_i$, $\mathbf{E}^t_i$ in \Cref{sec:minitraileq}.
+Also observe that in each equation, state bits that are shifted by a bit-wise rotation only contribute linearly.
+Because plaintext bits are XORed into each register at the same time bit-wise rotation is performed, this implies that plaintext bits resulting from \Cref{item:2} also only contribute linearly.
 In fact in all cases, it so happens that updating the equation to take plaintext variables into account simply involves XORing in the plaintext bit $M^t_i$.

### 2018-09-06 AC2018_camera_ready/morusAC_06_Discussion.tex
Camera-ready version

--- 2018-07-10 morusAC_06_Discussion.tex
@@
 
-\subsubsection{Keystream correlation.}
+\subsubsection{Keystream Correlation.}
 
@@
 
-\subsubsection{Data complexity.}
+\subsubsection{Data Complexity.}
 
@@
 
-\subsubsection{Design considerations.}
+\subsubsection{Design Considerations.}
 The existence of this trail does hint at some weakness in the
@@
 build a trail that ignores rotation constants may be surprising. This
-would have been prevented by adding a word-wise rotation to one of the
+would have been prevented by adding a bit-wise rotation to one of the
 state registers at the input of the ciphertext equation.

### 2018-09-06 AC2018_camera_ready/morusAC_07_IniFin.tex
Camera-ready version

--- 2018-05-10 morusAC_07_IniFin.tex
@@
 \label{Tbl:fin_diff}
-{\tiny
+{%
+\iflncsmargins
+\tiny
+\else
+\scriptsize
+\renewcommand{\arraystretch}{0.9}
+\fi
 \begin{tabular}{ccccccc} \hline
@@
       & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 0  & \\
-      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
-5     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-1}$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \multirow{3}{*}{$2^{-1}$}\\
+5     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & \\
       & {\tt 0020000000000080} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
@@
       & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
-      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \\
-6     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-3}$\\
+      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000400000000000} & {\tt 0000000000000000} & 1  & \multirow{3}{*}{$2^{-3}$}\\
+6     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1 & \\
       & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
@@
       & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
-      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \\
-7     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & $2^{-6}$\\
+      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \multirow{3}{*}{$2^{-6}$}\\
+7     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000004000000000} & 1  & \\
       & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
@@
       & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
-      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \\
-8     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & $2^{-10}$\\
+      & {\tt 0004400001000000} & {\tt 0000000000000000} & {\tt 0000000010000000} & {\tt 0000000000000000} & 4  & \multirow{3}{*}{$2^{-10}$}\\
+8     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & \\
       & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0020000000000080} & 2  & \\
@@
       & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
-      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
-9     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & $2^{-14}$\\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \multirow{3}{*}{$2^{-14}$}\\
+9     & {\tt 0400014000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & 4  & \\
       & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7  & \\
@@
       & {\tt 0000000000000000} & {\tt 0000000000100004} & {\tt 0000000000000000} & {\tt 0000000000000000} & 2  & \\
-      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
-10    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-20}$\\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \multirow{3}{*}{$2^{-20}$}\\
+10    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & \\
       & {\tt 0220000080000080} & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & 7  & \\
@@
       & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9  & \\
-      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  & \\
-11    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-28}$\\
+      & {\tt 0000000010000000} & {\tt 0000000000000000} & {\tt 0004400001000000} & {\tt 0000000000000000} & 4  &\multirow{3}{*}{$2^{-28}$} \\
+11    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & \\
       & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
@@
       & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & {\tt 0000000008000200} & 9  & \\
-      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \\
-12    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & $2^{-39}$\\
+      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \multirow{3}{*}{$2^{-39}$}\\
+12    & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000001000} & {\tt 0400014000000000} & 4  & \\
       & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
@@
       & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
-      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \\
-13    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 & $2^{-53}$\\
+      & {\tt 0004500005000400} & {\tt 0000000000000000} & {\tt 0040000100000040} & {\tt 4000000000000000} & 10 & \multirow{3}{*}{$2^{-53}$}\\
+13    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 &\\
       & {\tt 0000000000000000} & {\tt 0000000800000000} & {\tt 1000000000004000} & {\tt 0220000080000080} & 7  & \\
@@
       & {\tt 0000000008000200} & {\tt 0000100000100044} & {\tt 0000000200008000} & {\tt 0001000000000000} & 9  & \\
-      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 & \\
-14    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 & $2^{-69}$\\
+      & {\tt 0040000100000040} & {\tt 4000000000000000} & {\tt 0004500005000400} & {\tt 0000000000000000} & 10 & \multirow{3}{*}{$2^{-69}$}\\
+14    & {\tt 0400114000040000} & {\tt 0020000000000080} & {\tt 0004000000400000} & {\tt 0000800100005002} & 14 & \\
       & {\tt 0228000280020080} & {\tt 0000040000000000} & {\tt 2000008000202008} & {\tt 1000004000004021} & 18 & \\
@@
 The recovered state allows the attacker to mount a universal forgery attack under the same nonce.
-However, the key still cannot be recovered because the key is used both at the beginning and end of the finalization, which prevents the attacker from backtracking the state value to the initial state.
+However, the key still cannot be recovered because the key is used both at the beginning and end of the initialization, which prevents the attacker from backtracking the state value to the initial state.
 In this section, we show that meet-in-the-middle attacks allow the attacker to recover the key faster than exhaustive search for a relatively large number of steps, i.e., 10 out of 16 steps in \MORUS[1280].
@@
 
-\subsubsection{Full Diffusion Rounds.}
-We found that {\tt StepUpdate} was designed to have good diffusion in the forward direction.
-Thus, once the state is recovered, the attacker can perform the partial computation in the backward direction longer than the forward direction.
-We set $G_0$ and $G_1$ as follows.
-\begin{align*}
-G_0 &= \{1, 2, \cdots, 127\} & \textrm {Bit position 0 is unknown.}\\
-G_1 &= \{0, 1, \cdots, 7, 9, 10, \cdots, 127\} & \textrm {Bit position 8 is unknown.}
-\end{align*}
-Those will lead to 4 matching bits after the 4-step forward computation and the 6-step backward computation.
-The analysis of the diffusion is given in \Cref{Tbl:ini_diffuse}.
-In the end, $K_{128}$ can be recovered faster than the exhaustive search by 1 bit, i.e., with complexity $2^{127}$.
-
-\begin{table}[!tb]
-\begin{center}
+\begin{table}
 \caption{Analysis of the diffusion and matching bits over 10 steps.
@@
 \label{Tbl:ini_diffuse}
-{\tiny
+\centering
+{
+\iflncsmargins
+\tiny
+\else
+\footnotesize
+\renewcommand{\arraystretch}{0.9}
+\fi
 \begin{tabular}{ccccc} \hline
@@
                          & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
-                         & {\tt 0000000000000000} & {\tt 0000000000000001} & {\tt 0000000000000000} & {\tt 0000000000000001} \\
-$S^{-10}\oplus K_{128} $ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+\multirow{3}{*}{$S^{-10}\oplus K_{128}$} & {\tt 0000000000000000} & {\tt 0000000000000001} & {\tt 0000000000000000} & {\tt 0000000000000001} \\
+  & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
                          & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
@@
                      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
-                     & {\tt 0000000000000000} & {\tt 0000000000000100} & {\tt 0000000000000000} & {\tt 0000000000000100} \\
-$S^0\oplus K_{128} $ & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
+\multirow{3}{*}{$S^0\oplus K_{128} $} & {\tt 0000000000000000} & {\tt 0000000000000100} & {\tt 0000000000000000} & {\tt 0000000000000100} \\
+                     & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
                      & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} & {\tt 0000000000000000} \\
@@
 }
-\end{center}
 \end{table}
+
+\subsubsection{Full Diffusion Rounds.}
+We found that {\tt StepUpdate} was designed to have good diffusion in the forward direction.
+Thus, once the state is recovered, the attacker can perform the partial computation in the backward direction longer than the forward direction.
+We set $G_0$ and $G_1$ as follows.
+\begin{align*}
+G_0 &= \{1, 2, \cdots, 127\} & \textrm {Bit position 0 is unknown.}\\
+G_1 &= \{0, 1, \cdots, 7, 9, 10, \cdots, 127\} & \textrm {Bit position 8 is unknown.}
+\end{align*}
+Those will lead to 4 matching bits after the 4-step forward computation and the 6-step backward computation.
+The analysis of the diffusion is given in \Cref{Tbl:ini_diffuse}.
+In the end, $K_{128}$ can be recovered faster than the exhaustive search by 1 bit, i.e., with complexity $2^{127}$.
 

### 2018-09-06 AC2018_camera_ready/morusAC_01_Introduction.tex
some changes for camera-ready version

@@
 independently propose state-recovery and forgery attacks on \MORUS in a nonce-misuse setting with negligible data and time complexities.
-Finally, a keystream correlation similar in nature to our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist, which shares the same overall structure as \MORUS, but uses a very different, \cipher{AES}-based state update function.
+
+Finally, a keystream correlation similar in nature to our main attack was uncovered by Minaud \cite{sacryptMinaud14} on the authenticated cipher \cipher{AEGIS} \cite{AEGIS,sacryptWuP13}, another CAESAR finalist.
+\cipher{AEGIS} shares the same overall structure as \MORUS, but uses a very different state update function,
+based on the parallel application of \cipher{AES} rounds, rather than the shift/AND/XOR operations used in \MORUS.
+Similar to our attack, the approach in \cite{sacryptMinaud14} is to build a linear trail linking ciphertext bits, while canceling the
+contribution of inner state bits.
+How the trail is built depends primarily on the state update function, and how it lends itself to linear cryptanalysis.
+Because the state update function differs significantly between AEGIS and MORUS, the process used to build the trail is also quite different.
 

### 2018-09-06 AC2018_camera_ready/morusAC_03_MiniMorus.tex
some changes for camera-ready version

@@
 
-\section{Rotational Invariant Linear Combinations and \MiniMORUS}
+\section{Rotational Invariance and \MiniMORUS}
 \label{sec/introminimorus}
@@
 
-\subsection{Rotational Invariant Linear Combinations}
+\subsection{Rotationally Invariant Linear Combinations}
 
@@
 To build our linear trail, we start with a linear combinations of bits within a single register.
-\begin{definition}
+\begin{definition}[Rotational Invariance]
 Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
@@
 \]
-is said to be \emph{rotational invariant} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
+is said to be \emph{rotationally invariant} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
 \[
@@
 \end{definition}
-\emph{Example.} The following linear combination is rotational invariant for \MORUS[640], i.e. $w = 32$:
+\emph{Example.} The following linear combination is rotationally invariant for \MORUS[640], i.e. $w = 32$:
 \begin{equation}
@@
 The value of such a linear combination is unaffected by word-wise rotations, since those rotations always shift registers by a multiple of the word size.
-On the other hand, since bit-wise rotations always shift all four words within a register by the same amount, bit-wise rotations preserve the rotational invariant property.
-Moreover, the XOR of two rotational invariant linear combinations is also rotational invariant.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
+On the other hand, since bit-wise rotations always shift all four words within a register by the same amount, bit-wise rotations preserve the rotational invariance property.
+Moreover, the XOR of two rotationally invariant linear combinations is also rotationally invariant.%; and the same holds for the AND operation (if we extend the symmetric property to non-linear combinations in the natural way).
 
-This naturally leads to the idea of building a linear trail using only rotational invariant linear combinations, which is what we are going to do.
+This naturally leads to the idea of building a linear trail using only rotationally invariant linear combinations, which is what we are going to do.
 As a result, the effect of word-wise rotations can be ignored.
-Moreover, since all linear combinations we consider are going to be rotational invariant, they can be described by truncating the linear combination to the first word of a register.
-Indeed, an equivalent way of saying a linear combination is rotational invariant, is that it involves the same bits in each word within a register.
+Moreover, since all linear combinations we consider are going to be rotationally invariant, they can be described by truncating the linear combination to the first word of a register.
+Indeed, an equivalent way of saying a linear combination is rotationally invariant, is that it involves the same bits in each word within a register.
 For example, in the case of (\ref{eq:symmetric}) above, the four bits involved are the first bit of each of the four words.
@@
 
-Since the trail we are building is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous rotational invariant property.
+Since the trail we are building is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous rotational invariance property.
 

### 2018-09-06 AC2018_camera_ready/morusAC_04_Mini_trails.tex
some changes for camera-ready version

@@
 
-The equation for \MiniMORUS[640] is very similar, and is given in \Cref{sec:traileq}.
+The equation for \MiniMORUS[640] is very similar, and is given in the full version of this paper \cite{fullversion}.
 
@@
 Hence the linear combination of ciphertext bits on the left-hand side has a correlation of $2^{-16}$.
-The same holds for \MiniMORUS[640] (cf.
-\Cref{sec:traileq}).
+The same holds for \MiniMORUS[640].
 
@@
 To confirm that our analysis is correct, we ran experiments on an implementation of \MiniMORUS[1280] and \MiniMORUS[640].
-In both cases, we measured the correlation of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the correlation of the full trail itself (denoted by $\chi$).
+We consider two halves $\chi_1$ and $\chi_2$ of the full trail (depicted on \Cref{fig:twoapproximations}), as well as the full trail itself, denoted by $\chi$.
+In each case, we give the weight predicted by the analysis from the previous section, and the weight measured by our experiments.
+%In both cases, we measured the correlation of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the correlation of the full trail itself (denoted by $\chi$).
 Results are displayed on \Cref{tab:miniapproximations}.
-While our analysis from the previous section predicts a correlation of $2^{-16}$, experiments indicate a slightly better empirical correlation of $2^{-15.5}$ for \MORUS[640].
+While our analysis predicts a correlation of $2^{-16}$, experiments indicate a slightly better empirical correlation of $2^{-15.5}$ for \MORUS[640].
 The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.
@@
   \centerline{
-  \begin{tabular}{@{}llSSS@{}}
+  \begin{tabular}{@{}llSS@{}}
     \toprule
-    & & \multicolumn{3}{@{}c@{}}{Weight} \\
-    \cmidrule{3-5}
+    & & \multicolumn{2}{@{}c@{}}{Weight} \\
+    \cmidrule{3-4}
     \multicolumn{2}{@{}l}{Approximations for \MiniMORUS[640]}
-             & {Exp.} & {Bool.} & {Meas.} \\
+             & {Predicted} & {Measured} \\
     \midrule
     $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
-             & 7    & 7     & 7 \\
+             & 7     & 7 \\
     $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
-             & 9    & 9     & 9 \\
+             & 9     & 9 \\
     $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
-             & 16   & 16    & 15.5 \\
+             & 16    & 15.5 \\
     \midrule
@@
     $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
-             & 7    & 7     & 7 \\
+             & 7     & 7 \\
     $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
-             & 9    & 9     & 9 \\
+             & 9     & 9 \\
     $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
-             & 16   & 16    & 15.9 \\
+             & 16    & 15.9 \\
     \bottomrule

### 2018-09-06 AC2018_camera_ready/morusAC_05_Full_trails.tex
some changes for camera-ready version

@@
 %\subsection{Symmetrizing the Trail}
-\subsection{Making the Trail Rotational Invariant}
-In order to build a trail for the full \MORUS, we proceed exactly as we did for \MiniMORUS, following the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ rotational invariant, in the sense of \Cref{sec/introminimorus}.
+\subsection{Making the Trail Rotationally Invariant}
+In order to build a trail for the full \MORUS, we proceed exactly as we did for \MiniMORUS, following the same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \MiniMORUS to the four-word registers of full \MORUS, we make every term $S^t_{i,j}$ and $C^t_j$ rotationally invariant, in the sense of \Cref{sec/introminimorus}.
 That is, for every $S^t_{i,j}$ (resp. $C^t_j$) component in every trail fragment and every equation, we
@@
 
-Thus, translating the trail from one of the \MiniMORUS ciphers to the corresponding full \MORUS cipher amounts to making every linear combination rotational invariant---indeed, that was the point of introducing \MiniMORUS in the first place.
-Concretely, in order to build the full trail equation for \MORUS, we write rotational invariant versions of equations $\mathbf{A}^t_i$, $\mathbf{B}^t_i$, $\mathbf{C}^t_i$, $\mathbf{D}^t_i$, $\mathbf{E}^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same manner as before.
+Thus, translating the trail from one of the \MiniMORUS ciphers to the corresponding full \MORUS cipher amounts to making every linear combination rotationally invariant---indeed, that was the point of introducing \MiniMORUS in the first place.
+Concretely, in order to build the full trail equation for \MORUS, we write rotationally invariant versions of equations $\mathbf{A}^t_i$, $\mathbf{B}^t_i$, $\mathbf{C}^t_i$, $\mathbf{D}^t_i$, $\mathbf{E}^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same manner as before.
 This way, the biased linear combination on \MiniMORUS[1280] given in \Cref{sec:minitraileq}, namely:
@@
 \end{align*}
-ultimately yields the following biased rotational invariant linear combination on the full \MORUS[1280]:
+ultimately yields the following biased rotationally invariant linear combination on the full \MORUS[1280]:
 \begin{align*}
@@
 \end{align*}
-We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \MORUS[640].
+We refer the reader to the full version of this paper \cite{fullversion} for the corresponding linear combination on \MORUS[640].
 
@@
 
-The rotational invariant trail on full \MORUS may be intuitively understood as consisting of four copies of the original trail on \MiniMORUS.
+The rotationally invariant trail on full \MORUS may be intuitively understood as consisting of four copies of the original trail on \MiniMORUS.
 Indeed, the only difference between full \MORUS (for either version of \MORUS) and four independent copies of \MiniMORUS comes from word-wise rotations, which permute words within a register.
-But as observed in \Cref{sec/introminimorus}, word-wise rotations preserves rotational invariant property; and so, insofar as we only ever use rotational invariant linear combinations on all registers along the trail, word-wise rotations have no effect.
+But as observed in \Cref{sec/introminimorus}, word-wise rotations preserves the rotational invariance property; and so, insofar as we only ever use rotationally invariant linear combinations on all registers along the trail, word-wise rotations have no effect.
 
@@
 \Cref{fig:trailcollision}).
-While in \MiniMORUS, these terms cancel out, in the full \MORUS, when adding all rotational invariant copies of the trail, we end up with the sum:
+While in \MiniMORUS, these terms cancel out, in the full \MORUS, when adding up four copies of the trail to achieve rotational invariance, we end up with the sum:
 {
@@
 Therefore, when computing the full trail equation on \MORUS[640], we get that the weight of the trail is 73 (cf.
-\Cref{sec:traileq}).
+the full version of this paper for the full trail equation for \MORUS[640]).
 
@@
 
-Regarding \Cref{item:2}, recall that the full trail equation is a linear combination of (the rotational invariant version of) equations $\mathbf{A}^t_i$, $\mathbf{B}^t_i$, $\mathbf{C}^t_i$, $\mathbf{D}^t_i$, $\mathbf{E}^t_i$ in \Cref{sec:minitraileq}.
+Regarding \Cref{item:2}, recall that the full trail equation is a linear combination of (the rotationally invariant version of) equations $\mathbf{A}^t_i$, $\mathbf{B}^t_i$, $\mathbf{C}^t_i$, $\mathbf{D}^t_i$, $\mathbf{E}^t_i$ in \Cref{sec:minitraileq}.
 Also observe that in each equation, state bits that are shifted by a bit-wise rotation only contribute linearly.

### 2018-09-07 AC2018_camera_ready/morusAC_04_Mini_trails.tex
eliminated most warnings

@@
   \centering
-  \begin{subfigure}{.32\textwidth}
+  \begin{subfigure}{.325\textwidth}
   \centering
@@
       (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
-      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
+      (xor0) -- (lll0) (lll0) -- (W40) node[below,inner sep=2pt] {$i+b_0$}
       ;
@@
   \hfill
-  \begin{subfigure}{.32\textwidth}
+  \begin{subfigure}{.325\textwidth}
   \centering
@@
       (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
-      (tlll-1) -- (W-20) node[above] {$i$}
-      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
+      (tlll-1) -- (W-20) node[above,inner sep=2.5pt] {$i$}
+      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above,inner sep=2.5pt] {$i$}
       (xnd-1) -- (and-1)
-      (W40) node[below] {\phantom{$i$}}
+      (W40) node[below,inner sep=2pt] {\phantom{$i$}}
       ;
@@
   \hfill
-  \begin{subfigure}{.32\textwidth}
+  \begin{subfigure}{.325\textwidth}
   \tikzsetnextfilename{minimorus_part_gamma}
@@
     \draw[trail, gamma]
-      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
+      (W-21) node[above,inner sep=2.5pt] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
       (xnd1) -- (and1)
       (xor1) -- (txor1)
-      (txor1) -- (W-24) node[above] {$i$}
-      (lll1) -- (W41) node[below] {$i+b_1$}
+      (txor1) -- (W-24) node[above,inner sep=2.5pt] {$i$}
+      (lll1) -- (W41) node[below,inner sep=2pt] {$i+b_1$}
       ;
@@
   \end{subfigure}
-  \bigskip
+  \medskip
 
-  \begin{subfigure}{.32\textwidth}
+  \begin{subfigure}{.325\textwidth}
   \centering
@@
     \draw[trail, delta]
-      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
-      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
-      (lll4) -- (W44) node[below] {$i+b_4$}
+      (W-24) node[above,inner sep=2.5pt] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
+      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below,inner sep=2pt] {$i$}
+      (lll4) -- (W44) node[below,inner sep=2pt] {$i+b_4$}
       ;
@@
     \draw[trail, epsil]
-      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
-      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
-      (lll2) -- (W42) node[below] {$i+b_2$}
+      (W-22) node[above,inner sep=2.5pt] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
+      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below,inner sep=2pt] {$i$}
+      (lll2) -- (W42) node[below,inner sep=2pt] {$i+b_2$}
       ;
@@
   \hfill
-  \begin{subfigure}{.32\textwidth}
+  \begin{subfigure}{.325\textwidth}
     \small
@@
   \centering
-  \resizebox{!}{0.95\textheight}{
+  \resizebox{!}{0.91\textheight}{
   \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
@@
 
-    \begin{scope}[xshift=2.5cm]
+    \begin{scope}[xshift=2.25cm]
       \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
@@
 
-    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
+    \draw[dashed] (\S{4.125}{-1.5}) -- ++(\S{0}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});
 
-    \begin{scope}[yshift=-7.0cm]
+    \begin{scope}[yshift=-6.5cm]
       \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
@@
 
-    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
+    \begin{scope}[xshift=2.25cm,yshift=-6.5cm]
       \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }

### 2018-10-01 AC2018_camera_ready/morusAC_05_Full_trails.tex
eliminated most warnings
warnings
changed camera-ready version into full eprint version

@@
 
-\section{Trail for Full \MORUS}
+\section{Trail for Full \texorpdfstring{\MORUS}{MORUS}}
 \label{sec/fulltrails}
@@
 \end{align*}
-We refer the reader to the full version of this paper \cite{fullversion} for the corresponding linear combination on \MORUS[640].
+We refer the reader to
+\Cref{sec:traileq}
+%the full version of this paper \cite{fullversion}
+for the corresponding linear combination on \MORUS[640].
 
@@
   \centering
-  \begin{subfigure}{.4\textwidth}
+  \begin{subfigure}{.45\textwidth}
   \centering
@@
   \qquad
-  \begin{subfigure}{.4\textwidth}
+  \begin{subfigure}{.45\textwidth}
   \centering
@@
 Therefore, when computing the full trail equation on \MORUS[640], we get that the weight of the trail is 73 (cf.
-the full version of this paper for the full trail equation for \MORUS[640]).
+\Cref{sec:traileq}).
 

### 2018-10-01 AC2018_camera_ready/morusAC_07_IniFin.tex
warnings
changed camera-ready version into full eprint version

@@
 
-\section{Analysis on Initialization and Finalization of Reduced \MORUS}
+\section{Analysis on Initialization and Finalization of Reduced \texorpdfstring{\MORUS}{MORUS}}
 \label{sec/IniFin}
@@
 \item $F_1^{-1}$ is computed $2^{|G_1|-|G_2|}$ times and for each result we check the match with any entry in $T$.
-\item There are $2^{|G_0|-|G_2| + |G_1|-|G_2|}$ combinations, and the number of valid matches reduces to $2^{|G_0|-|G_2| + |G_1|-|G_2| - x}$ after matching the $x$ bits.
+\item The number of possible combinations is $2^{|G_0|-|G_2| + |G_1|-|G_2|}$, and the number of valid matches reduces to $2^{|G_0|-|G_2| + |G_1|-|G_2| - x}$ after matching the $x$ bits.
 \item Check the correctness of the guess by using one plaintext-ciphertext pair.

### 2018-11-27 slides-asiacrypt/tikz/trails-nopause.tex
working on late slides

\newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
\newcommand{\C}[1]{-.1,-.25-1.5*#1}
\renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}

% \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
\begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
  % \begin{scope}
    \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
    \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S_{\r}$}; }
    \draw (\C{0}|-\S{0}{0}) node[above] {$C$};

    \draw[beta, thick]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
                 (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
                 (\M{1}{0}) node[right] {$\beta_0$};
    \draw[alpha, thick] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
                 (\M{0}{0}) node[right] {$\alpha_{27}$};

     \draw[gamma, thick] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
                  (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
                  (\M{1}{2}) node[right] {$\gamma_0$};

    \draw[beta, thick]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
                 (\C{2}) -| (\S{2}{1}) node[below left] {31\!\!}
                 (\M{2}{0}) node[right] {$\beta_{31,}$};

    \draw[alpha, thick] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!}
                (\M{1}{1}) node[right] {$\alpha_{26,}$};

    \draw[delta, thick] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
                 (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
                 (\M{1}{3}) node[right] {$\delta_0$};

    \draw[gamma, thick] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
                 ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
                 (\M{2}{2}) node[right] {$\gamma_{13}$};

    \draw[beta, thick]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
                 ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
    \draw[beta, thick]  (\M{2}{0}) + (0.11,-0.02) node[right] {$_{, 13}$};

    \draw[alpha, thick] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
                 (\M{1}{1}) + (0.12,0) node[right] {$_{, 8}$};

    \draw[beta, thick]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
                 (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
                 (\M{3}{0}) node[right] {$\beta_{12}$};

    \draw[alpha, thick] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
                 (\M{2}{1}) node[right] {$\alpha_7$};
    \draw (\S{4}{2}) node[above] {\large  $\chi_1$: weight 7 (not 11)};

    \draw (\S{4}{1}) + (0,-0.25) node [below] {\large $C^0_{27} \oplus C^1_{0} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{7} \oplus C^2_{13} \oplus C^2_{31} \oplus C^3_{12} \to S^2_{2,0}$};

  \begin{scope}[xshift=2.5cm]
    \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
    \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S_{\r}$}; }
    \draw (\C{0}|-\S{1}{0}) node[above] {$C$};

    \draw[epsil, thick] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
                 (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
                 (\M{2}{3}) node[right] {$\varepsilon_0$};
    \draw[delta, thick] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
                 (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
                 (\M{2}{4}) node[right] {$\delta_7$};

    \draw[gamma, thick] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
                 (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
                 (\M{2}{2}) node[right] {$\gamma_7$};

    \draw[gamma, thick] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
                 ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
                 (\M{3}{2}) node[right] {$\gamma_{20}$};

    \draw[beta, thick]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
                (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
                (\M{2}{0}) node[right] {$\beta_7$};

    \draw[beta, thick]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
                 ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
                 (\M{3}{0}) node[right] {$\beta_{20,6}$};
    \draw[beta, thick]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
                 ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};

    \draw[beta, thick]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
                (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
                (\M{4}{0}) node[right] {$\beta_{19}$};

    \draw[alpha, thick] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
                 (\M{1}{0}) node[right] {$\alpha_{2}$};

    \draw[alpha, thick] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
                 (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
    \draw[alpha, thick] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
    \draw[alpha, thick] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};

    \draw[alpha, thick] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
                 (\M{3}{1}) node[right] {$\alpha_{14}$};

    \draw (\S{5}{2}) node[above] {\large $\chi_2$: weight 9 (not 13)};

    \draw (\S{1}{4}) + (0,0.5)  node [above] {\large $C^1_{2} \oplus C^2_{1} \oplus C^2_{7} \oplus C^2_{15} \oplus C^2_{27} \oplus C^3_{6} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19} \to  S^2_{2,0}$};
    % \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
    % \draw (\S{0}{2}) node[below] {\normalsize \MiniMORUS[640]};
  \end{scope}
\end{tikzpicture}

### 2018-11-29 slides-asiacrypt/tikz/VerticalMorus.tex
marias diagrams

\vspace{0.1cm}
\begin{tikzpicture}[xscale=1.2,yscale=1.2]
  \pgfmathsetmacro{\roundsep}{0.80}

  \foreach \w in {-1,...,5} {
    \foreach \r in {-1,...,5} {
      \coordinate (W\r\w) at (\w,-\r*\roundsep-.125);
    }
    \coordinate (W-2\w) at (\w,--2.375*\roundsep-.125);
  }
  \foreach \w in {0,...,4} {
    %\draw (W-2\w) -- (W4\w);
    \node[inner sep=1pt] (S\w) at ($(W0\w) +(0,.1+2)$) {$S_\w$};
  }
  \draw (W-1-1) node[inner sep=1pt] {\phantom{$M$}};
  \draw (W-1-1) node[above right,gray,box,minimum height=4.8ex,yshift=1ex,xshift=-.5ex] (filt) {\Filt};
  \foreach \w in {0,...,3} {
    \draw (W-2\w|-filt.south) +(0,.5ex+\w ex) coordinate[tee, name=t\w];
    \draw[next] (t\w) -- (t\w-|filt.east);
  }
  \draw (W22) node[gray,box,minimum width=5cm,minimum height=5cm,yshift=3ex] (upd) {\StateUpdate};
  \foreach \w in {0,...,4} {
    \draw[next] (W-2\w) -- (W-2\w|-upd.north);
    \draw[next] (W5\w|-upd.south) -- (W5\w);
  }
  \draw[next] (filt.south) -- ++(0,-2ex) node[below] {out};
  \draw[prev] (upd.west) -- (upd-|filt) -- ++(0,2ex) node[above] {in};
\end{tikzpicture}

### 2018-11-30 slides-asiacrypt/tikz/VerticalMorus.tex
finalize slides

@@
     }
-    \coordinate (W-2\w) at (\w,--2.375*\roundsep-.125);
+    \coordinate (W-2\w) at (\w,--2.375*\roundsep-.425);
   }
@@
     %\draw (W-2\w) -- (W4\w);
-    \node[inner sep=1pt] (S\w) at ($(W0\w) +(0,.1+2)$) {$S_\w$};
+    \node[inner sep=1pt] (S\w) at ($(W0\w) +(0,.1+1.7)$) {$S_\w$};
   }
   \draw (W-1-1) node[inner sep=1pt] {\phantom{$M$}};
-  \draw (W-1-1) node[above right,gray,box,minimum height=4.8ex,yshift=1ex,xshift=-.5ex] (filt) {\Filt};
+  \draw (W-1-1) node[above right,gray,box,minimum height=4.8ex,yshift=1.1ex,xshift=-.5ex] (filt) {\Filt};
   \foreach \w in {0,...,3} {
-    \draw (W-2\w|-filt.south) +(0,.5ex+\w ex) coordinate[tee, name=t\w];
+    \draw (W-2\w|-filt.south) +(0,.85ex+.7*\w ex) coordinate[tee, name=t\w];
     \draw[next] (t\w) -- (t\w-|filt.east);
